### 2024-05-27 Manuscript_ScoreReLiability.qmd
Initial commit

---
title: "Manuscript"
format: html
editor: visual
---

In light of discussions surrounding the replicability of psychological results, meta-analytic heterogeneity is understood as a crucial parameter that may inform some aspects of (non-)replicability. Heterogeneity in itself helps understand why some replications of a single phenomenon may be successful, while other are not. If heterogeneity grows, meaning the phenomenon’s effect size varies more strongly for unexplained reasons, the probability of observing an effect size around zero or even in the negative space grows larger as well. If we know the true size and heterogeneity of a phenomenon’s effect size, we would theoretically be able to establish a baseline of expected replication rate (reference).

Similarly, it has been argued that heterogeneity in effect sizes is an indicator of the theory’s “completeness” surrounding the phenomenon. Linden and Hönekopp argue that “low (as opposed to high) heterogeneity reflects a more advanced understanding of the subject matter being studied” (2021, p.2). Similarly, Schuetze and von Hippel (2023) argue that heterogeneity in effects is an indicator of a vague, poorly specified theory. With these perspectives, heterogeneity in effect sizes can be understood as an indicator of the extent, to which the science surrounding a phenomenon has evolved to a sufficient understanding of that phenomenon.

Re-analyses of multi-site direct replications have demonstrated that heterogeneity tends to be the norm, if a psychological effect overall manages to replicate. Large-scale attempts of direct replications, using identical protocols, such as the Many Labs studies or Registered Replication Reports (references), for the first time allow researchers to estimate heterogeneity undistorted by typical experimental factors. In re-analyses of these studies, Olsson-Collentine et al. (2020) identify a strong correlation between a studies effect size and its heterogeneity. Similarly, van Erp et al. (2017) or Stanley et al. (2018) estimate strong degrees of heterogeneity across psychological replications in general. In a separate re-analysis of large scale direct replications, Renkewitz et al. (2024) identify substantial heterogeneity in almost all projects where a non-zero effect could be identified. This aligns with the correlation found by Olsson-Collentine et al. (2020).

In both the initial reports of large scale replication attempts, as well as the re-analyses by Olsson-Collentine et al. (2020), standardized effect sizes (ES), such as Cohen’s d or Hedge’s g were used. For the remainder of the article, Cohen’s d, defined in equation 1, will be used as an exemplary estimate of ES, as its used across a wide range of contexts and well understood by a broad audience.

$$d = \frac{MD}{\sigma_X}$$

Here, $MD$ refers to the difference between the two groups of interest, while $\sigma_X$ is the total pooled standard deviation.

As is widely known, score reliability affects such ES. In the context of classical test (CTT) theory score reliability is defined as the ratio of true to observed score variance, as defined in equation 2.

$$\rho_{XX’} = \frac{\sigma^2_T}{\sigma^2_X}$$

In this equation $\sigma^2_T$ refers to the true variance in the sense of CTT, meaning the actual variance of the variable, undistorted by measurement error. In the same sense, $\sigma^2_X$ refers to the total variance of scores, including both the true variance and the random error variance. Lower score reliability in a data-set leads to a smaller ES, opposed to a similar data-set with higher score reliability (references). In the meta-analytic context, as score reliability is an aspect of a measuring instrument applied to a population, score reliability may not be identical across replications. If score reliability varies across replications, meaning it also carries heterogeneity, it clearly should affect heterogeneity in observed ES as well.

Previous discussions of heterogeneity in score reliability have exclusively discussed it as a parameter that inflates heterogeneity in observed ES, implying that, if score reliability was perfect across all replications, the actual heterogeneity would have been lower. In their discussion, Wiernik and Dahlke claim that „*Measurement error variance will impact the results of meta-analyses in three ways: by (a) biasing the mean effect size toward zero, (b) inflating effect-size heterogeneity and confounding moderator effects, and (c) confounding publication-bias and sensitivity analyses*” (p. 3, 2020). Additionally, more clearly, they state that “*If the studies included in a meta-analysis differ in their measures’ reliabilities, heterogeneity estimates will be artifactually inflated, erroneously suggesting larger potential moderator effects*” (p. 4, Wiernik & Dahlke, 2020). They base their claims largely on he work done by Hunter and Schmidt (2014), who claim that “*Variation in reliability across studies causes variation in the observed effect sizes above and beyond that produced by sampling error.*” (p. 302). Overall, both references imply that differences in score (un)reliability inevitably lead to an inflated heterogeneity in ES.

However, if information concerning score reliability is available, it is possible to correct the individual ES for its unreliability. This process is also known as attenuation correction and, while not without its criticisms (reference), is a widespread practice in Psychology (reference). If the claims above, made by Wiernik and Dahlke or Hunter and Schmidt, were to be true, it would imply that corrections of individual observed ESs for their unreliability, any subsequently performed meta-analysis should generate lower estimates of heterogeneity. Hunter & Schmidt even claim to have derived an equation which demonstrates how heterogeneity in score reliability inflates heterogeneity in ESs. “*If the level of reliability is independent of the true effect size across studies, then, to a close approximation*:” (p. 309)

$$Var(\delta_0) = [E(a)]^2 Var(\delta) + [E(\delta)]^2 Var(a)$$

Here, $\delta_0$ refers to the “true” ES, undistorted by sampling error but not corrected for measuring error, $\delta$ refers to the “true” ES, undistorted by sampling error and measurement error and $a$ refers to score reliability in Schmidt and Hunter's notation.

However, we propose that this understanding of how differences in score reliability affect the heterogeneity of observed ESs is incomplete. Hunter and Schmidt’s implication, that heterogeneity in score reliability inflates ES heterogeneity does not follow from equation 1. If equation 1 were to be true, it would imply that heterogeneity in $\delta_0$ is essentially a function where heterogeneity and mean value of $\delta$ are weighted by mean score reliability $a$ and its heterogeneity. In this scenario, it is easy to construct cases where heterogeneity in $\delta_0$ is inflated or deflated by heterogeneity in $a$. In table 1, two such cases are demonstrated. All else is held equal, only the mean score reliability $E[a]$ is changed from .7 to .8. True heterogeneity $Var[a]$ is .2, at $E[a] = .7$ this leads to a smaller $Var[\delta_0]$, while at $E[a] = .7$ this leads to a larger $Var[\delta_0]$.

Parameter combinations of ES and score reliability, according to Hunter & Schmidt (2014)

```{r, echo = FALSE}

m <- matrix(c(1,	.2,	.7,	.1,	.198,
              1,	.2,	.8,	.1,	.228),
            nrow = 2)

df <- as.data.frame(m)

names(df) <- c("$E[\\delta]$",
               "$Var[\\delta]$",
               "$E[a]$",
               "$Var[a]$",
               "$Var[\\delta\\_0]$")

knitr::kable(df)

```

As Hunter & Schmidt point out, equation 1 is only valid “*If the level of reliability is independent of the true effect size across studies \[…\]”* (2014, p. 309).

However, both parameters, ES and score reliability in their notation, can not be independent parameters by definition. ES is the “true” standardized mean difference, unaffected by sampling or measurement error. The ES is standardized using the total standard deviation, which is inflated by measurement error. Equation 1, demonstrating how Cohen’s d is computed demonstrates how the mean difference is standardized in this case. The ES on the other hand is standardized using the true standard deviation in the sense of CTT.

*\$\$* *= \\frac{MD}{\\sigma_T}\$\$*

In equation 4, is the true standard deviation, the square root of the true score variance found in equation 2. The fact that the true standard deviation parameter is found both in equation 4 and equation 2 demonstrates that ES and score reliability can not be independent variables. Therefore, the basic premise of equation 3 is not fulfilled, as the assumption of variable independence could not be fulfilled.

~~Hunter & Schmidt’s discussion of how score reliability affects heterogeneity in ESs is fatally flawed \[issues with the assumption underlying the equation\].~~

Alternatively, assuming that both the mean differences and pooled standard deviations are no constant values across replications, meaning both carry non-zero heterogeneity, we can describe observed ESs as a random variable following a ratio distribution. Such a ratio distribution is the probability distribution, constructed by dividing one random variable by a second variable: \$Z = \\frac{X}{Y}\$ (reference). If the distribution of the components is known, first order taylor approximation may be used to generate estimates of mean and variance of the ratio variable (reference). Using some additional assumptions, we use the Taylor-estimator of the ratio’s variance to demonstrate how differences in score reliability may affect heterogeneity in ES.

\$\$var\[Z\] \\approx \\frac{var\[X\]}{E\[Y\]\^2} - \\frac{2E\[X\]}{E\[Y\]\^3} cov\[X,Y\] + \\frac{E\[X\]\^2}{E\[Y\]\^4} var\[Y\]\$\$

Assuming that the random variable are uncorrelated, simplifies the equation to

\$\$var\[Z\] \\approx \\frac{var\[X\]}{E\[Y\]\^2} + \\frac{E\[X\]\^2}{E\[Y\]\^4} var\[Y\]\$\$

Substituting for ES, unstandardized mean difference MD and pooled standard deviation SD leads to

\$\$var\[ES\] \\approx \\frac{var\[MD\]}{E\[SD\]\^2} + \\frac{E\[MD\]\^2}{E\[SD\]\^4} var\[SD\]\$\$

Correcting ES for score reliability is technically done by correcting the pooled standard deviation: \$SD_c = \\sqrt{a V}\$ . Since is always between zero and one, in the case of imperfect score reliability, SD_c will always be smaller than total SD. Additionally, if the variance in SD was introduced by differences in score reliability, these would be removed by the attenuation correction. However, there are two parameters that may introduce variance in SD across replications.

Firstly, as discussed, differences in score reliability may introduce variation to SD. Secondly however, the underlying latent variable we attempted to measure may have been not identically distributed across replications. If the populations where the replications are collected from are differently spread out in terms of the variable of interest, true differences in SD would remain. Re-analyses of the ManyLabs and Registered Replication Reports have shown that such differences in true variance across replications are the norm, not the exception. In that case, some variance in SD_c would remain, albeit smaller than the variance in SD. We reformulate equation X in terms of SD_c

\$\$var\[ES_c\] \\approx \\frac{var\[MD\]}{E\[SD_c\]\^2} + \\frac{E\[MD\]\^2}{E\[SD_c\]\^4} var\[SD_c\]\$\$

The implication of this equation is, again, not straightforward. The expected value of SD_c is guaranteed to be smaller than the expected value of uncorrected SD. Similarly, the variance in SD_c is guaranteed to be smaller than the uncorrected variance in SD. The expected value is always placed in the denominator of the function, while the variance is placed in the numerator of the second term in equation X+1. This implies that the variance in uncorrected ES can only be larger than the variance in ES_c, if the reduction in mean SD_c due to the attenuation correction increases the total equation result faster than the reduction in variance of SD_c decreases the total equation result. To illustrate this point, we consider some parameter combinations in table 2.

*Table 2*

Parameter combinations, leading to varying degrees of heterogeneity in ES_c, based on the Taylor approximation to the variance of a ratio distribution

<table>
<tbody>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>E[MD]</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Var[MD}</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>E[SD]</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Var[SD]</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>E[SD_c]</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Var[SD_c]</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Var[ES]</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Var[ES_c]</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.8</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.1^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.08</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.087</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.8</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.05^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.08</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.069</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.8</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>0</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.08</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.063</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.9</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.1^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.08</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.064</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.7</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.1^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.08</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.124</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.2^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.5</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.1^2</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.08</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.32</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p> </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
</tbody>
</table>

 Table 2 demonstrates that in scenarios, where either score reliability was very high to begin with (.9 in this case), leading to very little change in E\[SD\] to E\[SD_c\] (row 5), or where the remaining variance in SD_c was comparably low (rows 3 and 4), the attenuation correction led to a reduction in ES heterogeneity. In most other scenarios, where score reliability was lower (.7 and .5 in rows 6 and 7) or remaining variance in SD_c sufficiently large (row 2), the attenuation correction actually increased the heterogeneity in ES. Note that in the table, we did not manipulate expected value or variance of MD or SD.

\[implications of this distribution concerning Var\[ES\] and how that relates to Var\[rel\]\]

Reformulating the approximation of variance in a ratio variable demonstrates how correcting for differences in reliability might affect the heterogeneity in ES.

Using analytical arguments from a reformulation of equation X we have demonstrated that differences in measuring quality often do not inflate estimates of heterogeneity. Instead, assuming that MD and SD are uncorrelated, we argue that in many circumstances differences in measuring quality deflate heterogeneity in standardized mean differences. This could be understood as a form of masking, where differences in measuring quality lead to a seemingly lower heterogeneity than we would have observed without any measurement error. Clearly, these arguments are in stark contrast with claims made by Hunter & Schmidt or Wiernik & Dahlke. To further illustrate this point, we turn towards an exemplary data-set, demonstrating that we can observe the deflation in heterogeneity in empirical data.

**Part C**

An exemplary data-set needs to fulfil the following conditions: i) it needs to contain several (direct/close or conceptual) replications of something that can be described as a two-group experimental effect; ii) the dependent variable needs to be measured using an identical scale across administrations, with multiple indicators/items/responses, so that estimates of score reliability can be derived; iii) there needs to be statistically significant heterogeneity in the standardized effect sizes; and iv) there needs to be statistically significant heterogeneity in the score reliability across administrations.

 While multi-site direct replication attempts, such as the Many Labs studies or Registered Replication Reports initially spring to mind, none of the studies replicated in these projects fulfil all four conditions. However, the Psychological Science Accelerator (PSA), even though their projects don’t have meta-analytic goals, provide data that may be transformed to fulfil all four conditions sufficiently. Specifically, we examined the data from PSACR002, one of the PSA’s studies concerning interventions and responses to the COVID-19 pandemic (reference). In the PSACR002 study, the authors assessed the effectiveness of reappraisal interventions on negative and positive emotion in participants grouped into four categories (reconstrual intervention, repurposing intervention, passive control, active control). The four groups may be collapsed into a single treatment group (reconstrual and repurposing) and a single control group (passive and active control). In total, Wang et al. (2021) assessed 21,644 individuals from 87 countries/regions. These countries/regions served as the study-level grouping variable for our meta-analyses.

 Wang et al. (2021) used several measures which may be promising for this study. Out of four measures they use to assess negative and positive emotions, three should fit conditions ii), ii) and iv); positive and negative emotional responses to ten photographs shown in the experimental trial, positive and negative state emotion after viewing the ten photographs and positive and negative anticipated emotions following the study. Over the following paragraphs, we will briefly describe the study design and measures used that are of interest to us. Details regarding the protocol, stimuli or additional measures used may be found at \[…\].

 After randomization to one of four conditions, participants in the two intervention conditions received instructions on how to “change one’s thinking to change one’s emotions”, with more detail pertaining to the respective intervention. In the active control condition, participants were asked to reflect on their emotions, while in the passive control condition participants were asked to react naturally. Afterwards, the participants completed practice trials where they viewed two negative photographs in relation to the COVID-19 pandemic (e.g. people in hazmat suits, exhausted doctors), rated those photos on a Likert-scale from 1 to 5 on two questions: “how negative did the photo make you feel” and “how positive did the photo make you feel” (anchors: 1 = not at all, 5 = extremely). Depending on the condition the individual was randomized to, reminders of the strategy to be used were displayed.

 In the experimental trials, the participants viewed 10, similarly negative, photographs of the COVID-19 pandemic and rated their emotions on each photograph using Likert-style responses to the positive/negative questions. After the experimental trials, individuals responded to five negative items from the Differential Emotions Scale on fear, anger, sadness, distrust, and stress. Similarly, they responded to five positive items from the same scale, on hope, gratitude, love, inspiration, and serenity. Averages of the five positive and negative items respectively were used as measures of state emotions after the experimental trials. Similarly, negative and positive anticipated emotions were assessed, by prefacing a question of “In the next week, to what extent, if at all, do you think you will feel each of the following?” and following up with the same items used to assess state emotions. All items or the measures used may be found in the appendix XXX, or at \[url…\].

*Table 3*

Meta-analytic results concerning uncorrected and corrected standardized effect sizes across 3x2 measures of emotional response

<table>
<tbody>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>Measure</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Valence</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>ES</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Tau ES</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>ES_c</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Tau ES_c</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>Photograph   Response</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Positive</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1.065   (.065)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.377 (sq   .036)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>1.219   (.080)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.468 (sq   .056)</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>Negative</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>-.663   (.040)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.220 (sq .014)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>-.747   (.047)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.260 (sq   .019)</p>
<p>   </p></td>
<td><p>  </p></td>
<td></td>
<td></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>Emotional   State</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Positive</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.213   (.018)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.052 (sq   .003)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.234   (.020)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.054 (sq   .003)</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>Negative</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>-.203   (.020)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.071 (sq   .003)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>-.221   (.022)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.078   (.004)</p>
<p>   </p></td>
<td><p>  </p></td>
<td></td>
<td></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>Anticipated   Emotion</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>Positive</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.159   (.015)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>0 (sq   .001)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.172   (.016)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>0 (.002)</p>
<p>   </p></td>
<td><p>  </p></td>
</tr>
<tr class="odd">
<td><p>  </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
<tr class="even">
<td><p>   </p></td>
<td><p>   </p>
<p>Negative</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>-.146   (.020)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.073 (sq   .003)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>-.145   (.022)</p>
<p>   </p></td>
<td><p>   </p></td>
<td><p>   </p>
<p>.079   (.004)</p>
<p>   </p></td>
<td><p>  </p></td>
<td></td>
<td></td>
</tr>
<tr class="odd">
<td><p> </p></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
<td></td>
</tr>
</tbody>
</table>

In table 3 the results of meta-analyses on uncorrected and correct standardized effect sizes are summarised. In line with previous analytical arguments and the exemplary table of conditions, we observe that for effects of substantial size, meaning the absolute values of uncorrected ES for both photograph response measures, the heterogeneity in ES is increased by about 25% for the positive measures and about 20% for the negative measures after correcting for attenuation. For all other measures, where the difference in treatment and control groups is less strongly pronounced in terms of mean ES, heterogeneity also increases after correcting for attenuation, but a lot less strongly.

 In figure 1 a), the meta-analytical results for the positive photograph response are visualised in a forest plot. Grey dots represent the uncorrected ESs for each country, with accompanying confidence interval represented by the bars surrounding the dots. The overlying black dots on the other hand represent the corrected ESs, with additional confidence intervals.  The position of the grey and black diamonds on the bottom display the meta-analytical estimate of mean ES, uncorrected (grey) and corrected (black), whereas the width of the diamonds represent the respective confidence interval. The forest plot shows that, as expected, the attenuation correction leads to larger ES and larger individual confidence intervals. Similarly, the meta-analytic estimate of mean ES is also larger, if score unreliability was corrected for, just like its accompanying confidence interval.

*Figure 1*

Forest plot and visualised underlying density for positive measures of Photographs

*Note:* In a), grey dots represent individual uncorrected ES with grey bars representing the accompanying 95%-confidence intervals; black dots represent individual corrected ES wit black bars representing accompanying 95%-confidence intervals. In b), the grey density in the background represents the implied underlying distribution of uncorrected ES, while the black density in the foreground represents the implied underlying distribution of corrected ES.

Part b) in figure 2 visualises the implied underlying “true” distribution of ESs, free of sampling error. Essentially, the meta-analytic estimates of mean ES and its heterogeneity are used to construct a density for a normal distribution with these specific parameters. In part b), again, the grey density represents the distribution of uncorrected ES, while the black density represents the distribution of ES corrected for score unreliability. The higher spike in the grey demonstrates that the heterogeneity in uncorrected ES is smaller than the heterogeneity for corrected ES. Generally, the distribution of corrected ES is positioned further to the right, as the average ES is larger, as can also be seen in the forest plot. Additionally, as the density is more “flat” and with a lower spike, it shows that the heterogeneity in corrected ES is larger as well.

**Discussion**

Using both analytical arguments as well as an exemplary empirical dataset, we have demonstrated that differences in score reliability do not necessarily inflate heterogeneity in effect sizes. Inferences derived from the equation supplied in Hunter and Schmidt (ref) also do not warrant such a conclusion. Alternatively, describing standardised effect sizes in terms of a ratio distribution and approximating its variance, we have demonstrated that differences in score reliability may very well inflate or deflate variance in ES. In data supplied by the PSA (ref.) we have demonstrated that, across all measures, correcting observed ES for (un)reliability led to larger heterogeneity in ES than before correction.

 These results fit in well with work recently published/preprinted by Olsson-Collentine et al. (2023). In a large simulation scheme, they demonstrate that differences in score reliability across administrations initially deflate heterogeneity in correlations. Only as the true heterogeneity in correlations grows larger, score reliability differences actually inflate heterogeneity as predicted by Wiernik and Dahlke (xxx). While correlations and standardized effect sizes are not identical, the way score (un)reliability affects these parameters is highly similar. Both correlations and ES are deflated by (un)reliability, implying that an attenuation correction increases these parameters.

 Heterogeneity in ES even in direct replications, where experimental factors are held as constant as possible, is substantial (Renkewitz, Fünderich, & Beinhauer, 2024). Previous assumptions on the role of score reliability differences, as discussed in Wiernik & Dahlke (xxx), has been that they “inflate” heterogeneity. If these previous assumptions were true, it would imply that the true heterogeneity in such projects may be considerably smaller and less alarming.

However, so far we have demonstrated that in several scenarios, differences in score reliability may in fact be masking, or “deflating” ES heterogeneity. This implies that the already substantial heterogeneity across direct replications may very well be even larger, if ES were corrected for (un)reliability. Understanding heterogeneity as an extent to which a theory or phenomenon is understood, heterogeneity that was actually deflated implies that the theory or phenomenon is understood even worse than initially assumed. If the score reliability varies across administrations because the measuring quality is not identical across populations (Beinhauer, Fünderich & Renkewitz, 2024), these differences additionally reflect the poor understanding of phenomenon or theory.

\+ and – of this study

 The empirical arguments presented are based on a single study. While, based on the combination of analytical and empirical arguments, we are convinced that these differences in score reliability across replications oftentimes mask true heterogeneity, whether these results actually generalize beyond this data-set remains to be seen. Unfortunately, the behavioural sciences are in dire need of open-data that resembles multi-site replications. As the majority of results discussed over the last years (e.g. ManyLabs or Registered Replication Reports) employ single-indicator scales as dependent measures, score reliability can not be easily estimated in order to replicate our analyses.

How to go from here

 The analytical arguments presented in equation X could be improved by properly disentangling the terms concerning variance of SD and expected value of SD. As it stands now, the polynomial structure of the equation makes singling out an individual term rather difficult. However, in order to more easily understand in what way differences in score reliability lead to an inflation or deflation of heterogeneity, solving the equation for both the variance and expected value of SD would be highly beneficial.

### 2024-05-27 E1_Data-Analysis.R

@@ -1,0 +1,540 @@
+## Score ReLiability
+
+# E1
+
+# Clean & extract data
+
+
+# library loading and installing as necessary
+
+
+# relevant libraries required for this script
+packages <- c("tidyverse", "here", "metafor")
+
+# check, whether library already installed or not - install and load as needed:
+apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
+  
+  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system
+  
+  if(pkg_avail){
+    require(x, character.only = TRUE)             # load the library, if already installed
+    
+  }else{
+    install.packages(x)                           # install the library, if missing
+    require(x, character.only = TRUE)             # load after installation
+  }
+})
+
+
+
+# load initial data-file
+# "wide, exclusions" - should be the correct, cleaned data-file
+psacr002_dat <- read.csv(here("Data/Downloaded/data_wide_exclusion_12.23.2020.csv"))
+
+# identify variables which contain responses on stimuli (positive and negative responses separate)
+pos_items <- names(psacr002_dat)[grep("positiveemotion_photo", names(psacr002_dat))]
+neg_items <- names(psacr002_dat)[grep("negativeemotion_photo", names(psacr002_dat))]
+
+pos_state_items <- names(psacr002_dat)[grep("C2_positiveemotion_state.*[^0-9]$", names(psacr002_dat))]
+neg_state_items <- names(psacr002_dat)[grep("C2_negativeemotion_state.*[^0-9]$", names(psacr002_dat))]
+
+pos_antc_items <- names(psacr002_dat)[grep("C2_positiveemotion_anticipated.*[^0-9]$", names(psacr002_dat))]
+neg_antc_items <- names(psacr002_dat)[grep("C2_negativeemotion_anticipated.*[^0-9]$", names(psacr002_dat))]
+
+# identify which countries contain sufficient (more than 20) respondents
+samples <- as.data.frame(table(psacr002_dat$country))$Var1[as.data.frame(table(psacr002_dat$country))$Freq >= 200]
+
+
+#### PHOTOGRAPHS
+
+# formulate data-file containing responses on positive photo items, information on the grouping
+#  variable, and the respondent's country
+psacr002_pos <- psacr002_dat %>% 
[…]

### 2024-05-27 E2_Graphics.R

@@ -1,0 +1,282 @@
+## Score ReLiability
+
+# E2
+
+# Graphics
+
+
+
+# library loading and installing as necessary
+
+
+# relevant libraries required for this script
+packages <- c("tidyverse", "here", "metafor", "ggpubr", "gridExtra")
+
+# check, whether library already installed or not - install and load as needed:
+apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
+  
+  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system
+  
+  if(pkg_avail){
+    require(x, character.only = TRUE)             # load the library, if already installed
+    
+  }else{
+    install.packages(x)                           # install the library, if missing
+    require(x, character.only = TRUE)             # load after installation
+  }
+})
+
+
+##### PHOTOGRAPHS
+
+# Read data-files containing estimates of SMDs and Standard Errors
+pos_SMDs.df <- read.csv(file = here("Data/Processed/SMDs_positive.csv"))
+neg_SMDs.df <- read.csv(file = here("Data/Processed/SMDs_negative.csv"))
+
+
+# fit a meta-analytic model to estimates of uncorrected Cohen's d, for positiveness-ratings
+pos_SMDs_rma_raw <- metafor::rma(yi = d_raw, 
+                                 sei = SE_d_raw, 
+                                 data = pos_SMDs.df,
+                                 measure = "GEN",
+                                 method = "REML")
+
+# fit a meta-analytic model to estimates of corrected Cohen's d, for positiveness-ratings
+pos_SMDs_rma_cor <- metafor::rma(yi = d_cor, 
+                                 sei = SE_d_cor, 
+                                 data = pos_SMDs.df,
+                                 measure = "GEN",
+                                 method = "REML")
+
+
+# fit a meta-analytic model to estimates of uncorrected Cohen's d, for negativeness-ratings
+neg_SMDs_rma_raw <- metafor::rma(yi = d_raw, 
+                                 sei = SE_d_raw, 
+                                 data = neg_SMDs.df,
+                                 measure = "GEN",
+                                 method = "REML")
+
+
[…]

### 2024-05-28 E0_Data-Download.R
add download script

@@ -1,0 +1,22 @@
+packages <- c("magrittr", "osfr", "here")
+
+# check, whether library already installed or not - install and load as needed:
+apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
+  
+  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system
+  
+  if(pkg_avail){
+    require(x, character.only = TRUE)             # load the library, if already installed
+    
+  }else{
+    install.packages(x)                           # install the library, if missing
+    require(x, character.only = TRUE)             # load after installation
+  }
+})
+
+
+
+
+# use osfr-package to download PSACR002-data, on IPD-level, pre-cleaned
+osfr::osf_retrieve_file("https://osf.io/rwksu") %>%
+  osfr::osf_download(path = here("Data/Downloaded/"))

### 2024-05-28 Manuscript_ScoreReLiability.qmd
text update add gitkeep

@@
 ---
-title: "Manuscript"
+title: "Score ReLiability? Exploring the role of score reliability for effect size heterogeneity"
 format: html
@@
 ---
+
+```{r, include = FALSE}
+# library loading and installing as necessary
+
+# relevant libraries required for this script
+packages <- c("ggplot2", "here", "metafor", "gridExtra", "knitr", "magrittr", "dplyr")
+
+# check, whether library already installed or not - install and load as needed:
+apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
+
+  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system
+
+  if(pkg_avail){
+    require(x, character.only = TRUE)             # load the library, if already installed
+
+  }else{
+    install.packages(x)                           # install the library, if missing
+    require(x, character.only = TRUE)             # load after installation
+  }
+})
+
+```
 
@@
 
+## Analytical Arguments
+
 However, we propose that this understanding of how differences in score reliability affect the heterogeneity of observed ESs is incomplete.
@@
 
-m <- matrix(c(1,	.2,	.7,	.1,	.198,
-              1,	.2,	.8,	.1,	.228),
-            nrow = 2)
+m <- matrix(c(1,	.2,	.7,	.1,
+              1,	.2,	.8,	.1),
+            nrow = 2, byrow = T)
 
@@
                "$E[a]$",
-               "$Var[a]$",
-               "$Var[\\delta\\_0]$")
+               "$Var[a]$")
+
+df$'$Var[\\delta\\_0]$' <- (df[,3]^2) * df[,2] + (df[,1]^2) * df[,4]
 
@@
 
-However, both parameters, ES and score reliability in their notation, can not be independent parameters by definition.
-ES is the “true” standardized mean difference, unaffected by sampling or measurement error.
-The ES is standardized using the total standard deviation, which is inflated by measurement error.
+However, both parameters, ES $\delta$ and score reliability $a$ in original notation can not be independent parameters by definition.
+ES $\delta$ is the “true” standardized mean difference, unaffected by sampling or measurement error.
+The ES $\delta_0$ is standardized using the total standard deviation, which is inflated by measurement error.
 Equation 1, demonstrating how Cohen’s d is computed demonstrates how the mean difference is standardized in this case.
-The ES on the other hand is standardized using the true standard deviation in the sense of CTT.
+The ES $\delta$ on the other hand is standardized using the true standard deviation in the sense of CTT.
 
-*\$\$* *= \\frac{MD}{\\sigma_T}\$\$*
+$$\delta = \frac{\Delta}{\sigma_T}$$
 
-In equation 4, is the true standard deviation, the square root of the true score variance found in equation 2.
-The fact that the true standard deviation parameter is found both in equation 4 and equation 2 demonstrates that ES and score reliability can not be independent variables.
+In equation 4, $\sigma_T$ is the true standard deviation, the square root of the true score variance found in equation 2.
+The fact that the true standard deviation parameter $\sigma_T$ is found both in equation 4 and equation 2 demonstrates that ES $\delta$ and score reliability $a$ can not be independent variables.
 Therefore, the basic premise of equation 3 is not fulfilled, as the assumption of variable independence could not be fulfilled.
 
-~~Hunter & Schmidt’s discussion of how score reliability affects heterogeneity in ESs is fatally flawed \[issues with the assumption underlying the equation\].~~
-
 Alternatively, assuming that both the mean differences and pooled standard deviations are no constant values across replications, meaning both carry non-zero heterogeneity, we can describe observed ESs as a random variable following a ratio distribution.
-Such a ratio distribution is the probability distribution, constructed by dividing one random variable by a second variable: \$Z = \\frac{X}{Y}\$ (reference).
+Such a ratio distribution is the probability distribution, constructed by dividing one random variable by a second variable: $Z = \frac{X}{Y}$ (reference).
 If the distribution of the components is known, first order taylor approximation may be used to generate estimates of mean and variance of the ratio variable (reference).
@@
 
-\$\$var\[Z\] \\approx \\frac{var\[X\]}{E\[Y\]\^2} - \\frac{2E\[X\]}{E\[Y\]\^3} cov\[X,Y\] + \\frac{E\[X\]\^2}{E\[Y\]\^4} var\[Y\]\$\$
+$$var[Z] \approx \frac{var[X]}{E[Y]^2} - \frac{2E[X]}{E[Y]^3} cov[X,Y] + \frac{E[X]^2}{E[Y]^4} var[Y]$$
 
@@
 
-\$\$var\[Z\] \\approx \\frac{var\[X\]}{E\[Y\]\^2} + \\frac{E\[X\]\^2}{E\[Y\]\^4} var\[Y\]\$\$
+$$var[Z] \approx \frac{var[X]}{E[Y]^2} + \frac{E[X]^2}{E[Y]^4} var[Y]$$
 
@@
 
-\$\$var\[ES\] \\approx \\frac{var\[MD\]}{E\[SD\]\^2} + \\frac{E\[MD\]\^2}{E\[SD\]\^4} var\[SD\]\$\$
+$$var[ES] \approx \frac{var[MD]}{E[SD]^2} + \frac{E[MD]^2}{E[SD]^4} var[SD]$$
 
-Correcting ES for score reliability is technically done by correcting the pooled standard deviation: \$SD_c = \\sqrt{a V}\$ .
-Since is always between zero and one, in the case of imperfect score reliability, SD_c will always be smaller than total SD.
-Additionally, if the variance in SD was introduced by differences in score reliability, these would be removed by the attenuation correction.
-However, there are two parameters that may introduce variance in SD across replications.
+Correcting ES for score reliability is technically done by correcting the pooled standard deviation: $SD_c = \sqrt{a V}$ .
+Since $a$ is always between zero and one, in the case of imperfect score reliability, $SD_c$ will always be smaller than total $SD$.
+Additionally, if the variance in $SD$ was introduced by differences in score reliability, these would be removed by the attenuation correction.
+However, there are two parameters that may introduce variance in $SD$ across replications.
 
@@
 Secondly however, the underlying latent variable we attempted to measure may have been not identically distributed across replications.
-If the populations where the replications are collected from are differently spread out in terms of the variable of interest, true differences in SD would remain.
+If the populations where the replications are collected from are differently spread out in terms of the variable of interest, true differences in $SD$ would remain.
 Re-analyses of the ManyLabs and Registered Replication Reports have shown that such differences in true variance across replications are the norm, not the exception.
-In that case, some variance in SD_c would remain, albeit smaller than the variance in SD.
-We reformulate equation X in terms of SD_c
+In that case, some variance in $SD_c$ would remain, albeit smaller than the variance in $SD$.
+We reformulate equation X in terms of $SD_c$
 
-\$\$var\[ES_c\] \\approx \\frac{var\[MD\]}{E\[SD_c\]\^2} + \\frac{E\[MD\]\^2}{E\[SD_c\]\^4} var\[SD_c\]\$\$
+$$var[ES_c] \approx \frac{var[MD]}{E[SD_c]^2} + \frac{E[MD]^2}{E[SD_c]^4} var[SD_c]$$
 
 The implication of this equation is, again, not straightforward.
-The expected value of SD_c is guaranteed to be smaller than the expected value of uncorrected SD.
-Similarly, the variance in SD_c is guaranteed to be smaller than the uncorrected variance in SD.
+The expected value of $SD_c$ is guaranteed to be smaller than the expected value of uncorrected $SD$.
+Similarly, the variance in $SD_c$ is guaranteed to be smaller than the uncorrected variance in $SD$.
 The expected value is always placed in the denominator of the function, while the variance is placed in the numerator of the second term in equation X+1.
-This implies that the variance in uncorrected ES can only be larger than the variance in ES_c, if the reduction in mean SD_c due to the attenuation correction increases the total equation result faster than the reduction in variance of SD_c decreases the total equation result.
+This implies that the variance in uncorrected ES can only be larger than the variance in $ES_c$, if the reduction in mean $SD_c$ due to the attenuation correction increases the total equation result faster than the reduction in variance of $SD_c$ decreases the total equation result.
 To illustrate this point, we consider some parameter combinations in table 2.
@@
 
-<table>
-<tbody>
-<tr class="odd">
-<td><p>  </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="even">
-<td><p>   </p></td>
-<td><p>   </p>
-<p>E[MD]</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Var[MD}</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>E[SD]</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Var[SD]</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>E[SD_c]</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Var[SD_c]</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Var[ES]</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Var[ES_c]</p>
-<p>   </p></td>
-<td><p>  </p></td>
-</tr>
-<tr class="odd">
-<td><p>  </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="even">
-<td><p>   </p></td>
-<td><p>   </p>
-<p>1</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.2^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>1</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.2^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.8</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.1^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.08</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.087</p>
-<p>   </p></td>
-<td><p>  </p></td>
-</tr>
-<tr class="odd">
-<td><p>  </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="even">
-<td><p>   </p></td>
-<td><p>   </p>
-<p>1</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.2^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>1</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.2^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.8</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.05^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.08</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.069</p>
-<p>   </p></td>
-<td><p>  </p></td>
-</tr>
-<tr class="odd">
-<td><p>  </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="even">
-<td><p>   </p></td>
-<td><p>   </p>
-<p>1</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.2^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>1</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.2^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.8</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>0</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.08</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.063</p>
-<p>   </p></td>
-<td><p>  </p></td>
-</tr>
-<tr class="odd">
-<td><p>  </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="even">
-<td><p>   </p></td>
-<td><p>   </p>
-<p>1</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.2^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>1</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.2^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.9</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.1^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.08</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.064</p>
-<p>   </p></td>
-<td><p>  </p></td>
-</tr>
-<tr class="odd">
-<td><p>  </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="even">
-<td><p>   </p></td>
-<td><p>   </p>
-<p>1</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.2^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>1</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.2^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.7</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.1^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.08</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.124</p>
-<p>   </p></td>
-<td><p>  </p></td>
-</tr>
-<tr class="odd">
-<td><p>  </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="even">
-<td><p>   </p></td>
-<td><p>   </p>
-<p>1</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.2^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>1</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.2^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.5</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.1^2</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.08</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.32</p>
-<p>   </p></td>
-<td><p>  </p></td>
-</tr>
-<tr class="odd">
-<td><p> </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-</tbody>
-</table>
+```{r, echo = FALSE}
 
- Table 2 demonstrates that in scenarios, where either score reliability was very high to begin with (.9 in this case), leading to very little change in E\[SD\] to E\[SD_c\] (row 5), or where the remaining variance in SD_c was comparably low (rows 3 and 4), the attenuation correction led to a reduction in ES heterogeneity.
-In most other scenarios, where score reliability was lower (.7 and .5 in rows 6 and 7) or remaining variance in SD_c sufficiently large (row 2), the attenuation correction actually increased the heterogeneity in ES.
+m2 <- matrix(c(1,	".2^2",	1,	".2^2",	.8,	".1^2",
+               1,	".2^2",	1,	".2^2",	.8,	".05^2",
+               1,	".2^2",	1,	".2^2",	.8,	0,
+               1,	".2^2",	1,	".2^2",	.9,	".1^2",
+               1,	".2^2",	1,	".2^2",	.7,	".1^2",
+               1,	".2^2",	1,	".2^2",	.5,	".1^2"),
+            nrow = 6, byrow = T)
+
+m_num <- matrix(c(1,	.2^2,	1,	.2^2,	.8,	.1^2,
+                  1,	.2^2,	1,	.2^2,	.8,	.05^2,
+                  1,	.2^2,	1,	.2^2,	.8,	0,
+                  1,	.2^2,	1,	.2^2,	.9,	.1^2,
+                  1,	.2^2,	1,	.2^2,	.7,	.1^2,
+                  1,	.2^2,	1,	.2^2,	.5,	.1^2),
+                nrow = 6, byrow = T)
+
+df2 <- as.data.frame(m2)
+
+names(df2) <- c("$E[\\Delta]$",
+                "$Var[\\Delta]$",
+                "$E[SD]$",
+                "$Var[SD]$",
+                "$E[SD_c]$",
+                "$Var[SD_c]$")
+
+df2$'$Var[\\delta_0]$' <- ((m_num[,2]) / (m_num[,3]^2)) + (((m_num[,1]^2) / (m_num[,3]^4) )* m_num[,4])
+df2$'$Var[\\delta]$' <- ((m_num[,2]) / (m_num[,5]^2)) + (((m_num[,1]^2) / (m_num[,5]^4) )* m_num[,6])
+
+knitr::kable(df2, digits = 3)
+
+```
+
+Table 2 demonstrates that in scenarios, where either score reliability was very high to begin with ($E[a] = .9$ in this case), leading to very little change in $E[SD]$ to $E[SD_c]$ (row 5), or where the remaining variance in $SD_c$ was comparably low (rows 3 and 4), the attenuation correction led to a reduction in ES heterogeneity.
+In most other scenarios, where score reliability was lower (.7 and .5 in rows 6 and 7) or remaining variance in $SD_c$ sufficiently large (row 2), the attenuation correction actually increased the heterogeneity in ES.
 Note that in the table, we did not manipulate expected value or variance of MD or SD.
-
-\[implications of this distribution concerning Var\[ES\] and how that relates to Var\[rel\]\]
 
@@
 
-**Part C**
+## Empirical Example
 
@@
 Over the following paragraphs, we will briefly describe the study design and measures used that are of interest to us.
-Details regarding the protocol, stimuli or additional measures used may be found at \[…\].
+Details regarding the protocol, stimuli or additional measures used may be found at […].
 
@@
 Similarly, negative and positive anticipated emotions were assessed, by prefacing a question of “In the next week, to what extent, if at all, do you think you will feel each of the following?” and following up with the same items used to assess state emotions.
-All items or the measures used may be found in the appendix XXX, or at \[url…\].
+All items or the measures used may be found in the appendix XXX, or at [url…].
 
@@
 
-<table>
-<tbody>
-<tr class="odd">
-<td><p>  </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="even">
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Measure</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Valence</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>ES</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Tau ES</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>ES_c</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Tau ES_c</p>
-<p>   </p></td>
-<td><p>  </p></td>
-</tr>
-<tr class="odd">
-<td><p>  </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="even">
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Photograph   Response</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Positive</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>1.065   (.065)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.377 (sq   .036)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>1.219   (.080)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.468 (sq   .056)</p>
-<p>   </p></td>
-<td><p>  </p></td>
-</tr>
-<tr class="odd">
-<td><p>  </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="even">
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Negative</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>-.663   (.040)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.220 (sq .014)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>-.747   (.047)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.260 (sq   .019)</p>
-<p>   </p></td>
-<td><p>  </p></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="odd">
-<td><p>  </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="even">
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Emotional   State</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Positive</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.213   (.018)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.052 (sq   .003)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.234   (.020)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.054 (sq   .003)</p>
-<p>   </p></td>
-<td><p>  </p></td>
-</tr>
-<tr class="odd">
-<td><p>  </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="even">
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Negative</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>-.203   (.020)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.071 (sq   .003)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>-.221   (.022)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.078   (.004)</p>
-<p>   </p></td>
-<td><p>  </p></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="odd">
-<td><p>  </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="even">
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Anticipated   Emotion</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Positive</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.159   (.015)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>0 (sq   .001)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.172   (.016)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>0 (.002)</p>
-<p>   </p></td>
-<td><p>  </p></td>
-</tr>
-<tr class="odd">
-<td><p>  </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="even">
-<td><p>   </p></td>
-<td><p>   </p>
-<p>Negative</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>-.146   (.020)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.073 (sq   .003)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>-.145   (.022)</p>
-<p>   </p></td>
-<td><p>   </p></td>
-<td><p>   </p>
-<p>.079   (.004)</p>
-<p>   </p></td>
-<td><p>  </p></td>
-<td></td>
-<td></td>
-</tr>
-<tr class="odd">
-<td><p> </p></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-<td></td>
-</tr>
-</tbody>
-</table>
+```{r, echo = FALSE}
+pos_SMDs.df <- read.csv(file = here("Data/Processed/SMDs_positive.csv"))
+neg_SMDs.df <- read.csv(file = here("Data/Processed/SMDs_negative.csv"))
+pos_state_SMDs.df <- read.csv(file = here("Data/Processed/SMDs_state_positive.csv"))
+neg_state_SMDs.df <- read.csv(file = here("Data/Processed/SMDs_state_negative.csv"))
+pos_antc_SMDs.df <- read.csv(file = here("Data/Processed/SMDs_antc_positive.csv"))
+neg_antc_SMDs.df <- read.csv(file = here("Data/Processed/SMDs_antc_negative.csv"))
+
+### photograph responses
+
+# fit a meta-analytic model to estimates of uncorrected Cohen's d, for positiveness-ratings
+pos_SMDs_rma_raw <- metafor::rma(yi = d_raw, 
+                                 sei = SE_d_raw, 
+                                 data = pos_SMDs.df,
+                                 measure = "GEN",
+                                 method = "REML")
+
+# fit a meta-analytic model to estimates of corrected Cohen's d, for positiveness-ratings
+pos_SMDs_rma_cor <- metafor::rma(yi = d_cor, 
+                                 sei = SE_d_cor, 
+                                 data = pos_SMDs.df,
+                                 measure = "GEN",
+                                 method = "REML")
+
+# fit a meta-analytic model to estimates of uncorrected Cohen's d, for negativeness-ratings
+neg_SMDs_rma_raw <- metafor::rma(yi = d_raw, 
+                                 sei = SE_d_raw, 
+                                 data = neg_SMDs.df,
+                                 measure = "GEN",
+                                 method = "REML")
+
+# fit a meta-analytic model to estimates of corrected Cohen's d, for negativeness-ratings
+neg_SMDs_rma_cor <- metafor::rma(yi = d_cor, 
+                                 sei = SE_d_cor, 
+                                 data = neg_SMDs.df,
+                                 measure = "GEN",
+                                 method = "REML")
+
+### state responses
+
+# fit a meta-analytic model to estimates of uncorrected Cohen's d, for positiveness-ratings
+pos_state_SMDs_rma_raw <- metafor::rma(yi = d_raw, 
+                                 sei = SE_d_raw, 
+                                 data = pos_state_SMDs.df,
+                                 measure = "GEN",
+                                 method = "REML")
+
+# fit a meta-analytic model to estimates of corrected Cohen's d, for positiveness-ratings
+pos_state_SMDs_rma_cor <- metafor::rma(yi = d_cor, 
+                                 sei = SE_d_cor, 
+                                 data = pos_state_SMDs.df,
+                                 measure = "GEN",
+                                 method = "REML")
+
+# fit a meta-analytic model to estimates of uncorrected Cohen's d, for negativeness-ratings
+neg_state_SMDs_rma_raw <- metafor::rma(yi = d_raw, 
+                                 sei = SE_d_raw, 
+                                 data = neg_state_SMDs.df,
+                                 measure = "GEN",
+                                 method = "REML")
+
+# fit a meta-analytic model to estimates of corrected Cohen's d, for negativeness-ratings
+neg_state_SMDs_rma_cor <- metafor::rma(yi = d_cor, 
+                                 sei = SE_d_cor, 
+                                 data = neg_state_SMDs.df,
+                                 measure = "GEN",
+                                 method = "REML")
+
+### anticipated respones
+
+# fit a meta-analytic model to estimates of uncorrected Cohen's d, for positiveness-ratings
+pos_antc_SMDs_rma_raw <- metafor::rma(yi = d_raw, 
+                                       sei = SE_d_raw, 
+                                       data = pos_antc_SMDs.df,
+                                       measure = "GEN",
+                                       method = "REML")
+
+# fit a meta-analytic model to estimates of corrected Cohen's d, for positiveness-ratings
+pos_antc_SMDs_rma_cor <- metafor::rma(yi = d_cor, 
+                                       sei = SE_d_cor, 
+                                       data = pos_antc_SMDs.df,
+                                       measure = "GEN",
+                                       method = "REML")
+
+# fit a meta-analytic model to estimates of uncorrected Cohen's d, for negativeness-ratings
+neg_antc_SMDs_rma_raw <- metafor::rma(yi = d_raw, 
+                                       sei = SE_d_raw, 
+                                       data = neg_antc_SMDs.df,
+                                       measure = "GEN",
+                                       method = "REML")
+
+# fit a meta-analytic model to estimates of corrected Cohen's d, for negativeness-ratings
+neg_antc_SMDs_rma_cor <- metafor::rma(yi = d_cor, 
+                                       sei = SE_d_cor, 
+                                       data = neg_antc_SMDs.df,
+                                       measure = "GEN",
+                                       method = "REML")
+
+df4 <- data.frame("Measure" = c(rep("Photograph Response", 2),
+                                rep("Emotional State", 2),
+                                rep("Anticipated Emotion",2)),
+                  "Valence" = rep(c("Positive", "Negative"), 3),
+                  "ES" = c(paste0(round(pos_SMDs_rma_raw$b[1], 3), " (", round(pos_SMDs_rma_raw$se, 3), ")"),
+                           paste0(round(neg_SMDs_rma_raw$b[1], 3), " (", round(neg_SMDs_rma_raw$se, 3), ")"),
+                           paste0(round(pos_state_SMDs_rma_raw$b[1], 3), " (", round(pos_state_SMDs_rma_raw$se, 3), ")"),
+                           paste0(round(neg_state_SMDs_rma_raw$b[1], 3), " (", round(neg_state_SMDs_rma_raw$se, 3), ")"),
+                           paste0(round(pos_antc_SMDs_rma_raw$b[1], 3), " (", round(pos_antc_SMDs_rma_raw$se, 3), ")"),
+                           paste0(round(neg_antc_SMDs_rma_raw$b[1], 3), " (", round(neg_antc_SMDs_rma_raw$se, 3), ")")
+                           ),
+                  "$\\tau_{ES}$"= c(paste0(round(sqrt(pos_SMDs_rma_raw$tau2), 3), " (sq. ", round(pos_SMDs_rma_raw$se.tau2, 3), ")"),
+                                    paste0(round(sqrt(neg_SMDs_rma_raw$tau2), 3), " (sq. ", round(neg_SMDs_rma_raw$se.tau2, 3), ")"),
+                                    paste0(round(sqrt(pos_state_SMDs_rma_raw$tau2), 3), " (sq. ", round(pos_state_SMDs_rma_raw$se.tau2, 3), ")"),
+                                    paste0(round(sqrt(neg_state_SMDs_rma_raw$tau2), 3), " (sq. ", round(neg_state_SMDs_rma_raw$se.tau2, 3), ")"),
+                                    paste0(round(sqrt(pos_antc_SMDs_rma_raw$tau2), 3), " (sq. ", round(pos_antc_SMDs_rma_raw$se.tau2, 3), ")"),
+                                    paste0(round(sqrt(neg_antc_SMDs_rma_raw$tau2), 3), " (sq. ", round(neg_antc_SMDs_rma_raw$se.tau2, 3), ")")
+                                    ),
+                  "$ES_c$" = c(paste0(round(pos_SMDs_rma_cor$b[1], 3), " (", round(pos_SMDs_rma_cor$se, 3), ")"),
+                               paste0(round(neg_SMDs_rma_cor$b[1], 3), " (", round(neg_SMDs_rma_cor$se, 3), ")"),
+                               paste0(round(pos_state_SMDs_rma_cor$b[1], 3), " (", round(pos_state_SMDs_rma_cor$se, 3), ")"),
+                               paste0(round(neg_state_SMDs_rma_cor$b[1], 3), " (", round(neg_state_SMDs_rma_cor$se, 3), ")"),
+                               paste0(round(pos_antc_SMDs_rma_cor$b[1], 3), " (", round(pos_antc_SMDs_rma_cor$se, 3), ")"),
+                               paste0(round(neg_antc_SMDs_rma_cor$b[1], 3), " (", round(neg_antc_SMDs_rma_cor$se, 3), ")")
+                               ),
+                  "$\\tau_{ES_c}$"= c(paste0(round(sqrt(pos_SMDs_rma_cor$tau2), 3), " (sq. ", round(pos_SMDs_rma_cor$se.tau2, 3), ")"),
+                                      paste0(round(sqrt(neg_SMDs_rma_cor$tau2), 3), " (sq. ", round(neg_SMDs_rma_cor$se.tau2, 3), ")"),
+                                      paste0(round(sqrt(pos_state_SMDs_rma_cor$tau2), 3), " (sq. ", round(pos_state_SMDs_rma_cor$se.tau2, 3), ")"),
+                                      paste0(round(sqrt(neg_state_SMDs_rma_cor$tau2), 3), " (sq. ", round(neg_state_SMDs_rma_cor$se.tau2, 3), ")"),
+                                      paste0(round(sqrt(pos_antc_SMDs_rma_cor$tau2), 3), " (sq. ", round(pos_antc_SMDs_rma_cor$se.tau2, 3), ")"),
+                                      paste0(round(sqrt(neg_antc_SMDs_rma_cor$tau2), 3), " (sq. ", round(neg_antc_SMDs_rma_cor$se.tau2, 3), ")")
+                                      )
+                  )
+
+```
+
+```{r, echo = FALSE}
+
+df4$Measure <- c(rep("Photograph Response", 2),
+                 rep("Emotional State", 2),
+                 rep("Anticipated Emotion",2))
+df4$Valence <- c(rep(c("Positive", "Negative"), 3))
+
+knitr::kable(df4)
+
+```
 
@@
 
+```{r, echo = FALSE}
+
+forest_plot_rel <- function(rma.fit_raw, rma.fit_cor, rma.data, ci.lvl = .975){
+  rma.data %>% 
+    mutate(cil_raw = d_raw - (1-(1-ci.lvl)/2) * (SE_d_raw),
+           ciu_raw = d_raw + (1-(1-ci.lvl)/2) * (SE_d_raw),
+           cil_cor = d_cor - (1-(1-ci.lvl)/2) * (SE_d_cor),
+           ciu_cor = d_cor + (1-(1-ci.lvl)/2) * (SE_d_cor)) %>% 
+    arrange(desc(d_raw)) %>% 
+    ggplot() +
+    # point estimate of raw d
+    geom_point(aes(x = d_raw, y = 1:nrow(rma.data)), colour = "darkgrey") +
+    # CI of point estimate of raw d
+    geom_segment(aes(x = cil_raw, y = 1:nrow(rma.data), xend = ciu_raw, yend = 1:nrow(rma.data)), colour = "darkgrey") +
+    geom_segment(aes(x = cil_raw, xend = cil_raw, y = (1:nrow(rma.data))+.3, yend = (1:nrow(rma.data))-.3), colour = "darkgrey") +
+    geom_segment(aes(x = ciu_raw, xend = ciu_raw, y =( 1:nrow(rma.data))+.3, yend = (1:nrow(rma.data))-.3), colour = "darkgrey") +
+    # point estiamte of corrected d
+    geom_point(aes(x = d_cor, y = 1:nrow(rma.data)), colour = "black") +
+    # CI of point estimate of corrected d
+    geom_segment(aes(x = cil_cor, y = 1:nrow(rma.data), xend = ciu_cor, yend = 1:nrow(rma.data)), colour = "black") +
+    geom_segment(aes(x = cil_cor, xend = cil_cor, y = (1:nrow(rma.data))+.3, yend = (1:nrow(rma.data))-.3), colour = "black") +
+    geom_segment(aes(x = ciu_cor, xend = ciu_cor, y = (1:nrow(rma.data))+.3, yend = (1:nrow(rma.data))-.3), colour = "black") +
+    # solid black line separating RMA-estimates from point estimates
+    geom_abline(slope = 0, intercept = -1, colour = "black") +
+    # adding diamond showing rma-estimate of raw d
+    geom_polygon(data = data.frame(x = c(rma.fit_raw$ci.ub, rma.fit_raw$b[1], rma.fit_raw$ci.lb, rma.fit_raw$b[1]),
+                                   y = c(-3, -3-.7, -3, -3+.7)),
+                 aes(x = x, y = y),
+                 fill = "darkgrey", colour = "darkgrey") +
+    # adding diamond showing rma-estimate of corrected d
+    geom_polygon(data = data.frame(x = c(rma.fit_cor$ci.ub, rma.fit_cor$b[1], rma.fit_cor$ci.lb, rma.fit_cor$b[1]),
+                                   y = c(-3, -3-.7, -3, -3+.7)),
+                 aes(x = x, y = y),
+                 colour = "black", fill = "black") +
+    # defining theme of plot (transparent background etc.)
+    theme(legend.position = "bottom", 
+          panel.background = element_rect(fill = "transparent"), 
+          plot.background = element_rect(fill = "transparent", colour = "transparent"), 
+          panel.grid.major.y = element_line(colour = "transparent"),
+          panel.grid.major.x = element_line(colour = "grey"),
+          panel.grid.minor = element_line(colour = "transparent"),
+          axis.ticks = element_line(colour = "grey"),
+          strip.background = element_rect(fill = "transparent"),
+          strip.text = element_text(size = 12)) +
+    scale_y_continuous(breaks = c(-3, 1:nrow(rma.data)), labels = c("RMA-estimate", rma.data$country)) +
+    labs(x = "Cohen's d",
+         y = "Country") 
+}
+
+underlying_density <- function(rma.fit_raw, rma.fit_cor){
+
+  x_raw <- seq(rma.fit_raw$b[1] - 3*sqrt(rma.fit_raw$tau2), rma.fit_raw$b[1] + 3*sqrt(rma.fit_raw$tau2), length.out = 1000)
+  density_values_raw <- dnorm(x_raw, mean = rma.fit_raw$b[1], sd = sqrt(rma.fit_raw$tau2)) 
+  df_raw <- data.frame(x_raw, density_values_raw)
+
+  x_cor <- seq(rma.fit_cor$b[1] - 3*sqrt(rma.fit_cor$tau2), rma.fit_cor$b[1] + 3*sqrt(rma.fit_cor$tau2), length.out = 1000)
+  density_values_cor <- dnorm(x_cor, mean = rma.fit_cor$b[1], sd = sqrt(rma.fit_cor$tau2)) 
+  df_cor <- data.frame(x_cor, density_values_cor)
+
+  df <- data.frame(df_raw, df_cor)
+
+  ggplot(df) +
+    geom_line(aes(x = x_raw, y = density_values_raw), colour = "darkgrey") +
+    geom_area(aes(x = x_raw, y = density_values_raw), fill = "darkgrey") +
+    geom_line(aes(x = x_cor, y = density_values_cor), colour = "black", alpha = .7) +
+    geom_area(aes(x = x_cor, y = density_values_cor), fill = "black", alpha = .7) +
+    labs(x = "Cohen's d", 
+         y = "Density") +
+    theme(legend.position = "none", 
+          panel.background = element_rect(fill = "transparent"), 
+          plot.background = element_rect(fill = "transparent", colour = "transparent"), 
+          panel.grid.major.y = element_line(colour = "transparent"),
+          panel.grid.major.x = element_line(colour = "black"),
+          panel.grid.minor = element_line(colour = "transparent"),
+          axis.ticks = element_line(colour = "black"))
+
+}
+
+fig1.p <- gridExtra::arrangeGrob(forest_plot_rel(pos_SMDs_rma_raw, pos_SMDs_rma_cor, pos_SMDs.df) +
+                                   labs(subtitle = "a)"),
+                                 underlying_density(pos_SMDs_rma_raw, pos_SMDs_rma_cor) +
+                                   labs(subtitle = "b)"),
+                                 layout_matrix = matrix(c(rep(1, 90), rep(2, 54)), byrow = FALSE, ncol = 16, nrow = 9))
+
+plot(fig1.p)
+
+```
+
 *Note:* In a), grey dots represent individual uncorrected ES with grey bars representing the accompanying 95%-confidence intervals; black dots represent individual corrected ES wit black bars representing accompanying 95%-confidence intervals.
@@
 
-**Discussion**
+## Discussion
 
@@
 
-\+ and – of this study
-
 The empirical arguments presented are based on a single study.
@@
 
-How to go from here
-
 The analytical arguments presented in equation X could be improved by properly disentangling the terms concerning variance of SD and expected value of SD.

### 2024-05-29 Manuscript_ScoreReLiability.qmd
text updates

@@
 If heterogeneity grows, meaning the phenomenon’s effect size varies more strongly for unexplained reasons, the probability of observing an effect size around zero or even in the negative space grows larger as well.
-If we know the true size and heterogeneity of a phenomenon’s effect size, we would theoretically be able to establish a baseline of expected replication rate (reference).
+If we know the true size and heterogeneity of a phenomenon’s effect size, we would theoretically be able to establish a baseline of expected replication rate, given no presence of publication bias or QRPs (reference).
 
@@
 (2020), standardized effect sizes (ES), such as Cohen’s d or Hedge’s g were used.
-For the remainder of the article, Cohen’s d, defined in equation 1, will be used as an exemplary estimate of ES, as its used across a wide range of contexts and well understood by a broad audience.
+For the remainder of the article, Cohen’s d, defined in equation [-@eq-SMD0], will be used as an exemplary estimate of ES, as its used across a wide range of contexts and well understood by a broad audience.
 
-$$d = \frac{MD}{\sigma_X}$$
+$$d = \frac{\Delta}{\sigma_X}$$ {#eq-SMD0}
 
-Here, $MD$ refers to the difference between the two groups of interest, while $\sigma_X$ is the total pooled standard deviation.
+Here, $\Delta$ refers to the difference in means between the two groups of interest, while $\sigma_X$ is the total pooled standard deviation.
 
 As is widely known, score reliability affects such ES.
-In the context of classical test (CTT) theory score reliability is defined as the ratio of true to observed score variance, as defined in equation 2.
+In the context of classical test theory (CTT) score reliability is defined as the ratio of true to observed score variance, as defined in equation [-@eq-alpha].
 
-$$\rho_{XX’} = \frac{\sigma^2_T}{\sigma^2_X}$$
+$$a = \frac{\sigma^2_T}{\sigma^2_X}$$ {#eq-alpha}
 
@@
 In the same sense, $\sigma^2_X$ refers to the total variance of scores, including both the true variance and the random error variance.
-Lower score reliability in a data-set leads to a smaller ES, opposed to a similar data-set with higher score reliability (references).
+In line with Hunter & Schmidt (2014), score reliability is denoted with $a$.
+Lower score reliability in a data-set leads to a smaller ES, opposed to a similar data-set with identical $\Delta$ but higher score reliability (references).
 In the meta-analytic context, as score reliability is an aspect of a measuring instrument applied to a population, score reliability may not be identical across replications.
@@
 This process is also known as attenuation correction and, while not without its criticisms (reference), is a widespread practice in Psychology (reference).
-If the claims above, made by Wiernik and Dahlke or Hunter and Schmidt, were to be true, it would imply that corrections of individual observed ESs for their unreliability, any subsequently performed meta-analysis should generate lower estimates of heterogeneity.
+If the claims above, made by Wiernik and Dahlke (2020) or Hunter and Schmidt (2014), were to be true, it would imply that corrections of individual observed ESs for their unreliability, any subsequently performed meta-analysis should generate lower estimates of heterogeneity.
 Hunter & Schmidt even claim to have derived an equation which demonstrates how heterogeneity in score reliability inflates heterogeneity in ESs.
-“*If the level of reliability is independent of the true effect size across studies, then, to a close approximation*:” (p.
+“*If the level of reliability is independent of the true effect size across studies, then, to a close approximation*:” (2014, p.
 309)
 
-$$Var(\delta_0) = [E(a)]^2 Var(\delta) + [E(\delta)]^2 Var(a)$$
+$$Var(\delta_0) = [E(a)]^2 Var(\delta) + [E(\delta)]^2 Var(a)$$ {#eq-HS}
 
-Here, $\delta_0$ refers to the “true” ES, undistorted by sampling error but not corrected for measuring error, $\delta$ refers to the “true” ES, undistorted by sampling error and measurement error and $a$ refers to score reliability in Schmidt and Hunter's notation.
+Here, $\delta_0$ refers to the “true” ES, undistorted by sampling error but not corrected for measuring error, while $\delta$ refers to the “true” ES, undistorted by sampling error and measurement error.
 
@@
 However, we propose that this understanding of how differences in score reliability affect the heterogeneity of observed ESs is incomplete.
-Hunter and Schmidt’s implication, that heterogeneity in score reliability inflates ES heterogeneity does not follow from equation 1.
-If equation 1 were to be true, it would imply that heterogeneity in $\delta_0$ is essentially a function where heterogeneity and mean value of $\delta$ are weighted by mean score reliability $a$ and its heterogeneity.
+Hunter and Schmidt’s implication, that heterogeneity in score reliability inflates ES heterogeneity does not follow from equation [-@eq-HS].
+If equation [-@eq-HS] were to be true, it would imply that heterogeneity in $\delta_0$ is essentially a function in which heterogeneity and mean value of $\delta$ are weighted by mean score reliability $a$ and its heterogeneity.
 In this scenario, it is easy to construct cases where heterogeneity in $\delta_0$ is inflated or deflated by heterogeneity in $a$.
@@
 
-As Hunter & Schmidt point out, equation 1 is only valid “*If the level of reliability is independent of the true effect size across studies \[…\]”* (2014, p. 309).
-
-However, both parameters, ES $\delta$ and score reliability $a$ in original notation can not be independent parameters by definition.
+Hunter & Schmidt point out that equation [-@eq-HS] is only valid “*If the level of reliability is independent of the true effect size across studies \[…\]”* (2014, p.
+309).
+However, both parameters, ES $\delta$ and score reliability $a$ in original notation can not, by definition, be independent parameters.
 ES $\delta$ is the “true” standardized mean difference, unaffected by sampling or measurement error.
-The ES $\delta_0$ is standardized using the total standard deviation, which is inflated by measurement error.
-Equation 1, demonstrating how Cohen’s d is computed demonstrates how the mean difference is standardized in this case.
+The ES $\delta_0$ is standardized using the total standard deviation, which is inflated by measurement error, see equation [-@eg-SMD0].
 The ES $\delta$ on the other hand is standardized using the true standard deviation in the sense of CTT.
 
-$$\delta = \frac{\Delta}{\sigma_T}$$
+$$\delta = \frac{\Delta}{\sigma_T}$$ {#eq-SMD}
 
-In equation 4, $\sigma_T$ is the true standard deviation, the square root of the true score variance found in equation 2.
+In equation [-@eq-SMD], $\sigma_T$ is the true standard deviation, the square root of the true score variance found in equation 2.
 The fact that the true standard deviation parameter $\sigma_T$ is found both in equation 4 and equation 2 demonstrates that ES $\delta$ and score reliability $a$ can not be independent variables.
-Therefore, the basic premise of equation 3 is not fulfilled, as the assumption of variable independence could not be fulfilled.
+Therefore, the basic premise of equation [-@eq-HS] is not fulfilled, as the assumption of variable independence could not be fulfilled.
 
@@
 If the distribution of the components is known, first order taylor approximation may be used to generate estimates of mean and variance of the ratio variable (reference).
-Using some additional assumptions, we use the Taylor-estimator of the ratio’s variance to demonstrate how differences in score reliability may affect heterogeneity in ES.
+Using some additional assumptions, we use the Taylor-estimator of the ratio’s variance in [-@eq-TaylorZ] to demonstrate how differences in score reliability may affect heterogeneity in ES.
 
-$$var[Z] \approx \frac{var[X]}{E[Y]^2} - \frac{2E[X]}{E[Y]^3} cov[X,Y] + \frac{E[X]^2}{E[Y]^4} var[Y]$$
+$$var[Z] \approx \frac{var[X]}{E[Y]^2} - \frac{2E[X]}{E[Y]^3} cov[X,Y] + \frac{E[X]^2}{E[Y]^4} var[Y]$$ {#eq-TaylorZ}
 
-Assuming that the random variable are uncorrelated, simplifies the equation to
+Assuming that the random variable are uncorrelated, simplifies equation [-@eq-TaylorZ] to equation [-@eq-TaylorZs]
 
-$$var[Z] \approx \frac{var[X]}{E[Y]^2} + \frac{E[X]^2}{E[Y]^4} var[Y]$$
+$$var[Z] \approx \frac{var[X]}{E[Y]^2} + \frac{E[X]^2}{E[Y]^4} var[Y]$$ {#eq-TaylorZs}
 
-Substituting for ES, unstandardized mean difference MD and pooled standard deviation SD leads to
+Substituting for ES, unstandardized mean difference MD and pooled standard deviation SD leads to equation [-@eq-TaylorES0]
 
-$$var[ES] \approx \frac{var[MD]}{E[SD]^2} + \frac{E[MD]^2}{E[SD]^4} var[SD]$$
+$$var[\delta_0] \approx \frac{var[\Delta]}{E[\sigma]^2} + \frac{E[\Delta]^2}{E[\sigma]^4} var[\sigma]$$ {#eq-TaylorES0}
 
-Correcting ES for score reliability is technically done by correcting the pooled standard deviation: $SD_c = \sqrt{a V}$ .
-Since $a$ is always between zero and one, in the case of imperfect score reliability, $SD_c$ will always be smaller than total $SD$.
-Additionally, if the variance in $SD$ was introduced by differences in score reliability, these would be removed by the attenuation correction.
-However, there are two parameters that may introduce variance in $SD$ across replications.
+Correcting ES for score reliability is technically done by estimating true score standard deviation $\sigma_T$ from total score standardd eviation $\sigma$: $\sigma_T = \sqrt{a V}$.
 
-Firstly, as discussed, differences in score reliability may introduce variation to SD.
+Since $a$ is always between zero and one, in the case of imperfect score reliability, $\sigma_T$ will always be smaller than total $\sigma$.
+Additionally, if the variance in $\sigma$ was introduced by differences in score reliability, these would be removed by the attenuation correction.
+However, in total, there are two parameters that may introduce variance in $\sigma$ across replications.
+
+Firstly, as discussed, differences in score reliability may introduce variation to $\sigma$.
 Secondly however, the underlying latent variable we attempted to measure may have been not identically distributed across replications.
-If the populations where the replications are collected from are differently spread out in terms of the variable of interest, true differences in $SD$ would remain.
+If the populations where the replications are collected from are differently spread out in terms of the variable of interest, true differences in $\sigma_T$.
 Re-analyses of the ManyLabs and Registered Replication Reports have shown that such differences in true variance across replications are the norm, not the exception.
-In that case, some variance in $SD_c$ would remain, albeit smaller than the variance in $SD$.
-We reformulate equation X in terms of $SD_c$
+In that case, some variance in $\sigma_T$ would remain, albeit smaller than the variance in $\sigma$.
+We reformulate equation [-@eq-TaylorES0] in terms of $\sigma_T$
 
-$$var[ES_c] \approx \frac{var[MD]}{E[SD_c]^2} + \frac{E[MD]^2}{E[SD_c]^4} var[SD_c]$$
+$$var[\delta] \approx \frac{var[\Delta]}{E[\sigma_T]^2} + \frac{E[\Delta]^2}{E[\sigma_T]^4} var[\sigma_T]$$ {#eq-TaylorES}
 
 The implication of this equation is, again, not straightforward.
-The expected value of $SD_c$ is guaranteed to be smaller than the expected value of uncorrected $SD$.
-Similarly, the variance in $SD_c$ is guaranteed to be smaller than the uncorrected variance in $SD$.
-The expected value is always placed in the denominator of the function, while the variance is placed in the numerator of the second term in equation X+1.
-This implies that the variance in uncorrected ES can only be larger than the variance in $ES_c$, if the reduction in mean $SD_c$ due to the attenuation correction increases the total equation result faster than the reduction in variance of $SD_c$ decreases the total equation result.
+The expected value of $\sigma_T$ is guaranteed to be smaller than the expected value of uncorrected $\sigma$.
+Similarly, the variance in $\sigma_T$ is guaranteed to be smaller than the uncorrected variance in $\sigma$.
+The expected value is always placed in the denominator of the function, while the variance is placed in the numerator of the second term in equation [-@eq-TaylorES].
+This implies that the variance in uncorrected ES can only be larger than the variance in $\delta_c$, if the reduction in mean $\sigma_T$ due to the attenuation correction increases the total equation result faster than the reduction in variance of $\sigma_T$ decreases the total equation result.
 To illustrate this point, we consider some parameter combinations in table 2.
@@
 
-Table 2 demonstrates that in scenarios, where either score reliability was very high to begin with ($E[a] = .9$ in this case), leading to very little change in $E[SD]$ to $E[SD_c]$ (row 5), or where the remaining variance in $SD_c$ was comparably low (rows 3 and 4), the attenuation correction led to a reduction in ES heterogeneity.
-In most other scenarios, where score reliability was lower (.7 and .5 in rows 6 and 7) or remaining variance in $SD_c$ sufficiently large (row 2), the attenuation correction actually increased the heterogeneity in ES.
-Note that in the table, we did not manipulate expected value or variance of MD or SD.
+Table 2 demonstrates that in scenarios, where either score reliability was very high to begin with ($E[a] = .9$ in this case), leading to very little change in $E[\sigma]$ to $E[\sigma_c]$ (row 5), or where the remaining variance in $\sigma_c$ was comparably low (rows 3 and 4), the attenuation correction led to a reduction in ES heterogeneity.
+In most other scenarios, where score reliability was lower (.7 and .5 in rows 6 and 7) or remaining variance in $\sigma_c$ sufficiently large (row 2), the attenuation correction actually increased the heterogeneity in ES.
+Note that in the table, we did not manipulate expected value or variance of $\Delta$ or $\sigma$.
 
@@
 Using analytical arguments from a reformulation of equation X we have demonstrated that differences in measuring quality often do not inflate estimates of heterogeneity.
-Instead, assuming that MD and SD are uncorrelated, we argue that in many circumstances differences in measuring quality deflate heterogeneity in standardized mean differences.
+Instead, assuming that $\Delta$ and $\sigma$ are uncorrelated, we argue that in many circumstances differences in measuring quality deflate heterogeneity in standardized mean differences.
 This could be understood as a form of masking, where differences in measuring quality lead to a seemingly lower heterogeneity than we would have observed without any measurement error.
@@
 Over the following paragraphs, we will briefly describe the study design and measures used that are of interest to us.
-Details regarding the protocol, stimuli or additional measures used may be found at […].
+Details regarding the protocol, stimuli or additional measures used may be found at \[…\].
 
@@
 Similarly, negative and positive anticipated emotions were assessed, by prefacing a question of “In the next week, to what extent, if at all, do you think you will feel each of the following?” and following up with the same items used to assess state emotions.
-All items or the measures used may be found in the appendix XXX, or at [url…].
+All items or the measures used may be found in the appendix XXX, or at \[url…\].
 

### 2024-10-22 E0.5_Data-Extraction.R
change structure to two analysis phases

@@ -1,0 +1,47 @@
+### DGPs / ÖGP 2024 Wien ###
+
+
+###################################################################################################
+# This script is used purely for data cleaning, initial manipulation and extraction to single     #
+#  files per scale                                                                                #
+# Raw data won't be made public, as long as no agreement from authors is obtained.                #
+###################################################################################################
+
+
+# library loading and installing as necessary
+
+
+
+packages <- c("tidyverse", "here", "data.table", "haven")
+
+# check, whether library already installed or not - install and load as needed:
+apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
+  
+  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system
+  
+  if(pkg_avail){
+    require(x, character.only = TRUE)             # load the library, if already installed
+    
+  }else{
+    install.packages(x)                           # install the library, if missing
+    require(x, character.only = TRUE)             # load after installation
+  }
+})
+
+
+
+# Extraction Scripts are saved in directory "Extraction_Scripts". These contain R-code, which
+#  cleans and extracts data for each individual scale.
+Extraction_Scripts <- list.files(here("Extraction_Scripts"), full.names = TRUE)
+
+
+# BEWARE: This may take a while! Additionally, this will wipe your environment.
+# extract data from all original data files - this results in a single csv for each scale.
+sapply(Extraction_Scripts, FUN = function(x){
+  tryCatch(
+    source(x),
+    
+    error = function(e)(cat("ERROR: ", conditionMessage(e), " - ", x, "\n"))
+  )
+}
+) # (in folder "Data/Extracted (Project) Data")

### 2024-10-22 E0_Data-Download.R

@@ -15,8 +15,28 @@ apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
 })
 
 
+### Many Labs 1
 
+# download the zipped-directory of Many Labs 1 data, from the OSF
+osfr::osf_retrieve_file("https://osf.io/nqg97") %>% 
+  osfr::osf_download(path = here("Data/Downloaded Data"), conflicts = "overwrite")
+# the data-file is named very generically "Datasets.zip" - it is renamed here to avoid confusion
+file.rename(here("Data/Downloaded Data/Datasets.zip"), here("Data/Downloaded Data/ML1_Datasets.zip"))
+
+# unzip the directory and extract only the specific data-file required for anayses
+unzip(here("Data/Downloaded Data/ML1_Datasets.zip"),
+      files = "Data/CleanedDataset.sav",
+      exdir = here("Data/Original Data/ManyLabs1"),
+      junkpaths = TRUE)
+
+### PSACR001
+
+# use osfr-package to download PSACR001-data, on IPD-level, pre-cleaned
+osfr::osf_retrieve_file("https://osf.io/jecmr") %>%
+  osfr::osf_download(path = here("Data/Original Data/PSACR001"))
+
+### PSACR002
 
 # use osfr-package to download PSACR002-data, on IPD-level, pre-cleaned
 osfr::osf_retrieve_file("https://osf.io/rwksu") %>%
-  osfr::osf_download(path = here("Data/Downloaded/"))
+  osfr::osf_download(path = here("Data/Original Data/PSACR002"))

### 2024-10-22 E1_Data-Analysis.R

@@ -1,540 +1,0 @@
-## Score ReLiability
-
-# E1
-
-# Clean & extract data
-
-
-# library loading and installing as necessary
-
-
-# relevant libraries required for this script
-packages <- c("tidyverse", "here", "metafor")
-
-# check, whether library already installed or not - install and load as needed:
-apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
-  
-  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system
-  
-  if(pkg_avail){
-    require(x, character.only = TRUE)             # load the library, if already installed
-    
-  }else{
-    install.packages(x)                           # install the library, if missing
-    require(x, character.only = TRUE)             # load after installation
-  }
-})
-
-
-
-# load initial data-file
-# "wide, exclusions" - should be the correct, cleaned data-file
-psacr002_dat <- read.csv(here("Data/Downloaded/data_wide_exclusion_12.23.2020.csv"))
-
-# identify variables which contain responses on stimuli (positive and negative responses separate)
-pos_items <- names(psacr002_dat)[grep("positiveemotion_photo", names(psacr002_dat))]
-neg_items <- names(psacr002_dat)[grep("negativeemotion_photo", names(psacr002_dat))]
-
-pos_state_items <- names(psacr002_dat)[grep("C2_positiveemotion_state.*[^0-9]$", names(psacr002_dat))]
-neg_state_items <- names(psacr002_dat)[grep("C2_negativeemotion_state.*[^0-9]$", names(psacr002_dat))]
-
-pos_antc_items <- names(psacr002_dat)[grep("C2_positiveemotion_anticipated.*[^0-9]$", names(psacr002_dat))]
-neg_antc_items <- names(psacr002_dat)[grep("C2_negativeemotion_anticipated.*[^0-9]$", names(psacr002_dat))]
-
-# identify which countries contain sufficient (more than 20) respondents
-samples <- as.data.frame(table(psacr002_dat$country))$Var1[as.data.frame(table(psacr002_dat$country))$Freq >= 200]
-
-
-#### PHOTOGRAPHS
-
-# formulate data-file containing responses on positive photo items, information on the grouping
-#  variable, and the respondent's country
-psacr002_pos <- psacr002_dat %>% 
[…]

### 2024-10-22 E1_Data-Analysis_demonstrate.R

@@ -1,0 +1,246 @@
+## Score ReLiability
+
+# E1
+
+# Clean & extract data
+
+
+# library loading and installing as necessary
+
+
+# relevant libraries required for this script
+packages <- c("tidyverse", "here", "metafor")
+
+# check, whether library already installed or not - install and load as needed:
+apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
+  
+  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system
+  
+  if(pkg_avail){
+    require(x, character.only = TRUE)             # load the library, if already installed
+    
+  }else{
+    install.packages(x)                           # install the library, if missing
+    require(x, character.only = TRUE)             # load after installation
+  }
+})
+
+
+
+data_files <- list.files(here("Data/Extracted (Project) Data"), full.names = TRUE)
+
+
+data.list <- lapply(data_files, read.csv)
+
+
+MASC_names <- substr(list.files(here("Data/Extracted (Project) Data")), 1, nchar(list.files(here("Data/Extracted (Project) Data")))-4) 
+
+
+# generate aggregates for each lab/country/source within a single meta-analysis. This is done by
+#  running an lapply-loop across all MASCs (meta-analyses, see MetaPipeX, Fünderich et al. 2024)
+#  and, nested within each loop, a second lapply-loop across all labs/countries/sources (samples)
+agg_L <- lapply(data.list, FUN = function(x){
+  
+  # loop across labs/countries/sources (samples) within a single MASC
+  MASC_MD_L <- lapply(unique(x$source), FUN = function(lab){
+    
+    # data for respective country
+    d <- x[x$source == lab,]
+    
+    # data for respective treatment (1) or control (0) subgroup
+    d1 <- na.omit(d[d$group == 1, -c(grep("group", names(d)), grep("source", names(d)))])
+    d0 <- na.omit(d[d$group == 0, -c(grep("group", names(d)), grep("source", names(d)))])
+    
+    # sample size
+    n1 <- nrow(d1)
+    n0 <- nrow(d0)
+    
+    # covariance matrix
+    C1 <- cov(d1)
+    C0 <- cov(d0)
[…]

### 2024-10-22 E1_Data-Analysis_demonstrate_graphics.R

@@ -1,0 +1,152 @@
+## Score ReLiability
+
+# E1.5
+
+# Clean & extract data
+
+
+# library loading and installing as necessary
+
+
+# relevant libraries required for this script
+packages <- c("magrittr", "dplyr", "here", "patchwork", "ggplot2")
+
+# check, whether library already installed or not - install and load as needed:
+apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
+  
+  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system
+  
+  if(pkg_avail){
+    require(x, character.only = TRUE)             # load the library, if already installed
+    
+  }else{
+    install.packages(x)                           # install the library, if missing
+    require(x, character.only = TRUE)             # load after installation
+  }
+})
+
+B_alpha_rma_df <- read.csv(here("Data/Processed/Reliability_analysis.csv"))
+
+ES_rma_df <- read.csv(here("Data/Processed/Aggregates_ES_analysis.csv"))
+
+agg_L <- readRDS(here("Data/Processed/Aggregates_simple.csv"))
+
+MASC_names <- substr(list.files(here("Data/Extracted (Project) Data")), 1, nchar(list.files(here("Data/Extracted (Project) Data")))-4) 
+
+# generate plots for meta-analyses of corrected AND uncorrected Cohen's d (slide 7)
+plots2 <- lapply((1:length(agg_L)), FUN = function(idx){
+  
+  # name of MASC
+  name <- MASC_names[idx]
+  
+  # get aggregates per sample
+  y <- agg_L[[idx]]
+  
+  # get rma-results separated
+  rma_raw <- ES_rma_df[ES_rma_df$MASC == name & ES_rma_df$corr == 0,]
+  rma_corr <- ES_rma_df[ES_rma_df$MASC == name & ES_rma_df$corr == 1,]
+  
+  # get idea of minimum and maximum observations in MASC, to draw the respective distribution
+  minmaxsequence_raw <- seq(from = min(y$d_raw), to = max(y$d_raw), length.out = 100)
+  minmaxsequence_corr <- seq(from = min(y$d_corr), to = max(y$d_corr), length.out = 100)
+  
+  # Freedman-Diacons rule for binwidth:
+  bw_raw <- (2 * IQR(y$d_raw) * (length(y$d_raw)^-(1/3)))
[…]

### 2024-10-22 temp_results.qmd
change names, add combined forest plot

---
title: "temp - Results"
format: html
editor: visual
---

## Resuls

```{r, echo = FALSE}
# relevant libraries required for this script
packages <- c("magrittr", "dplyr", "here", "patchwork", "ggplot2")

# check, whether library already installed or not - install and load as needed:
apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {

  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system

  if(pkg_avail){
    require(x, character.only = TRUE)             # load the library, if already installed

  }else{
    install.packages(x)                           # install the library, if missing
    require(x, character.only = TRUE)             # load after installation
  }
})

B_alpha_rma_df <- read.csv(here("Data/Processed/Reliability_analysis.csv"))

ES_rma_df <- read.csv(here("Data/Processed/Aggregates_ES_analysis.csv"))

agg_L <- readRDS(here("Data/Processed/Aggregates_simple.csv"))

MASC_names <- substr(list.files(here("Data/Extracted (Project) Data")), 1, nchar(list.files(here("Data/Extracted (Project) Data")))-4)

forest_plot_rel <- function(rma_df, aggregates, ci.lvl = .975){
  rma_raw <- rma_df[rma_df$corr == 0,]
  rma_corr <- rma_df[rma_df$corr == 1,]

  aggregates %>% 
    mutate(cil_raw = d_raw - (1-(1-ci.lvl)/2) * (SE_d),
           ciu_raw = d_raw + (1-(1-ci.lvl)/2) * (SE_d),
           cil_corr = d_corr - (1-(1-ci.lvl)/2) * (SE_d_corr),
           ciu_corr = d_corr + (1-(1-ci.lvl)/2) * (SE_d_corr)) %>% 
    arrange(desc(d_raw)) %>% 
    ggplot() +
    # point estimate of raw d
    geom_point(aes(x = d_raw, y = 1:nrow(aggregates)), colour = "darkgrey") +
    # CI of point estimate of raw d
    geom_segment(aes(x = cil_raw, y = 1:nrow(aggregates), xend = ciu_raw, yend = 1:nrow(aggregates)), colour = "darkgrey") +
    geom_segment(aes(x = cil_raw, xend = cil_raw, y = (1:nrow(aggregates))+.3, yend = (1:nrow(aggregates))-.3), colour = "darkgrey") +
    geom_segment(aes(x = ciu_raw, xend = ciu_raw, y =( 1:nrow(aggregates))+.3, yend = (1:nrow(aggregates))-.3), colour = "darkgrey") +
    # point estiamte of corrected d
    geom_point(aes(x = d_corr, y = 1:nrow(aggregates)), colour = "black") +
    # CI of point estimate of corrected d
    geom_segment(aes(x = cil_corr, y = 1:nrow(aggregates), xend = ciu_corr, yend = 1:nrow(aggregates)), colour = "black") +
    geom_segment(aes(x = cil_corr, xend = cil_corr, y = (1:nrow(aggregates))+.3, yend = (1:nrow(aggregates))-.3), colour = "black") +
    geom_segment(aes(x = ciu_corr, xend = ciu_corr, y = (1:nrow(aggregates))+.3, yend = (1:nrow(aggregates))-.3), colour = "black") +
    # solid black line separating RMA-estimates from point estimates
    geom_abline(slope = 0, intercept = -1, colour = "black") +
    # adding diamond showing rma-estimate of raw d
    geom_polygon(data = data.frame(x = c(rma_raw$ci.ub, rma_raw$mu, rma_raw$ci.lb, rma_raw$mu),
                                   y = c(-3, -3-.7, -3, -3+.7)),
                 aes(x = x, y = y),
                 fill = "darkgrey", colour = "darkgrey") +
    # adding diamond showing rma-estimate of corrected d
    geom_polygon(data = data.frame(x = c(rma_corr$ci.ub, rma_corr$mu, rma_corr$ci.lb, rma_corr$mu),
                                   y = c(-3, -3-.7, -3, -3+.7)),
                 aes(x = x, y = y),
                 colour = "black", fill = "black") +
    # defining theme of plot (transparent background etc.)
    theme(legend.position = "bottom", 
          panel.background = element_rect(fill = "transparent"), 
          plot.background = element_rect(fill = "transparent", colour = "transparent"), 
          panel.grid.major.y = element_line(colour = "transparent"),
          panel.grid.major.x = element_line(colour = "grey"),
          panel.grid.minor = element_line(colour = "transparent"),
          axis.ticks = element_line(colour = "grey"),
          strip.background = element_rect(fill = "transparent"),
          strip.text = element_text(size = 12)) +
    scale_y_continuous(breaks = c(-3, 1:nrow(aggregates)), labels = c("RMA-estimate", aggregates$source)) +
    labs(x = "Cohen's d",
         y = "Country") 
}

```

```{r}
knitr::kable(ES_rma_df)
```

Estimating Cohen's d using equations X and XX lead to the estimates displayed in the forest plots found in Figure 1. Both estimates both uncorrected and corrected for imperfect score reliability are generated. The differences in these estimates are shown in the two different colours, where gray dots and confidence intervals represent uncorrected estimates and black dots and confidence intervals represent corrected estimates. Additionally, each forest plot visualises the meta-analytic estimate of mean effect size for both raw and corrected ES, again, differentiated in grey and black colours.

```{r, echo = FALSE}
plots3 <- lapply((1:length(agg_L)), FUN = function(idx){

  # name of MASC
  name <- MASC_names[idx]

  # get aggregates per sample
  y <- agg_L[[idx]]

  # get rma-results separated
  rma_raw <- ES_rma_df[ES_rma_df$MASC == name & ES_rma_df$corr == 0,]
  rma_corr <- ES_rma_df[ES_rma_df$MASC == name & ES_rma_df$corr == 1,]

  forest_plot_rel(ES_rma_df[ES_rma_df$MASC == name,], y)

})

# combine plots using pathwork-infrastructure
combined_plot3 <- Reduce('+', plots3) + plot_layout(ncol = 2) &
  plot_annotation(theme = theme(plot.background = element_rect(fill = "transparent", colour = "transparent")))
combined_plot3
```

First and foremost, figure 1 demonstrates how correcting ES for imperfect score reliability consistently affects these estimates. 
Across all phenomena, for each individual sample, the absolute ES grows larger as we correct for score reliability. 
Negative ES grow smaller while positive ES grow larger, all ES move farther away from zero. This is consistent with the expectations from equation XX. 
Correcting for reliability implies that the mean value remains unchanged, but the denominator, pooled standard deviation, can only grow smaller. 
Therefore the absolute value of ES can only grow larger. 
From Figure 1, we also see that the confidence intervals grow larger, as we correct for score reliability. 
This is also consistent and follows from equation XX, as the absolute effect size can only grow larger in the numerator, the standard error is bound to grow. 
Since correcting for score reliability makes use of an additional uncertain estimate, it is sensible that the standard error grows larger. 
Similarly, the meta-analytic estimate grows more uncertain, as we correct for imperfect score reliability, highlighted in the wider diamond shape.

While Figure 1 demonstrates the consistency of how reliability attenuation affects individual estimates of ES, as well as its meta-analytic average, it does not inform us about how meta-analytic heterogeneity is affected.
Figure 2 demonstrates how ES heterogeneity is affected by correcting for imperfect reliability.
Here, the underlying distribution of ES, implied by the meta-analytic models, are visualised across the four phenomena.
Similar to Figure 1, in Figure 2 a grey curve represents the distribution for the raw, uncorrected ES. The black cuve represents the distribution of ES, corrected for score reliability.
ES heterogeneity is essentially the variance (standard deviation) of "true" ES (true in the sense of free from sampling error). 
In Figure 2, the height and width of the distributions can easily inform the reader on how ES heterogeneity has changed due to the correction procedure.
A curve with a higher peak and less spread corresponds to a smaller extent of ES heterogeneity. 
A lower peak with larger spread on the other hand indicates larger ES heterogeneity.

```{r, echo = FALSE}

plots2 <- lapply((1:length(agg_L)), FUN = function(idx){

  # name of MASC
  name <- MASC_names[idx]

  # get aggregates per sample
  y <- agg_L[[idx]]

  # get rma-results separated
  rma_raw <- ES_rma_df[ES_rma_df$MASC == name & ES_rma_df$corr == 0,]
  rma_corr <- ES_rma_df[ES_rma_df$MASC == name & ES_rma_df$corr == 1,]

  # get idea of minimum and maximum observations in MASC, to draw the respective distribution
  minmaxsequence_raw <- seq(from = min(y$d_raw), to = max(y$d_raw), length.out = 100)
  minmaxsequence_corr <- seq(from = min(y$d_corr), to = max(y$d_corr), length.out = 100)

  # Freedman-Diacons rule for binwidth:
  bw_raw <- (2 * IQR(y$d_raw) * (length(y$d_raw)^-(1/3)))
  bw_corr <- (2 * IQR(y$d_corr) * (length(y$d_corr)^-(1/3)))

  # make plot
  p <- ggplot() +
    scale_shape_identity() +
    # geom_point(aes(x = y$d_raw, y = 0), colour = "darkgrey", shape = 108, size = 3) +
    # geom_point(aes(x = y$d_corr, y = 0), colour = "black", shape = 108, size = 3) +
    # geom_histogram(aes(x = y$d_raw), colour = NULL, fill = "darkgrey", binwidth = bw_raw, alpha = .5) +
    # geom_histogram(aes(x = y$d_corr), colour = NULL, fill = "black", binwidth = bw_raw, alpha = .5) +
    # geom_point(aes(x = rma_raw$mu, y = 0), colour = "darkgrey", shape = 108, size = 8) +
    # geom_point(aes(x = rma_corr$mu, y = 0), colour = "black", shape = 108, size = 8) +
    geom_line(aes(x = minmaxsequence_raw,
                  y = dnorm(minmaxsequence_raw, mean = rma_raw$mu, sd = sqrt(rma_raw$tau^2))),
              linewidth = 1, colour = "darkgrey") +
    geom_line(aes(x = minmaxsequence_corr,
                  y = dnorm(minmaxsequence_corr, mean = rma_corr$mu, sd = sqrt(rma_corr$tau^2))),
              linewidth = 1, colour = "black") +
    theme(axis.title = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.y = element_blank(),
          panel.background = element_rect(fill = "transparent"), 
          plot.background = element_rect(fill = "transparent", colour = "transparent"), 
          panel.grid.major.y = element_line(colour = "transparent"),
          panel.grid.major.x = element_line(colour = "grey"),
          panel.grid.minor = element_line(colour = "transparent"),
          axis.ticks.x = element_line(colour = "grey"),
          strip.background = element_rect(fill = "transparent")) +
    labs(subtitle = name)

  # this is actually an artefact, we don't need this, heterogeneity is never truly zero in this selection
  p

})

# combine plots using pathwork-infrastructure
combined_plot2 <- Reduce('+', plots2) + plot_layout(ncol = 2) &
  plot_annotation(theme = theme(plot.background = element_rect(fill = "transparent", colour = "transparent")))
combined_plot2

```

Figure 2 demonstrates that, across all four phenomana, correcting ES for imperfect score reliability did not reduce heterogeneity. 
Instead, ES heterogeneity was increased in every single case, albeit the change was rather small in three out of four cases. 
These results stand in stark contrast with the claims made by Wiernik & Dahlke (ref) and Hunter & Schmidt (ref).
Correcting ES for imperfect score reliability, thereby implying that score reliability is identical across all samples, does not lead to a reduction in ES heterogeneity.

To assess whether actual differences in score reliability existed, heterogeneity in score reliability is assessed across all four projects, making use of Reliability Generalization Meta-Analysis.

### 2024-10-22 E1_Data-Analysis_demonstrate.R

@@ -33,6 +33,8 @@ data_files <- list.files(here("Data/Extracted (Project) Data"), full.names = TRU
 data.list <- lapply(data_files, read.csv)
 
 
+data.list[[4]]$source <- substr(data.list[[4]]$source, start = nchar(data.list[[4]]$source) - 2, stop = nchar(data.list[[4]]$source))
+
 MASC_names <- substr(list.files(here("Data/Extracted (Project) Data")), 1, nchar(list.files(here("Data/Extracted (Project) Data")))-4) 
 
 

### 2024-10-22 E1_Data-Analysis_demonstrate_graphics.R

@@ -150,3 +150,26 @@ forest_plot_rel <- function(rma_df, aggregates, ci.lvl = .975){
 
 forest_plot_rel(ES_rma_df[c(1,2),], agg_L[[1]])
 
+
+
+plots3 <- lapply((1:length(agg_L)), FUN = function(idx){
+  
+  # name of MASC
+  name <- MASC_names[idx]
+  
+  # get aggregates per sample
+  y <- agg_L[[idx]]
+  
+  # get rma-results separated
+  rma_raw <- ES_rma_df[ES_rma_df$MASC == name & ES_rma_df$corr == 0,]
+  rma_corr <- ES_rma_df[ES_rma_df$MASC == name & ES_rma_df$corr == 1,]
+  
+  forest_plot_rel(ES_rma_df[ES_rma_df$MASC == name,], y)
+  
+})
+
+# combine plots using pathwork-infrastructure
+combined_plot3 <- Reduce('+', plots3) + plot_layout(ncol = 2) &
+  plot_annotation(theme = theme(plot.background = element_rect(fill = "transparent", colour = "transparent")))
+combined_plot3
+

### 2024-10-22 E1_Data-Analysis_demonstrate.R
add function library expand results add variance rma

@@ -145,15 +145,6 @@ agg_L <- lapply(data.list, FUN = function(x){
 
 saveRDS(agg_L, file = here("Data/Processed/Aggregates_simple.csv"))
 
-# function to estimate heterogeneity in Cronbach's Alpha, if transformation was used
-var_Bonnett_backtransformed <- function(rma_obj){
-  (((-exp(rma_obj$b[1]))^2) * rma_obj$tau2) + (.5*((-exp(rma_obj$b[1]))^2)*(rma_obj$tau2^2)) + ((-exp(rma_obj$b[1])) * (-exp(rma_obj$b[1])) * (rma_obj$tau2^2))
-}
-
-# function to estimate mean value of Cronbach's Alpha, if transformation was used
-mean_Bonnett_backtransformed <- function(rma_obj){
-  1 - exp(rma_obj$b[1]) + ((-exp(rma_obj$b[1])) / 2) * rma_obj$tau2
-}
 
 
 # perform meta-analyses of transformed score reliability
@@ -189,9 +180,11 @@ B_alpha_rma <- lapply(agg_L, FUN = function(x){
   # return data-frame with all important aggregates
   return(data.frame(tau2_BA = BA_rma$tau2,
                     mu_BA = BA_rma$b[1],
-                    tau2_alpha = var_Bonnett_backtransformed(BA_rma),
+                    tau_alpha = sqrt(var_Bonnett_backtransformed(BA_rma)),
                     mu_alpha = mean_Bonnett_backtransformed(BA_rma),
+                    QE = BA_rma$QE,
                     k = nrow(x),
+                    QEp = BA_rma$QEp,
                     I2 = BA_rma$I2,
                     H2 = BA_rma$H2))
   

### 2024-10-22 E1_Data-Analysis_demonstrate_graphics.R

@@ -25,6 +25,9 @@ apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
   }
 })
 
+source(here("ReLiability_Function-library.R"))
+
+
 B_alpha_rma_df <- read.csv(here("Data/Processed/Reliability_analysis.csv"))
 
 ES_rma_df <- read.csv(here("Data/Processed/Aggregates_ES_analysis.csv"))
@@ -96,57 +99,6 @@ ggsave(filename = here("Graphics/densities_four.png"),
        plot = last_plot(),
        width = 11, height = 5)
 
-
-
-forest_plot_rel <- function(rma_df, aggregates, ci.lvl = .975){
-  rma_raw <- rma_df[rma_df$corr == 0,]
-  rma_corr <- rma_df[rma_df$corr == 1,]
-  
-  aggregates %>% 
-    mutate(cil_raw = d_raw - (1-(1-ci.lvl)/2) * (SE_d),
-           ciu_raw = d_raw + (1-(1-ci.lvl)/2) * (SE_d),
-           cil_corr = d_corr - (1-(1-ci.lvl)/2) * (SE_d_corr),
-           ciu_corr = d_corr + (1-(1-ci.lvl)/2) * (SE_d_corr)) %>% 
-    arrange(desc(d_raw)) %>% 
-    ggplot() +
-    # point estimate of raw d
-    geom_point(aes(x = d_raw, y = 1:nrow(aggregates)), colour = "darkgrey") +
-    # CI of point estimate of raw d
-    geom_segment(aes(x = cil_raw, y = 1:nrow(aggregates), xend = ciu_raw, yend = 1:nrow(aggregates)), colour = "darkgrey") +
-    geom_segment(aes(x = cil_raw, xend = cil_raw, y = (1:nrow(aggregates))+.3, yend = (1:nrow(aggregates))-.3), colour = "darkgrey") +
-    geom_segment(aes(x = ciu_raw, xend = ciu_raw, y =( 1:nrow(aggregates))+.3, yend = (1:nrow(aggregates))-.3), colour = "darkgrey") +
-    # point estiamte of corrected d
-    geom_point(aes(x = d_corr, y = 1:nrow(aggregates)), colour = "black") +
-    # CI of point estimate of corrected d
-    geom_segment(aes(x = cil_corr, y = 1:nrow(aggregates), xend = ciu_corr, yend = 1:nrow(aggregates)), colour = "black") +
[…]

### 2024-10-22 E2_Data-Analysis_explore.R

@@ -1,0 +1,159 @@
+## Score ReLiability
+
+# E1.5
+
+# Clean & extract data
+
+
+# library loading and installing as necessary
+
+
+# relevant libraries required for this script
+packages <- c("magrittr", "dplyr", "boot", "here")
+
+# check, whether library already installed or not - install and load as needed:
+apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
+  
+  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system
+  
+  if(pkg_avail){
+    require(x, character.only = TRUE)             # load the library, if already installed
+    
+  }else{
+    install.packages(x)                           # install the library, if missing
+    require(x, character.only = TRUE)             # load after installation
+  }
+})
+
+source(here("ReLiability_Function-library.R"))
+
+
+B_alpha_rma_df <- read.csv(here("Data/Processed/Reliability_analysis.csv"))
+
+ES_rma_df <- read.csv(here("Data/Processed/Aggregates_ES_analysis.csv"))
+
+agg_L <- readRDS(here("Data/Processed/Aggregates_simple.csv"))
+
+MASC_names <- substr(list.files(here("Data/Extracted (Project) Data")), 1, nchar(list.files(here("Data/Extracted (Project) Data")))-4) 
+
+
+
+
+data_files <- list.files(here("Data/Extracted (Project) Data"), full.names = TRUE)
+
+
+data.list <- lapply(data_files, read.csv)
+
+
+data.list[[4]]$source <- substr(data.list[[4]]$source, start = nchar(data.list[[4]]$source) - 2, stop = nchar(data.list[[4]]$source))
+
+
+# generate estimates of ln-error score variance and its associated standard error using bootstrapping
+varE_est.L <- lapply(seq_along(data.list), FUN = function(x){
+  
+  # sometimes calculation of SE might lead to issues, as negative variances can not be log-transformed
+  #  therefore, function is run within a tryCatch-environment, so the script does not crash
+  # apply_Bootstrap_SE_Project.specific is the project-specific function
+  tryCatch(apply_Bootstrap_SE_Project.specific(na.omit(data.list[[x]])),
+           
+           # print error to console
+           error = function(e)(cat("ERROR: ", conditionMessage(e), " - ",
+                                   substr(names(data.list), 
+                                          (regexpr("Project) Data/", names(data.list)) + 14), 
[…]

### 2024-10-22 ReLiability_Function-library.R

@@ -1,0 +1,204 @@
+
+
+## function library
+
+
+
+
+
+# function to estimate heterogeneity in Cronbach's Alpha, if transformation was used
+var_Bonnett_backtransformed <- function(rma_obj){
+  (((-exp(rma_obj$b[1]))^2) * rma_obj$tau2) + (.5*((-exp(rma_obj$b[1]))^2)*(rma_obj$tau2^2)) + ((-exp(rma_obj$b[1])) * (-exp(rma_obj$b[1])) * (rma_obj$tau2^2))
+}
+
+# function to estimate mean value of Cronbach's Alpha, if transformation was used
+mean_Bonnett_backtransformed <- function(rma_obj){
+  1 - exp(rma_obj$b[1]) + ((-exp(rma_obj$b[1])) / 2) * rma_obj$tau2
+}
+
+
+
+bt_var_m <- function(rma_obj){
+  exp(rma_obj$b[1] + (.5*rma_obj$tau2))
+}
+
+bt_var_v <- function(rma_obj){
+  (exp(rma_obj$tau2) - 1) * exp((2 * rma_obj$b[1]) + rma_obj$tau2)
+}
+
+
+
+
+
+## a function required for the boot-package, which allows to estimate the error score
+##  variance in a bootstrapped sample
+bootstrap_SE_varE <- function(data, indices){
+  
+  # select subset of data.
+  d <- data[indices,]
+  
+  d1 <- d[d$group == 1, -(which(names(d) %in% c("source", "group")))]
+  d0 <- d[d$group == 0, -(which(names(d) %in% c("source", "group")))]
+  
+  mv1 <- rowMeans(d1)
+  mv0 <- rowMeans(d0)
+  
+  sd1 <- sqrt(mean((mv1 - mean(mv1))^2))
+  sd0 <- sqrt(mean((mv0 - mean(mv0))^2))
+  
+  n1 <- length(d1)
+  n0 <- length(d0)
+  
+  C1 <- cov(d1)
+  j1 <- dim(C1)[1]
+  alpha1 <- (1 - sum(diag(C1))/sum(C1)) * (j1/(j1 - 1))
+  
+  # compute Cronbach's Alpha
+  C0 <- cov(d0)
+  j0 <- dim(C0)[1]
+  alpha0 <- (1 - sum(diag(C0))/sum(C0)) * (j0/(j0 - 1))
+  
[…]

### 2024-10-24 E2_Data-Analysis_explore.R
test rma of SD

@@ -54,7 +54,7 @@ varE_est.L <- lapply(seq_along(data.list), FUN = function(x){
   # sometimes calculation of SE might lead to issues, as negative variances can not be log-transformed
   #  therefore, function is run within a tryCatch-environment, so the script does not crash
   # apply_Bootstrap_SE_Project.specific is the project-specific function
-  tryCatch(apply_Bootstrap_SE_Project.specific(na.omit(data.list[[x]])),
+  tryCatch(apply_Bootstrap_SE_Project.specific(na.omit(data.list[[x]]), component = "E"),
            
            # print error to console
            error = function(e)(cat("ERROR: ", conditionMessage(e), " - ",
@@ -89,40 +89,19 @@ names(varE_rma.list) <- names(data.list)
 
 
 # generate estimates of ln-observed score variance for each sample & projects
-varX_est.L <- lapply(data.list, FUN = function(data){
+varX_est.L <- lapply(seq_along(data.list), FUN = function(x){
   
-  # apply over labs in MASC
-  df <- apply(as.matrix(seq_along(unique(data$source))), MARGIN = 1, FUN = function(x){
-    
-    # remove source-col & omit NAs
-    d <- na.omit(data[data$source == unique(data$source)[x],-grep("source", names(data))])
-    
-    d1 <- d[d$group == 1, -(which(names(d) %in% c("source", "group")))]
-    d0 <- d[d$group == 0, -(which(names(d) %in% c("source", "group")))]
-    
-    mv1 <- rowMeans(d1)
-    mv0 <- rowMeans(d0)
-    
-    sd1 <- sqrt(mean((mv1 - mean(mv1))^2))
-    sd0 <- sqrt(mean((mv0 - mean(mv0))^2))
-    
-    n1 <- length(d1)
-    n0 <- length(d0)
-    
-    # pooled variance
-    var_X <- ((n1-1)*sd1^2 + (n0-1)*sd0^2)/(n1+n0-2)
-    
-    # store SE, ln(varX) & varX
-    return(data.frame(SE = sqrt((2/(nrow(d) - 1))), 
-                      var.emp = log(var_X)))
-    
-  })
-  
-  # return formatted data.frame
-  df.formatted <- data.frame(SE = sapply(df, FUN = function(x){x$SE}),
-                             var.est = sapply(df, FUN = function(x){x$var.emp}),
-                             source = unique(data$source))
-  
+  # sometimes calculation of SE might lead to issues, as negative variances can not be log-transformed
[…]

### 2024-10-24 ReLiability_Function-library.R

@@ -69,7 +69,33 @@ bootstrap_SE_varE <- function(data, indices){
   # estimate errorcsore variance sigma^2_E
   var_E <- var_X - var_T
   
-  return(var_E)
+  return(sqrt(var_E))
+  
+}
+
+## a function required for the boot-package, which allows to estimate the error score
+##  variance in a bootstrapped sample
+bootstrap_SE_varX <- function(data, indices){
+  
+  # select subset of data.
+  d <- data[indices,]
+  
+  d1 <- d[d$group == 1, -(which(names(d) %in% c("source", "group")))]
+  d0 <- d[d$group == 0, -(which(names(d) %in% c("source", "group")))]
+  
+  mv1 <- rowMeans(d1)
+  mv0 <- rowMeans(d0)
+  
+  sd1 <- sqrt(mean((mv1 - mean(mv1))^2))
+  sd0 <- sqrt(mean((mv0 - mean(mv0))^2))
+  
+  n1 <- length(d1)
+  n0 <- length(d0)
+
+  # pooled standard deviation
+  var_X <- ((n1-1)*sd1^2 + (n0-1)*sd0^2)/(n1+n0-2)
+  
+  return(sqrt(var_X))
   
 }
 
@@ -77,9 +103,14 @@ bootstrap_SE_varE <- function(data, indices){
 # Function to estimate bootstrapped Standard Errors for either true or error score variance component,
 #  specifically made for empirical data (e.g. containing missing data)
 # expects data in a data.frame
-apply_Bootstrap_SE_Project.specific <- function(data, R = 100){
+apply_Bootstrap_SE_Project.specific <- function(data, R = 100, component = "E"){
   
-  stat.f <- bootstrap_SE_varE
+  if(component == "E"){
+    stat.f <- bootstrap_SE_varE
+  }else{
+    stat.f <- bootstrap_SE_varX
+  }
+  
  
   
   # suppress messages, as coefficientalpha-package can be quite "noisy"
@@ -128,12 +159,18 @@ apply_Bootstrap_SE_Project.specific <- function(data, R = 100){
       # estimate errorcsore variance sigma^2_E
       var_E <- var_X - var_T
       
+      if(component == "E"){
+        var_res <- var_E
+      }else{
+        var_res <- var_X
+      }
+      
       # return a data.frame, containing bootstrapped estimates of the standard error of the log-transformed
       #  variance component, as well as the bootstrapped mean of the log-transformed variance component. 
       # additionaly, return the empirical estimate using the full replication sample
[…]

### 2025-01-24 Manuscript_MD_ScoreReLiability.qmd
manuscript draft quarto

---
title: "Manuscript"
format: docx
editor: 
  markdown: 
    wrap: 72
---

# Score ReLiability – Can we explain effect size heterogeneity using score reliability?

Large-scale collaborative replication efforts have sparked discussions
surrounding the replicability of psychological phenomena. However,
attempts to estimate a phenomenon’s replicability or to predict whether
a future replication will replicate successfully, heterogeneity is a
crucial parameter that needs to be assessed. In the meta-analytic
context, heterogeneity describes the variation in effect sizes, free of
sampling error, if these are not stable or constant across replications.
Therefore, the presence of heterogeneity implies that some replications
of a single phenomenon may be successful, while others are not. If
heterogeneity grows, meaning the phenomenon’s effect size varies more
strongly for unexplained reasons, the probability of observing an effect
size around zero or even in the negative space grows larger as well. If
we know the mean size and heterogeneity of a phenomenon’s effect size,
we would theoretically be able to establish a baseline of expected
replication rate (reference).

Similarly, it has been argued that heterogeneity in effect sizes is an
indicator of the theory’s “completeness” surrounding the phenomenon.
Linden and Hönekopp argue that “low (as opposed to high) heterogeneity
reflects a more advanced understanding of the subject matter being
studied” (2021, p.2). Similarly, Schuetze and von Hippel (2023) argue
that heterogeneity in effects is an indicator of a vague, poorly
specified theory.

Re-analyses of multi-site direct replications have demonstrated that, if
a psychological effect size can be distinguished from zero,
heterogeneity is present across most phenomena (ref). Large-scale
attempts of direct replications, using identical protocols, such as the
Many Labs studies or Registered Replication Reports (references), for
the first time allow researchers to estimate heterogeneity undistorted
by differences in experimental designs. In re-analyses of these studies,
Olsson-Collentine et al. (2020) identify a positive correlation between
a phenomenon’s effect size and its heterogeneity. Similarly, van Erp et
al. (2017) or Stanley et al. (2018) estimate strong degrees of
heterogeneity across conceptual replications in Psychology. In a
separate re-analysis of large scale direct replications, Renkewitz et
al. (2025) identify substantial heterogeneity in almost all projects
where a non-zero effect could be identified. This aligns with the
correlation found by Olsson-Collentine et al. (2020).

## Effect sizes and score reliability in meta-analysis

In both the initial reports of large scale replication attempts, as well
as the re-analyses by Olsson-Collentine et al. (2020), standardized
effect sizes (from here-on abbreviated as ES) such as Cohen’s d or
Hedge’s g were used. For the remainder of the article, Cohen’s d,
defined in Equation 1, will be used as an exemplary estimate of ES, as
it is used across a wide range of contexts and well understood by a
broad audience.

$$d = \frac{MD}{\sigma_X}$${#eq-d}

Here, MD refers to the difference between two groups of interest, while
$\sigma_X$ is the total pooled standard deviation. Since $d$ constitutes
an observed effect size it is affected by sampling error. In terms of
heterogeneity, we are not interested in the variance of $d$, but the
variance of the true, underlying ES $\delta$. The standard error of
Cohen’s $d$, which is used to weight individual estimates in a
random-effects meta-analysis (reference) is computed as defined in
Equation 2.

$$SE_d=\sqrt{\frac{n_1+n_2}{n_1n_2}+\frac{d^2}{2\left(n_1+n_2\right)}}$${#eq-SEd}

As is widely known, score reliability affects ES. In the context of
classical test theory (CTT) score reliability ρXX’ is defined as the
ratio of true to total score variance, as defined in Equation 3.

$$\rho_{XX’} = \frac{\sigma^2_T}{\sigma^2_X}=\frac{\sigma^2_T}{\sigma^2_T + \sigma^2_E}$${#eq-rel}

In this equation $\sigma_T^2$ refers to the true score variance in the
sense of CTT, meaning the actual variance of the variable, undistorted
by random measurement error. In the same sense, $\sigma_X^2$ refers to
the total variance of scores, including both the true variance and the
random error variance $\sigma_E^2$. This implies that, if true score
variance is assumed to be constant across replications, a lower score
reliability can only occur due to a larger error score variance. Total
score variance would also be larger, therein leading to a smaller ES, as
opposed to a similar data-set where a higher score reliability is
achieved through a smaller error score variance. However, score
reliability is an aspect of a measuring instrument applied to a
population. In the meta-analytic context, as studies tend to be
replicated in different populations, neither true nor error score
variance can be expected to always be identical across replications.
This most likely leads to heterogeneity in score reliability, which, as
discussed, is bound to affect ES heterogeneity as well.

Previous
discussions of heterogeneity in score reliability have exclusively
discussed it as a parameter that inflates heterogeneity in ES, implying
that, if score reliability was perfect across all replications, the
actual heterogeneity would have been lower. In their discussion, Wiernik
and Dahlke claim that „*Measurement error variance will impact the
results of meta-analyses in three ways: by (a) biasing the mean effect
size toward zero, (b) inflating effect-size heterogeneity and
confounding moderator effects, and (c) confounding publication-bias and
sensitivity analyses*” (p. 3, 2020). Additionally, more clearly, they
state that “*If the studies included in a meta-analysis differ in their
measures’ reliabilities, heterogeneity estimates will be artifactually
inflated, erroneously suggesting larger potential moderator effects*” (p.
4, Wiernik & Dahlke, 2020). They base their claims largely on work done
by Hunter and Schmidt (2014), who claim that “*Variation in reliability
across studies causes variation in the observed effect sizes above and
beyond that produced by sampling error.*” (p. 302). Overall, both
references imply that differences in score (un)reliability inevitably
lead to an inflated heterogeneity in ES.

As both sources Hunter and
Schmidt (2014) and Wiernik and Dahlke (2020) will be referred to
repeatedly, we abbreviate these references as H&S and W&D respectively.

### Attenuation correction

If information concerning score reliability is
available, it is possible to correct the individual ES for its
unreliability. This process is also known as attenuation correction and,
while not without its criticisms (reference), is a widespread practice
in Psychology (reference). Equation 4 describes a simple attenuation
correction procedure.

$$d_c = \frac{d}{\sqrt{\hat{\rho}_{xx’}}}$${#eq-dc}

Here, $d_c$ describes the attenuation-corrected estimate of $d$, which
is corrected by dividing it by the estimate of score reliability
$\hat{\rho}_{xx’}$. As described in W&D or Lord & Novick (references),
this is essentially the same, as if the corrected estimate of $d_c$ was
constructed using an estimate of true score standard deviation, as in
Equation 5.

$$d_c = \frac{MD}{\hat{\sigma}_{T}}=\frac{MD}{\sqrt{\hat{\rho}_{xx’} \hat{\sigma}_X^2}}$${#eq-dct}

$\hat{\sigma}_T$ can be estimated by simply rearranging the terms from
equation 2. Thereby, an estimate of ES is generated, under the premise
that score reliability would have been perfect ($\rho_{xx'} = 1$). As
the corrected estimate of Cohen’s $d$ is larger and requires an
additional unknown parameter, an estimate of score reliability, the
larger uncertainty in this parameter should be reflected in a larger
standard error. The standard error for $d_c$ is defined in equation 6.

$$SE_{d_c} = \sqrt{SE^2_d \frac{d_c}{d}}$${#eq-SEdc}

The claims made by W&D imply that corrections of individual observed ESs
for their unreliability should lead to lower estimates of heterogeneity
in meta-analyses of corrected ES. If heterogeneity in score reliability
adds to ES heterogeneity, attenuation correction procedures should
eliminate that additional variation, as all corrected ES come with
identical (perfect) score reliability.

Re-analyses of the collaborative multi-site replication efforts have
uncovered differences in score reliability across replications of the
same phenomena (ref McShaw etc.). The claims made in W&D and H&S imply
that the ES heterogeneity uncovered in these projects therefore could
potentially be explained by these differences in score reliability. The
same argument has been proposed by both original authors of phenomena
that did not replicate or carry substantial heterogeneity (references)
as well as the authors of the multi-site replication attempts (refs).
Over the following pages, we assess for the first time, whether the
differences in score reliability do indeed explain, and thereby reduce,
the ES heterogeneity observed across the different multi-site
replication efforts. Even though only a limited number of phenomena
allow for such an analysis, we find no evidence for a reduction in ES
heterogeneity after taking score reliability into account. Subsequently,
we discuss why the expectation formulated in H&S does not allow for such
claims. Alternatively, we explore to what extent defining ES as a random
ratio variable helps understand how ES heterogeneity is affected by
differences in score reliability. We close with a discussion on what
future meta-analysts can expect to observe, when dealing with issues of
(un)reliability in assessments of meta-analyses of ES.

## Re-analysis of archival data Data

We have attempted to collect all openly available multi-site direct
replications on psychological phenomena (Fünderich et al. 2024). The
data-sets of about 50 phenomena , largely stemming from the efforts of
the Many Labs studies, Registered Replication Reports or the
Psychological Science Accelerator, are made openly available in a
standardised format at \[osf.io\], the DRIPHT repository.

The ManyLabs projects were collaborative efforts to replicate several
psychological phenomena across different research sites, employing
identical protocols. From five published projects, the data for the
first three and the fifth Many Labs projects was made publicly available
when the DRIPHT repository was set up (references). The Registered
Replication Reports are similar collaborative efforts, with the sole
distinction that for each report a single phenomenon was replicated
several times. The Registered Reports 3-10 were added to the DRIPHT
repository, as they employed experimental designs with at least two
groups. Lastly, some projects from the Psychological Science Accelerator
were added to the DRIPHT repository. Different to Many Labs or
Registered Replication Projects, these collaborative efforts, while also
distributed across the globe, do not formally attempt to perform direct
replications, but focus on original research or conceptual replications.
However, as for each project, across all sites the same protocol is
used, and the subjects are distributed across different countries, a
data structure similar to that of a multi-site replication study
emerges. What makes all these collaborative efforts valuable for this
manuscript is the fact that all phenomena collected in the
DRIPHT-repository are making use of two-group designs and are therefore
easily assessed using standardised effect sizes.

While replication data on a decent number of psychological phenomena is
available in the DRIPHT-repository, to demonstrate how incorporating
score reliability affects heterogeneity in ES, selected phenomena need
to fulfil three conditions: (1) As we focus on the use of ES $d$ d, the
phenomenon needs to be assessed using a two-group design. This is the
case for all projects in the DRIPHT repository; (2) Score reliability
estimates can be derived, this implies that the phenomenon needs to be
measured using either several indicators forming a single scale or by
repeatedly measuring across several timepoints; (3) Score reliability
can only attenuate effect sizes that are not zero in the first place.
Therefore, we focus on phenomena where meta-analytic mean ES can be
statistically distinguished from zero.

As all phenomena in the collection fulfil the first condition, 50
phenomena have been catalogued where the effect is studied by comparing
the outcome across two separate groups. While the majority of designs
make use of some form of control-treatment manipulation, randomly
assigning participants to a condition (20 phenomena), some phenomena
discuss the effect of pre-existing differences, e.g. biological sex (2
phenomena). From the collection of 50 phenomena, 22 fulfil the second
condition, so that estimates of score reliability can be derived. For
these phenomena, the dependent variable used to construct the effect
size is measured by employing more than a single indicator. At the same
time, however, this implies that the remaining 28 phenomena did not make
use of more than a single indicator to measure the dependent variable.
Therefore, for the majority of phenomena it is not possible to assess in
how far measurement quality is sufficiently high in terms of e.g.
internal consistency or low random error variance.

From the 22 phenomena that employed more than a single indicator, 19
made use of Likert-style items, where respondents would indicate their
agreement to some kind of statement. The scale-length varied from 3 to
10-points. The items measuring the remaining 3 phenomena were coded
dichotomously, as responses were right or wrong (2 cases), or expressed
agreement vs disagreement dichotomously (1 case). Subsequently, the 22
phenomena need to be identified where the meta-analytic mean effect size
is statistically significant different from zero. This will be done as a
first step of this manuscript’s analysis procedure. Detailed information
concerning the different phenomena and how they were measured can be
found at \[osf-link\], while a brief summary can be found in Table 1.

To assess whether ES heterogeneity can be reduced by correcting for
score reliability, estimates of meta-analytic heterogeneity are compared
across two situations: first, heterogeneity of raw ES, as computed in
equation 1 is assessed. Secondly, ES are corrected for imperfect
reliability. Computing the heterogeneity of these corrected ES allows
for an assessment, whether the heterogeneity did indeed shrink compared
to the first situation, as predicted in W&D and H&S.

### Methods

For each replication of a phenomenon, Cohen’s $d$ is computed as an
estimate of raw ES, according to Equation 1, including its standard
error (Equation 2). Subsequently, a random-effects meta-analysis is
performed using metafor version X.XX. To estimate meta-analytic mean ES
and heterogeneity, the REML-estimatore is used (ref). To identify which
phenomena pass criterion (3), using a Wald-type significance test we
assess for which phenomenon the meta-analytic mean ES is significantly
different from zero. For this hypothesis test, as all other hypothesis
tests in this manuscript, a significance level of .05 is used.

Additionally, estimates of score reliability are derived. While not
without its criticism (references), Cronbach’s Alpha is used to estimate
score reliability. It is often noted that tau-equivalence is an
unrealistic assumption to hold against real-world data, implying that
other estimates of score reliability, such as McDonald’s Omega
(reference), Guttman’s Lambda 2 or 4 (reference), or the Greatest Lower
Bound (reference) might be better suited. However, all scales employed
in this project are evaluated by computing the simple mean in responses
across items for each individual. Computing the mean of several items
implicitly assumes tau-equivalence, as each item contributes equally to
the individual’s test score. Therefore, Cronbach’s Alpha is the better
choice to estimate score reliability, as it avoids this mismatch in
assumptions between how the score is computed and how score reliability
is estimated. Using these estimates of Cronbach’s Alpha, the individual
estimates of Cohen’s d are corrected, according to Equation 4, including
the corrected standard errors (Equation 5). Thereby, estimates of ES are
computed, which are corrected for imperfect reliability. Again, a
random-effects meta-analysis is run using metafor, generating an
estimate of heterogeneity in corrected ES. Different indicators of
heterogeneity are readily available, such as $I^2$, $H^2$ or the
coefficient of variation. However, the descriptions found in H&S and W&D
discuss the absolute amount of heterogeneity in terms of variance
($\tau^2$) or standard deviation ($\tau$). These parameters discuss the
variability of the standardized ES in the population in terms of
variance or standard deviation. Therefore, we also make use of absolute
heterogeneity here.

Subsequently, in order to assess whether ES heterogeneity was indeed
reduced by the attenuation correction procedure, the estimates of
heterogeneity in uncorrected ES and corrected ES are compared. Using
Cochran’s Q-test, we assess whether the estimates of heterogeneity are
statistically significantly different from 0. Only for those projects
where we have sufficient confidence that the individual estimates of
heterogeneity are larger than zero can we actually begin to interpret
whether there was any change.

To identify whether a reduction in ES heterogeneity is indeed
accompanied by differences in score reliability, the estimates of score
reliability are assessed meta-analytically as well. To do so, a
Reliability Generalization Meta-Analysis is performed (references). This
entails that individual estimates of score reliability are adequately
transformed using Bonett’s transformation:
$T_{\hat{\rho}} = ln(1 - \hat{\rho})$. This transformation has
variance-stabilising properties and therefore allows for adequate
inferences concerning heterogeneity in score reliability (references).
Cochran’s Q-test is used to identify statistically significant
heterogeneity in transformed score reliability. Estimates derived from a
Reliability Generalization Meta-Analysis using these transformations can
be back-transformed to the original score reliability scale
(references):

$$\mu_{\rho{xx'}} \approx 1 - exp(\mu_{T\left[\rho_{XX^\prime}\right]})-\frac{1}{2} exp(\mu_{T\left[\rho_{xx'}\right]})\tau_{T\left[\rho_{XX^\prime}\right]}^2$${#eq-rel_back_mu}

$$\tau_{\rho_{xx'}}^2 \approx exp(\mu_{T\left[\rho_{xx'}\right]})2 \tau_{T\left[\rho_{xx'}\right]}^2+exp(\mu_{T\left[\rho_{xx'}\right]})^2\tau_{T\left[\rho_{xx'}\right]}^4$${#eq-rel_back_tau}

These procedures are followed separately for each phenomenon. The
statistical programming language R, Version X.XXX (reference) is used
for all data manipulation and statistical analysis.

### Results

Equations 1, 2, 4, and 5 were used to generate estimates of standardized
ES, both uncorrected and corrected, with their corresponding standard
errors. Generally, this leads to larger (absolute) ES and larger
standard errors. As an example from the 12 selected phenomena, the
results from this procedure on Nosek’s Explicit Art sex differences
phenomenon are displayed in the forest plot in Figure 1. Here, grey dots
represent the uncorrected ES for each sample, while black dots represent
the corrected ES in each sample. The bars surrounding these dots show
the respective 95%-Cis.

In Figure 1 it becomes apparent that the reliability attenuation
procedure leads to an increase in the individual absolute effect size,
as the black dots are moved further away from zero. Simultaneously, the
standard errors grew larger after the attenuation correction, which
leads to larger 95%-Cis. This is most easily observed for the
mturk-sample, which already had a rather large confidence interval to
begin with. Additionally, in the diamond at the bottom of the figure, we
can see that the meta-analytic estimate mean ES is also larger after the
attenuation correction took place. In Table 2, the estimates and tests
on the meta-analytic average ES, both corrected and uncorrected, can be
found.

*Table 2: Meta-analytic average  ES* \| ------- \| ------- \|
------- \| ------- \| ------- \| \| MASC \|
{\hat{\mathbit{\mu}}}*{*\mathbit{\delta}\mathbit{o}} \| p \|
{\hat{\mathbit{\mu}}}\_\mathbit{\delta} \| p \| \|
Albarracin_Priming_SAT \| .127 (.053) \| .016 \| .158 (.066) \| .016 \|
\| Alter_Analytic_Processing \| -.155 (.040) \| \<.001 \| -.174 (.044)
\| \<.001 \| \| Carter_Flag_Priming \| .018 (.029) \| .538 \| .026
(.036) \| .465 \| \| Caruso_Currency_Priming \| -.019 (.026) \| .457 \|
-.022 (.029) \| .459 \| \| Finkel_Exit_Forgiveness \| -.050 (.049) \|
.314 \| -.056 (.056) \| .316 \| \| Finkel_Neglect_Forgiveness \| -.051
(.055) \| .355 \| -.059 (.063) \| .347 \| \| Giessner_Vertical_Position
\| .026 (.023) \| .248 \| .027 (.025) \| .267 \| \|
Hart_Criminal_Intentionality \| .170 (.078) \| .030 \| .184 (.085) \|
.030 \| \| Hart_Detailed_Processing \| .082 (.058) \| .157 \| .096
(.067) \| .147 \| \| Hart_Intention_Attribution \| -.011 (.058) \| .856
\| -.021 (.080) \| .795 \| \| Husnu_Imagined_Contact \| .116 (.032) \|
\<.001 \| .129 (.035) \| \<.001 \| \| Nosek_Explicit_Art \| .360 (.047)
\| \<.001 \| .383 (.050) \| \<.001 \| \| Nosek_Explicit_Math \| .401
(.029) \| \<.001 \| .415 (.030) \| \<.001 \| \| PSACR001_anxiety_int \|
.251 (.022) \| \<.001 \| .263 (.023) \| \<.001 \| \| PSACR001_behav_int
\| .035 (.020) \| .079 \| .040 (.023) \| .088 \| \| PSACR002_neg_photo
\| -.663 (.040) \| \<.001 \| -.747 (.047) \| \<.001 \| \|
Shnabel_Willingness_Reconcile_Rev \| -.533 (.101) \| \<.001 \| -.582
(.109) \| \<.001 \| \| Shnabel_Willingness_Reconcile_RPP \| -.585 (.062)
\| \<.001 \| -.655 (.067) \| \<.001 \| \| Srull_Behaviour_Hostility \|
-.051 (.029) \| .079 \| -.062 (.034) \| .071 \| \|
Srull_Ronald_Hostility \| .063 (.029) \| .028 \| .074 (.033) \| .026 \|
\| Tversky_Directionality_Similarity1 \| -.111 (.034) \| .001 \| -.121
(.037) \| \<.001 \| \| Zhong_Desirability_Cleaning \| -.031 (.024) \|
.194 \| -.042 (.036) \| .240 \|

Table 2 demonstrates that what we observed in Figure 1 holds across all
22 phenomena. All meta-analytic effect sizes are larger after applying
an attenuation correction procedure. Similarly, the uncertainty
quantified in the estimates’ standard error is larger. Additionally,
Table 2 highlights which phenomena pass criterion (3) – the effect size
must be statistically significantly distinguishable from zero. 12
phenomena pass this criterion, highlighted in boldface. It may be
remarked that the attenuation correction procedure, in this data,
appears to have no influence on whether a phenomenon passes the
criterion. The differences in p-value are rather small and lead to no
difference in conclusion, regarding a significance level of .05.

*Table 3: Tests for ES Heterogeneity* \| \| Corrected ES \| \|
Corrected ES \| NA \| NA \| NA \| NA \| \| ------- \| ------- \| -------
\| ------- \| ------- \| ------- \| ------- \| ------- \| \| MASC \|
$\tau$ \| QE \| p \| \| $\tau$ \| QE \| p \| \| Albarracin_Priming_SAT
\| 0 \| 5.987 (8) \| .649 \| \| 0 \| 6.055 (8) \| .641 \| \|
Alter_Analytic_Processing \| .025 \| 15.706 (20) \| .735 \| \| .018 \|
15.701 (20) \| .735 \| \| Hart_Criminal_Intentionality \| .171 \| 18.96
(11) \| .062 \| \| .185 \| 18.81 (11) \| .065 \| \|
Husnu_Imagined_Contact \| .087 \| 47.581 (35) \| .076 \| \| .094 \|
47.351 (35) \| .079 \| \| Nosek_Explicit_Art \| .178 \| 62.771 (35) \|
.003 \| \| .19 \| 63.074 (35) \| .003 \| \| Nosek_Explicit_Math \| 0 \|
39.169 (35) \| .288 \| \| 0 \| 39.107 (35) \| .29 \| \|
PSACR001_anxiety_int \| .085 \| 74.655 (48) \| .008 \| \| .088 \| 74.251
(48) \| .009 \| \| PSACR002_neg_photo \| .22 \| 219.709 (36) \| \<.001
\| \| .26 \| 237.115 (36) \| \<.001 \| \| Shnabel\_ Reconcile_Rev \|
.233 \| 23.367 (7) \| .001 \| \| .253 \| 23.182 (7) \| .002 \| \|
Shnabel\_ Reconcile_RPP \| .073 \| 8.472 (7) \| .293 \| \| .071 \| 8.244
(7) \| .312 \| \| Srull_Ronald_Hostility \| .046 \| 24.756 (21) \| .258
\| \| .053 \| 24.593 (21) \| .265 \| \| Tversky\_ Similarity1 \| .001 \|
59.434 (60) \| .496 \| \| .001 \| 59.102 (60) \| .509 \|

Cases incompatible with claims made in W&D and H&S are highlighted using
boldface.

However, what Figure 1 and Table 2 can not inform us about is in how far
the heterogeneity has changed after correcting for (un)reliability. In
Table 3, the estimates and tests of heterogeneity concerning uncorrected
and corrected ES on the 12 remaining phenomena, which passed all
criteria (1) – (3), can be found. In the table, the estimate of
heterogeneity $\tau$, with the accompanying QE-test statistic, degrees
of freedom and its p-value are reported, separately for uncorrected and
corrected ES.

Most importantly, Table 3 demonstrates that of 12 phenomena assessed,
for seven phenomena we find patterns incompatible with claims made in
W&D and H&S. For those seven phenomena, the reliability attenuation
correction led to an increase in absolute ES heterogeneity $\tau$. For
the remaining five phenomena, estimates of ES heterogeneity were zero
before and after the correction procedure, leading to no change at all.
This leaves three phenomena where the attenuation correction did indeed
lead to a reduction of heterogeneity. However, it is crucial to note
that the hypothesis tests identified statistically significant
heterogeneity in only four out of 12 phenomena and for none of those a
reduction in heterogeneity took place. In all four cases of
statistically significant heterogeneity, the attenuation correction
procedure led to an increase in heterogeneity, contrary to the W&D and
H&S predictions.

To make sure that actual differences in score reliability exist,
heterogeneity in score reliability is assessed, making use of
Reliability Generalization Meta-Analysis (reference). Table 4 summarises
the results across all 12 phenomena. Overall, scales for almost all
phenomena appear have produced scores with an average score reliability
larger than .7. Only two phenomena (Albarracin_Priming_SAT and
Alter_Analytic_Processing) come with lower score reliability. However,
for those phenomena we could not identify statistically significant
heterogeneity in effect sizes in Table 2 in the first place. From the
phenomena where the ES heterogeneity grew larger as a result from the
attenuation correction procedure, for six out of seven phenomena, we
observe statistically significant heterogeneity in score reliability.
Only the Shnabel_Willingness_Reconcile_Rev project produced scores where
no heterogeneity in score reliability was identified. However, for this
project, it is crucial to note that only a small number of replications
(8), was available for analysis. Since the power of Cochran’s Q-test for
heterogeneity largely depends on the number of replications, this result
is inconclusive.

Despite this one inconclusive result, for the majority of phenomena
where ES heterogeneity is larger after applying an attenuation
correction, a lack of differences in score reliability can not be made
responsible for results contrary to claims made in W&D and H&S. All
seven phenomena came with an average score reliability from .74 to .9
and for six out of seven phenomena we identified statistically
significant heterogeneity in score reliability.

*Table 4: Reliability Generalization Meta-Analysis MASC* \| MASC \|
\mathbit{\mu}*{*\mathbit{\rho}{\mathbit{xx}\prime}} \|
\mathbit{\tau}*{*\mathbit{\rho}{\mathbit{xx}\prime}} \| QE \| p \| \|
------- \| ------- \| ------: \| ------- \| ------- \| \|
Albarracin_Priming_SAT \| .61 \[.557 : .657\] \| .057 \| 26.019 (9) \|
.001 \| \| Alter_Analytic_Processing \| .64 \[.604 : .673\] \| .056 \|
38.384 (20) \| .008 \| \| Hart_Criminal_Intentionality \| .849 \[.811 :
.879\] \| .054 \| 66.17 (11) \| \<.001 \| \| Husnu_Imagined_Contact \|
.814 \[.802 : .825\] \| .023 \| 66.389 (35) \| .001 \| \|
Nosek_Explicit_Art \| .881 \[.869 : .891\] \| .029 \| 142.551 (35) \|
\<.001 \| \| Nosek_Explicit_Math \| .936 \[.932 :.939\] \| .007 \| 59.43
(35) \| .006 \| \| PSACR001_anxiety_int \| .903 \[.89 : .914\] \| .041
\| 494.822 (48) \| \<.001 \| \| PSACR002_neg_photo \| .791 \[.774 :
.806\] \| .047 \| 350.852 (36) \| \<.001 \| \|
Shnabel_Willingness_Reconcile_Rev \| .832 \[.818 : .845\] \| 0 \| 5.053
(7) \| .653 \| \| Shnabel_Willingness_Reconcile_RPP \| .768 \[.737:
.796\] \| .032 \| 17.535 (7) \| .014 \| \| Srull_Ronald_Hostility \|
.739 \[.716 : .76\] \| .047 \| 91.722 (21) \| \<.001 \| \|
Tversky_Directionality_Similarity1 \| .854 \[.841 : .866\] \| .039 \|
164.933 (60) \| \<.001 \|

## Alternative discussion of how score reliability affects ES heterogeneity

In this section, we explore some facets on why we disagree with claims
stated in W&D and H&S. Additionally, we explore some alternative
explanations that might help understand why correcting ES for score
reliability does not lead to an inevitable reduction in heterogeneity.

### Misinterpreted equation in Hunter & Schmidt (XXXX)

Hunter & Schmidt claim to have derived an equation which demonstrates
how heterogeneity in score reliability inflates heterogeneity in ESs.
“If the level of reliability is independent of the true effect size
across studies, then, to a close approximation:” (p. 309)

$$\tau^2_{\delta_0} = [\mu_{\rho_{xx‘}}]^2 \tau^2_{\delta} + [\mu_{\delta}]^2 \tau^2_{\rho_{xx‘}}$${#eq-HS}

We adjusted Equation 9 using the notation used throughout this text.
Here, $\delta_0$ refers to the ES, undistorted by sampling error but not
corrected for measuring error, $\delta$ refers to the ES, undistorted by
sampling error and measurement error and ρxx‘ refers to score
reliability.

Equation 9 implies that heterogeneity in $\delta_0$ is essentially a
function where heterogeneity and mean value of $\delta$ are weighted by
mean score reliability ρxx‘ and its heterogeneity. While it may not be
self-evident in the equation itself, it is easy to construct cases where
heterogeneity in $\delta_0$ is inflated or deflated by heterogeneity in
ρxx‘, contradicting Hunter & Schmidt’s interpretation of this equation.
In Figure 2, the results of employing equation 9 to compute the
heterogeneity (variance) in corrected ES is highlighted. All else held
equal, only the individual level of mean reliability \$
\mu*{*\rho{xx‘}}\$ is varied from .5 to .8. Mean corrected ES
$\mu_{\delta}$ is held constant at 1, its variance $\tau^2_{\delta}$
held constant at .04 and the variance of score reliability
$\tau^2_{\rho_{xx‘}}$ held constant at .02.

*Figure 2*

In Figure 2, the red line corresponds to the variance of corrected ES.

If everything else is held constant, the variance in uncorrected ES can
be a function only of the mean level of score reliability
$\mu_{rho_{xx‘}}$, according to H&S equation. In that case, Figure 2
demonstrates that correcting for imperfect score reliability does not
inevitably lead to lower heterogeneity, as the line representing
heterogeneity in uncorrected ES does cross the red line, representing
the heterogeneity in corrected ES. Depending on the mean level of score
reliability, heterogeneity in uncorrected ES can be higher or lower than
heterogeneity in corrected ES.

However, Equation 9 is additionally flawed, as it violates a crucial
definition. As Hunter & Schmidt point out, Equation 9 is only valid “If
the level of reliability is independent of the true effect size across
studies \[…\]” (2014, p. 309). However, both parameters, ES $\delta$ and
score reliability ρxx‘, can not be independent parameters by definition.
ES $\delta$ is the “true” standardized mean difference, unaffected by
sampling or measurement error. The ES $\delta_0$ is standardized using
the total standard deviation, which is inflated by measurement error.
Equation 1, demonstrating how Cohen’s d is computed demonstrates how the
mean difference is standardized in this case. As demonstrated further
above in equation 4, ES $\delta$ is standardized using the true standard
deviation in the sense of CTT. The fact that the true standard deviation
parameter $\sigma_T$ is found both in equation 4 and equation 2
demonstrates that ES $\delta$ and score reliability $\rho_{xx‘}$ can not
be independent variables. Therefore, the basic premise of equation 9 is
not fulfilled, as the assumption of variable independence could not be
fulfilled.

### Describing ES as a random ratio variable

However, realising that an ES, as defined in Equation 1, is actually a
ratio, allows for a different analytical description of how differences
in score reliability can affect ES heterogeneity. If the phenomenon is
heterogeneous and score reliability varies across replications, it seems
sensible to assume that both numerator and denominator (MD and
$\sigma_X$) from Equation 1 vary across replications. In that case, an
ES can be described as a random variable stemming from a ratio
distribution. A ratio distribution is a probability distribution
constructed by dividing one random variable by a second random variable:
$Z = \frac{X}{Y}$ (reference). If the distribution of the components is
known, first order taylor approximation may be used to generate
estimates of mean and variance of the ratio variable (reference). Here
we use the Taylor-estimator of the ratio’s variance to demonstrate how
differences in score reliability may affect heterogeneity in ES.

$$\tau^2[Z] \approx \frac{\tau^2[X]}{\mu[Y]^2} - \frac{2\mu[X]}{\mu[Y]^3} cov[X,Y] + \frac{\mu[X]^2}{\mu[Y]^4} \tau^2[Y]$${#eq-XYZ_cov}

Assuming that the random variables are uncorrelated, simplifies the
equation to

$$\tau^2[Z] \approx \frac{\tau^2[X]}{\mu[Y]^2} + \frac{\mu[X]^2}{\mu[Y]^4} \tau^2[Y]$${#eq-XYZ}

Substituting for ES, unstandardized mean difference MD and pooled
standard deviation $\sigma_X$ leads to

$$\tau^2[\delta_0] \approx \frac{\tau^2[MD]}{\mu[\sigma_x]^2} + \frac{\mu[MD]^2}{\mu[\sigma_x]^4} var[SD]$${#eq-ratio_d0}

As stated in Equation 5, correcting ES for score reliability can be
described through a correction of the pooled standard deviation:
$\sigma_T = \sqrt{\rho_{xx’} \sigma^2_x}$. Since $\rho_{xx'}$ is always
between zero and one, in the case of imperfect score reliability,
$\sigma_T$ will always be smaller than total $\sigma_x$. Additionally,
if the heterogeneity in $\sigma_X$ was introduced purely by differences
in measurement precision – the error score variance $\sigma_E^2$, this
would be removed by the attenuation correction. However, alternatively,
the underlying latent variable we attempted to measure may not be
identically distributed across replications. In that case, true score
variance $\sigma^2_T$ would vary across replications, leading to
differences in both score reliability, corrected ES and raw ES.
Importantly, this implies that even if measurements of perfect score
reliability were taken across the different replications, heterogeneity
in ES would still persist.

In that case, some heterogeneity in $\sigma_T$ would remain, albeit
heterogeneity found in $\sigma_x$. Equation X essentially demonstrates
how mean and heterogeneity of MD, paired with mean and heterogeneity of
total score standard deviation $\sigma_x$ $\sigma_x$ can be used to
estimate heterogeneity in raw ES $\delta_0$. Similarly, we can adjust
Equation X to describe how, instead of total score standard deviation,
mean and variance in true score standard deviation $\sigma_T$ \sigma\_T
affect heterogeneity in corrected ES $\delta$.

$$\tau^2[\delta] \approx \frac{\tau^2[MD]}{\mu[\sigma_T]^2} + \frac{\mu[MD]^2}{\mu[\sigma_T]^4} \tau^2[\sigma_T]$${#eq-ratio_d}

The differences between the two Equations 12 and 13 demonstrates how ES
heterogeneity is affected by an attenuation correction procedure.
According to the quotes in H&S and W&D, Equation 12 should lead to a
larger value of $\tau^2\left[\delta_0\right]$, compared to the estimate
of $\tau^2\left[\delta\right]$ from Equation 13. This, however, does not
follow from any equations shown throughout the text.

What is guaranteed is that the expected value of $\sigma_T$ is smaller
than the expected value of $\sigma_X$. Similarly, the variance in
$\sigma_T$ is guaranteed to be smaller than the uncorrected variance in
$\sigma_X$. However, since the expected value of either standard
deviation is placed in the denominator of the function and the variance
of either is placed in the numerator of the function, Equation 9 is not
sufficient to explain how heterogeneity in ES changes due to the
attenuation correction. Instead, we introduce two additional metrics.

$$R_1 = \frac{\mu_{\sigma^2_T}}{\mu_{\sigma^2_X}}$${#eq-R1}

$R_1$ describes the relative size of mean true score variance, compared
to the mean observed score variance. This informs us, how much of the
mean total score variance can be attributed to mean true score variance.
Therefore, this metric is equivalent to the average score reliability.
Large values indicate that correcting the individual score variances (or
standard deviations for that matter) using the attenuation correction
procedure should lead to smaller change. Small values indicate the
opposite, large amounts of random error score variance on average lead
to large changes in mean score variance (or mean standard deviation).

$$R_2 = \frac{CV_{\sigma^2_T}}{CV_{\sigma^2_X}}$${#eq-R2}
The metric $R_2$ in
Equation 15 introduces the Coefficient of Variation
($CV = \frac{\tau}{\mu}$). The CV describes the amount of heterogeneity
of a parameters, relative to its mean value. Here, the CV describes how
much heterogeneity can be found in either true or observed score
variance, relative to how large either score variance is on average.
Since any variance component is bound to be larger than zero, its
heterogeneity can not be independent of its mean value. This means that
a smaller score variance inevitably needs to come with a smaller
heterogeneity, everything else held constant. Therefore, the absolute
heterogeneity $\tau$ is less informative for score variance
(components). The CV on the other hand can describe how much
heterogeneity is found in the score variance, relative to the mean
value. By conditioning on the mean value, the CV might therefore be a
more informative tool about the amount of heterogeneity in a score
variance (component).

The metric $R_2$ therefore describes, in how far the relative amount of
heterogeneity in the score variance after the attenuation correction,
the $CV_{\sigma^2_T}$, is different compared to the relative amount of
heterogeneity before the attenuation correction, the $CV_{\sigma^2_X}$.
A value smaller than 1 implies that the $CV_{\sigma^2_T}$ is smaller
than the $CV_{\sigma^2_X}$. The relative amount of heterogeneity is
reduced after the attenuation correction procedure. This would imply
that a substantial amount of the relative heterogeneity observed in
$\sigma^2_X$ was in fact due to differences in random error variance, or
differences in the measurement precision. As these differences are
corrected for, the relative heterogeneity in the score variance is
reduced. A value larger than 1, on the other hand, implies that the
$CV_{\sigma^2_T}$ is larger than the $CV_{\sigma^2_X}$. In that case,
the heterogeneity in score reliability can be less strongly attributed
to differences in random error variance but were actually due to
substantial heterogeneity in the true score variances $\sigma^2_X$.

In this section, we attempt to understand under which circumstances the
heterogeneity in ES grows larger as the result of an attenuation
correction procedure, contrary to claims made in W&D and H&S. To do so,
we propose the following inequality, involving Equations 8 and 9.
$$\tau^2_{\delta} > \tau^2_{\delta_0} \quad  \quad \quad \quad \quad 
\frac{\tau_{MD}^2}{\mu_{\sigma_T}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_T}^4} \tau^2_{\sigma_T} > \frac{\tau_{MD}^2}{\mu_{\sigma_x}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_x}^4} \tau^2_{\sigma_x}$${#eq-ineq_init}

Using the metrics $R_1$ and $R_2$, defined in Equations 14 and 15, we
attempt to understand under which circumstances the inequality defined
in Equation 16 holds true. However, while the metrics make use of mean
and heterogeneity of the score variances, Equation 15 makes use of mean
and heterogeneity of the standard deviations. In order to incorporate
the metrics, it is necessary to reparameterised the equation. Using the
delta method to approximate mean value and heterogeneity of variance
from those of the standard deviations, we know that (ref):
$$\mu_{\sigma^2_X} \approx \mu^2_{\sigma_X} \quad \quad  and \quad \quad 
CV_{\sigma^2} \approx 2 CV_{\sigma}$${#eq-delta_method}

While Equation 16 only explicitly contains parameters for the observed
score variance/standard deviation, the same can be done using the true
score variance/standard deviation. Using the approximates defined in
Equation 17 for observed and true score variance, we arrive at a new
inequality in Equation 18.

$$\frac{\tau_{MD}^2}{\mu_{\sigma^2_t}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_t^2}} 4 CV^2_{\sigma^2_t} \mu_{\sigma^2_t} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} 4 CV^2_{\sigma^2_x} \mu_{\sigma^2_x}$${#eq-ineq_CV}

18 Equation 18 describes that the heterogeneity in corrected ES is
larger than the heterogeneity in uncorrected ES, using parameters of the
distributions of mean differences MD, true score variances\*
$\sigma^2_T$ and observed score variances $\sigma^2_X$. By introducing
the metrics $R_1$ and $R_2$ to this inequality, we can begin to
disentangle which circumstances need to be fulfilled for the inequality
to hold. Rearranging the terms in Equations 14 and 15, and subsequently
entering these terms into Equation 18 leads to the following
$$ \frac{\tau_{MD}^2}{R_1 \mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} 4 R^2_2 CV^2_{\sigma^2_X} R_1 \mu_{\sigma^2_x} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} 4 CV^2_{\sigma^2_x} \mu_{\sigma^2_x}$${#eq-ineq_R}

Equation 15, again, describes the inequality that heterogeneity in
corrected ES is larger than the heterogeneity in uncorrected ES,
incorporating the metrics\* $R_1$ and $R_2$. This equation alone is not
sufficient to identify the relevant circumstances required for the
inequality to hold. However, rearranging the terms from Equation 19
leads to Equation 20:

$$ \tau_{MD}^2\left(\frac{1}{R_1} - 1\right) + \mu_{MD}^2 4 CV^2_{\sigma^2_X} \left(\frac{R^2_2}{R_1}  - 1\right) > 0$${#eq-ineq_fin}

The circumstances under which the inequality described in Equation 20
holds, are circumstances where the claims made in W&D and H&S are
directly contradicted, as in those cases the ES heterogeneity is larger
after applying an attenuation correction procedure. Concerning the terms
in Equation 20, we know that all terms left-hand of the inequality,
outside of the brackets ($\tau^2_{MD}$, $\mu^2_{MD}$ and
$CV^2_{\sigma^2_X}$), are bound to be positive. For the inequality to
hold, we need the left-hand side of the equation to remain positive,
larger than zero. We can distil two scenarios, under which this equation
should hold: (a) one of the terms inside the brackets
($\frac{1}{R_1} - 1$ and $\frac{R^2_2}{R_1}  - 1$) is positive and large
enough, so that the second term not containing that same bracket is
positively dominated by the first term; or (b) both terms inside the
brackets need to be positive.

Generally, we know that both $R_1$ and $R_2$ are bound to be positive,
as both contain different, strictly positive parameters of the
distributions of true and observed score variance. Additionally, we know
that $R_1$ is equivalent to the average score reliability, and therefore
bound between $0\le R_1\le1$. Therefore, the term inside the first
bracket is bound to be positive ($\frac{1}{R_1} – 1$). Since $R_2$ is
the ratio of CVs, or relative heterogeneity in true and total score
variance, it can very well be larger than 1, it is only restricted to be
positive. If $R_1$ is smaller than $R_2^2$, then the term inside the
second bracket turns negative. Generally, this means that the inequality
can only be violated if the squared reduction in relative heterogeneity
of score variance ($R_2^2$) is larger than the average score
reliability.

Concerning the two scenarios, for scenario (a) to hold true, since the
term involving only $R_1$ in the first bracket can only be positive, the
second term involving both metrics would need to be negative. For the
first term to positively dominate the second term, however, we would
need to make additional assumptions about the size of the parameters
outside of the brackets, specifically about $\tau^2_{MD}$ and
$\mu_{MD}$. This is something we would prefer to avoid, as it involves
additional assumptions about the individual raw effect sizes of the
different phenomena. Additionally, it can be said that as long as the
term in the second bracket involving both metrics is positive, scenario
(b) is true by default. This occurs any time $R_1 > R^2_2$. For example,
if the average score reliability is about .8, if $R_2$ is at least .89
or larger, the inequality holds true and ES heterogeneity grows larger
as a consequence of the attenuation correction procedure, contrary to
claims made in W&D and H&S. This means that for the exemplary score
reliability of .8, a reduction in relative heterogeneity in score
variance of 11% ($1 – R_2$) or more is necessary for scenario (b) to no
longer be true. The smaller the average score reliability, the smaller
the $R_2$-metric may be for the inequality to still hold true.

Keeping this in mind, even if $R^2_2$ is smaller than .8, it does not
immediately imply that ES heterogeneity is smaller in $\delta$. For
example, it is unlikely that an $R^2_2 = .79$ leads to a term that is
not positively dominated by the first term involving only $R_1$.
However, where exactly the cut-off is, when ES heterogeneity starts
shrinking due to the attenuation correction procedure, can not be easily
defined without additional assumptions about the distribution of $MD$.

Generally speaking, as long as the relative heterogeneity in
$\sigma^2_X$ does not substantially exceed the relative heterogeneity in
$\sigma^2_T$, heterogeneity in corrected ES will remain larger than
heterogeneity in uncorrected ES. Additionally, the proportion of
relative heterogeneity quantified in $R_2$ needs to lie below
$\sqrt{R_1}$, the square root of the average score reliability, for the
second term to be negative. Only if the second term is negative AND no
longer positively dominated by the first term, can the heterogeneity in
corrected ES shrink as a result of the attenuation correction procedure.

### Re-analysis of archival data

Table 3 showed that for all phenomena, where we identified statistically
significant ES heterogeneity, the heterogeneity was even larger after
the attenuation correction procedure. Subsequently, in Equation 16 we
demonstrate that if the $R_2$ metric is larger than
$\sqrt{\mu[\rho_{xx’}]}$, ES heterogeneity is bound to be larger after
applying said procedure. Table 5 reports the meta-analytic estimates of
observed and true score variances, including their relative
heterogeneity and metrics $R_1$ and $R_2$ across all 12 phenomena. In
Table 5 we see that for ten out of 12 phenomena, the $R_2$ metric is
larger than 1. This implies that for all measurements concerning these
phenomena, the relative heterogeneity in observed score variances could
not be substantially explained by differences in measurement quality.
Correcting for heterogeneity in error score variance has in fact only
increased the extent of relative heterogeneity present in the score
variances. As demonstrated in Equation 20, this implies that the ES
heterogeneity grows even larger, as we correct for differences in score
reliability.

In Table 3 we also found some phenomena where the heterogeneity did in
fact shrink after applying the attenuation correction procedure. This is
not reflected by what we observe in Table 5, as $R_2 > R_1$ for almost
all phenomena, which would imply that across all phenomena we expected
ES heterogeneity to grow as a consequence of the attenuation correction
procedure. However, it is crucial to realise that we did not find
statistically significant heterogeneity for any of the phenomena where
ES heterogeneity was reduced. This implies, that we actually could not
statistically distinguish the ES heterogeneity from zero. Therefore,
what seems to be a reduction in ES heterogeneity was most likely caused
by estimation issues, as we could not accurately estimate the extent of
ES heterogeneity, either before or after the attenuation correction.

*Table 5: Re-analysis concerning metrics $R_1$ and $R_2$*

| MASC | \mathbit{\mu}*{*\mathbit{\sigma}\mathbit{X}\^\mathbf{2}} | \mathbit{\mu}*{*\mathbit{\sigma}\mathbit{T}\^\mathbf{2}} | {\mathbit{CV}}*{*\mathbit{\sigma}\mathbit{X}\^\mathbf{2}} | {\mathbit{CV}}*{*\mathbit{\sigma}\mathbit{T}\^\mathbf{2}} | \mathbit{R}\_\mathbf{1} | \mathbit{R}\_\mathbf{2} |
|-----------|----------:|----------:|----------:|----------:|----------:|----------:|
| Albarracin_Priming_SAT | 0.024 | 0.015 | 0.155 | 0.25 | 0.618 | 1.614 |
| Alter_Analytic_Processing | 0.166 | 0.105 | 0 | NA | 0.632 | NA |
| Hart_Criminal_Intentionality | 2.195 | 1.865 | 0.354 | 0.41 | 0.85 | 1.16 |
| Husnu_Imagined_Contact | 3.336 | 2.725 | 0.145 | 0.177 | 0.817 | 1.217 |
| Nosek_Explicit_Art | 0.927 | 0.822 | 0.218 | 0.245 | 0.887 | 1.124 |
| Nosek_Explicit_Math | 1.74 | 1.629 | 0.059 | 0.063 | 0.936 | 1.058 |
| PSACR001_anxiety_int | 1.149 | 1.04 | 0.219 | 0.237 | 0.905 | 1.085 |
| PSACR002_neg_photo | 0.603 | 0.482 | 0.218 | 0.271 | 0.799 | 1.243 |
| Shnabel_Willingness_Reconcile_Rev | 1.055 | 0.878 | 0 | NA | 0.832 | NA |
| Shnabel_Willingness_Reconcile_RPP | 0.9 | 0.698 | 0.216 | 0.277 | 0.775 | 1.283 |
| Srull_Ronald_Hostility | 1.87 | 1.38 | 0.181 | 0.23 | 0.738 | 1.267 |
| Tversky_Directionality_Similarity1 | 4.474 | 3.836 | 0.275 | 0.32 | 0.857 | 1.16 |

Lastly, for two out of the 12 phenomena, we find a $CV_{\sigma^2_X}$ of
0, as the estimate of absolute heterogeneity $\tau_{\sigma^2_X}$ was 0
as well. For these phenomena, it is therefore also impossible to compute
the $R_2$ metric. While for the Alter_Analytic_Processing phenomenon we
did not identify statistically significant ES heterogeneity in the first
place, we did find some for the Shnabel_Willingness_Reconcile_Rev
phenomenon. At the same time, we did not identify statistically
significant heterogeneity in score reliability. With zero heterogeneity
in score variances, Equation 13 implies that the ES heterogeneity is
bound to grow larger due to the attenuation procedure. Also, this
phenomenon was assessed with a particularly low number of replications
(8). Most likely, adequate power to detect heterogeneity in score
reliability and observed score variance requires a larger number of
replications than adequate power to detect heterogeneity. In that case,
the results from Table 5, concerning the
Shnabel_Willingness_Reconcile_Rev phenomenon might alternatively be
explained by power-issues. Similarly, while Table 3 does not imply
specific expectations concerning the presence of score variance
heterogeneity in the Alter_Analytic_Processing phenomenon, the low
number of replications (20) makes it hard to distinguish whether the
estimate of $\tau_{\sigma^2_X} = 0$ is actually sensible or can also be
attributed to low power.

## Discussion

Across twelve exemplary archival datasets, we demonstrated that claims
made in H&S and W&D do not hold in empirical observations. Even though
removing the impact of score reliability on ES heterogeneity by means of
attenuation correction procedures, they do not lead to a reduction in
heterogeneity as claimed in H&S and W&D. Inferences derived from the
equation supplied in H&S (ref), Equation 5, also do not warrant such a
conclusion. Alternatively, describing standardised effect sizes in terms
of a ratio distribution and approximating its variance, we demonstrated
that differences in score reliability may both inflate or deflate
variance in ES, depending on how the relative heterogeneity in score
variance changes due to the attenuation correction procedure. The metric
$R_2$, compared with the metric $R_1$ or the average score reliability,
informs us on whether relative score variance heterogeneity decreases or
increases with the attenuation correction.

These results fit in with work recently published/preprinted by
Olsson-Collentine et al. (2023). In a large simulation scheme, they
demonstrate that differences in score reliability across administrations
typically deflate heterogeneity in uncorrected correlations. Only as the
true heterogeneity in correlations grows larger, score reliability
differences actually inflate heterogeneity as claimed in W&D. While
correlations and standardized effect sizes are not identical, the way
score (un)reliability affects these parameters is highly similar. Both
correlations and ES are deflated by (un)reliability, implying that an
attenuation correction increases mean values of these parameters, while
the impact on standard deviations used for standardisation purposes
needs to be checked via $R_2$.

The analytical arguments presented indicate that an attenuation
correction procedure, even if no differences between score reliability
exist, lead to larger ES heterogeneity. It appears that, as the
meta-analytical mean ES grows larger due to the correction procedure, so
does the ES heterogeneity, if homogeneity is present. If heterogeneity
is present on the other hand, and $R_2^2 < R_1$, still a reduction in ES
heterogeneity is not inevitable. It can be argued that, for the
differences in ES to grow smaller, the reduction in relative
heterogeneity in score variance needs to be substantially large, so the
increase in ES heterogeneity due to the larger mean ES can be
“outpaced”. However, we did not observe this pattern in the 12 phenomena
that allowed for such an analysis.

Heterogeneity in ES even in direct replications, where experimental
factors are held as constant as possible, is already substantial
(Renkewitz, Fünderich, & Beinhauer, 2024). However, so far we have
demonstrated that in several scenarios, differences in score reliability
may in fact be masking, or “deflating” ES heterogeneity. This implies
that the already substantial heterogeneity across direct replications
may very well be even larger, if ES were corrected for (un)reliability.
Understanding heterogeneity as an extent to which a theory or phenomenon
is understood, deflated estimates of heterogeneity imply that the theory
or phenomenon is understood even worse than initially assumed.

The empirical arguments presented are based on a rather small number of
non-representative data-sets. Based on the combination of analytical and
empirical arguments, we are convinced that these differences in score
reliability across replications oftentimes mask true heterogeneity. In
realistic meta-analytic studies of conceptual replications, data stems
from a variety of sources. Across different populations, different
research settings or designs are used with varying or adjusted research
instruments. It is unlikely that, across these replications, we can
expect either the measuring quality or the true score variance to be
stable. However, as demonstrated above, differences in score reliability
and true score variance tend to lead to larger corrected ES
heterogeneity, implying that observed ES heterogeneity was reduced by
differences in score reliability. However, whether these results
actually generalize beyond these data-sets remains to be seen.
Unfortunately, the behavioural sciences are in dire need of open-data
that resemble multi-site replications. As the majority of results
discussed over the last years (in e.g. ManyLabs or Registered
Replication Reports) employ single-indicator scales as dependent
measures, score reliability can not be easily estimated in order to
replicate our analyses.

The analytical arguments presented following Equation 17 rested on the
use of the delta-method, as Equation 17 contained parameters of the
score standard deviations’ distributions, while the $R_1$ and $R_2$
metrics were designed to discuss score variances. The delta-method rests
on a number of assumptions: (1) approximations derived from the
delta-method require for the transformed variable to follow a marginal
normal distribution (ref). Since the variable under discussion here is
the standard deviation, which is bounded to be larger than zero and has
a non-normal sampling distribution (ref), this assumption is most likely
violated; (2) the derivative of the transformation function needs to be
available. As the transformation function is essentially just the square
root, this assumption should hold up; (3) as the delta-method
approximates the parameters of the transformed variable using the
first-order Taylor-expansion, this implies that higher-order terms are
required to be negligible. In practice, this implies that the variance
of the untransformed variable needs to be small. However, as this entire
procedure discusses the influence of variance in standard deviations on
ES heterogeneity, introduced by differences in score reliability, this
assumption is probably violated as well. The violation of these
assumptions likely introduce a bias in estimates derived from these
approximations (ref). However, the inequality derived with use of the
delta-method is not used to actually derive any estimates, but to
discuss how an attenuation correction procedure can affect ES
heterogeneity. Therefore, we believe that the violation of these
assumptions is defensible in this case, as the analytical arguments
derived from the inequality should still hold up, nevertheless.

Similarly, all Equations following Equation 10 rest on the assumption
that mean difference ($MD$) and true or total score variance
($\sigma^2_T$ and $\sigma^2_X$) are independent. While this assumption
is in line with assumptions underlying CTT, it is by no means guaranteed
that it holds in these data-sets. If this assumption was violated,
whether the relationship was positive or negative, indicated by a
positive or negative covariance in Equation 10, would most likely have
an impact on how strongly ES heterogeneity is affected by the
attenuation correction procedure, possibly even the direction of how it
is affected. However, currently we have no reason to assume that there
is a systematic relationship between mean difference and standard
deviation. As this relationship would also affect the distribution of ES
like Cohen’s d itself, strong violations of this assumption would go far
beyond invalidating the claims we derived in Equation 16. Instead,
assumptions of the meta-analytic tests concerning the size ES and the
presence of ES heterogeneity would be violated, making any research on
change in ES heterogeneity due to attenuation correction obsolete.

The work done by Hunter and Schmidt (2014) was crucial in informing and
guiding methodology development in the field of meta-analyses.
Similarly, Wiernik and Dahlke (2020) raise a number of important points
that have been neglected in the application of meta-analytic research
over the last decades. We strongly agree with their ideas that
underappreciated (differences in un-)reliability introduce substantial
biases in meta-analytic estimates and tests in Psychology, which need to
be corrected. Future applications of meta-analyses on standardized ES
need to take score reliability into account to arrive at correct
estimates and inferences. However, the meta-analyst should not expect
that correcting ES for score reliability will reduce the extent of
heterogeneity identified.

### 2025-01-24 E0.5_Data-Extraction.R

@@ -12,7 +12,7 @@
 
 
 
-packages <- c("tidyverse", "here", "data.table", "haven")
+packages <- c("tidyverse", "here", "data.table", "haven", "readxl")
 
 # check, whether library already installed or not - install and load as needed:
 apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {

### 2025-01-24 E0_Data-Download.R

@@ -1,4 +1,17 @@
-packages <- c("magrittr", "osfr", "here")
+### Score ReLiability ###
+
+
+
+###################################################################################################
+# This script is used purely for downloading the data from the respective OSF-repositories.       #
+####################################################################################################
+
+
+# library loading and installing as necessary
+
+
+# relevant libraries required for this script
+packages <- c("magrittr", "osfr", "here", "zip", "data.table")
 
 # check, whether library already installed or not - install and load as needed:
 apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
@@ -15,6 +28,7 @@ apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
 })
 
 
+
 ### Many Labs 1
 
 # download the zipped-directory of Many Labs 1 data, from the OSF
@@ -29,14 +43,189 @@ unzip(here("Data/Downloaded Data/ML1_Datasets.zip"),
       exdir = here("Data/Original Data/ManyLabs1"),
       junkpaths = TRUE)
 
-### PSACR001
+### Many Labs 2
+
+# use osfr-package to download Many Labs 2 data, on IPD-level
+# Many Labs 2 data comes in two separate data-files, containing data on two "slates" (projects were
+#  distributed across slates). These are ML2_S1.csv (slate 1) and ML2_S2.csv (slate 2)
+osfr::osf_retrieve_file("https://osf.io/cwjp3") %>% 
+  osfr::osf_download(path = here("Data/Original Data/ManyLabs2"), conflicts = "overwrite")
+osfr::osf_retrieve_file("https://osf.io/jg9hc") %>% 
+  osfr::osf_download(path = here("Data/Original Data/ManyLabs2"), conflicts = "overwrite")
+
+### Many Labs 3
+
+# use osfr-package to download Many Labs 3 data, on IPD-level
+osfr::osf_retrieve_file("https://osf.io/bxw8j") %>%
+  osfr::osf_download(path = here("Data/Downloaded Data"), conflicts = "overwrite")
+# Data is downloaded in a .zip file to the Downloaded Data sub-folder
+
+# Unzip the zip folder and transfer the data to the Original Data sub-folder
+unzip(here("Data/Downloaded Data/Ml3 Final Data.zip"),
[…]

### 2025-01-24 E1_Data-Analysis_demonstrate.R

@@ -26,6 +26,7 @@ apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
 })
 
 
+source(here("ReLiability_Function-library.R"))
 
 data_files <- list.files(here("Data/Extracted (Project) Data"), full.names = TRUE)
 
@@ -38,10 +39,28 @@ data.list[[4]]$source <- substr(data.list[[4]]$source, start = nchar(data.list[[
 MASC_names <- substr(list.files(here("Data/Extracted (Project) Data")), 1, nchar(list.files(here("Data/Extracted (Project) Data")))-4) 
 
 
+effect_index <- MASC_names %in% c("Albarracin_Priming_SAT", "Alter_Analytic_Processing",
+                                  "Carter_Flag_Priming", "Caruso_Currency_Priming",
+                                  "Dijksterhuis_trivia", "Finkel_Exit_Forgiveness", 
+                                  "Finkel_Neglect_Forgiveness",
+                                  "Giessner_Vertical_Position", "Hart_Criminal_Intentionality",    
+                                  "Hart_Detailed_Processing", "Hart_Intention_Attribution",       
+                                  "Husnu_Imagined_Contact", "Nosek_Explicit_Art",
+                                  "Nosek_Explicit_Math", "PSACR001_anxiety_int", 
+                                  "PSACR001_behav_int", 
+                                  "PSACR002_neg_photo", "Shnabel_Willingness_Reconcile_Rev",
+                                  "Shnabel_Willingness_Reconcile_RPP", "Srull_Behaviour_Hostility",
+                                  "Srull_Ronald_Hostility", "Tversky_Directionality_Similarity1", 
+                                  "Zhong_Desirability_Cleaning"
+)
+
+
 # generate aggregates for each lab/country/source within a single meta-analysis. This is done by
 #  running an lapply-loop across all MASCs (meta-analyses, see MetaPipeX, Fünderich et al. 2024)
 #  and, nested within each loop, a second lapply-loop across all labs/countries/sources (samples)
-agg_L <- lapply(data.list, FUN = function(x){
[…]

### 2025-01-24 E1_Data-Analysis_demonstrate_graphics.R

@@ -34,7 +34,8 @@ ES_rma_df <- read.csv(here("Data/Processed/Aggregates_ES_analysis.csv"))
 
 agg_L <- readRDS(here("Data/Processed/Aggregates_simple.csv"))
 
-MASC_names <- substr(list.files(here("Data/Extracted (Project) Data")), 1, nchar(list.files(here("Data/Extracted (Project) Data")))-4) 
+MASC_names <- unique(ES_rma_df$MASC)
+
 
 # generate plots for meta-analyses of corrected AND uncorrected Cohen's d (slide 7)
 plots2 <- lapply((1:length(agg_L)), FUN = function(idx){
@@ -43,19 +44,19 @@ plots2 <- lapply((1:length(agg_L)), FUN = function(idx){
   name <- MASC_names[idx]
   
   # get aggregates per sample
-  y <- agg_L[[idx]]
+  y <- na.omit(agg_L[[idx]])
   
   # get rma-results separated
   rma_raw <- ES_rma_df[ES_rma_df$MASC == name & ES_rma_df$corr == 0,]
   rma_corr <- ES_rma_df[ES_rma_df$MASC == name & ES_rma_df$corr == 1,]
   
-  # get idea of minimum and maximum observations in MASC, to draw the respective distribution
+  # # get idea of minimum and maximum observations in MASC, to draw the respective distribution
   minmaxsequence_raw <- seq(from = min(y$d_raw), to = max(y$d_raw), length.out = 100)
   minmaxsequence_corr <- seq(from = min(y$d_corr), to = max(y$d_corr), length.out = 100)
-  
-  # Freedman-Diacons rule for binwidth:
-  bw_raw <- (2 * IQR(y$d_raw) * (length(y$d_raw)^-(1/3)))
-  bw_corr <- (2 * IQR(y$d_corr) * (length(y$d_corr)^-(1/3)))
+  # 
+  # # Freedman-Diacons rule for binwidth:
+  # bw_raw <- (2 * IQR(y$d_raw) * (length(y$d_raw)^-(1/3)))
+  # bw_corr <- (2 * IQR(y$d_corr) * (length(y$d_corr)^-(1/3)))
   
   # make plot
   p <- ggplot() +
@@ -90,19 +91,175 @@ plots2 <- lapply((1:length(agg_L)), FUN = function(idx){
 })
 
 # combine plots using pathwork-infrastructure
[…]

### 2025-01-24 E2_Data-Analysis_explore.R

@@ -48,13 +48,33 @@ data.list <- lapply(data_files, read.csv)
 data.list[[4]]$source <- substr(data.list[[4]]$source, start = nchar(data.list[[4]]$source) - 2, stop = nchar(data.list[[4]]$source))
 
 
+effect_index <- MASC_names %in% c("Albarracin_Priming_SAT", "Alter_Analytic_Processing",
+                                  "Carter_Flag_Priming", "Caruso_Currency_Priming",
+                                  "Dijksterhuis_trivia", "Finkel_Exit_Forgiveness", 
+                                  "Finkel_Neglect_Forgiveness",
+                                  "Giessner_Vertical_Position", "Hart_Criminal_Intentionality",    
+                                  "Hart_Detailed_Processing", "Hart_Intention_Attribution",       
+                                  "Husnu_Imagined_Contact", "Nosek_Explicit_Art",
+                                  "Nosek_Explicit_Math", "PSACR001_anxiety_int", 
+                                  "PSACR001_behav_int", 
+                                  "PSACR002_neg_photo", "Shnabel_Willingness_Reconcile_Rev",
+                                  "Shnabel_Willingness_Reconcile_RPP", "Srull_Behaviour_Hostility",
+                                  "Srull_Ronald_Hostility", "Tversky_Directionality_Similarity1", 
+                                  "Zhong_Desirability_Cleaning"
+)
+
+
+nn_eff_idx <- which(ES_rma_df$pval[ES_rma_df$corr == 0] <= .05)
+
+
+
 # generate estimates of ln-error score variance and its associated standard error using bootstrapping
-varE_est.L <- lapply(seq_along(data.list), FUN = function(x){
+varE_est.L <- lapply(which(effect_index)[nn_eff_idx], FUN = function(x){
   
   # sometimes calculation of SE might lead to issues, as negative variances can not be log-transformed
   #  therefore, function is run within a tryCatch-environment, so the script does not crash
   # apply_Bootstrap_SE_Project.specific is the project-specific function
-  tryCatch(apply_Bootstrap_SE_Project.specific(na.omit(data.list[[x]]), component = "E"),
[…]

### 2025-01-24 test_delta_assumptions.R

@@ -1,0 +1,207 @@
+
+
+
+
+
+# generate estimates of ln-observed score variance for each sample & projects
+varX_est.L <- lapply(which(effect_index)[nn_eff_idx], FUN = function(x){
+  
+  # sometimes calculation of SE might lead to issues, as negative variances can not be log-transformed
+  #  therefore, function is run within a tryCatch-environment, so the script does not crash
+  # apply_Bootstrap_SE_Project.specific is the project-specific function
+  tryCatch(apply_Bootstrap_SE_Project.specific(na.omit(data.list[[x]]), component = "X", R = 100),
+           
+           # print error to console
+           error = function(e)(cat("ERROR: ", conditionMessage(e), " - ",
+                                   substr(names(data.list), 
+                                          (regexpr("Project) Data/", names(data.list)) + 14), 
+                                          (nchar(names(data.list))-4))[x], 
+                                   " - ", x, "\n")))
+})
+
+
+
+
+# Perform random-effects meta-analysis on estimates of ln-observed score variance using metafor
+varX_rma.list <- lapply(seq_along(varX_est.L), FUN = function(x){
+  
+  # at times estimates of ln-varX & SE may be NA, as negative estimates can't be transformed
+  #  therefore, function is nested in tryCatch, so it doesn't break down with errors
+  tryCatch(metafor::rma(measure = "GEN", method = "REML", 
+                        yi = varX_est.L[[x]]$var.est, 
+                        sei = varX_est.L[[x]]$SE),
+           
+           # print error to console
+           error = function(e)(cat("ERROR: ", conditionMessage(e), " - ",
+                                   substr(names(data.list), 
+                                          (regexpr("Project) Data/", names(data.list)) + 14), 
+                                          (nchar(names(data.list))-4))[x], 
+                                   " - ", x, "\n")))
+  
+})
+
+
+
+vars_rma_L <- lapply(1:length(varE_rma.list), FUN = function(x){
+  
+  data.frame(mu_X = bt_var_m(varX_rma.list[[x]]),
+             tau2_X = bt_var_v(varX_rma.list[[x]]))
+  
+})
+
+
+
+
+
+
+
+
+
+apply_Bootstrap_SE_Project.specific2 <- function(data, R = 100, component = "E"){
+  
+  if(component == "E"){
[…]

### 2025-02-06 Manuscript_MD_ScoreReLiability.qmd
update MS

@@
 
+```{r, include = FALSE}
+# library loading and installing as necessary
+
+# relevant libraries required for this script
+packages <- c("ggplot2", "here", "metafor", "gridExtra", "knitr", "magrittr", "dplyr")
+
+# check, whether library already installed or not - install and load as needed:
+apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
+
+  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system
+
+  if(pkg_avail){
+    require(x, character.only = TRUE)             # load the library, if already installed
+
+  }else{
+    install.packages(x, repos = "https://ftp.gwdg.de/pub/misc/cran/")                           # install the library, if missing
+    require(x, character.only = TRUE)             # load after installation
+  }
+})
+
+source(here("ReLiability_Function-library.R"))
+
+ES_rma_df <- read.csv(here("Data/Processed/Aggregates_ES_analysis.csv"))
+
+agg_L <- readRDS(here("Data/Processed/Aggregates_simple.csv"))
+
+MASC_names <- unique(ES_rma_df$MASC)
+
+```
+
 # Score ReLiability – Can we explain effect size heterogeneity using score reliability?
@@
 Large-scale collaborative replication efforts have sparked discussions
-surrounding the replicability of psychological phenomena. However,
+surrounding the replicability of psychological phenomena. However, for
 attempts to estimate a phenomenon’s replicability or to predict whether
@@
 Hedge’s g were used. For the remainder of the article, Cohen’s d,
-defined in Equation 1, will be used as an exemplary estimate of ES, as
-it is used across a wide range of contexts and well understood by a
-broad audience.
+defined in @eq-d, will be used as an exemplary estimate of ES, as it is
+used across a wide range of contexts and well understood by a broad
+audience.
 
-$$d = \frac{MD}{\sigma_X}$${#eq-d}
+$$d = \frac{MD}{\sigma_X}$$ {#eq-d}
 
@@
 random-effects meta-analysis (reference) is computed as defined in
-Equation 2.
+@eq-SEd.
 
-$$SE_d=\sqrt{\frac{n_1+n_2}{n_1n_2}+\frac{d^2}{2\left(n_1+n_2\right)}}$${#eq-SEd}
+$$SE_d=\sqrt{\frac{n_1+n_2}{n_1n_2}+\frac{d^2}{2\left(n_1+n_2\right)}}$$ {#eq-SEd}
 
 As is widely known, score reliability affects ES. In the context of
-classical test theory (CTT) score reliability ρXX’ is defined as the
-ratio of true to total score variance, as defined in Equation 3.
+classical test theory (CTT) score reliability $\rho_{XX’}$ is defined as
+the ratio of true to total score variance, as defined in @eq-rel.
 
-$$\rho_{XX’} = \frac{\sigma^2_T}{\sigma^2_X}=\frac{\sigma^2_T}{\sigma^2_T + \sigma^2_E}$${#eq-rel}
+$$\rho_{XX’} = \frac{\sigma^2_T}{\sigma^2_X}=\frac{\sigma^2_T}{\sigma^2_T + \sigma^2_E}$$ {#eq-rel}
 
@@
 
-Previous
-discussions of heterogeneity in score reliability have exclusively
-discussed it as a parameter that inflates heterogeneity in ES, implying
-that, if score reliability was perfect across all replications, the
-actual heterogeneity would have been lower. In their discussion, Wiernik
-and Dahlke claim that „*Measurement error variance will impact the
-results of meta-analyses in three ways: by (a) biasing the mean effect
-size toward zero, (b) inflating effect-size heterogeneity and
-confounding moderator effects, and (c) confounding publication-bias and
-sensitivity analyses*” (p. 3, 2020). Additionally, more clearly, they
-state that “*If the studies included in a meta-analysis differ in their
-measures’ reliabilities, heterogeneity estimates will be artifactually
-inflated, erroneously suggesting larger potential moderator effects*” (p.
-4, Wiernik & Dahlke, 2020). They base their claims largely on work done
-by Hunter and Schmidt (2014), who claim that “*Variation in reliability
-across studies causes variation in the observed effect sizes above and
-beyond that produced by sampling error.*” (p. 302). Overall, both
-references imply that differences in score (un)reliability inevitably
-lead to an inflated heterogeneity in ES.
+Previous discussions of heterogeneity in score reliability have
+exclusively discussed it as a parameter that inflates heterogeneity in
+ES, implying that, if score reliability was perfect across all
+replications, the actual heterogeneity would have been lower. In their
+discussion, Wiernik and Dahlke claim that „*Measurement error variance
+will impact the results of meta-analyses in three ways: by (a) biasing
+the mean effect size toward zero, (b) inflating effect-size
+heterogeneity and confounding moderator effects, and (c) confounding
+publication-bias and sensitivity analyses*” (p. 3, 2020). Additionally,
+more clearly, they state that “*If the studies included in a
+meta-analysis differ in their measures’ reliabilities, heterogeneity
+estimates will be artifactually inflated, erroneously suggesting larger
+potential moderator effects*” (p. 4, Wiernik & Dahlke, 2020). They base
+their claims largely on work done by Hunter and Schmidt (2014), who
+claim that “*Variation in reliability across studies causes variation in
+the observed effect sizes above and beyond that produced by sampling
+error.*” (p. 302). Overall, both references imply that differences in
+score (un)reliability inevitably lead to an inflated heterogeneity in
+ES.
 
-As both sources Hunter and
-Schmidt (2014) and Wiernik and Dahlke (2020) will be referred to
-repeatedly, we abbreviate these references as H&S and W&D respectively.
+As both sources Hunter and Schmidt (2014) and Wiernik and Dahlke (2020)
+will be referred to repeatedly, we abbreviate these references as H&S
+and W&D respectively.
 
@@
 
-If information concerning score reliability is
-available, it is possible to correct the individual ES for its
-unreliability. This process is also known as attenuation correction and,
-while not without its criticisms (reference), is a widespread practice
-in Psychology (reference). Equation 4 describes a simple attenuation
-correction procedure.
+If information concerning score reliability is available, it is possible
+to correct the individual ES for its unreliability. This process is also
+known as attenuation correction and, while not without its criticisms
+(reference), is a widespread practice in Psychology (reference). @eq-dc
+describes a simple attenuation correction procedure.
 
-$$d_c = \frac{d}{\sqrt{\hat{\rho}_{xx’}}}$${#eq-dc}
+$$d_c = \frac{d}{\sqrt{\hat{\rho}_{xx’}}}$$ {#eq-dc}
 
@@
 constructed using an estimate of true score standard deviation, as in
-Equation 5.
+@eq-dct.
 
-$$d_c = \frac{MD}{\hat{\sigma}_{T}}=\frac{MD}{\sqrt{\hat{\rho}_{xx’} \hat{\sigma}_X^2}}$${#eq-dct}
+$$d_c = \frac{MD}{\hat{\sigma}_{T}}=\frac{MD}{\sqrt{\hat{\rho}_{xx’} \hat{\sigma}_X^2}}$$ {#eq-dct}
 
 $\hat{\sigma}_T$ can be estimated by simply rearranging the terms from
-equation 2. Thereby, an estimate of ES is generated, under the premise
-that score reliability would have been perfect ($\rho_{xx'} = 1$). As
-the corrected estimate of Cohen’s $d$ is larger and requires an
-additional unknown parameter, an estimate of score reliability, the
-larger uncertainty in this parameter should be reflected in a larger
-standard error. The standard error for $d_c$ is defined in equation 6.
+@eq-rel. Thereby, an estimate of ES is generated, under the premise that
+score reliability would have been perfect ($\rho_{xx'} = 1$). As the
+corrected estimate of Cohen’s $d$ is larger and requires an additional
+unknown parameter, an estimate of score reliability, the larger
+uncertainty in this parameter should be reflected in a larger standard
+error. The standard error for $d_c$ is defined in @eq-SEdc.
 
-$$SE_{d_c} = \sqrt{SE^2_d \frac{d_c}{d}}$${#eq-SEdc}
+$$SE_{d_c} = \sqrt{SE^2_d \frac{d_c}{d}}$$ {#eq-SEdc}
 
@@
 differences in score reliability. We close with a discussion on what
-future meta-analysts can expect to observe, when dealing with issues of
+future meta-analysts can expect, when dealing with issues of
 (un)reliability in assessments of meta-analyses of ES.
 
-## Re-analysis of archival data Data
+## Re-analysis of archival data
 
@@
 standardised format at \[osf.io\], the DRIPHT repository.
+
+### Data
 
@@
 score reliability affects heterogeneity in ES, selected phenomena need
-to fulfil three conditions: (1) As we focus on the use of ES $d$ d, the
+to fulfil three conditions: (1) As we focus on the use of ES $d$, the
 phenomenon needs to be assessed using a two-group design. This is the
@@
 For each replication of a phenomenon, Cohen’s $d$ is computed as an
-estimate of raw ES, according to Equation 1, including its standard
-error (Equation 2). Subsequently, a random-effects meta-analysis is
-performed using metafor version X.XX. To estimate meta-analytic mean ES
-and heterogeneity, the REML-estimatore is used (ref). To identify which
+estimate of raw ES, according to @eq-d, including its standard error
+(@eq-SEd). Subsequently, a random-effects meta-analysis is performed
+using metafor version X.XX. To estimate meta-analytic mean ES and
+heterogeneity, the REML-estimatore is used (ref). To identify which
 phenomena pass criterion (3), using a Wald-type significance test we
@@
 is estimated. Using these estimates of Cronbach’s Alpha, the individual
-estimates of Cohen’s d are corrected, according to Equation 4, including
-the corrected standard errors (Equation 5). Thereby, estimates of ES are
+estimates of Cohen’s d are corrected, according to @eq-dc, including the
+corrected standard errors (@eq-SEdc). Thereby, estimates of ES are
 computed, which are corrected for imperfect reliability. Again, a
@@
 
-$$\mu_{\rho{xx'}} \approx 1 - exp(\mu_{T\left[\rho_{XX^\prime}\right]})-\frac{1}{2} exp(\mu_{T\left[\rho_{xx'}\right]})\tau_{T\left[\rho_{XX^\prime}\right]}^2$${#eq-rel_back_mu}
+$$\mu_{\rho{xx'}} \approx 1 - exp(\mu_{T\left[\rho_{XX^\prime}\right]})-\frac{1}{2} exp(\mu_{T\left[\rho_{xx'}\right]})\tau_{T\left[\rho_{XX^\prime}\right]}^2$$ {#eq-rel_back_mu}
 
-$$\tau_{\rho_{xx'}}^2 \approx exp(\mu_{T\left[\rho_{xx'}\right]})2 \tau_{T\left[\rho_{xx'}\right]}^2+exp(\mu_{T\left[\rho_{xx'}\right]})^2\tau_{T\left[\rho_{xx'}\right]}^4$${#eq-rel_back_tau}
+$$\tau_{\rho_{xx'}}^2 \approx exp(\mu_{T\left[\rho_{xx'}\right]})2 \tau_{T\left[\rho_{xx'}\right]}^2+exp(\mu_{T\left[\rho_{xx'}\right]})^2\tau_{T\left[\rho_{xx'}\right]}^4$$ {#eq-rel_back_tau}
 
@@
 
-Equations 1, 2, 4, and 5 were used to generate estimates of standardized
-ES, both uncorrected and corrected, with their corresponding standard
-errors. Generally, this leads to larger (absolute) ES and larger
-standard errors. As an example from the 12 selected phenomena, the
-results from this procedure on Nosek’s Explicit Art sex differences
-phenomenon are displayed in the forest plot in Figure 1. Here, grey dots
-represent the uncorrected ES for each sample, while black dots represent
-the corrected ES in each sample. The bars surrounding these dots show
-the respective 95%-Cis.
+[Equations @eq-d], [-@eq-SEd], [-@eq-dc], and [-@eq-SEdc] were used to
+generate estimates of standardized ES, both uncorrected and corrected,
+with their corresponding standard errors. Generally, this leads to
+larger (absolute) ES and larger standard errors. As an example from the
+12 selected phenomena, the results from this procedure on Nosek’s
+Explicit Art sex differences phenomenon are displayed in the forest plot
+in Figure 1. Here, grey dots represent the uncorrected ES for each
+sample, while black dots represent the corrected ES in each sample. The
+bars surrounding these dots show the respective 95%-Cis.
 
-In Figure 1 it becomes apparent that the reliability attenuation
+In @fig-forest it becomes apparent that the reliability attenuation
 procedure leads to an increase in the individual absolute effect size,
@@
 can see that the meta-analytic estimate mean ES is also larger after the
-attenuation correction took place. In Table 2, the estimates and tests
-on the meta-analytic average ES, both corrected and uncorrected, can be
-found.
+attenuation correction took place. In @tbl-meanES, the estimates and
+tests on the meta-analytic average ES, both corrected and uncorrected,
+can be found.
 
-*Table 2: Meta-analytic average  ES* \| ------- \| ------- \|
-------- \| ------- \| ------- \| \| MASC \|
-{\hat{\mathbit{\mu}}}*{*\mathbit{\delta}\mathbit{o}} \| p \|
-{\hat{\mathbit{\mu}}}\_\mathbit{\delta} \| p \| \|
-Albarracin_Priming_SAT \| .127 (.053) \| .016 \| .158 (.066) \| .016 \|
-\| Alter_Analytic_Processing \| -.155 (.040) \| \<.001 \| -.174 (.044)
-\| \<.001 \| \| Carter_Flag_Priming \| .018 (.029) \| .538 \| .026
-(.036) \| .465 \| \| Caruso_Currency_Priming \| -.019 (.026) \| .457 \|
--.022 (.029) \| .459 \| \| Finkel_Exit_Forgiveness \| -.050 (.049) \|
-.314 \| -.056 (.056) \| .316 \| \| Finkel_Neglect_Forgiveness \| -.051
-(.055) \| .355 \| -.059 (.063) \| .347 \| \| Giessner_Vertical_Position
-\| .026 (.023) \| .248 \| .027 (.025) \| .267 \| \|
-Hart_Criminal_Intentionality \| .170 (.078) \| .030 \| .184 (.085) \|
-.030 \| \| Hart_Detailed_Processing \| .082 (.058) \| .157 \| .096
-(.067) \| .147 \| \| Hart_Intention_Attribution \| -.011 (.058) \| .856
-\| -.021 (.080) \| .795 \| \| Husnu_Imagined_Contact \| .116 (.032) \|
-\<.001 \| .129 (.035) \| \<.001 \| \| Nosek_Explicit_Art \| .360 (.047)
-\| \<.001 \| .383 (.050) \| \<.001 \| \| Nosek_Explicit_Math \| .401
-(.029) \| \<.001 \| .415 (.030) \| \<.001 \| \| PSACR001_anxiety_int \|
-.251 (.022) \| \<.001 \| .263 (.023) \| \<.001 \| \| PSACR001_behav_int
-\| .035 (.020) \| .079 \| .040 (.023) \| .088 \| \| PSACR002_neg_photo
-\| -.663 (.040) \| \<.001 \| -.747 (.047) \| \<.001 \| \|
-Shnabel_Willingness_Reconcile_Rev \| -.533 (.101) \| \<.001 \| -.582
-(.109) \| \<.001 \| \| Shnabel_Willingness_Reconcile_RPP \| -.585 (.062)
-\| \<.001 \| -.655 (.067) \| \<.001 \| \| Srull_Behaviour_Hostility \|
--.051 (.029) \| .079 \| -.062 (.034) \| .071 \| \|
-Srull_Ronald_Hostility \| .063 (.029) \| .028 \| .074 (.033) \| .026 \|
-\| Tversky_Directionality_Similarity1 \| -.111 (.034) \| .001 \| -.121
-(.037) \| \<.001 \| \| Zhong_Desirability_Cleaning \| -.031 (.024) \|
-.194 \| -.042 (.036) \| .240 \|
+```{r, echo = FALSE}
+#| label: fig-forest
+#| fig-cap: Forest Plot Nosek_Explicit_Art
+forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
+```
 
-Table 2 demonstrates that what we observed in Figure 1 holds across all
-22 phenomena. All meta-analytic effect sizes are larger after applying
-an attenuation correction procedure. Similarly, the uncertainty
+```{r, echo = FALSE}
+#| label: tbl-meanES
+#| tbl-cap: Meta-analytic average  ES
+knitr::kable(x = read.csv(here("Tables/Mean_ES.csv")))
+```
+
+@tbl-meanES demonstrates that what we observed in Figure 1 holds across
+all 22 phenomena. All meta-analytic effect sizes are larger after
+applying an attenuation correction procedure. Similarly, the uncertainty
 quantified in the estimates’ standard error is larger. Additionally,
-Table 2 highlights which phenomena pass criterion (3) – the effect size
-must be statistically significantly distinguishable from zero. 12
+@tbl-meanES highlights which phenomena pass criterion (3) – the effect
+size must be statistically significantly distinguishable from zero. 12
 phenomena pass this criterion, highlighted in boldface. It may be
@@
 
-*Table 3: Tests for ES Heterogeneity* \| \| Corrected ES \| \|
-Corrected ES \| NA \| NA \| NA \| NA \| \| ------- \| ------- \| -------
-\| ------- \| ------- \| ------- \| ------- \| ------- \| \| MASC \|
-$\tau$ \| QE \| p \| \| $\tau$ \| QE \| p \| \| Albarracin_Priming_SAT
-\| 0 \| 5.987 (8) \| .649 \| \| 0 \| 6.055 (8) \| .641 \| \|
-Alter_Analytic_Processing \| .025 \| 15.706 (20) \| .735 \| \| .018 \|
-15.701 (20) \| .735 \| \| Hart_Criminal_Intentionality \| .171 \| 18.96
-(11) \| .062 \| \| .185 \| 18.81 (11) \| .065 \| \|
-Husnu_Imagined_Contact \| .087 \| 47.581 (35) \| .076 \| \| .094 \|
-47.351 (35) \| .079 \| \| Nosek_Explicit_Art \| .178 \| 62.771 (35) \|
-.003 \| \| .19 \| 63.074 (35) \| .003 \| \| Nosek_Explicit_Math \| 0 \|
-39.169 (35) \| .288 \| \| 0 \| 39.107 (35) \| .29 \| \|
-PSACR001_anxiety_int \| .085 \| 74.655 (48) \| .008 \| \| .088 \| 74.251
-(48) \| .009 \| \| PSACR002_neg_photo \| .22 \| 219.709 (36) \| \<.001
-\| \| .26 \| 237.115 (36) \| \<.001 \| \| Shnabel\_ Reconcile_Rev \|
-.233 \| 23.367 (7) \| .001 \| \| .253 \| 23.182 (7) \| .002 \| \|
-Shnabel\_ Reconcile_RPP \| .073 \| 8.472 (7) \| .293 \| \| .071 \| 8.244
-(7) \| .312 \| \| Srull_Ronald_Hostility \| .046 \| 24.756 (21) \| .258
-\| \| .053 \| 24.593 (21) \| .265 \| \| Tversky\_ Similarity1 \| .001 \|
-59.434 (60) \| .496 \| \| .001 \| 59.102 (60) \| .509 \|
+```{r, echo = FALSE}
+#| label: tbl-tauES
+#| tbl-cap: Tests for ES Heterogeneity
+knitr::kable(x = read.csv(here("Tables/Heterogeneity_ES.csv")))
+```
 
@@
 
-However, what Figure 1 and Table 2 can not inform us about is in how far
-the heterogeneity has changed after correcting for (un)reliability. In
-Table 3, the estimates and tests of heterogeneity concerning uncorrected
-and corrected ES on the 12 remaining phenomena, which passed all
-criteria (1) – (3), can be found. In the table, the estimate of
-heterogeneity $\tau$, with the accompanying QE-test statistic, degrees
-of freedom and its p-value are reported, separately for uncorrected and
-corrected ES.
+However, what @fig-forest and @tbl-meanES can not inform us about is in
+how far the heterogeneity has changed after correcting for
+(un)reliability. In @tbl-tauES, the estimates and tests of heterogeneity
+concerning uncorrected and corrected ES on the 12 remaining phenomena,
+which passed all criteria (1) – (3), can be found. In the table, the
+estimate of heterogeneity $\tau$, with the accompanying QE-test
+statistic, degrees of freedom and its p-value are reported, separately
+for uncorrected and corrected ES.
 
-Most importantly, Table 3 demonstrates that of 12 phenomena assessed,
+Most importantly, @tbl-tauES demonstrates that of 12 phenomena assessed,
 for seven phenomena we find patterns incompatible with claims made in
@@
 heterogeneity in score reliability is assessed, making use of
-Reliability Generalization Meta-Analysis (reference). Table 4 summarises
-the results across all 12 phenomena. Overall, scales for almost all
-phenomena appear have produced scores with an average score reliability
-larger than .7. Only two phenomena (Albarracin_Priming_SAT and
-Alter_Analytic_Processing) come with lower score reliability. However,
-for those phenomena we could not identify statistically significant
-heterogeneity in effect sizes in Table 2 in the first place. From the
-phenomena where the ES heterogeneity grew larger as a result from the
-attenuation correction procedure, for six out of seven phenomena, we
-observe statistically significant heterogeneity in score reliability.
-Only the Shnabel_Willingness_Reconcile_Rev project produced scores where
-no heterogeneity in score reliability was identified. However, for this
-project, it is crucial to note that only a small number of replications
-(8), was available for analysis. Since the power of Cochran’s Q-test for
-heterogeneity largely depends on the number of replications, this result
-is inconclusive.
+Reliability Generalization Meta-Analysis (reference). @tbl-RGMA
+summarises the results across all 12 phenomena. Overall, scales for
+almost all phenomena appear have produced scores with an average score
+reliability larger than .7. Only two phenomena (Albarracin_Priming_SAT
+and Alter_Analytic_Processing) come with lower score reliability.
+However, for those phenomena we could not identify statistically
+significant heterogeneity in effect sizes in @tbl-meanES in the first
+place. From the phenomena where the ES heterogeneity grew larger as a
+result from the attenuation correction procedure, for six out of seven
+phenomena, we observe statistically significant heterogeneity in score
+reliability. Only the Shnabel_Willingness_Reconcile_Rev project produced
+scores where no heterogeneity in score reliability was identified.
+However, for this project, it is crucial to note that only a small
+number of replications (8), was available for analysis. Since the power
+of Cochran’s Q-test for heterogeneity largely depends on the number of
+replications, this result is inconclusive.
 
@@
 
-*Table 4: Reliability Generalization Meta-Analysis MASC* \| MASC \|
-\mathbit{\mu}*{*\mathbit{\rho}{\mathbit{xx}\prime}} \|
-\mathbit{\tau}*{*\mathbit{\rho}{\mathbit{xx}\prime}} \| QE \| p \| \|
-------- \| ------- \| ------: \| ------- \| ------- \| \|
-Albarracin_Priming_SAT \| .61 \[.557 : .657\] \| .057 \| 26.019 (9) \|
-.001 \| \| Alter_Analytic_Processing \| .64 \[.604 : .673\] \| .056 \|
-38.384 (20) \| .008 \| \| Hart_Criminal_Intentionality \| .849 \[.811 :
-.879\] \| .054 \| 66.17 (11) \| \<.001 \| \| Husnu_Imagined_Contact \|
-.814 \[.802 : .825\] \| .023 \| 66.389 (35) \| .001 \| \|
-Nosek_Explicit_Art \| .881 \[.869 : .891\] \| .029 \| 142.551 (35) \|
-\<.001 \| \| Nosek_Explicit_Math \| .936 \[.932 :.939\] \| .007 \| 59.43
-(35) \| .006 \| \| PSACR001_anxiety_int \| .903 \[.89 : .914\] \| .041
-\| 494.822 (48) \| \<.001 \| \| PSACR002_neg_photo \| .791 \[.774 :
-.806\] \| .047 \| 350.852 (36) \| \<.001 \| \|
-Shnabel_Willingness_Reconcile_Rev \| .832 \[.818 : .845\] \| 0 \| 5.053
-(7) \| .653 \| \| Shnabel_Willingness_Reconcile_RPP \| .768 \[.737:
-.796\] \| .032 \| 17.535 (7) \| .014 \| \| Srull_Ronald_Hostility \|
-.739 \[.716 : .76\] \| .047 \| 91.722 (21) \| \<.001 \| \|
-Tversky_Directionality_Similarity1 \| .854 \[.841 : .866\] \| .039 \|
-164.933 (60) \| \<.001 \|
+```{r, echo = FALSE}
+#| label: tbl-RGMA
+#| tbl-cap: Reliability Generalization Meta-Analysis
+knitr::kable(x = read.csv(here("Tables/Results_RMA_alpha.csv")))
+```
 
@@
 
-$$\tau^2_{\delta_0} = [\mu_{\rho_{xx‘}}]^2 \tau^2_{\delta} + [\mu_{\delta}]^2 \tau^2_{\rho_{xx‘}}$${#eq-HS}
+$$\tau^2_{\delta_0} = [\mu_{\rho_{xx‘}}]^2 \tau^2_{\delta} + [\mu_{\delta}]^2 \tau^2_{\rho_{xx‘}}$$ {#eq-HS}
 
-We adjusted Equation 9 using the notation used throughout this text.
-Here, $\delta_0$ refers to the ES, undistorted by sampling error but not
+We adjusted @eq-HS using the notation used throughout this text. Here,
+$\delta_0$ refers to the ES, undistorted by sampling error but not
 corrected for measuring error, $\delta$ refers to the ES, undistorted by
@@
 
-Equation 9 implies that heterogeneity in $\delta_0$ is essentially a
+@eq-HS implies that heterogeneity in $\delta_0$ is essentially a
 function where heterogeneity and mean value of $\delta$ are weighted by
@@
 ρxx‘, contradicting Hunter & Schmidt’s interpretation of this equation.
-In Figure 2, the results of employing equation 9 to compute the
-heterogeneity (variance) in corrected ES is highlighted. All else held
-equal, only the individual level of mean reliability \$
-\mu*{*\rho{xx‘}}\$ is varied from .5 to .8. Mean corrected ES
-$\mu_{\delta}$ is held constant at 1, its variance $\tau^2_{\delta}$
-held constant at .04 and the variance of score reliability
-$\tau^2_{\rho_{xx‘}}$ held constant at .02.
+In @fig-hs, the results of employing @eq-HS to compute the heterogeneity
+(variance) in corrected ES is highlighted. All else held equal, only the
+individual level of mean reliability \$ \mu\*{\*\\rho{xx‘}}\$ is varied
+from .5 to .8. Mean corrected ES $\mu_{\delta}$ is held constant at 1,
+its variance $\tau^2_{\delta}$ held constant at .04 and the variance of
+score reliability $\tau^2_{\rho_{xx‘}}$ held constant at .02.
 
-*Figure 2*
+```{r, echo = FALSE}
+#| label: fig-HS
+#| fig-cap: Implications by Hunter & Schmidt's @eg-HS
 
-In Figure 2, the red line corresponds to the variance of corrected ES.
+mu_delta <- 1
+var_delta <- .2^2
+mu_alpha <- seq(from = .5, to = .8, length.out = 100)
+var_alpha <- .14^2
+
+var_delta0 <- mu_alpha^2 * var_delta + mu_delta^2 * var_alpha
+
+ggplot() +
+  geom_hline(yintercept = .04, colour = "red", linewidth = 1, linetype = "dashed") +
+  geom_line(aes(y = var_delta0, x = mu_alpha), linewidth = 2) +
+  theme(panel.background = element_rect(fill = "transparent"), 
+        plot.background = element_rect(fill = "transparent", colour = "transparent"), 
+        panel.grid.major.y = element_line(colour = "grey"),
+        panel.grid.major.x = element_line(colour = "grey"),
+        panel.grid.minor = element_line(colour = "transparent"),
+        axis.ticks = element_line(colour = "grey"),
+        strip.background = element_rect(fill = "transparent")) +
+  labs(y = expression(tau^2~"["~delta[0]~"]"),
+       x = expression(mu~"["~rho[xx]~"]"))
+```
+
+In @fig-hs, the red line corresponds to the variance of corrected ES.
 
@@
 
-However, Equation 9 is additionally flawed, as it violates a crucial
-definition. As Hunter & Schmidt point out, Equation 9 is only valid “If
-the level of reliability is independent of the true effect size across
+However, @eq-hs is additionally flawed, as it violates a crucial
+definition. As Hunter & Schmidt point out, @eq-hs is only valid “If the
+level of reliability is independent of the true effect size across
 studies \[…\]” (2014, p. 309). However, both parameters, ES $\delta$ and
@@
 mean difference is standardized in this case. As demonstrated further
-above in equation 4, ES $\delta$ is standardized using the true standard
+above in @eg-dct, ES $\delta$ is standardized using the true standard
 deviation in the sense of CTT. The fact that the true standard deviation
-parameter $\sigma_T$ is found both in equation 4 and equation 2
-demonstrates that ES $\delta$ and score reliability $\rho_{xx‘}$ can not
-be independent variables. Therefore, the basic premise of equation 9 is
-not fulfilled, as the assumption of variable independence could not be
+parameter $\sigma_T$ is found both in @eq-dct and @eq-rel demonstrates
+that ES $\delta$ and score reliability $\rho_{xx‘}$ can not be
+independent variables. Therefore, the basic premise of @eq-hs is not
+fulfilled, as the assumption of variable independence could not be
 fulfilled.
@@
 
-However, realising that an ES, as defined in Equation 1, is actually a
-ratio, allows for a different analytical description of how differences
-in score reliability can affect ES heterogeneity. If the phenomenon is
+However, realising that an ES, as defined in @eq-d, is actually a ratio,
+allows for a different analytical description of how differences in
+score reliability can affect ES heterogeneity. If the phenomenon is
 heterogeneous and score reliability varies across replications, it seems
 sensible to assume that both numerator and denominator (MD and
-$\sigma_X$) from Equation 1 vary across replications. In that case, an
-ES can be described as a random variable stemming from a ratio
-distribution. A ratio distribution is a probability distribution
-constructed by dividing one random variable by a second random variable:
-$Z = \frac{X}{Y}$ (reference). If the distribution of the components is
-known, first order taylor approximation may be used to generate
-estimates of mean and variance of the ratio variable (reference). Here
-we use the Taylor-estimator of the ratio’s variance to demonstrate how
-differences in score reliability may affect heterogeneity in ES.
+$\sigma_X$) from @eq-d vary across replications. In that case, an ES can
+be described as a random variable stemming from a ratio distribution. A
+ratio distribution is a probability distribution constructed by dividing
+one random variable by a second random variable: $Z = \frac{X}{Y}$
+(reference). If the distribution of the components is known, first order
+taylor approximation may be used to generate estimates of mean and
+variance of the ratio variable (reference). Here we use the
+Taylor-estimator of the ratio’s variance to demonstrate how differences
+in score reliability may affect heterogeneity in ES.
 
-$$\tau^2[Z] \approx \frac{\tau^2[X]}{\mu[Y]^2} - \frac{2\mu[X]}{\mu[Y]^3} cov[X,Y] + \frac{\mu[X]^2}{\mu[Y]^4} \tau^2[Y]$${#eq-XYZ_cov}
+$$\tau^2[Z] \approx \frac{\tau^2[X]}{\mu[Y]^2} - \frac{2\mu[X]}{\mu[Y]^3} cov[X,Y] + \frac{\mu[X]^2}{\mu[Y]^4} \tau^2[Y]$$ {#eq-XYZ_cov}
 
@@
 
-$$\tau^2[Z] \approx \frac{\tau^2[X]}{\mu[Y]^2} + \frac{\mu[X]^2}{\mu[Y]^4} \tau^2[Y]$${#eq-XYZ}
+$$\tau^2[Z] \approx \frac{\tau^2[X]}{\mu[Y]^2} + \frac{\mu[X]^2}{\mu[Y]^4} \tau^2[Y]$$ {#eq-XYZ}
 
@@
 
-$$\tau^2[\delta_0] \approx \frac{\tau^2[MD]}{\mu[\sigma_x]^2} + \frac{\mu[MD]^2}{\mu[\sigma_x]^4} var[SD]$${#eq-ratio_d0}
+$$\tau^2[\delta_0] \approx \frac{\tau^2[MD]}{\mu[\sigma_x]^2} + \frac{\mu[MD]^2}{\mu[\sigma_x]^4} var[SD]$$ {#eq-ratio_d0}
 
-As stated in Equation 5, correcting ES for score reliability can be
+As stated in @eq-dct, correcting ES for score reliability can be
 described through a correction of the pooled standard deviation:
@@
 
-$$\tau^2[\delta] \approx \frac{\tau^2[MD]}{\mu[\sigma_T]^2} + \frac{\mu[MD]^2}{\mu[\sigma_T]^4} \tau^2[\sigma_T]$${#eq-ratio_d}
+$$\tau^2[\delta] \approx \frac{\tau^2[MD]}{\mu[\sigma_T]^2} + \frac{\mu[MD]^2}{\mu[\sigma_T]^4} \tau^2[\sigma_T]$$ {#eq-ratio_d}
 
-The differences between the two Equations 12 and 13 demonstrates how ES
-heterogeneity is affected by an attenuation correction procedure.
-According to the quotes in H&S and W&D, Equation 12 should lead to a
-larger value of $\tau^2\left[\delta_0\right]$, compared to the estimate
-of $\tau^2\left[\delta\right]$ from Equation 13. This, however, does not
+The differences between the two [Equations @eq-ratio_d0] and
+[-@eq-ratio_d] demonstrates how ES heterogeneity is affected by an
+attenuation correction procedure. According to the quotes in H&S and
+W&D, @eq-ratio_d0 should lead to a larger value of
+$\tau^2\left[\delta_0\right]$, compared to the estimate of
+$\tau^2\left[\delta\right]$ from @eq-ratio_d. This, however, does not
 follow from any equations shown throughout the text.
@@
 deviation is placed in the denominator of the function and the variance
-of either is placed in the numerator of the function, Equation 9 is not
+of either is placed in the numerator of the function, @eq-hs is not
 sufficient to explain how heterogeneity in ES changes due to the
@@
 
-$$R_1 = \frac{\mu_{\sigma^2_T}}{\mu_{\sigma^2_X}}$${#eq-R1}
+$$R_1 = \frac{\mu_{\sigma^2_T}}{\mu_{\sigma^2_X}}$$ {#eq-R1}
 
@@
 
-$$R_2 = \frac{CV_{\sigma^2_T}}{CV_{\sigma^2_X}}$${#eq-R2}
-The metric $R_2$ in
-Equation 15 introduces the Coefficient of Variation
-($CV = \frac{\tau}{\mu}$). The CV describes the amount of heterogeneity
-of a parameters, relative to its mean value. Here, the CV describes how
-much heterogeneity can be found in either true or observed score
-variance, relative to how large either score variance is on average.
-Since any variance component is bound to be larger than zero, its
-heterogeneity can not be independent of its mean value. This means that
-a smaller score variance inevitably needs to come with a smaller
-heterogeneity, everything else held constant. Therefore, the absolute
-heterogeneity $\tau$ is less informative for score variance
-(components). The CV on the other hand can describe how much
-heterogeneity is found in the score variance, relative to the mean
-value. By conditioning on the mean value, the CV might therefore be a
-more informative tool about the amount of heterogeneity in a score
-variance (component).
+$$R_2 = \frac{\tau^2_{\sigma^2_T}}{\tau^2_{\sigma^2_X}}$$ {#eq-R2}
 
-The metric $R_2$ therefore describes, in how far the relative amount of
-heterogeneity in the score variance after the attenuation correction,
-the $CV_{\sigma^2_T}$, is different compared to the relative amount of
-heterogeneity before the attenuation correction, the $CV_{\sigma^2_X}$.
-A value smaller than 1 implies that the $CV_{\sigma^2_T}$ is smaller
-than the $CV_{\sigma^2_X}$. The relative amount of heterogeneity is
-reduced after the attenuation correction procedure. This would imply
-that a substantial amount of the relative heterogeneity observed in
-$\sigma^2_X$ was in fact due to differences in random error variance, or
-differences in the measurement precision. As these differences are
-corrected for, the relative heterogeneity in the score variance is
-reduced. A value larger than 1, on the other hand, implies that the
-$CV_{\sigma^2_T}$ is larger than the $CV_{\sigma^2_X}$. In that case,
-the heterogeneity in score reliability can be less strongly attributed
-to differences in random error variance but were actually due to
-substantial heterogeneity in the true score variances $\sigma^2_X$.
+The metric $R_2$ on the other hand describes, in how far the attenuation
+correction procedure has successfully reduced heterogeneity in score
+variance $\tau^2_{\sigma^2_X}$. A value of 1 indicates that the
+heterogeneity in true score variance $\tau^2_{\sigma^2_T}$, the score
+variance after the attenuation correction, is essentially just as large
+as the heterogeneity initally observed. Smaller values indicate how much
+heterogeneity is "left", after applying an attenuation correction,
+relative to the inital heterogeneity. This means that a value of .7
+indicates that about 70% of heterogeneity in score variance remains,
+even after correcting for differences in score reliability. As only 30%
+of score variance heterogeneity could be removed, this would imply that
+differences in error score variance $\sigma^2_E$ were responsible for
+less than a third of the heterogeneity in score variances found. On the
+contrary, more than two thirds of heterogeneity could be attributed to
+actual differences in how the underlying true scores are distributed
+across samples.
 
@@
 correction procedure, contrary to claims made in W&D and H&S. To do so,
-we propose the following inequality, involving Equations 8 and 9.
+we propose the following inequality, involving [Equations @eq-ratio_d0]
+and [-@eq-ratio_d].
+
 $$\tau^2_{\delta} > \tau^2_{\delta_0} \quad  \quad \quad \quad \quad 
-\frac{\tau_{MD}^2}{\mu_{\sigma_T}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_T}^4} \tau^2_{\sigma_T} > \frac{\tau_{MD}^2}{\mu_{\sigma_x}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_x}^4} \tau^2_{\sigma_x}$${#eq-ineq_init}
+\frac{\tau_{MD}^2}{\mu_{\sigma_T}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_T}^4} \tau^2_{\sigma_T} > \frac{\tau_{MD}^2}{\mu_{\sigma_x}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_x}^4} \tau^2_{\sigma_x}$$ {#eq-ineq_init}
 
-Using the metrics $R_1$ and $R_2$, defined in Equations 14 and 15, we
-attempt to understand under which circumstances the inequality defined
-in Equation 16 holds true. However, while the metrics make use of mean
-and heterogeneity of the score variances, Equation 15 makes use of mean
-and heterogeneity of the standard deviations. In order to incorporate
-the metrics, it is necessary to reparameterised the equation. Using the
-delta method to approximate mean value and heterogeneity of variance
-from those of the standard deviations, we know that (ref):
+Using the metrics $R_1$ and $R_2$, defined in [Equations @eq-R1] and
+[-@eq-R2], we attempt to understand under which circumstances the
+inequality defined in @eq-ineq_init holds true. However, while the
+metrics make use of mean and heterogeneity of the score variances,
+@eq-ineq_init makes use of mean and heterogeneity of the standard
+deviations. In order to incorporate the metrics, it is necessary to
+reparameterised the equation. Using the delta method to approximate mean
+value and heterogeneity of variance from those of the standard
+deviations, we know that (ref):
 $$\mu_{\sigma^2_X} \approx \mu^2_{\sigma_X} \quad \quad  and \quad \quad 
-CV_{\sigma^2} \approx 2 CV_{\sigma}$${#eq-delta_method}
+CV_{\sigma^2} \approx 2 CV_{\sigma}$$ {#eq-delta_method}
 
-While Equation 16 only explicitly contains parameters for the observed
-score variance/standard deviation, the same can be done using the true
-score variance/standard deviation. Using the approximates defined in
-Equation 17 for observed and true score variance, we arrive at a new
-inequality in Equation 18.
+While @eq-delta_method only explicitly contains parameters for the
+observed score variance/standard deviation, the same can be done using
+the true score variance/standard deviation. Using the approximates
+defined in @eq-delta_method for observed and true score variance, we
+arrive at a new inequality in @eq-ineq_CV.
 
-$$\frac{\tau_{MD}^2}{\mu_{\sigma^2_t}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_t^2}} 4 CV^2_{\sigma^2_t} \mu_{\sigma^2_t} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} 4 CV^2_{\sigma^2_x} \mu_{\sigma^2_x}$${#eq-ineq_CV}
+$$\frac{\tau_{MD}^2}{\mu_{\sigma^2_t}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_t^2}} \frac{1}{4} CV^2_{\sigma^2_t} \mu_{\sigma^2_t} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{1}{4} CV^2_{\sigma^2_x} \mu_{\sigma^2_x}$$ {#eq-ineq_CV}
 
-18 Equation 18 describes that the heterogeneity in corrected ES is
-larger than the heterogeneity in uncorrected ES, using parameters of the
+@eq-ineq_CV describes that the heterogeneity in corrected ES is larger
+than the heterogeneity in uncorrected ES, using parameters of the
 distributions of mean differences MD, true score variances\*
@@
 disentangle which circumstances need to be fulfilled for the inequality
-to hold. Rearranging the terms in Equations 14 and 15, and subsequently
-entering these terms into Equation 18 leads to the following
-$$ \frac{\tau_{MD}^2}{R_1 \mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} 4 R^2_2 CV^2_{\sigma^2_X} R_1 \mu_{\sigma^2_x} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} 4 CV^2_{\sigma^2_x} \mu_{\sigma^2_x}$${#eq-ineq_R}
+to hold. Rearranging the terms in [Equations @eq-R1] and [-@eq-R2], and
+subsequently entering these terms into Equation 18 leads to the
+following
+$$ \frac{\tau_{MD}^2}{R_1 \mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} \frac{1}{4} R^2_2 CV^2_{\sigma^2_X} R_1 \mu_{\sigma^2_x} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{1}{4} CV^2_{\sigma^2_x} \mu_{\sigma^2_x}$$ {#eq-ineq_R}
 
-Equation 15, again, describes the inequality that heterogeneity in
+@eq-ineq_R, again, describes the inequality that heterogeneity in
 corrected ES is larger than the heterogeneity in uncorrected ES,
-incorporating the metrics\* $R_1$ and $R_2$. This equation alone is not
+incorporating the metrics $R_1$ and $R_2$. This equation alone is not
 sufficient to identify the relevant circumstances required for the
-inequality to hold. However, rearranging the terms from Equation 19
-leads to Equation 20:
+inequality to hold. However, rearranging the terms from @eq-ineq_R leads
+to @eq-ineq_fin:
 
-$$ \tau_{MD}^2\left(\frac{1}{R_1} - 1\right) + \mu_{MD}^2 4 CV^2_{\sigma^2_X} \left(\frac{R^2_2}{R_1}  - 1\right) > 0$${#eq-ineq_fin}
+$$ \tau_{MD}^2\left(\frac{1}{R_1} - 1\right) + \mu_{MD}^2 \frac{1}{4} CV^2_{\sigma^2_X} \left(\frac{R_2}{R_1^3}  - 1\right) > 0$$ {#eq-ineq_fin}
 
@@
 after applying an attenuation correction procedure. Concerning the terms
-in Equation 20, we know that all terms left-hand of the inequality,
+in @eq-ineq_fin, we know that all terms left-hand of the inequality,
 outside of the brackets ($\tau^2_{MD}$, $\mu^2_{MD}$ and
@@
 hold, we need the left-hand side of the equation to remain positive,
-larger than zero. We can distil two scenarios, under which this equation
-should hold: (a) one of the terms inside the brackets
+larger than zero. We can distill two scenarios, under which this
+equation should hold: (a) one of the terms inside the brackets
 ($\frac{1}{R_1} - 1$ and $\frac{R^2_2}{R_1}  - 1$) is positive and large
@@
 bound between $0\le R_1\le1$. Therefore, the term inside the first
-bracket is bound to be positive ($\frac{1}{R_1} – 1$). Since $R_2$ is
-the ratio of CVs, or relative heterogeneity in true and total score
-variance, it can very well be larger than 1, it is only restricted to be
-positive. If $R_1$ is smaller than $R_2^2$, then the term inside the
-second bracket turns negative. Generally, this means that the inequality
-can only be violated if the squared reduction in relative heterogeneity
-of score variance ($R_2^2$) is larger than the average score
-reliability.
+bracket is bound to be positive ($\frac{1}{R_1} – 1$). Similalry, $R_2$,
+as the ratio of true and error score variance heterogeneity, is bound
+between $0\le R_2\le1$. If $R_1^3$ is smaller than $R_2$, then the term
+inside the second bracket turns negative. Generally, this means that the
+inequality can only be violated if the reduction in relative
+heterogeneity of score variance ($R_2$) is larger than the the average
+score reliability to the power of 3.
 
@@
 term in the second bracket involving both metrics is positive, scenario
-(b) is true by default. This occurs any time $R_1 > R^2_2$. For example,
-if the average score reliability is about .8, if $R_2$ is at least .89
+(b) is true by default. This occurs any time $R_1^3 > R_2$. For example,
+if the average score reliability is about .8, if $R_2$ is at least .512
 or larger, the inequality holds true and ES heterogeneity grows larger
@@
 reliability of .8, a reduction in relative heterogeneity in score
-variance of 11% ($1 – R_2$) or more is necessary for scenario (b) to no
-longer be true. The smaller the average score reliability, the smaller
-the $R_2$-metric may be for the inequality to still hold true.
+variance of 49.8% ($1 – R_2$) or more is necessary for scenario (b) to
+no longer be true. The smaller the average score reliability, the
+smaller the $R_2$-metric may be for the inequality to still hold true.
 
-Keeping this in mind, even if $R^2_2$ is smaller than .8, it does not
+Keeping this in mind, even if $R_2$ is smaller than .512, it does not
 immediately imply that ES heterogeneity is smaller in $\delta$. For
-example, it is unlikely that an $R^2_2 = .79$ leads to a term that is
-not positively dominated by the first term involving only $R_1$.
-However, where exactly the cut-off is, when ES heterogeneity starts
-shrinking due to the attenuation correction procedure, can not be easily
-defined without additional assumptions about the distribution of $MD$.
+example, it is unlikely that an $R_2 = .51$ leads to a term that is not
+positively dominated by the first term involving only $R_1$. However,
+where exactly the cut-off is, when ES heterogeneity starts shrinking due
+to the attenuation correction procedure, can not be easily defined
+without additional assumptions about the distribution of $MD$.
 
-Generally speaking, as long as the relative heterogeneity in
-$\sigma^2_X$ does not substantially exceed the relative heterogeneity in
-$\sigma^2_T$, heterogeneity in corrected ES will remain larger than
-heterogeneity in uncorrected ES. Additionally, the proportion of
-relative heterogeneity quantified in $R_2$ needs to lie below
-$\sqrt{R_1}$, the square root of the average score reliability, for the
-second term to be negative. Only if the second term is negative AND no
-longer positively dominated by the first term, can the heterogeneity in
-corrected ES shrink as a result of the attenuation correction procedure.
+Generally speaking, the attenuation correction leads to a reduction both
+in mean value of $\sigma^2_X$ and in its heterogeneity. While a
+reduction in the former typically leads to an increase in heterogeneity
+of $d_c$, the latter has the potential to compensate for that increase
+and thereby reduce that heterogeneity after the attenuation correction.
+This reduction in ES heterogeneity can only occur, if the reduction in
+score variance heterogeneity is substantially stronger than the
+reduction in mean score variance. An extreme version of such a situation
+might occur, if the true score variance is identical across all samples.
+This would imply that all heterogeneity in score variance identified can
+be attributed to differences in how much random noise is found in the
+scores across the samples. In such a case, $R_2$ would be zero and a
+reduction in ES heterogeneity due to the attenuation correction would be
+guaranteed. This also implies that the lower $R_2$ and the larger the
+average score reliability $R_1$, the more likely it is that the
+attenuation correction actually does lead to a reduction in ES
+heterogeneity.
 
@@
 
-In Table 3 we also found some phenomena where the heterogeneity did in
-fact shrink after applying the attenuation correction procedure. This is
-not reflected by what we observe in Table 5, as $R_2 > R_1$ for almost
-all phenomena, which would imply that across all phenomena we expected
-ES heterogeneity to grow as a consequence of the attenuation correction
-procedure. However, it is crucial to realise that we did not find
-statistically significant heterogeneity for any of the phenomena where
-ES heterogeneity was reduced. This implies, that we actually could not
-statistically distinguish the ES heterogeneity from zero. Therefore,
+In @tbl-tauES we also found some phenomena where the heterogeneity did
+in fact shrink after applying the attenuation correction procedure. This
+is not reflected by what we observe in @tbl-R1R2, as $R_2 > R_1$ for
+almost all phenomena, which would imply that across all phenomena we
+expected ES heterogeneity to grow as a consequence of the attenuation
+correction procedure. However, it is crucial to realise that we did not
+find statistically significant heterogeneity for any of the phenomena
+where ES heterogeneity was reduced. This implies, that we actually could
+not statistically distinguish the ES heterogeneity from zero. Therefore,
 what seems to be a reduction in ES heterogeneity was most likely caused
@@
 
-*Table 5: Re-analysis concerning metrics $R_1$ and $R_2$*
-
-| MASC | \mathbit{\mu}*{*\mathbit{\sigma}\mathbit{X}\^\mathbf{2}} | \mathbit{\mu}*{*\mathbit{\sigma}\mathbit{T}\^\mathbf{2}} | {\mathbit{CV}}*{*\mathbit{\sigma}\mathbit{X}\^\mathbf{2}} | {\mathbit{CV}}*{*\mathbit{\sigma}\mathbit{T}\^\mathbf{2}} | \mathbit{R}\_\mathbf{1} | \mathbit{R}\_\mathbf{2} |
-|-----------|----------:|----------:|----------:|----------:|----------:|----------:|
-| Albarracin_Priming_SAT | 0.024 | 0.015 | 0.155 | 0.25 | 0.618 | 1.614 |
-| Alter_Analytic_Processing | 0.166 | 0.105 | 0 | NA | 0.632 | NA |
-| Hart_Criminal_Intentionality | 2.195 | 1.865 | 0.354 | 0.41 | 0.85 | 1.16 |
-| Husnu_Imagined_Contact | 3.336 | 2.725 | 0.145 | 0.177 | 0.817 | 1.217 |
-| Nosek_Explicit_Art | 0.927 | 0.822 | 0.218 | 0.245 | 0.887 | 1.124 |
-| Nosek_Explicit_Math | 1.74 | 1.629 | 0.059 | 0.063 | 0.936 | 1.058 |
-| PSACR001_anxiety_int | 1.149 | 1.04 | 0.219 | 0.237 | 0.905 | 1.085 |
-| PSACR002_neg_photo | 0.603 | 0.482 | 0.218 | 0.271 | 0.799 | 1.243 |
-| Shnabel_Willingness_Reconcile_Rev | 1.055 | 0.878 | 0 | NA | 0.832 | NA |
-| Shnabel_Willingness_Reconcile_RPP | 0.9 | 0.698 | 0.216 | 0.277 | 0.775 | 1.283 |
-| Srull_Ronald_Hostility | 1.87 | 1.38 | 0.181 | 0.23 | 0.738 | 1.267 |
-| Tversky_Directionality_Similarity1 | 4.474 | 3.836 | 0.275 | 0.32 | 0.857 | 1.16 |
+```{r, echo = FALSE}
+#| label: tbl-R1R2
+#| tbl-cap: Re-analysis concerning metrics $R_1$ and $R_2$
+knitr::kable(read.csv(here("Tables/Variances_analysis.csv")))
+```
 

### 2025-02-06 E1_Data-Analysis_demonstrate_graphics.R

@@ -178,21 +178,24 @@ combined_plot3
 
 
 
-res_tab <- ES_rma_df[ES_rma_df$corr == 0,][nn_eff_idx,] %>% 
-  select(MASC, mu, se, pval, tau, QE, k, QEp) %>% 
-  mutate(mu_str = paste0(round(mu, 3), " (", round(se, 3), ")"),
-         QE_str = paste0(round(QE, 3), " (", k-1, ")"))
-
-MD_tab <- knitr::kable(res_tab %>% 
-               select(MASC, mu_str, pval, tau, QE_str, QEp), 
-             digits = 3, 
-             col.names = c("MASC", "$\\mu_{ES}$ (SE)", "p", "$\\tau$", "$Q_E$ (df)", "p"))
-
-writeLines(MD_tab, "C:/Users/Lukas/Downloads/markdown_table.txt")
-write.csv(res_tab %>% 
-            select(MASC, mu_str, pval, tau, QE_str, QEp), 
-          "C:/Users/Lukas/Downloads/markdown_table.csv",
-          row.names = F)
+# res_tab <- ES_rma_df[ES_rma_df$corr == 0,][nn_eff_idx,] %>% 
+#   select(MASC, mu, se, pval, tau, QE, k, QEp) %>% 
+#   mutate(mu_str = paste0(round(mu, 3), " (", round(se, 3), ")"),
+#          QE_str = paste0(round(QE, 3), " (", k-1, ")"))
+# 
+# MD_tab <- knitr::kable(res_tab %>% 
+#                select(MASC, mu_str, pval, tau, QE_str, QEp), 
+#              digits = 3, 
+#              col.names = c("MASC", "$\\mu_{ES}$ (SE)", "p", "$\\tau$", "$Q_E$ (df)", "p"))
+# 
+# writeLines(MD_tab, "C:/Users/Lukas/Downloads/markdown_table.txt")
+# write.csv(res_tab %>% 
+#             mutate_if(is.numeric, round, 3) %>% 
+#             mutate(pval = ifelse(pval < .001, yes = "<.001", no = pval),
+#                    QEp = ifelse(pval < .001, yes = "<.001", no = QEp)) %>% 
+#             select(MASC, mu_str, pval, tau, QE_str, QEp), 
+#           "C:/Users/Lukas/Downloads/markdown_table.csv",
+#           row.names = F)
 
[…]

### 2025-02-06 E2_Data-Analysis_explore.R

@@ -172,7 +172,7 @@ vars_rma_df <- do.call(rbind, vars_rma_L) %>%
 
 
 write.csv(vars_rma_df %>% 
-            select(MASC, mu_X, mu_T, CV_X, CV_T, R1, R2), here("Data/Processed/Variances_analysis.csv"), row.names = FALSE)
+            select(MASC, mu_X, mu_T, CV_X, CV_T, R1, R2), here("Tables/Variances_analysis.csv"), row.names = FALSE)
 
 
 agg_L <- lapply(agg_L, FUN = function(x){

### 2025-02-27 Manuscript_MD_ScoreReLiability.qmd
init laptop change

@@
 dichotomously, as responses were right or wrong (2 cases), or expressed
-agreement vs disagreement dichotomously (1 case). Subsequently, the 22
+agreement vs disagreement dichotomously (1 case). Subsequently, those
 phenomena need to be identified where the meta-analytic mean effect size
@@
 concerning the different phenomena and how they were measured can be
-found at \[osf-link\], while a brief summary can be found in Table 1.
+found at \[osf-link\], while a brief summary can be found in Table A.
 
@@
 using metafor version X.XX. To estimate meta-analytic mean ES and
-heterogeneity, the REML-estimatore is used (ref). To identify which
+heterogeneity, the REML-estimator is used (ref). To identify which
 phenomena pass criterion (3), using a Wald-type significance test we
@@
 Additionally, estimates of score reliability are derived. While not
-without its criticism (references), Cronbach’s Alpha is used to estimate
+without criticism (references), Cronbach’s Alpha is used to estimate
 score reliability. It is often noted that tau-equivalence is an
@@
 Bound (reference) might be better suited. However, all scales employed
-in this project are evaluated by computing the simple mean in responses
-across items for each individual. Computing the mean of several items
-implicitly assumes tau-equivalence, as each item contributes equally to
-the individual’s test score. Therefore, Cronbach’s Alpha is the better
-choice to estimate score reliability, as it avoids this mismatch in
-assumptions between how the score is computed and how score reliability
-is estimated. Using these estimates of Cronbach’s Alpha, the individual
-estimates of Cohen’s d are corrected, according to @eq-dc, including the
-corrected standard errors (@eq-SEdc). Thereby, estimates of ES are
-computed, which are corrected for imperfect reliability. Again, a
-random-effects meta-analysis is run using metafor, generating an
-estimate of heterogeneity in corrected ES. Different indicators of
-heterogeneity are readily available, such as $I^2$, $H^2$ or the
-coefficient of variation. However, the descriptions found in H&S and W&D
-discuss the absolute amount of heterogeneity in terms of variance
-($\tau^2$) or standard deviation ($\tau$). These parameters discuss the
-variability of the standardized ES in the population in terms of
-variance or standard deviation. Therefore, we also make use of absolute
-heterogeneity here.
+in this project are scored by computing the simple mean or sum of
+responses across items for each individual. Computing the mean or sum of
+several items implicitly assumes tau-equivalence, as each item
+contributes equally to the individual’s test score. Therefore,
+Cronbach’s Alpha is the better choice to estimate score reliability, as
+it avoids a mismatch in assumptions between how the score is computed
+and how score reliability is estimated. Using these estimates of
+Cronbach’s Alpha, the individual estimates of Cohen’s d are corrected,
+according to @eq-dc, including the corrected standard errors (@eq-SEdc).
+Thereby, estimates of ES are computed, which are corrected for imperfect
+reliability. Again, a random-effects meta-analysis is run using metafor,
+generating an estimate of heterogeneity in corrected ES. Different
+indicators of heterogeneity are readily available, such as $I^2$, $H^2$
+or the coefficient of variation. However, the descriptions found in H&S
+and W&D discuss the absolute amount of heterogeneity in terms of
+variance ($\tau^2$) or standard deviation ($\tau$). These parameters
+discuss the variability of the standardized ES in the population in
+terms of variance or standard deviation. Therefore, we also make use of
+absolute heterogeneity here.
 
@@
 standard errors grew larger after the attenuation correction, which
-leads to larger 95%-Cis. This is most easily observed for the
+leads to larger 95%-CIs. This is most easily observed for the
 mturk-sample, which already had a rather large confidence interval to
@@
 function where heterogeneity and mean value of $\delta$ are weighted by
-mean score reliability ρxx‘ and its heterogeneity. While it may not be
-self-evident in the equation itself, it is easy to construct cases where
-heterogeneity in $\delta_0$ is inflated or deflated by heterogeneity in
-ρxx‘, contradicting Hunter & Schmidt’s interpretation of this equation.
-In @fig-hs, the results of employing @eq-HS to compute the heterogeneity
-(variance) in corrected ES is highlighted. All else held equal, only the
-individual level of mean reliability \$ \mu\*{\*\\rho{xx‘}}\$ is varied
-from .5 to .8. Mean corrected ES $\mu_{\delta}$ is held constant at 1,
-its variance $\tau^2_{\delta}$ held constant at .04 and the variance of
-score reliability $\tau^2_{\rho_{xx‘}}$ held constant at .02.
+mean score reliability $\rho_{xx‘}$ and its heterogeneity. While it may
+not be self-evident in the equation itself, it is easy to construct
+cases where heterogeneity in $\delta_0$ is inflated or deflated by
+heterogeneity in $\rho_{xx‘}$, contradicting Hunter & Schmidt’s
+interpretation of this equation. In @fig-HS, the results of employing
+@eq-HS to compute the heterogeneity (variance) in corrected ES is
+highlighted. All else held equal, only the individual level of mean
+reliability $\mu_{\rho_{xx‘}}$ is varied from .5 to .8. Mean corrected
+ES $\mu_{\delta}$ is held constant at 1, its variance $\tau^2_{\delta}$
+held constant at .04 and the variance of score reliability
+$\tau^2_{\rho_{xx‘}}$ held constant at .02.
 
@@
 #| label: fig-HS
-#| fig-cap: Implications by Hunter & Schmidt's @eg-HS
+#| fig-cap: Implications by Hunter & Schmidt's @eq-HS
 
@@
 
-In @fig-hs, the red line corresponds to the variance of corrected ES.
+In @fig-HS, the dashed red line corresponds to the variance of corrected
+ES.
 
@@
 be a function only of the mean level of score reliability
-$\mu_{rho_{xx‘}}$, according to H&S equation. In that case, Figure 2
+$\mu_{\rho_{xx‘}}$, according to H&S equation. In that case, Figure 2
 demonstrates that correcting for imperfect score reliability does not
@@
 heterogeneity in uncorrected ES does cross the red line, representing
-the heterogeneity in corrected ES. Depending on the mean level of score
-reliability, heterogeneity in uncorrected ES can be higher or lower than
-heterogeneity in corrected ES.
-
-However, @eq-hs is additionally flawed, as it violates a crucial
-definition. As Hunter & Schmidt point out, @eq-hs is only valid “If the
+the heterogeneity in corrected ES. This implies that, as long as mean
+score reliability is below .72, an attenuation correction would lead to
+a reduction in ES heterogeneity. As mean score reliability grows larger
+than .72, however, the heterogeneity in uncorrected ES grows larger than
+the heterogeneity in corrected ES - an attenuation correction would
+therefore only increase the extent of ES heterogeneity. This
+demonstrates that, depending on the mean level of score reliability,
+heterogeneity in uncorrected ES can be higher or lower than
+heterogeneity in corrected ES. As heterogeneity in corrected ES is not
+always smaller, the interpretationsof H&S concerning this equation is
+erroneous.
+
+However, @eq-HS is additionally flawed, as it violates a crucial
+definition. As Hunter & Schmidt point out, @eq-HS is only valid “If the
 level of reliability is independent of the true effect size across
@@
 that ES $\delta$ and score reliability $\rho_{xx‘}$ can not be
-independent variables. Therefore, the basic premise of @eq-hs is not
+independent variables. Therefore, the basic premise of @eq-HS is not
 fulfilled, as the assumption of variable independence could not be
@@
 the underlying latent variable we attempted to measure may not be
-identically distributed across replications. In that case, true score
+distributed identically across replications. In that case, true score
 variance $\sigma^2_T$ would vary across replications, leading to
@@
 In that case, some heterogeneity in $\sigma_T$ would remain, albeit
-heterogeneity found in $\sigma_x$. Equation X essentially demonstrates
-how mean and heterogeneity of MD, paired with mean and heterogeneity of
-total score standard deviation $\sigma_x$ $\sigma_x$ can be used to
-estimate heterogeneity in raw ES $\delta_0$. Similarly, we can adjust
-Equation X to describe how, instead of total score standard deviation,
-mean and variance in true score standard deviation $\sigma_T$ \sigma\_T
+heterogeneity found in $\sigma_x$ was reduced. @eq-ratio_d0 essentially
+demonstrates how mean and heterogeneity of MD, paired with mean and
+heterogeneity of total score standard deviation $\sigma_x$ can be used
+to estimate heterogeneity in raw ES $\delta_0$. Similarly, we can adjust
+Equation @eq-ratio_d0 to describe how, instead of total score standard
+deviation, mean and variance in true score standard deviation $\sigma_T$
 affect heterogeneity in corrected ES $\delta$.
@@
 $\tau^2\left[\delta_0\right]$, compared to the estimate of
-$\tau^2\left[\delta\right]$ from @eq-ratio_d. This, however, does not
-follow from any equations shown throughout the text.
-
-What is guaranteed is that the expected value of $\sigma_T$ is smaller
-than the expected value of $\sigma_X$. Similarly, the variance in
-$\sigma_T$ is guaranteed to be smaller than the uncorrected variance in
-$\sigma_X$. However, since the expected value of either standard
-deviation is placed in the denominator of the function and the variance
-of either is placed in the numerator of the function, @eq-hs is not
-sufficient to explain how heterogeneity in ES changes due to the
-attenuation correction. Instead, we introduce two additional metrics.
+$\tau^2\left[\delta\right]$ from @eq-ratio_d. However, the equations
+shown throughout this text do not agree with this conclusion.
+
+Instead, what is guaranteed is that the expected value of $\sigma_T$ is
+smaller than the expected value of $\sigma_X$. Similarly, the
+heterogeneity in $\sigma_T$ is guaranteed to be smaller than the
+uncorrected heterogeneity in $\sigma_X$. However, since the expected
+value of either standard deviation is placed in the denominator of the
+function and the variance of either is placed in the numerator of the
+function, @eq-HS is not sufficient to explain how heterogeneity in ES
+changes due to the attenuation correction. Instead, we introduce two
+additional metrics.
 
@@
 mean total score variance can be attributed to mean true score variance.
-Therefore, this metric is equivalent to the average score reliability.
-Large values indicate that correcting the individual score variances (or
-standard deviations for that matter) using the attenuation correction
-procedure should lead to smaller change. Small values indicate the
-opposite, large amounts of random error score variance on average lead
-to large changes in mean score variance (or mean standard deviation).
+Therefore, this metric is close to the average score reliability. Even
+though formally $R_1$ and the meta-analytic mean score reliability
+$\mu_{\rho_{xx'}}$ are not the same value, in practice we observe little
+to no differences between the two parameters. Large values indicate that
+correcting the individual score variances (or standard deviations for
+that matter) using the attenuation correction procedure should lead to
+very little change, as most of the variation observed were due to
+genuine differences in true scores. Small values indicate the opposite,
+the observed score variance was due to a larger amount of random error
+score variance, leading to a stronger correction.
 
@@
 variance after the attenuation correction, is essentially just as large
-as the heterogeneity initally observed. Smaller values indicate how much
-heterogeneity is "left", after applying an attenuation correction,
-relative to the inital heterogeneity. This means that a value of .7
+as the heterogeneity initially observed. Smaller values indicate how
+much heterogeneity is "left", after applying an attenuation correction,
+relative to the initial heterogeneity. This means that a value of .7
 indicates that about 70% of heterogeneity in score variance remains,
@@
 
+Both metrics $R_1$ and $R_2$ make use of parameters of the distribution
+of true and total score variance, not standard deviations. This choice
+was made to facilitate the role of score reliability in the subsequent
+discussion, as score reliability itself is defined by variances, not
+standard deviations (see @eq-rel).
+
 In this section, we attempt to understand under which circumstances the
@@
 deviations. In order to incorporate the metrics, it is necessary to
-reparameterised the equation. Using the delta method to approximate mean
+reparameterise the equation. Using the delta method to approximate mean
 value and heterogeneity of variance from those of the standard
 deviations, we know that (ref):
-$$\mu_{\sigma^2_X} \approx \mu^2_{\sigma_X} \quad \quad  and \quad \quad 
-CV_{\sigma^2} \approx 2 CV_{\sigma}$$ {#eq-delta_method}
+$$\mu^2_{\sigma_X} \approx \mu_{\sigma^2_X} \quad \quad  and \quad \quad 
+\tau^2_{\sigma_X} \approx \frac{\tau^2_{\sigma_X^2}}{4\mu_{\sigma^2_X}}$$ {#eq-delta_method}
 
@@
 
-$$\frac{\tau_{MD}^2}{\mu_{\sigma^2_t}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_t^2}} \frac{1}{4} CV^2_{\sigma^2_t} \mu_{\sigma^2_t} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{1}{4} CV^2_{\sigma^2_x} \mu_{\sigma^2_x}$$ {#eq-ineq_CV}
+$$\frac{\tau_{MD}^2}{\mu_{\sigma^2_t}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_t^2}} \frac{\tau^2_{\sigma_t^2}}{4\mu_{\sigma^2_t}} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4\mu_{\sigma^2_x}}$$ {#eq-ineq_CV}
 
@@
 than the heterogeneity in uncorrected ES, using parameters of the
-distributions of mean differences MD, true score variances\*
-$\sigma^2_T$ and observed score variances $\sigma^2_X$. By introducing
-the metrics $R_1$ and $R_2$ to this inequality, we can begin to
-disentangle which circumstances need to be fulfilled for the inequality
-to hold. Rearranging the terms in [Equations @eq-R1] and [-@eq-R2], and
-subsequently entering these terms into Equation 18 leads to the
-following
-$$ \frac{\tau_{MD}^2}{R_1 \mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} \frac{1}{4} R^2_2 CV^2_{\sigma^2_X} R_1 \mu_{\sigma^2_x} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{1}{4} CV^2_{\sigma^2_x} \mu_{\sigma^2_x}$$ {#eq-ineq_R}
+distributions of mean difference MD, true score variance $\sigma^2_T$
+and observed score variance $\sigma^2_X$. By introducing the metrics
+$R_1$ and $R_2$ to this inequality, we can begin to disentangle which
+conditions need to be fulfilled for the inequality to hold. Rearranging
+the terms in [Equations @eq-R1] and [-@eq-R2], and subsequently entering
+these terms into Equation 18 leads to the following
+$$\frac{\tau_{MD}^2}{R_1 \mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} \frac{R_2 \tau^2_{\sigma_x^2}}{4 R_1 \mu_{\sigma^2_x}} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4\mu_{\sigma^2_x}}$$ {#eq-ineq_R}
 
@@
 bound between $0\le R_1\le1$. Therefore, the term inside the first
-bracket is bound to be positive ($\frac{1}{R_1} – 1$). Similalry, $R_2$,
+bracket is bound to be positive ($\frac{1}{R_1} – 1$). Similarly, $R_2$,
 as the ratio of true and error score variance heterogeneity, is bound
@@
 (b) is true by default. This occurs any time $R_1^3 > R_2$. For example,
-if the average score reliability is about .8, if $R_2$ is at least .512
-or larger, the inequality holds true and ES heterogeneity grows larger
-as a consequence of the attenuation correction procedure, contrary to
-claims made in W&D and H&S. This means that for the exemplary score
-reliability of .8, a reduction in relative heterogeneity in score
-variance of 49.8% ($1 – R_2$) or more is necessary for scenario (b) to
-no longer be true. The smaller the average score reliability, the
-smaller the $R_2$-metric may be for the inequality to still hold true.
+assuming the average score reliability is about .8, if $R_2$ is at least
+.512 or larger, the inequality holds true and ES heterogeneity grows
+larger as a consequence of the attenuation correction procedure,
+contrary to claims made in W&D and H&S. This means that for the
+exemplary score reliability of .8, a reduction in relative heterogeneity
+in score variance of 48.8% ($(1 – R_2)100\%$) or more is necessary for
+scenario (b) to no longer be true. The smaller the average score
+reliability, the smaller the $R_2$-metric may be for the inequality to
+still hold true.
 
@@
 immediately imply that ES heterogeneity is smaller in $\delta$. For
-example, it is unlikely that an $R_2 = .51$ leads to a term that is not
+example, it is unlikely that an $R_2 = .5$ leads to a term that is not
 positively dominated by the first term involving only $R_1$. However,
-where exactly the cut-off is, when ES heterogeneity starts shrinking due
-to the attenuation correction procedure, can not be easily defined
+where exactly the cut-off is, where ES heterogeneity starts shrinking
+due to the attenuation correction procedure, can not be easily defined
 without additional assumptions about the distribution of $MD$.
@@
 
-Table 3 showed that for all phenomena, where we identified statistically
-significant ES heterogeneity, the heterogeneity was even larger after
-the attenuation correction procedure. Subsequently, in Equation 16 we
-demonstrate that if the $R_2$ metric is larger than
-$\sqrt{\mu[\rho_{xx’}]}$, ES heterogeneity is bound to be larger after
-applying said procedure. Table 5 reports the meta-analytic estimates of
-observed and true score variances, including their relative
-heterogeneity and metrics $R_1$ and $R_2$ across all 12 phenomena. In
-Table 5 we see that for ten out of 12 phenomena, the $R_2$ metric is
-larger than 1. This implies that for all measurements concerning these
-phenomena, the relative heterogeneity in observed score variances could
-not be substantially explained by differences in measurement quality.
-Correcting for heterogeneity in error score variance has in fact only
-increased the extent of relative heterogeneity present in the score
-variances. As demonstrated in Equation 20, this implies that the ES
+@tbl-tauES showed that for all phenomena, where we identified
+statistically significant ES heterogeneity, the heterogeneity was even
+larger after the attenuation correction procedure. Subsequently, in
+Equation 16 we demonstrate that if the $R_2$ metric is larger than
+$R_1^3$, ES heterogeneity is bound to be larger after applying said
+procedure. @tbl-R1R2 reports the meta-analytic estimates of observed and
+true score variances, including their heterogeneity and metrics $R_1$
+and $R_2$ across all 12 phenomena. In @tbl-R1R2 we see that for ten out
+of 12 phenomena, the $R_2$ metric is larger than the $R_1$ metric.
+Additionally, we observe that across nine phenomena, $R_2$ is larger
+than .97. This implies that of the heterogeneity observed in score
+variances, 97% or more could not be explained by differences in random
+error introduced by the measuring instrument, but by genuine differences
+in the underlying distribution. Next to the two exceptional phenomena,
+where zero heterogeneity in score variance was found
+(Alter_Analytic_Processing and Shnabel_Willingness_Reconcile_Rev), only
+for the Srull_Ronald_Hostility phenomenon some heterogeneity could be
+attributed to differences in random error variance, albeit less than
+14%.
+
+This implies that for all measurements concerning these phenomena, the
+heterogeneity in observed score variances could not be substantially
+explained by differences in measurement quality. For no phenomenon, if
+differences in score variance could be found, was $R_2$ smaller than
+$R_1^3$. As demonstrated in Equation 20, this implies that the ES
 heterogeneity grows even larger, as we correct for differences in score
 reliability.
+
+```{r, echo = FALSE}
+#| label: tbl-R1R2
+#| tbl-cap: Re-analysis concerning metrics $R_1$ and $R_2$
+knitr::kable(read.csv(here("Tables/Variances_analysis.csv")))
+```
 
@@
 correction procedure. However, it is crucial to realise that we did not
-find statistically significant heterogeneity for any of the phenomena
+find statistically significant ES heterogeneity for any of the phenomena
 where ES heterogeneity was reduced. This implies, that we actually could
@@
 
-```{r, echo = FALSE}
-#| label: tbl-R1R2
-#| tbl-cap: Re-analysis concerning metrics $R_1$ and $R_2$
-knitr::kable(read.csv(here("Tables/Variances_analysis.csv")))
-```
-
-Lastly, for two out of the 12 phenomena, we find a $CV_{\sigma^2_X}$ of
-0, as the estimate of absolute heterogeneity $\tau_{\sigma^2_X}$ was 0
-as well. For these phenomena, it is therefore also impossible to compute
+Lastly, for two out of the 12 phenomena, we find a $\tau_{\sigma^2_X}$
+of 0. For these phenomena, it is therefore also impossible to compute
 the $R_2$ metric. While for the Alter_Analytic_Processing phenomenon we
@@
 heterogeneity as claimed in H&S and W&D. Inferences derived from the
-equation supplied in H&S (ref), Equation 5, also do not warrant such a
+equation supplied in H&S (ref), @eq-HS, also do not warrant such a
 conclusion. Alternatively, describing standardised effect sizes in terms
@@
 $R_2$, compared with the metric $R_1$ or the average score reliability,
-informs us on whether relative score variance heterogeneity decreases or
-increases with the attenuation correction.
+informs us in how far score variance heterogeneity decreases due to the
+attenuation correction.
 
@@
 does the ES heterogeneity, if homogeneity is present. If heterogeneity
-is present on the other hand, and $R_2^2 < R_1$, still a reduction in ES
+is present on the other hand and $R_2 < R_1^3$, still a reduction in ES
 heterogeneity is not inevitable. It can be argued that, for the
@@
 or phenomenon is understood even worse than initially assumed.
+
+The data-sets available to this project fall into two categories: direct
+multi-site replications of a specific phenomenon, employing identical
+protocols across samples; a large-scale collaborative effort to assess a
+single phenomenon globally, distributed across a large number of
+independent laboratories. For both categories, rather arbitrary sampling
+schemes were used. The majority of respondents are students at the
+respective universities, taking part in the studies voluntarily or
+because they were required as part of their study's programme (at times
+compensation was awarded). This is not the same as random sampling.
+These samples are put together arbitrarily, it is unlikely that the
+underlying distribution of true scores is meaningful for a properly
+specified population. The between-studies randomness introduced by these
+undefined sampling schemes manifests itself in true score variance
+heterogeneity $\tau^2_{\sigma^2_X}$. Between-studies randomness
+introduces by issues of measurement quality manifest themselves in error
+score variance heterogeneity $\tau^2_{\sigma^2_X }$. This implies that
+it should be of little surprise that the differences in true score
+variance largely exceed those in error score variance.
+
+Without sampling schemes that assure that the samples are constructed to
+be representative for some kind of population, we have no reason to
+assume that there should be few differences in the dispersion of true
+scores. Even though the measuring instruments are translated to
+different languages, which is bound to introduce differences in how much
+random noise is added to scores across samples, we have no reason to
+assume that these differences are larger than those introduced by
+employing no sampling schemes. Therein, we believe that, as the
+differences in score reliability are largely introduced by failure of
+ensuring appropriate sampling mechanisms, not by differences in how well
+the instrument captures the underlying construct, attenuation correction
+can not be expected to remedy ES heterogeneity.
+
+However, this also means that results contrary to what we observed in
+this project are not unthinkable. Valid sampling schemes need to be
+considered, ensuring that the samples are representative for populations
+that can be meaningfully compared. In that case, we are convinced that
+true score variance should be more stable across samples and differences
+in the instrument's measuring quality are bound to have a much larger
+impact. Not only would an attenuation correction in such a case actually
+have the potential of reducing ES heterogeneity, the variability
+underlying a psychological phenomenon could be meaningfully assessed
+across samples - something that would greatly help attempts to identify
+universal and non-universal patterns of psychological phenomena. As it
+stands, current approaches to multi-site direct replications and other
+collaborative efforts suffer strongly from the lack of appropriate
+sampling schemes.
 
@@
 derived from the inequality should still hold up, nevertheless.
+Additionally, we used the empirical examples available to assess in how
+far estimates derived using the delta-method approximations diverge from
+estimates derived without it. The differences identified are rather
+miniscule at the third decimal, reinforcing our belief that these
+assumption violations are defensible.
 

### 2025-03-05 Manuscript_MD_ScoreReLiability.qmd
update manuscript intro

@@
 title: "Manuscript"
-format: docx
+format: html
 editor: 
@@
 
-# Score ReLiability – Can we explain effect size heterogeneity using score reliability?
+# Liability or reliability? Exploring the role of effect size attenuation on meta-analytic heterogeneity
 
 Large-scale collaborative replication efforts have sparked discussions
-surrounding the replicability of psychological phenomena. However, for
-attempts to estimate a phenomenon’s replicability or to predict whether
-a future replication will replicate successfully, heterogeneity is a
-crucial parameter that needs to be assessed. In the meta-analytic
-context, heterogeneity describes the variation in effect sizes, free of
-sampling error, if these are not stable or constant across replications.
-Therefore, the presence of heterogeneity implies that some replications
-of a single phenomenon may be successful, while others are not. If
-heterogeneity grows, meaning the phenomenon’s effect size varies more
-strongly for unexplained reasons, the probability of observing an effect
-size around zero or even in the negative space grows larger as well. If
-we know the mean size and heterogeneity of a phenomenon’s effect size,
-we would theoretically be able to establish a baseline of expected
-replication rate (reference).
+surrounding the replicability of psychological phenomena. For attempts
+to estimate a phenomenon’s replicability or to predict whether a future
+replication will replicate successfully, heterogeneity is a crucial
+parameter that needs to be assessed. In the meta-analytic context,
+heterogeneity describes the variation in effect sizes, free of sampling
+error, if these are not constant across replications. Therefore, the
+presence of heterogeneity implies that some replications of a single
+phenomenon may be successful, while others are not. If heterogeneity
+grows, meaning the phenomenon’s effect size varies more strongly for
+unexplained reasons, the probability of observing an effect size around
+zero or even in the negative space grows larger as well. If we know the
+mean size and heterogeneity of a phenomenon’s effect size, we would
+theoretically be able to establish an expected replication rate
+(reference).
 
@@
 reflects a more advanced understanding of the subject matter being
-studied” (2021, p.2). Similarly, Schuetze and von Hippel (2023) argue
-that heterogeneity in effects is an indicator of a vague, poorly
+studied” (2021, p.2). Alternatively, Schuetze and von Hippel (2023)
+argue that heterogeneity in effects is an indicator of a vague, poorly
 specified theory.
 
-Re-analyses of multi-site direct replications have demonstrated that, if
-a psychological effect size can be distinguished from zero,
-heterogeneity is present across most phenomena (ref). Large-scale
-attempts of direct replications, using identical protocols, such as the
-Many Labs studies or Registered Replication Reports (references), for
-the first time allow researchers to estimate heterogeneity undistorted
-by differences in experimental designs. In re-analyses of these studies,
-Olsson-Collentine et al. (2020) identify a positive correlation between
-a phenomenon’s effect size and its heterogeneity. Similarly, van Erp et
-al. (2017) or Stanley et al. (2018) estimate strong degrees of
-heterogeneity across conceptual replications in Psychology. In a
-separate re-analysis of large scale direct replications, Renkewitz et
-al. (2025) identify substantial heterogeneity in almost all projects
-where a non-zero effect could be identified. This aligns with the
-correlation found by Olsson-Collentine et al. (2020).
+Large-scale attempts of direct replications, using identical protocols,
+such as the Many Labs studies or Registered Replication Reports
+(references), for the first time allow researchers to estimate
+heterogeneity undistorted by selection processes induced by publication
+bias. In re-analyses of these studies, Olsson-Collentine et al. (2020)
+identify a positive correlation between a phenomenon’s effect size and
+its heterogeneity. Similarly, van Erp et al. (2017) or Stanley et al.
+(2018) estimate strong degrees of heterogeneity across conceptual
+replications in Psychology. In a separate re-analysis of large scale
+direct replications, Renkewitz et al. (2025) identify substantial
+heterogeneity in almost all projects where a non-zero effect could be
+identified. This aligns with the correlation found by Olsson-Collentine
+et al. (2020).
+
+One source for this heterogeneity, which is not a typical moderator tied
+to a specific theory, might be impairments in score reliability which
+vary in their extent across replications. This is an argument put
+forward repeatedly by original researchers and replicators alike.
+Literature discussing the role of score reliability for standardized
+effect sizes claims that differences in score reliability should
+inflate, and therefore increase, heterogeneity. Over the following
+pages, we assess whether score reliability can in fact be made
+responsible for heterogeneity identified across phenomena. Even though
+only a limited number of phenomena allow for such an analysis, we find
+no empirical evidence for the argument that differences in score
+reliability inflate effect size heterogeneity. Therefore, in the last
+part of this text we turn towards an alternative discussion using
+analytical arguments on how differences in measurement quality affect
+standardized effect sizes and their differences.
 
@@
 
-The claims made by W&D imply that corrections of individual observed ESs
+The claims made in W&D imply that corrections of individual observed ESs
 for their unreliability should lead to lower estimates of heterogeneity
@@
 that the ES heterogeneity uncovered in these projects therefore could
-potentially be explained by these differences in score reliability. The
-same argument has been proposed by both original authors of phenomena
-that did not replicate or carry substantial heterogeneity (references)
-as well as the authors of the multi-site replication attempts (refs).
-Over the following pages, we assess for the first time, whether the
-differences in score reliability do indeed explain, and thereby reduce,
-the ES heterogeneity observed across the different multi-site
-replication efforts. Even though only a limited number of phenomena
-allow for such an analysis, we find no evidence for a reduction in ES
-heterogeneity after taking score reliability into account. Subsequently,
-we discuss why the expectation formulated in H&S does not allow for such
-claims. Alternatively, we explore to what extent defining ES as a random
-ratio variable helps understand how ES heterogeneity is affected by
-differences in score reliability. We close with a discussion on what
-future meta-analysts can expect, when dealing with issues of
-(un)reliability in assessments of meta-analyses of ES.
+potentially be explained by these differences in score reliability.
+Similarly, attempting to explain the substantial extent of ES
+heterogeneity identified in psychological phenomena, both the original
+authors as well as the multi-site replication authors claim that
+differences in score reliability might be responsible (references).
+
+Therefore, in the following section we assess to what extent this
+intuition can be . The re-analysis of archival data is
 
@@
 phenomena have been catalogued where the effect is studied by comparing
-the outcome across two separate groups. While the majority of designs
-make use of some form of control-treatment manipulation, randomly
-assigning participants to a condition (20 phenomena), some phenomena
-discuss the effect of pre-existing differences, e.g. biological sex (2
-phenomena). From the collection of 50 phenomena, 22 fulfil the second
-condition, so that estimates of score reliability can be derived. For
-these phenomena, the dependent variable used to construct the effect
-size is measured by employing more than a single indicator. At the same
-time, however, this implies that the remaining 28 phenomena did not make
-use of more than a single indicator to measure the dependent variable.
-Therefore, for the majority of phenomena it is not possible to assess in
-how far measurement quality is sufficiently high in terms of e.g.
-internal consistency or low random error variance.
+the outcome across two separate groups. From the collection of 50
+phenomena, 22 fulfil the second condition, so that estimates of score
+reliability can be derived. For the 22 phenomena, the dependent variable
+used to construct the effect size is measured by employing more than a
+single indicator. At the same time, however, this implies that the
+remaining 28 phenomena did not make use of more than a single indicator
+to measure the dependent variable. Therefore, for the majority of
+phenomena it is not possible to assess in how far measurement quality is
+sufficiently high in terms of e.g. internal consistency or low random
+error variance.
 
@@
 dichotomously, as responses were right or wrong (2 cases), or expressed
-agreement vs disagreement dichotomously (1 case). Subsequently, those
-phenomena need to be identified where the meta-analytic mean effect size
-is statistically significant different from zero. This will be done as a
-first step of this manuscript’s analysis procedure. Detailed information
-concerning the different phenomena and how they were measured can be
-found at \[osf-link\], while a brief summary can be found in Table A.
+agreement vs disagreement dichotomously (1 case). While the majority of
+designs make use of some form of control-treatment manipulation,
+randomly assigning participants to a condition (20 phenomena), some
+phenomena discuss the effect of pre-existing differences, e.g.
+biological sex (2 phenomena). Subsequently, those phenomena need to be
+identified where the meta-analytic mean effect size is statistically
+significant different from zero. This will be done as a first step of
+this manuscript’s analysis procedure. Detailed information concerning
+the different phenomena and how they were measured can be found at
+\[osf-link\], while a brief summary can be found in Table 1.
 
@@
 heterogeneity in $\rho_{xx‘}$, contradicting Hunter & Schmidt’s
-interpretation of this equation. In @fig-HS, the results of employing
+interpretation of this equation. In @fig-hs, the results of employing
 @eq-HS to compute the heterogeneity (variance) in corrected ES is
@@
 #| label: fig-HS
-#| fig-cap: Implications by Hunter & Schmidt's @eq-HS
+#| fig-cap: Implications by Hunter & Schmidt's @eg-HS
 
@@
 
-In @fig-HS, the dashed red line corresponds to the variance of corrected
-ES.
+In @fig-hs, the red line corresponds to the variance of corrected ES.
 
@@
 heterogeneity in uncorrected ES does cross the red line, representing
-the heterogeneity in corrected ES. This implies that, as long as mean
-score reliability is below .72, an attenuation correction would lead to
-a reduction in ES heterogeneity. As mean score reliability grows larger
-than .72, however, the heterogeneity in uncorrected ES grows larger than
-the heterogeneity in corrected ES - an attenuation correction would
-therefore only increase the extent of ES heterogeneity. This
-demonstrates that, depending on the mean level of score reliability,
-heterogeneity in uncorrected ES can be higher or lower than
-heterogeneity in corrected ES. As heterogeneity in corrected ES is not
-always smaller, the interpretationsof H&S concerning this equation is
-erroneous.
-
-However, @eq-HS is additionally flawed, as it violates a crucial
-definition. As Hunter & Schmidt point out, @eq-HS is only valid “If the
+the heterogeneity in corrected ES. Depending on the mean level of score
+reliability, heterogeneity in uncorrected ES can be higher or lower than
+heterogeneity in corrected ES.
+
+However, @eq-hs is additionally flawed, as it violates a crucial
+definition. As Hunter & Schmidt point out, @eq-hs is only valid “If the
 level of reliability is independent of the true effect size across
@@
 that ES $\delta$ and score reliability $\rho_{xx‘}$ can not be
-independent variables. Therefore, the basic premise of @eq-HS is not
+independent variables. Therefore, the basic premise of @eq-hs is not
 fulfilled, as the assumption of variable independence could not be
@@
 function and the variance of either is placed in the numerator of the
-function, @eq-HS is not sufficient to explain how heterogeneity in ES
+function, @eq-hs is not sufficient to explain how heterogeneity in ES
 changes due to the attenuation correction. Instead, we introduce two
@@
 
-The data-sets available to this project fall into two categories: direct
-multi-site replications of a specific phenomenon, employing identical
-protocols across samples; a large-scale collaborative effort to assess a
-single phenomenon globally, distributed across a large number of
-independent laboratories. For both categories, rather arbitrary sampling
-schemes were used. The majority of respondents are students at the
-respective universities, taking part in the studies voluntarily or
-because they were required as part of their study's programme (at times
-compensation was awarded). This is not the same as random sampling.
-These samples are put together arbitrarily, it is unlikely that the
-underlying distribution of true scores is meaningful for a properly
-specified population. The between-studies randomness introduced by these
-undefined sampling schemes manifests itself in true score variance
-heterogeneity $\tau^2_{\sigma^2_X}$. Between-studies randomness
-introduces by issues of measurement quality manifest themselves in error
-score variance heterogeneity $\tau^2_{\sigma^2_X }$. This implies that
-it should be of little surprise that the differences in true score
-variance largely exceed those in error score variance.
-
-Without sampling schemes that assure that the samples are constructed to
-be representative for some kind of population, we have no reason to
-assume that there should be few differences in the dispersion of true
-scores. Even though the measuring instruments are translated to
-different languages, which is bound to introduce differences in how much
-random noise is added to scores across samples, we have no reason to
-assume that these differences are larger than those introduced by
-employing no sampling schemes. Therein, we believe that, as the
-differences in score reliability are largely introduced by failure of
-ensuring appropriate sampling mechanisms, not by differences in how well
-the instrument captures the underlying construct, attenuation correction
-can not be expected to remedy ES heterogeneity.
-
-However, this also means that results contrary to what we observed in
-this project are not unthinkable. Valid sampling schemes need to be
-considered, ensuring that the samples are representative for populations
-that can be meaningfully compared. In that case, we are convinced that
-true score variance should be more stable across samples and differences
-in the instrument's measuring quality are bound to have a much larger
-impact. Not only would an attenuation correction in such a case actually
-have the potential of reducing ES heterogeneity, the variability
-underlying a psychological phenomenon could be meaningfully assessed
-across samples - something that would greatly help attempts to identify
-universal and non-universal patterns of psychological phenomena. As it
-stands, current approaches to multi-site direct replications and other
-collaborative efforts suffer strongly from the lack of appropriate
-sampling schemes.
-
 The empirical arguments presented are based on a rather small number of
@@
 replicate our analyses.
+
+Paragraph on implications -\> data collected using no specific sampling
+schemes - little surprise that the radomness induced by sampling
+schemes, which manifests itself in true score variance heterogeneity,
+largely surpasses the heterogeneity induced by imperfect measurement.
 
@@
 derived from the inequality should still hold up, nevertheless.
-Additionally, we used the empirical examples available to assess in how
-far estimates derived using the delta-method approximations diverge from
-estimates derived without it. The differences identified are rather
-miniscule at the third decimal, reinforcing our belief that these
-assumption violations are defensible.
 

### 2025-03-05 Manuscript_MD_ScoreReLiability.qmd

--- 2025-02-27 Manuscript_MD_ScoreReLiability.qmd
@@
 dichotomously, as responses were right or wrong (2 cases), or expressed
-agreement vs disagreement dichotomously (1 case). Subsequently, those
-phenomena need to be identified where the meta-analytic mean effect size
-is statistically significant different from zero. This will be done as a
-first step of this manuscript’s analysis procedure. Detailed information
-concerning the different phenomena and how they were measured can be
-found at \[osf-link\], while a brief summary can be found in Table A.
 

### 2025-03-06 Manuscript_MD_ScoreReLiability.qmd
fix direction in scripts

--- 2025-03-05 Manuscript_MD_ScoreReLiability.qmd
@@
 Therefore, in the following section we assess to what extent this
-intuition can be . The re-analysis of archival data is
+intuition is supported by empirical evidence. The re-analysis of
+archival data serves to shed light on the following two research
+questions central to this text:
+
+-   Can score reliability be used to explain heterogeneity in
+    standardized effect sizes?
+
+-   How does score reliability affect ES heterogeneity?
 
@@
 
-We have attempted to collect all openly available multi-site direct
+We have collected a large number of openly available multi-site direct
 replications on psychological phenomena (Fünderich et al. 2024). The
@@
 randomly assigning participants to a condition (20 phenomena), some
-phenomena discuss the effect of pre-existing differences, e.g.
+phenomena refer to the effect of pre-existing differences, e.g.
 biological sex (2 phenomena). Subsequently, those phenomena need to be
@@
 different from zero. For this hypothesis test, as all other hypothesis
-tests in this manuscript, a significance level of .05 is used.
+tests in this manuscript, a one-tailed significance level of .05 is
+used.
 

### 2025-03-06 E0_Data-Download.R

@@ -174,7 +174,7 @@ zip::unzip(here("Data/Downloaded Data/RRR8_Meta-analysis_final.zip"),
 
 # Additionally, in one of the lab-directories, there seems to be an erroneous data.file, which we remove here
 #  by hand, to enable further analyses down the line.
-file.remove("C:/Users/Lukas/Documents/Git_RStudio/Reliability-Generalizations/Data/Original Data/RRR8/Meta-analysis/Krahmer_data/Rentzelas_data_complete.csv")
+file.remove(here("Data/Original Data/RRR8/Meta-analysis/Krahmer_data/Rentzelas_data_complete.csv"))
 
 
 
@@ -191,9 +191,9 @@ unzip(here("Data/Downloaded Data/SW_Script_and_Data.zip"),
 
 # Unfortunately, in RRR9 one of the data-files (Vanpaemel) consists of a larger number of emptly columns,
 #  which complicate further analyses. Therefore, these columns are removed and the data-file is replaced here
-RRR9_Vanpaemel <- read.csv("C:/Users/Lukas/Documents/Git_RStudio/Reliability-Generalizations/Data/Original Data/RRR9/SW_Script_and_Data/Vanpaemel_Final.csv")
+RRR9_Vanpaemel <- read.csv(here("Data/Original Data/RRR9/SW_Script_and_Data/Vanpaemel_Final.csv"))
 RRR9_Vanpaemel <- RRR9_Vanpaemel[,names(RRR9_Vanpaemel)[!names(RRR9_Vanpaemel) %in% names(RRR9_Vanpaemel)[grep("^X", names(RRR9_Vanpaemel))]]]
-write.csv(RRR9_Vanpaemel, "C:/Users/Lukas/Documents/Git_RStudio/Reliability-Generalizations/Data/Original Data/RRR9/SW_Script_and_Data/Vanpaemel_Final.csv",
+write.csv(RRR9_Vanpaemel, here("Data/Original Data/RRR9/SW_Script_and_Data/Vanpaemel_Final.csv"),
           row.names = FALSE)
 
 ### RRR10 - Mazar et al.

### 2025-03-06 E1_Data-Analysis_demonstrate.R

@@ -39,7 +39,7 @@ data.list[[4]]$source <- substr(data.list[[4]]$source, start = nchar(data.list[[
 MASC_names <- substr(list.files(here("Data/Extracted (Project) Data")), 1, nchar(list.files(here("Data/Extracted (Project) Data")))-4) 
 
 
-effect_index <- MASC_names %in% c("Albarracin_Priming_SAT", "Alter_Analytic_Processing",
+effect_index <- MASC_names %in% c("Albarracin_Priming_SAT", 
                                   "Carter_Flag_Priming", "Caruso_Currency_Priming",
                                   "Dijksterhuis_trivia", "Finkel_Exit_Forgiveness", 
                                   "Finkel_Neglect_Forgiveness",
@@ -248,7 +248,10 @@ names(d_rma_full_L) <- MASC_names[effect_index]
 
 ES_rma_df <- data.frame(mu = unlist(lapply(d_rma_full_L, FUN = function(x){c(x$rma_raw$b[1], x$rma_cor$b[1])})),
                         se = unlist(lapply(d_rma_full_L, FUN = function(x){c(x$rma_raw$se, x$rma_cor$se)})),
-                        pval = unlist(lapply(d_rma_full_L, FUN = function(x){c(x$rma_raw$pval, x$rma_cor$pval)})),
+                        pval = unlist(lapply(d_rma_full_L, FUN = function(x){c(
+                          if(x$rma_raw$b[1] > 0){x$rma_raw$pval/2}else{1 - (x$rma_raw$pval/2)}, 
+                          if(x$rma_cor$b[1] > 0){x$rma_cor$pval/2}else{1 - (x$rma_cor$pval/2)}
+                          )})),
                         ci.lb = unlist(lapply(d_rma_full_L, FUN = function(x){c(x$rma_raw$ci.lb, x$rma_cor$ci.lb)})),
                         ci.ub = unlist(lapply(d_rma_full_L, FUN = function(x){c(x$rma_raw$ci.ub, x$rma_cor$ci.ub)})),
                         tau = sqrt(unlist(lapply(d_rma_full_L, FUN = function(x){c(x$rma_raw$tau2, x$rma_cor$tau2)}))),

### 2025-03-06 E2_Data-Analysis_explore.R

@@ -48,7 +48,7 @@ data.list <- lapply(data_files, read.csv)
 data.list[[4]]$source <- substr(data.list[[4]]$source, start = nchar(data.list[[4]]$source) - 2, stop = nchar(data.list[[4]]$source))
 
 
-effect_index <- MASC_names %in% c("Albarracin_Priming_SAT", "Alter_Analytic_Processing",
+effect_index <- MASC_names %in% c("Albarracin_Priming_SAT", 
                                   "Carter_Flag_Priming", "Caruso_Currency_Priming",
                                   "Dijksterhuis_trivia", "Finkel_Exit_Forgiveness", 
                                   "Finkel_Neglect_Forgiveness",

### 2025-03-12 Manuscript_MD_ScoreReLiability.qmd
init rework results move tab_meanES combine tab_rel tab_tauES

@@
 the outcome across two separate groups. From the collection of 50
-phenomena, 22 fulfil the second condition, so that estimates of score
-reliability can be derived. For the 22 phenomena, the dependent variable
+phenomena, 21 fulfil the second condition, so that estimates of score
+reliability can be derived. For the 21 phenomena, the dependent variable
 used to construct the effect size is measured by employing more than a
 single indicator. At the same time, however, this implies that the
-remaining 28 phenomena did not make use of more than a single indicator
+remaining 29 phenomena did not make use of more than a single indicator
 to measure the dependent variable. Therefore, for the majority of
@@
 
-From the 22 phenomena that employed more than a single indicator, 19
+From the 21 phenomena that employed more than a single indicator, 19
 made use of Likert-style items, where respondents would indicate their
 agreement to some kind of statement. The scale-length varied from 3 to
-10-points. The items measuring the remaining 3 phenomena were coded
-dichotomously, as responses were right or wrong (2 cases), or expressed
-agreement vs disagreement dichotomously (1 case). While the majority of
+10-points. The items measuring the remaining 2 phenomena were coded
+dichotomously, as responses were right or wrong. While the majority of
 designs make use of some form of control-treatment manipulation,
@@
 identified where the meta-analytic mean effect size is statistically
-significant different from zero. This will be done as a first step of
+significant different from zero. This will be done as a preliminary step in
 this manuscript’s analysis procedure. Detailed information concerning
@@
 \[osf-link\], while a brief summary can be found in Table 1.
+
+[Equations @eq-d], [-@eq-SEd], [-@eq-dc], and [-@eq-SEdc] were used to
+generate estimates of standardized ES, both uncorrected and corrected,
+with their corresponding standard errors. Generally, this leads to
+larger (absolute) ES and larger standard errors. As an example from the
+21 selected phenomena, the results from this procedure on Nosek’s
+Explicit Art sex differences phenomenon are displayed in the forest plot
+in @fig-forest. Here, grey dots represent the uncorrected ES for each
+sample, while black dots represent the corrected ES in each sample. The
+bars surrounding these dots show the respective 95%-Cis.
+
+In @fig-forest it becomes apparent that the reliability attenuation
+procedure leads to an increase in the individual absolute effect size,
+as the black dots are moved further away from zero. Simultaneously, the
+standard errors grew larger after the attenuation correction, which
+leads to larger 95%-CIs. This is most easily observed for the
+mturk-sample, which already had a rather large confidence interval to
+begin with. Additionally, in the diamond at the bottom of the figure, we
+can see that the meta-analytic estimate mean ES is also larger after the
+attenuation correction took place. In @tbl-meanES, the estimates and
+tests on the meta-analytic average ES, both corrected and uncorrected,
+can be found.
+
+```{r, echo = FALSE}
+#| label: fig-forest
+#| fig-cap: Forest Plot Nosek_Explicit_Art
+forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
+```
+
+```{r, echo = FALSE}
+#| label: tbl-meanES
+#| tbl-cap: Meta-analytic average  ES
+knitr::kable(x = read.csv(here("Tables/Mean_ES.csv")))
+```
+
+@tbl-meanES demonstrates that what we observed in Figure 1 holds across
+all 21 phenomena. All meta-analytic effects are larger in their absolute size after
+applying an attenuation correction procedure. Similarly, the uncertainty
+quantified in the estimates’ standard error is larger. Additionally,
+@tbl-meanES highlights which phenomena pass criterion (3) – the effect
+size must be statistically significantly distinguishable from zero. 12
+phenomena pass this criterion, highlighted in boldface. It may be
+remarked that the attenuation correction procedure, in this data,
+appears to have no influence on whether a phenomenon passes the
+criterion. The differences in p-value are rather small and lead to no
+difference in conclusion, regarding a significance level of .05.
 
@@
 
-[Equations @eq-d], [-@eq-SEd], [-@eq-dc], and [-@eq-SEdc] were used to
-generate estimates of standardized ES, both uncorrected and corrected,
-with their corresponding standard errors. Generally, this leads to
-larger (absolute) ES and larger standard errors. As an example from the
-12 selected phenomena, the results from this procedure on Nosek’s
-Explicit Art sex differences phenomenon are displayed in the forest plot
-in Figure 1. Here, grey dots represent the uncorrected ES for each
-sample, while black dots represent the corrected ES in each sample. The
-bars surrounding these dots show the respective 95%-Cis.
-
-In @fig-forest it becomes apparent that the reliability attenuation
-procedure leads to an increase in the individual absolute effect size,
-as the black dots are moved further away from zero. Simultaneously, the
-standard errors grew larger after the attenuation correction, which
-leads to larger 95%-CIs. This is most easily observed for the
-mturk-sample, which already had a rather large confidence interval to
-begin with. Additionally, in the diamond at the bottom of the figure, we
-can see that the meta-analytic estimate mean ES is also larger after the
-attenuation correction took place. In @tbl-meanES, the estimates and
-tests on the meta-analytic average ES, both corrected and uncorrected,
-can be found.
-
-```{r, echo = FALSE}
-#| label: fig-forest
-#| fig-cap: Forest Plot Nosek_Explicit_Art
-forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
-```
-
-```{r, echo = FALSE}
-#| label: tbl-meanES
-#| tbl-cap: Meta-analytic average  ES
-knitr::kable(x = read.csv(here("Tables/Mean_ES.csv")))
-```
-
-@tbl-meanES demonstrates that what we observed in Figure 1 holds across
-all 22 phenomena. All meta-analytic effect sizes are larger after
-applying an attenuation correction procedure. Similarly, the uncertainty
-quantified in the estimates’ standard error is larger. Additionally,
-@tbl-meanES highlights which phenomena pass criterion (3) – the effect
-size must be statistically significantly distinguishable from zero. 12
-phenomena pass this criterion, highlighted in boldface. It may be
-remarked that the attenuation correction procedure, in this data,
-appears to have no influence on whether a phenomenon passes the
-criterion. The differences in p-value are rather small and lead to no
-difference in conclusion, regarding a significance level of .05.
-
 ```{r, echo = FALSE}
@@
 #| tbl-cap: Tests for ES Heterogeneity
-knitr::kable(x = read.csv(here("Tables/Heterogeneity_ES.csv")))
+#knitr::kable(x = read.csv(here("Tables/Heterogeneity_ES.csv")))
+
+tab_hetES <- read.csv(here("Tables/Heterogeneity_ES.csv"))
+tab_hetrel <- read.csv(here("Tables/Results_RMA_alpha.csv"))
+
+tab_both <- data.frame(tab_hetES$MASC,
+                       tab_hetrel[,c("mu_alpha_str", "tau_alpha", "QE_str", "QEp")],
+                       tab_hetES[,c("tau_raw", "QE_raw_str", "QEp_raw", "tau_cor", "QE_cor_str", "QEp_cor")])
+
+knitr::kable(tab_both)
 ```

### 2025-03-14 Manuscript_MD_ScoreReLiability.qmd
update results in manuscript

@@
 identified where the meta-analytic mean effect size is statistically
-significant different from zero. This will be done as a preliminary step in
-this manuscript’s analysis procedure. Detailed information concerning
+significant different from zero. This will be done as a preliminary step
+in this manuscript’s analysis procedure. Detailed information concerning
 the different phenomena and how they were measured can be found at
@@
 @tbl-meanES demonstrates that what we observed in Figure 1 holds across
-all 21 phenomena. All meta-analytic effects are larger in their absolute size after
-applying an attenuation correction procedure. Similarly, the uncertainty
-quantified in the estimates’ standard error is larger. Additionally,
-@tbl-meanES highlights which phenomena pass criterion (3) – the effect
-size must be statistically significantly distinguishable from zero. 12
-phenomena pass this criterion, highlighted in boldface. It may be
-remarked that the attenuation correction procedure, in this data,
+all 21 phenomena. All meta-analytic effects are larger in their absolute
+size after applying an attenuation correction procedure. Similarly, the
+uncertainty quantified in the estimates’ standard error is larger.
+Additionally, @tbl-meanES highlights which phenomena pass criterion (3)
+– the effect size must be statistically significantly distinguishable
+from zero. 12 phenomena pass this criterion, highlighted in boldface. It
+may be remarked that the attenuation correction procedure, in this data,
 appears to have no influence on whether a phenomenon passes the
@@
 tab_hetrel <- read.csv(here("Tables/Results_RMA_alpha.csv"))
+tab_muES <- read.csv(here("Tables/Mean_ES.csv"))
 
@@
 
-knitr::kable(tab_both)
+knitr::kable(tab_both,
+             col.names = c("MASC", 
+                           "$\\hat{\\mu}_{\\alpha}$ (95%-CI)", 
+                           "$\\hat{\\tau}_{\\alpha}$", 
+                           "QE (df)", "p", 
+                           "$\\hat{\\tau}_{\\delta_0}$", 
+                           "QE (df)", "p",
+                           "$\\hat{\\tau}_{\\delta}$", 
+                           "QE (df)", "p"),
+             escape = FALSE)
+# 
+# tab_one <- data.frame(tab_muES,
+#                       tab_hetrel[,c("mu_alpha_str", "tau_alpha", "QE_str", "QEp")])
 ```
 
-Cases incompatible with claims made in W&D and H&S are highlighted using
-boldface.
-
-However, what @fig-forest and @tbl-meanES can not inform us about is in
-how far the heterogeneity has changed after correcting for
-(un)reliability. In @tbl-tauES, the estimates and tests of heterogeneity
-concerning uncorrected and corrected ES on the 12 remaining phenomena,
-which passed all criteria (1) – (3), can be found. In the table, the
-estimate of heterogeneity $\tau$, with the accompanying QE-test
-statistic, degrees of freedom and its p-value are reported, separately
-for uncorrected and corrected ES.
-
-Most importantly, @tbl-tauES demonstrates that of 12 phenomena assessed,
-for seven phenomena we find patterns incompatible with claims made in
-W&D and H&S. For those seven phenomena, the reliability attenuation
-correction led to an increase in absolute ES heterogeneity $\tau$. For
-the remaining five phenomena, estimates of ES heterogeneity were zero
-before and after the correction procedure, leading to no change at all.
-This leaves three phenomena where the attenuation correction did indeed
-lead to a reduction of heterogeneity. However, it is crucial to note
-that the hypothesis tests identified statistically significant
-heterogeneity in only four out of 12 phenomena and for none of those a
-reduction in heterogeneity took place. In all four cases of
-statistically significant heterogeneity, the attenuation correction
-procedure led to an increase in heterogeneity, contrary to the W&D and
-H&S predictions.
-
-To make sure that actual differences in score reliability exist,
-heterogeneity in score reliability is assessed, making use of
-Reliability Generalization Meta-Analysis (reference). @tbl-RGMA
-summarises the results across all 12 phenomena. Overall, scales for
-almost all phenomena appear have produced scores with an average score
-reliability larger than .7. Only two phenomena (Albarracin_Priming_SAT
-and Alter_Analytic_Processing) come with lower score reliability.
-However, for those phenomena we could not identify statistically
-significant heterogeneity in effect sizes in @tbl-meanES in the first
-place. From the phenomena where the ES heterogeneity grew larger as a
-result from the attenuation correction procedure, for six out of seven
-phenomena, we observe statistically significant heterogeneity in score
-reliability. Only the Shnabel_Willingness_Reconcile_Rev project produced
-scores where no heterogeneity in score reliability was identified.
-However, for this project, it is crucial to note that only a small
-number of replications (8), was available for analysis. Since the power
-of Cochran’s Q-test for heterogeneity largely depends on the number of
-replications, this result is inconclusive.
-
-Despite this one inconclusive result, for the majority of phenomena
-where ES heterogeneity is larger after applying an attenuation
-correction, a lack of differences in score reliability can not be made
-responsible for results contrary to claims made in W&D and H&S. All
-seven phenomena came with an average score reliability from .74 to .9
-and for six out of seven phenomena we identified statistically
-significant heterogeneity in score reliability.
-
-```{r, echo = FALSE}
-#| label: tbl-RGMA
-#| tbl-cap: Reliability Generalization Meta-Analysis
-knitr::kable(x = read.csv(here("Tables/Results_RMA_alpha.csv")))
-```
+@tbl-tauES contains the results from performing an RG-MA on the
+Cronbach's Alpha coefficients of the scales used to measure the 12
+remaining phenomena. Additionally, it contains estimates of
+heterogeneity for both uncorrected ES ($\delta_0$) and corrected ES
+($\delta$), which are central to assess the research questions posed in
+section XXX. Most importantly, @tbl-tauES demonstrates that for all but
+one of the phenomena, the attenuation correction procedure barely has an
+impact on the extent of ES heterogeneity. For all phenomena but
+Srull_Behaviour_Hostility, the heterogeneity in ES before and after the
+correction changes very little. This implies that for 11 out of 12
+phenomena, differences in score reliability can not be made responsible
+for heterogeneity in standardized effect sizes. Only for the
+Srull_Behaviour_Hostility phenomenon, accounting for score reliability
+actually does reduce, and therefore explain, ES heterogeneity.
+
+Secondly, in @tbl-tauES we find that for ten out of twelve phenomena, ES
+heterogeneity grows as a result from the attenuation correction
+procedure. Even though these differences tend to be small, it is clear
+that in all those cases, not only did accounting for score reliability
+not explain any ES heterogeneity, we find even more than initially
+expected. This is incompatible with claims made in W&D and H&S, the
+attenuation correction increased ES heterogeneity. Concerning the
+remaining two phenomena, for the Shnabel_Willingness_Reconcile_Rev, the
+reduction in heterogeneity is miniscule and the estimate of
+heterogeneity is not statistically significant in the first place. Most
+likely, the difference in ES heterogeneity estimates can be explained by
+issues of estimation, as only eight replications are available. Again,
+only for the Srull_Behaviour_Hostility phenomenon, we find a substantial
+reduction in heterogeneity due to the attenuation correction procedure.
+However, for the majority of phenomena it appears that differences in
+score reliability do not inflate heterogeneity in uncorrected effect
+sizes, but instead masks heterogeneity, that can be unveiled by
+accounting for score reliability.
 

### 2025-03-17 Manuscript_MD_ScoreReLiability.qmd
Update HunterSchmidt Section

@@
 
-In this section, we explore some facets on why we disagree with claims
-stated in W&D and H&S. Additionally, we explore some alternative
-explanations that might help understand why correcting ES for score
-reliability does not lead to an inevitable reduction in heterogeneity.
+In the previous section we demonstrated that assumptions formulated in
+the literature regarding meta-scientific methodology are false. We found
+a large number of phenomena where, contrary to popular claims, score
+reliability does not explain ES heterogeneity. Over the following
+paragraphs, we discuss how the true relationship between these two
+parameters might look like. Therein, we also identify under which
+circumstances an attenuation correction can actually reduce ES
+heterogeneity - and where it is more likely to inflate it.
 
@@
 “If the level of reliability is independent of the true effect size
-across studies, then, to a close approximation:” (p. 309)
+across studies, then, to a close approximation:” (p. 309). However, as
+can be easily verified, the equation presented in Hunter & Schmidt
+(ref.), Equation X on p. 309, does not imply that assumption across all
+conditions.
 
@@
 We adjusted @eq-HS using the notation used throughout this text. Here,
-$\delta_0$ refers to the ES, undistorted by sampling error but not
-corrected for measuring error, $\delta$ refers to the ES, undistorted by
-sampling error and measurement error and ρxx‘ refers to score
-reliability.
+$\delta_0$, carrying a 0 in the index, refers to the uncorrected ES,
+undistorted by sampling error but not corrected for measuring error. On
+the other hand $\delta$, no index, refers to the corrected ES,
+undistorted by sampling error and measurement error. Lastly,
+$\rho_{XX'}$ refers to score reliability.
 
 @eq-HS implies that heterogeneity in $\delta_0$ is essentially a
-function where heterogeneity and mean value of $\delta$ are weighted by
-mean score reliability $\rho_{xx‘}$ and its heterogeneity. While it may
-not be self-evident in the equation itself, it is easy to construct
-cases where heterogeneity in $\delta_0$ is inflated or deflated by
-heterogeneity in $\rho_{xx‘}$, contradicting Hunter & Schmidt’s
-interpretation of this equation. In @fig-hs, the results of employing
-@eq-HS to compute the heterogeneity (variance) in corrected ES is
-highlighted. All else held equal, only the individual level of mean
-reliability $\mu_{\rho_{xx‘}}$ is varied from .5 to .8. Mean corrected
-ES $\mu_{\delta}$ is held constant at 1, its variance $\tau^2_{\delta}$
-held constant at .04 and the variance of score reliability
-$\tau^2_{\rho_{xx‘}}$ held constant at .02.
+function in which heterogeneity and mean value of $\delta$ are weighted
+by mean score reliability $\rho_{xx‘}$ and its heterogeneity. Using
+selected values for the parameters in @eq-HS, we demonstrate how the
+heterogeneity in standardized ES can be both inflated and deflated by
+the presence of differences in score reliability. In @fig-hs, the
+results of employing @eq-HS to compute the heterogeneity (variance) in
+uncorrected ES $\delta_0$ is highlighted. All else held equal, only the
+individual level of mean reliability $\mu_{\rho_{xx‘}}$ is varied from
+.5 to .8. Mean corrected ES $\mu_{\delta_0}$ is held constant at 1, its
+heterogeneity $\tau_{\delta_0}$ held constant at .2 and the
+heterogeneity of score reliability $\tau_{\rho_{xx‘}}$ held constant at
+.12.
 
@@
 #| label: fig-HS
-#| fig-cap: Implications by Hunter & Schmidt's @eg-HS
-
-mu_delta <- 1
-var_delta <- .2^2
+#| fig-cap: Implications by Hunter & Schmidt's @eg-HS \n Note: The red line corresponds to the variance of uncorrected ES $\delta_0$.
+
+library(ggplot2)
+
+mu_delta0 <- 1
+var_delta0 <- .2^2
 mu_alpha <- seq(from = .5, to = .8, length.out = 100)
-var_alpha <- .14^2
-
-var_delta0 <- mu_alpha^2 * var_delta + mu_delta^2 * var_alpha
+var_alpha <- .12^2
+mu_delta <- mu_delta0/sqrt(mu_alpha)
+
+# var_delta0 <- mu_alpha^2 * var_delta + mu_delta^2 * var_alpha
+var_delta <- (var_delta0 - mu_delta^2 * var_alpha) / (mu_alpha^2)
 
@@
   geom_hline(yintercept = .04, colour = "red", linewidth = 1, linetype = "dashed") +
-  geom_line(aes(y = var_delta0, x = mu_alpha), linewidth = 2) +
+  geom_line(aes(y = var_delta, x = mu_alpha), linewidth = 2) +
   theme(panel.background = element_rect(fill = "transparent"), 
@@
         strip.background = element_rect(fill = "transparent")) +
-  labs(y = expression(tau^2~"["~delta[0]~"]"),
+  labs(y = expression(tau^2~"["~delta~"]"),
        x = expression(mu~"["~rho[xx]~"]"))
@@
 
-In @fig-hs, the red line corresponds to the variance of corrected ES.
-
-If everything else is held constant, the variance in uncorrected ES can
-be a function only of the mean level of score reliability
+If everything else is held constant, the variance in corrected ES
+$\delta$ can be a function only of the mean level of score reliability
 $\mu_{\rho_{xx‘}}$, according to H&S equation. In that case, Figure 2
@@
 inevitably lead to lower heterogeneity, as the line representing
-heterogeneity in uncorrected ES does cross the red line, representing
-the heterogeneity in corrected ES. Depending on the mean level of score
+heterogeneity in corrected ES does cross the red line, representing the
+heterogeneity in uncorrected ES. Depending on the mean level of score
 reliability, heterogeneity in uncorrected ES can be higher or lower than
@@
 level of reliability is independent of the true effect size across
-studies \[…\]” (2014, p. 309). However, both parameters, ES $\delta$ and
-score reliability ρxx‘, can not be independent parameters by definition.
-ES $\delta$ is the “true” standardized mean difference, unaffected by
-sampling or measurement error. The ES $\delta_0$ is standardized using
-the total standard deviation, which is inflated by measurement error.
-Equation 1, demonstrating how Cohen’s d is computed demonstrates how the
-mean difference is standardized in this case. As demonstrated further
-above in @eg-dct, ES $\delta$ is standardized using the true standard
-deviation in the sense of CTT. The fact that the true standard deviation
-parameter $\sigma_T$ is found both in @eq-dct and @eq-rel demonstrates
-that ES $\delta$ and score reliability $\rho_{xx‘}$ can not be
-independent variables. Therefore, the basic premise of @eq-hs is not
-fulfilled, as the assumption of variable independence could not be
-fulfilled.
+studies \[…\]” (2014, p. 309). However, if we assume that true score
+variance $\sigma^2_T$ varies across administrations, it is unlikely that
+this assumption can hold. As demonstrated further above in @eg-dct, ES
+$\delta$ is standardized using the true standard deviation $\sigma_T$ in
+the sense of CTT. The fact that this parameter is found both in @eq-dct
+and @eq-rel demonstrates that ES $\delta$ and score reliability
+$\rho_{xx‘}$ are unlikely to constitute truly independent variables, as
+long as true score variance is not identical across administrations.
 

### 2025-03-18 Manuscript_MD_ScoreReLiability.qmd
change & shorten first discussion

@@
 
-$$\tau^2[\delta_0] \approx \frac{\tau^2[MD]}{\mu[\sigma_x]^2} + \frac{\mu[MD]^2}{\mu[\sigma_x]^4} var[SD]$$ {#eq-ratio_d0}
+$$\tau^2[\delta_0] \approx \frac{\tau^2[MD]}{\mu[\sigma_x]^2} + \frac{\mu[MD]^2}{\mu[\sigma_x]^4} var[\sigma_x]$$ {#eq-ratio_d0}
 
@@
 function, @eq-hs is not sufficient to explain how heterogeneity in ES
-changes due to the attenuation correction. Instead, we introduce two
-additional metrics.
+changes due to the attenuation correction.
+
+In this section, we attempt to understand under which circumstances the
+heterogeneity in ES grows larger as the result of an attenuation
+correction procedure, contrary to claims made in W&D and H&S. To do so,
+we propose the following inequality, involving [Equations @eq-ratio_d0]
+and [-@eq-ratio_d].
+
+$$\tau^2_{\delta} > \tau^2_{\delta_0} \quad  \quad \quad \quad \quad  \frac{\tau_{MD}^2}{\mu_{\sigma_T}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_T}^4} \tau^2_{\sigma_T} > \frac{\tau_{MD}^2}{\mu_{\sigma_x}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_x}^4} \tau^2_{\sigma_x}$$ {#eq-ineq_init}
+
+As long as the inequality defined in @eq-ineq_init holds true, an
+attenuation correction procedure would lead to larger estimates of ES
+heterogeneity. Conditions under which @eq-ineq_init does not hold are
+conditions under which the claims made in W&D and H&S are met. However,
+the inequality in this form does not convey sufficient information about
+the conditions under which it can potentially hold. In Appendix A, we
+rearranged @eq-ineq_init, so we end up with a form that we can use to
+actually derive conditions under which ES heterogeneity can be inflated
+or deflated by differences in score reliability.
+
+To do so, we introduce two additional metrics.
 
@@
 though formally $R_1$ and the meta-analytic mean score reliability
-$\mu_{\rho_{xx'}}$ are not the same value, in practice we observe little
-to no differences between the two parameters. Large values indicate that
-correcting the individual score variances (or standard deviations for
-that matter) using the attenuation correction procedure should lead to
-very little change, as most of the variation observed were due to
-genuine differences in true scores. Small values indicate the opposite,
-the observed score variance was due to a larger amount of random error
-score variance, leading to a stronger correction.
+$\mu_{\rho_{xx'}}$ do not share not the same value, in practice we
+observe little to no differences between the two parameters. Large
+values indicate that correcting the individual score variances (or
+standard deviations for that matter) using the attenuation correction
+procedure should lead to very little change in ES, as most of the
+variation observed were due to genuine differences in true scores. Small
+values indicate the opposite, the observed score variance was due to a
+larger amount of random error score variance, leading to a stronger
+correction.
 
@@
 
-Both metrics $R_1$ and $R_2$ make use of parameters of the distribution
-of true and total score variance, not standard deviations. This choice
-was made to facilitate the role of score reliability in the subsequent
-discussion, as score reliability itself is defined by variances, not
-standard deviations (see @eq-rel).
+Concerning the re-arranged inequality from Appendix A, we identify one
+crucial condition described in @eq-ineq_short.
 
-In this section, we attempt to understand under which circumstances the
-heterogeneity in ES grows larger as the result of an attenuation
-correction procedure, contrary to claims made in W&D and H&S. To do so,
-we propose the following inequality, involving [Equations @eq-ratio_d0]
-and [-@eq-ratio_d].
+$$R_1^3 \leq R_2$$ {#eq-ineq_short}
 
-$$\tau^2_{\delta} > \tau^2_{\delta_0} \quad  \quad \quad \quad \quad 
-\frac{\tau_{MD}^2}{\mu_{\sigma_T}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_T}^4} \tau^2_{\sigma_T} > \frac{\tau_{MD}^2}{\mu_{\sigma_x}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_x}^4} \tau^2_{\sigma_x}$$ {#eq-ineq_init}
+As long as $R_1^3$ is smaller or equal compared to $R_2$, @eq-ineq_init
+is bound to hold true. This means that as long as the average score
+reliability, to the power of 3, does not exceed the reduction in score
+variance heterogeneity due to the attenuation correction, @eq-ineq_init
+is not violated. For example, assuming the average score reliability is
+about .8, as long as $R_2$ is at least .512 or larger, the inequality
+holds true and ES heterogeneity grows larger as a consequence of the
+attenuation correction procedure. In such cases, the claims made in W&D
+and H&S could not hold.
 
-Using the metrics $R_1$ and $R_2$, defined in [Equations @eq-R1] and
-[-@eq-R2], we attempt to understand under which circumstances the
-inequality defined in @eq-ineq_init holds true. However, while the
-metrics make use of mean and heterogeneity of the score variances,
-@eq-ineq_init makes use of mean and heterogeneity of the standard
-deviations. In order to incorporate the metrics, it is necessary to
-reparameterise the equation. Using the delta method to approximate mean
-value and heterogeneity of variance from those of the standard
-deviations, we know that (ref):
-$$\mu^2_{\sigma_X} \approx \mu_{\sigma^2_X} \quad \quad  and \quad \quad 
-\tau^2_{\sigma_X} \approx \frac{\tau^2_{\sigma_X^2}}{4\mu_{\sigma^2_X}}$$ {#eq-delta_method}
+$R_1$ essentially describes the reduction in mean score variance due to
+the attenuation correction, and $R_2$ essentially described the
+reduction in heterogeneity of score variance due to the attenuation
+correction. Therefore, @eq-ineq_init states that for such a correction
+procedure to truly reduce ES heterogeneity, the reduction in the
+heterogeneity of score variance needs to be substantially larger than
+the reduction in its mean values. The mean value reduction carries
+stronger weight, as $R_1$ is taken to the power of 3 in @eq-ineq_init.
 
-While @eq-delta_method only explicitly contains parameters for the
-observed score variance/standard deviation, the same can be done using
-the true score variance/standard deviation. Using the approximates
-defined in @eq-delta_method for observed and true score variance, we
-arrive at a new inequality in @eq-ineq_CV.
+As long as there are substantial differences in the true score variance
+across administrations, it is unlikely that an attenuation correction
+can actually produce such a low $R_2$. Generally speaking, $R_1$, and
+therefore the reduction in mean score variance, largely depends on the
+relative measurement precision. If measurements were made more
+precisely, average score reliability and $R_1$ would be higher. On the
+other hand $R_2$, and therefore the reduction in heterogeneity of score
+variance, depends on how different the measurement precision was at the
+individual administrations, but also how different the individual
+administrations sites are regarding their true score variance. If the
+administration sites carry large differences in true score variance
+$R_2$ is unlikely to be substantially smaller than $R_1$.
 
-$$\frac{\tau_{MD}^2}{\mu_{\sigma^2_t}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_t^2}} \frac{\tau^2_{\sigma_t^2}}{4\mu_{\sigma^2_t}} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4\mu_{\sigma^2_x}}$$ {#eq-ineq_CV}
-
-@eq-ineq_CV describes that the heterogeneity in corrected ES is larger
-than the heterogeneity in uncorrected ES, using parameters of the
-distributions of mean difference MD, true score variance $\sigma^2_T$
-and observed score variance $\sigma^2_X$. By introducing the metrics
-$R_1$ and $R_2$ to this inequality, we can begin to disentangle which
-conditions need to be fulfilled for the inequality to hold. Rearranging
-the terms in [Equations @eq-R1] and [-@eq-R2], and subsequently entering
-these terms into Equation 18 leads to the following
-$$\frac{\tau_{MD}^2}{R_1 \mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} \frac{R_2 \tau^2_{\sigma_x^2}}{4 R_1 \mu_{\sigma^2_x}} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4\mu_{\sigma^2_x}}$$ {#eq-ineq_R}
-
-@eq-ineq_R, again, describes the inequality that heterogeneity in
-corrected ES is larger than the heterogeneity in uncorrected ES,
-incorporating the metrics $R_1$ and $R_2$. This equation alone is not
-sufficient to identify the relevant circumstances required for the
-inequality to hold. However, rearranging the terms from @eq-ineq_R leads
-to @eq-ineq_fin:
-
-$$ \tau_{MD}^2\left(\frac{1}{R_1} - 1\right) + \mu_{MD}^2 \frac{1}{4} CV^2_{\sigma^2_X} \left(\frac{R_2}{R_1^3}  - 1\right) > 0$$ {#eq-ineq_fin}
-
-The circumstances under which the inequality described in Equation 20
-holds, are circumstances where the claims made in W&D and H&S are
-directly contradicted, as in those cases the ES heterogeneity is larger
-after applying an attenuation correction procedure. Concerning the terms
-in @eq-ineq_fin, we know that all terms left-hand of the inequality,
-outside of the brackets ($\tau^2_{MD}$, $\mu^2_{MD}$ and
-$CV^2_{\sigma^2_X}$), are bound to be positive. For the inequality to
-hold, we need the left-hand side of the equation to remain positive,
-larger than zero. We can distill two scenarios, under which this
-equation should hold: (a) one of the terms inside the brackets
-($\frac{1}{R_1} - 1$ and $\frac{R^2_2}{R_1}  - 1$) is positive and large
-enough, so that the second term not containing that same bracket is
-positively dominated by the first term; or (b) both terms inside the
-brackets need to be positive.
-
-Generally, we know that both $R_1$ and $R_2$ are bound to be positive,
-as both contain different, strictly positive parameters of the
-distributions of true and observed score variance. Additionally, we know
-that $R_1$ is equivalent to the average score reliability, and therefore
-bound between $0\le R_1\le1$. Therefore, the term inside the first
-bracket is bound to be positive ($\frac{1}{R_1} – 1$). Similarly, $R_2$,
-as the ratio of true and error score variance heterogeneity, is bound
-between $0\le R_2\le1$. If $R_1^3$ is smaller than $R_2$, then the term
-inside the second bracket turns negative. Generally, this means that the
-inequality can only be violated if the reduction in relative
-heterogeneity of score variance ($R_2$) is larger than the the average
-score reliability to the power of 3.
-
-Concerning the two scenarios, for scenario (a) to hold true, since the
-term involving only $R_1$ in the first bracket can only be positive, the
-second term involving both metrics would need to be negative. For the
-first term to positively dominate the second term, however, we would
-need to make additional assumptions about the size of the parameters
-outside of the brackets, specifically about $\tau^2_{MD}$ and
-$\mu_{MD}$. This is something we would prefer to avoid, as it involves
-additional assumptions about the individual raw effect sizes of the
-different phenomena. Additionally, it can be said that as long as the
-term in the second bracket involving both metrics is positive, scenario
-(b) is true by default. This occurs any time $R_1^3 > R_2$. For example,
-assuming the average score reliability is about .8, if $R_2$ is at least
-.512 or larger, the inequality holds true and ES heterogeneity grows
-larger as a consequence of the attenuation correction procedure,
-contrary to claims made in W&D and H&S. This means that for the
-exemplary score reliability of .8, a reduction in relative heterogeneity
-in score variance of 48.8% ($(1 – R_2)100\%$) or more is necessary for
-scenario (b) to no longer be true. The smaller the average score
-reliability, the smaller the $R_2$-metric may be for the inequality to
-still hold true.
-
-Keeping this in mind, even if $R_2$ is smaller than .512, it does not
-immediately imply that ES heterogeneity is smaller in $\delta$. For
-example, it is unlikely that an $R_2 = .5$ leads to a term that is not
-positively dominated by the first term involving only $R_1$. However,
-where exactly the cut-off is, where ES heterogeneity starts shrinking
-due to the attenuation correction procedure, can not be easily defined
-without additional assumptions about the distribution of $MD$.
-
-Generally speaking, the attenuation correction leads to a reduction both
-in mean value of $\sigma^2_X$ and in its heterogeneity. While a
-reduction in the former typically leads to an increase in heterogeneity
-of $d_c$, the latter has the potential to compensate for that increase
-and thereby reduce that heterogeneity after the attenuation correction.
-This reduction in ES heterogeneity can only occur, if the reduction in
-score variance heterogeneity is substantially stronger than the
-reduction in mean score variance. An extreme version of such a situation
-might occur, if the true score variance is identical across all samples.
-This would imply that all heterogeneity in score variance identified can
-be attributed to differences in how much random noise is found in the
-scores across the samples. In such a case, $R_2$ would be zero and a
-reduction in ES heterogeneity due to the attenuation correction would be
-guaranteed. This also implies that the lower $R_2$ and the larger the
-average score reliability $R_1$, the more likely it is that the
-attenuation correction actually does lead to a reduction in ES
-heterogeneity.
+***\[maybe move to discussion further below\]*** The majority of studies
+in the social sciences make use of convenience sampling. Often,
+respondents are collected from a local student pool, where students
+themselves would choose which studies to partake in. In such cases, we
+believe there is little reason for us to expect homogeneity in true
+score variance across administration sites. As participants across
+administrations self-select into the study, no real sampling process
+takes place - implying that we have no reason to assume that the true
+score variance is stable across these administration sites. However,
+since researchers are well aware of the importance of measurement
+precision, which is also held stable more easily, we believe there is
+little reason for $R_1$ to be extremely low.
 

### 2025-03-18 E1_Data-Analysis_demonstrate.R
Add Strack

@@ -51,7 +51,7 @@ effect_index <- MASC_names %in% c("Albarracin_Priming_SAT",
                                   "PSACR002_neg_photo", "Shnabel_Willingness_Reconcile_Rev",
                                   "Shnabel_Willingness_Reconcile_RPP", "Srull_Behaviour_Hostility",
                                   "Srull_Ronald_Hostility", "Tversky_Directionality_Similarity1", 
-                                  "Zhong_Desirability_Cleaning"
+                                  "Strack_Facial_Feedback", "Zhong_Desirability_Cleaning"
 )
 
 

### 2025-03-18 E2_Data-Analysis_explore.R

@@ -60,7 +60,7 @@ effect_index <- MASC_names %in% c("Albarracin_Priming_SAT",
                                   "PSACR002_neg_photo", "Shnabel_Willingness_Reconcile_Rev",
                                   "Shnabel_Willingness_Reconcile_RPP", "Srull_Behaviour_Hostility",
                                   "Srull_Ronald_Hostility", "Tversky_Directionality_Similarity1", 
-                                  "Zhong_Desirability_Cleaning"
+                                  "Strack_Facial_Feedback", "Zhong_Desirability_Cleaning"
 )
 
 

### 2025-03-19 Manuscript_MD_ScoreReLiability.qmd
start on second re-analysis

@@
 likely, the difference in ES heterogeneity estimates can be explained by
-issues of estimation, as only eight replications are available. Again,
-only for the Srull_Behaviour_Hostility phenomenon, we find a substantial
-reduction in heterogeneity due to the attenuation correction procedure.
-However, for the majority of phenomena it appears that differences in
-score reliability do not inflate heterogeneity in uncorrected effect
-sizes, but instead masks heterogeneity, that can be unveiled by
-accounting for score reliability.
+issues of estimation, as only eight replications are available.
+Similarly, only for the Srull_Behaviour_Hostility phenomenon, we find a
+substantial reduction in heterogeneity due to the attenuation correction
+procedure. The estimates of ES heterogeneity are not statistically
+significant, however, which hampers our ability to meaningfully
+interpret and compare these estimates. For the majority of phenomena it
+appears that differences in score reliability do not inflate
+heterogeneity in uncorrected effect sizes, but instead mask
+heterogeneity, that can be unveiled by accounting for score reliability.
 
@@
 larger after the attenuation correction procedure. Subsequently, in
-Equation 16 we demonstrate that if the $R_2$ metric is larger than
+Appendix A we demonstrate that if the $R_2$ metric is larger than
 $R_1^3$, ES heterogeneity is bound to be larger after applying said
@@
 Additionally, we observe that across nine phenomena, $R_2$ is larger
-than .97. This implies that of the heterogeneity observed in score
+than .96. This implies that of the heterogeneity observed in score
 variances, 97% or more could not be explained by differences in random
 error introduced by the measuring instrument, but by genuine differences
-in the underlying distribution. Next to the two exceptional phenomena,
-where zero heterogeneity in score variance was found
-(Alter_Analytic_Processing and Shnabel_Willingness_Reconcile_Rev), only
-for the Srull_Ronald_Hostility phenomenon some heterogeneity could be
-attributed to differences in random error variance, albeit less than
-14%.
+in the underlying distribution.
 
-This implies that for all measurements concerning these phenomena, the
-heterogeneity in observed score variances could not be substantially
-explained by differences in measurement quality. For no phenomenon, if
-differences in score variance could be found, was $R_2$ smaller than
-$R_1^3$. As demonstrated in Equation 20, this implies that the ES
-heterogeneity grows even larger, as we correct for differences in score
-reliability.
+For one of the phenomena, however, we observe a large disparity compared
+to the rest of the results. For the Shnabel_Willingness_Reconcile_RPP
+phenomenon, $R_2$ is very low, below .25. This makes it even lower than
+its $R_1^3 = .574$. This would imply that for this phenomenon, the
+conditions for a reduction in ES heterogeneity would be met. However, in
+@tbl-tauES, we find no such reduction. What becomes apparent, however,
+is that for this phenomenon we are working with the low number of 9
+administration sites. Estimation precision for heterogeneity in variance
+components, even more so than for ES, largely depends on the sample
+size. Beinhauer et al. (ref) come to the conclusion that at 10 items,
+which the WIllingness to Reconcile scale consists of, a sample size
+around 33 is required to achieve sufficient power for a statistical test
+of score variance heterogeneity. Most likely, the low size led to
+implausible estimates in both true and total score variance
+heterogeneity and subsequently to an implausible estimate of $R_2$.
+Meta-analytic estimates, especially those concerning score variance
+components such as $R_2$, therefore should be interpreted with caution,
+if the sample size was particularly low.
+
+Similarly, for the two exceptional phenomena where the attenuation
+correction procedure actually led to an increase in ES heterogeneity
+(Srull_Behavior_Hostility and Shnabel_Willingness_Reconcile_RPP), the
+significance test for the ES heterogeneity was not significant, both
+before and after the correction. The results regarding $R_2$ indicate
+that the majority of score variance heterogeneity was due to differences
+in true score variance $\sigma^2_T$. Both phenomena were examined using
+a rather low number of administration sites (22 and 9 respectively),
+implying that issues of estimation precision might explain the
+inconclusive results. Lastly, for one out of the 12 phenomena, the
+Albarracin_Priming_SAT phenomenon, we find a $\tau_{\sigma^2_X}$ of 0.
+Therefore, the $R_2$ metric could not be computed, as the denominator
+would have been zero. However, we also did not identify statistically
+significant ES heterogeneity in the first place. This phenomenon was
+assessed with a particularly low number of replications (8). As
+discussed, adequate power to detect heterogeneity in score reliability
+and observed score variance requires a larger number of replications
+than adequate power to detect heterogeneity.
+
+Generally, @tbl-tauES implies that for almmost all measurements
+concerning these phenomena, the heterogeneity in observed score
+variances could not be substantially explained by differences in
+measurement quality. Only for one phenomenon, where differences in score
+variance could be found, $R_2$ was smaller than $R_1^3$. As discussed,
+for this phenomenon, the number of administration sites was too low to
+accurately estimate heterogeneity in score variance components. For all
+remaining phenomena, $R_2$ was larger than $R_1$, implying that an
+attenuation correction procedure could not reduce ES heterogeneity. The
+results in @tbl-tauES agree with this conclusion, as no phenomena could
+be identified where statistically significant ES heterogeneity was
+accompanied by a reduction in differences due to the correction.
 
@@
 ```
-
-In @tbl-tauES we also found some phenomena where the heterogeneity did
-in fact shrink after applying the attenuation correction procedure. This
-is not reflected by what we observe in @tbl-R1R2, as $R_2 > R_1$ for
-almost all phenomena, which would imply that across all phenomena we
-expected ES heterogeneity to grow as a consequence of the attenuation
-correction procedure. However, it is crucial to realise that we did not
-find statistically significant ES heterogeneity for any of the phenomena
-where ES heterogeneity was reduced. This implies, that we actually could
-not statistically distinguish the ES heterogeneity from zero. Therefore,
-what seems to be a reduction in ES heterogeneity was most likely caused
-by estimation issues, as we could not accurately estimate the extent of
-ES heterogeneity, either before or after the attenuation correction.
-
-Lastly, for two out of the 12 phenomena, we find a $\tau_{\sigma^2_X}$
-of 0. For these phenomena, it is therefore also impossible to compute
-the $R_2$ metric. While for the Alter_Analytic_Processing phenomenon we
-did not identify statistically significant ES heterogeneity in the first
-place, we did find some for the Shnabel_Willingness_Reconcile_Rev
-phenomenon. At the same time, we did not identify statistically
-significant heterogeneity in score reliability. With zero heterogeneity
-in score variances, Equation 13 implies that the ES heterogeneity is
-bound to grow larger due to the attenuation procedure. Also, this
-phenomenon was assessed with a particularly low number of replications
-(8). Most likely, adequate power to detect heterogeneity in score
-reliability and observed score variance requires a larger number of
-replications than adequate power to detect heterogeneity. In that case,
-the results from Table 5, concerning the
-Shnabel_Willingness_Reconcile_Rev phenomenon might alternatively be
-explained by power-issues. Similarly, while Table 3 does not imply
-specific expectations concerning the presence of score variance
-heterogeneity in the Alter_Analytic_Processing phenomenon, the low
-number of replications (20) makes it hard to distinguish whether the
-estimate of $\tau_{\sigma^2_X} = 0$ is actually sensible or can also be
-attributed to low power.
 

### 2025-03-22 Manuscript_MD_ScoreReLiability.qmd
add methods to second analysis

@@
 $R_1^3$, ES heterogeneity is bound to be larger after applying said
-procedure. @tbl-R1R2 reports the meta-analytic estimates of observed and
-true score variances, including their heterogeneity and metrics $R_1$
-and $R_2$ across all 12 phenomena. In @tbl-R1R2 we see that for ten out
-of 12 phenomena, the $R_2$ metric is larger than the $R_1$ metric.
+procedure. Here, we will re-analyse the data, estimating $R_1$ and $R_2$
+for each phenomenon respectively. This allows us to assess in how far
+the estimates of ES heterogeneity align with our claims regarding the
+$R_1$ and $R_2$ metrics.
+
+#### Methods
+
+Estimation of $R_1$ and $R_2$ require estimates of meta-analytic mean
+and heterogeneity of both the total score variance $\sigma^2_X$ and the
+true score variance $\sigma^2_T$. Elsewhere, we proposed the method
+boot-err (ref.) as an alternative method to test for and estimate the
+differences in error score variance $\sigma^2_E$. In this text, we
+deliver detailed instructions on how estimates for mean and
+heterogeneity of total score variance, and the score variance components
+can be derived. For the total score variance $\sigma^2_X$, all we need
+to do is to appropriately transform the observed estimates using the
+variance-stabilizing properties of the natural log. Subsequently, we use
+bootstrapping to derive robust standard errors for these transformed
+observations. Using a random-efffects meta-analysis, we derive estimates
+of mean and heterogeneity, which need to be backtransformed to the
+original scale. Appropriate equation to back-transform estimates can be
+found in Appendix B.
+
+For the true score variance $\sigma^2_T$, we found that directly
+estimating its heterogeneity, even if an appropriate
+variance-stabilizing transformation is used, is severely biased.
+Instead, we can estimate the error score variance $\sigma^2_E$ using an
+estimate of score reliability. Performing the boot-err, which, again,
+implies transformations using the natural log, deriving standard errors
+using Bootstrapping, running a random-effects meta-analysis and
+back-transforming the estimates, nets us estimates of mean and
+heterogeneity of $\sigma^2_E$. Since we already have the same estimates
+available for the total score variance, we can estimate meta-analytic
+mean and heterogeneity of true score variance.
+
+$$\hat{\mu}_{\sigma^2_T} = \hat{\mu}_{\sigma^2_X} - \hat{\mu}_{\sigma^2_E} \ \ \ \ \ \ and \ \ \ \ \  \hat{\tau}^2_{\sigma^2_T} = \hat{\tau}^2_{\sigma^2_X} - \hat{\tau}^2_{\sigma^2_E}$$ {#eq-meantauT}
+
+@eq-meantauT provides estimates of meta-analytic mean of true score
+variance and heterogeneity. These estimates are largely unbiased, unlike
+estimates stemming from using the boot-err to directly run meta-analytic
+models on true score variance estimates. Interested readers are referred
+to Beinhauer et al. (ref.)
+
+#### Results
+
+@tbl-R1R2 reports the meta-analytic estimates of observed and true score
+variances, including their heterogeneity and metrics $R_1$ and $R_2$
+across all 12 phenomena. In @tbl-R1R2 we see that for ten out of 12
+phenomena, the $R_2$ metric is larger than the $R_1$ metric.
 Additionally, we observe that across nine phenomena, $R_2$ is larger

### 2025-03-22 Manuscript_MD_ScoreReLiability.qmd
added bullet point rant

@@
 
+-   examing heterogeneity in standardized mean differences before and
+    after correction for differences in score reliability using an
+    attenuation correction.
+
+-   Find that a) in the majority of cases, the change in ES
+    heterogeneity is small to non-existent, implying that differences in
+    score reliability do not serve an appropriate explanation of ES
+    heterogeneity.
+
+-   Find that b) in the majority of cases, the direction of change in ES
+    heterogeneity does not align with assumptions formulated in W&D and
+    H&S. In close to all cases (all cases if we only look at
+    statistically significant differences in ES), the attenuation
+    correction leads to increases estimates of ES heterogeneity. This
+    implies that W&D and H&S generally misunderstand how score
+    reliability heterogeneity affects ESs.
+
+-   Using analytical arguments, we derived a general rule on under which
+    conditions we can expect an attenuation correction to reduce ES
+    heterogeneity. These conditions depend on how both the mean value
+    and the heterogeneity of score variance is reduced by correction for
+    sore reliability. At the very least, $R_1^3 < R_2$. This means that
+    as long as the reduction in score variance heterogeneity is larger
+    than the reduction in mean score variance, ES heterogeneity will
+    grow as we follow the attenuation correction procedure.
+
+-   We know what an acceptable average score reliability looks like in
+    the area of Psychology. Researchers largely target a score
+    reliability of at least .7, preferably .8. An average score
+    reliability of .7 implies that 70% of the variance observed in
+    scores can actually be attributed to true underlying differences.
+    However, from our derived rule, we know that at an average score
+    reliability of .7, the heterogeneity in score variances across
+    administration sites that can be attributed to differences in the
+    true score variance $\sigma^2_T$ may be at most .343 or 34.3%.
+    Accordingly, at least 65.7%, almost two thirds of the score variance
+    heterogeneity, needs to be due to differences in error score
+    variance, differences in the measurement precision.
+
+-   However, in the examples we re-analysed throughout this text, as
+    well as the majority of research studies done in the behavioural
+    sciences, no proper sampling techniques are employed. Most samples
+    where these instruments are applied come to be through convenience
+    sampling, where most often students self-select into studies.
+    Without a valid sampling scheme, we have essentially have no basis
+    to assume anything about the true score variance. Most importantly,
+    however, we have no reason to assume that the true score variance
+    could potentially be homogeneous or stable across administration
+    sites. Instead, it is very likely that the convenience sampling
+    procedures produce samples that vary extremely strongly in how they
+    are put together, leading to large degrees of true score variance
+    heterogeneity. At the same time, most researchers are aware and
+    trained in assessing score reliability.
+
+-   Taken together, these facets lead to what we observed in the
+    phenomena examined in this text: Score reliability in general is
+    acceptable, leading to a small, but constant reduction in mean score
+    variance if we apply an attenuation correction. At the same time,
+    there are substantial differences in true score variance, implying
+    that most of the score variance heterogeneity can not be attributed
+    to differences in error score variance. Since only differences in
+    error score variance can be corrected for, the attenuation
+    correction has a very small effect on heterogeneity estimates in
+    score variance. Therefore, as indicated by our rule, an attenuation
+    correction in these setting can not deliver a reduction in ES
+    heterogeneity. This also means that, in such settings, score
+    reliability does not explain any ES heterogeneity, if anything, it
+    masks it to a small degree.
+
+-   If we, in the space of behavioural sciences, keep making use of
+    convenience samples simply because they are so convenient, we can
+    not expect anything to change. Most likely, ES heterogeneity will
+    remain a substantial factor hampering the interpretability of
+    original and replication studies. We could learn more about
+    standardized effect sizes and why our phenomena vary across
+    settings, populations or continents by not following the current
+    status quo of convenience samples. Until this status quo changes,
+    however, we will keep wasting valuable resources by attempts of
+    discussing ES heterogeneity or assessing uninformative differences
+    in score reliability.
+
 Across twelve exemplary archival datasets, we demonstrated that claims

### 2025-03-25 Manuscript_MD_ScoreReLiability.qmd
change bullet points into text

--- 2025-03-22 Manuscript_MD_ScoreReLiability.qmd
@@
 
-Across twelve exemplary archival datasets, we demonstrated that claims
-made in H&S and W&D do not hold in empirical observations. Even though
-removing the impact of score reliability on ES heterogeneity by means of
-attenuation correction procedures, they do not lead to a reduction in
-heterogeneity as claimed in H&S and W&D. Inferences derived from the
-equation supplied in H&S (ref), @eq-HS, also do not warrant such a
-conclusion. Alternatively, describing standardised effect sizes in terms
-of a ratio distribution and approximating its variance, we demonstrated
-that differences in score reliability may both inflate or deflate
-variance in ES, depending on how the relative heterogeneity in score
-variance changes due to the attenuation correction procedure. The metric
-$R_2$, compared with the metric $R_1$ or the average score reliability,
-informs us in how far score variance heterogeneity decreases due to the
-attenuation correction.
+We have examined meta-analytic heterogeneity in standardized mean
+difference. Since score reliability is often discussed as a source for
+such heterogeneity, we estimated the heterogeneity before and after
+correcting for differences in reliability coefficients using an
+attenuation correction procedure. We found that a) in the majority of
+cases, the change in ES heterogeneity is small to non-existent. This
+implies that, across the data-sets we got to examine, differences in
+score reliability do not serve as an appropriate explanation for ES
+heterogeneity. We also found that b) in the majority of cases, the
+direction of change in ES heterogeneity does not align with assumptions
+formulated in W&D and H&S. In almost all cases (every single case, if we
+restrict ourself to phenomena where we found statistically significant
+differences in ES), the attenuation correction led to increased
+estimates of ES heterogeneity. This implies that the claims found in W&D
+and H&S stem from a general misunderstanding of how score reliability
+can affect ESs.
+
+Using analytical arguments, we derived a broad rule, outlining under
+which conditions we can expect an attenuation correction to reduce ES
+heterogeneity. In general, score reliability affects ESs by inflating
+the score variance, as the (square root of the) score variance is used
+to standardize the mean difference. The conditions we found depend on
+how both the mean value and the heterogeneity of score variance are
+reduced by correcting for imperfect score reliability. We introduced two
+novel metrics: $R_1$, describing the fraction of mean true score
+variance in the mean total score variance, therefore mirroring the
+average score reliability coefficient, and $R_2$, describing the
+fraction of true score variance heterogeneity in the total score
+variance heterogeneity. We find that, as long as $R_1^3 < R_2$, an
+attenuation correction will increase ES heterogeneity.
+
+Since $R_1$ is carrying the power of 3, it implies that the reduction in
+mean score variance is weighted more heavily than the reduction in score
+variance heterogeneity. This is best described through an example: We
+know what an acceptable average score reliability looks like in the area
+of Psychology. Researchers largely target a score reliability of at
+least .7, preferably .8. An average score reliability of .7 implies that
+70% of the variance observed in scores can actually be attributed to
+true underlying differences. From our derived rule, we know that, as
+long as $R_2$ is larger or equal to $.7^3 = .343$, the attenuation
+correction procedure can only increase ES heterogeneity. If we want to
+explain ES heterogeneity through score reliability, even though only 30%
+of the variance observed in score can be attributed to random
+measurement error, at least 65.7%, almost two thirds, of the
+heterogeneity in score variance would need to be attributable to
+differences in the error score variance. In other words, at most 34.3%
+of heterogeneity in score variances may be due to actual differences in
+how the latent variable is distributed (in terms of spread) across the
+administration sites.
+
+This also means that ES heterogeneity is best reduced, if average score
+reliability ($R_1$) is substantially large. In such a scenario, the
+$R_2$ is more likely to be low enough for the inequality to no longer
+hold. The larger the average score reliability, the more likely it is
+that the attenuation correction procedure can actually reduce ES
+heterogeneity. However, as the correction would only be rather small in
+such cases, it remains to be seen whether the reduction in ES
+heterogeneity would be sufficient to claim that differences in score
+reliability actually posed a relevant moderator in such a case.
+
+In the examples we re-analysed throughout this text, as well as the
+majority of research studies done in the behavioural sciences, no proper
+sampling techniques are employed. Most samples where these instruments
+are applied come to be through convenience sampling, where students
+self-select into studies. Without a valid sampling scheme, we
+essentially have have no basis to assume anything about the true score
+variance. Most importantly, however, we have no reason to assume that
+the true score variance could potentially be homogeneous or stable
+across administration sites. Instead, it is very likely that the
+convenience sampling procedures produce samples that vary extremely
+strongly in how they are put together, leading to large degrees of true
+score variance heterogeneity. At the same time, score reliability is a
+low-hanging fruit for researchers to optimise most researchers are aware
+and trained in assessing score reliability.
+
+Taken together, these facets lead to what we observed throughout this
+text: Reliability coefficients are in an acceptable range, leading to a
+small, but constant reduction in mean score variance if we apply an
+attenuation correction. At the same time, there is substantial variation
+in true score variance, implying that most of the score variance
+heterogeneity can not be attributed to differences in error score
+variance. Since only differences in error score variance can be
+corrected for, the attenuation correction has a very small effect on
+heterogeneity estimates in score variance. Therefore, as indicated by
+our rule, an attenuation correction in these setting can rarely deliver
+a reduction in ES heterogeneity. This also means that, in such settings,
+score reliability does not explain any ES heterogeneity. If anything, it
+masks it to a small degree.
+
+If we, in the space of behavioural sciences, keep making use of
+convenience samples simply because they are so convenient, we can not
+expect anything to change. Most likely, ES heterogeneity will remain a
+substantial factor hampering the interpretability of original and
+replication studies. We could learn more about standardized effect sizes
+and why our phenomena vary across settings, populations or continents by
+not following the current status quo of convenience samples. Until this
+status quo changes, however, we will keep wasting valuable resources by
+attempts of discussing ES heterogeneity or assessing uninformative
+differences in score reliability.
 
@@
 correlations and standardized effect sizes are not identical, the way
-score (un)reliability affects these parameters is highly similar. Both
-correlations and ES are deflated by (un)reliability, implying that an
-attenuation correction increases mean values of these parameters, while
-the impact on standard deviations used for standardisation purposes
-needs to be checked via $R_2$.
-
-The analytical arguments presented indicate that an attenuation
-correction procedure, even if no differences between score reliability
-exist, lead to larger ES heterogeneity. It appears that, as the
-meta-analytical mean ES grows larger due to the correction procedure, so
-does the ES heterogeneity, if homogeneity is present. If heterogeneity
-is present on the other hand and $R_2 < R_1^3$, still a reduction in ES
-heterogeneity is not inevitable. It can be argued that, for the
-differences in ES to grow smaller, the reduction in relative
-heterogeneity in score variance needs to be substantially large, so the
-increase in ES heterogeneity due to the larger mean ES can be
-“outpaced”. However, we did not observe this pattern in the 12 phenomena
-that allowed for such an analysis.
+score (un)reliability affects these parameters is highly similar.
 
@@
 demonstrated that in several scenarios, differences in score reliability
-may in fact be masking, or “deflating” ES heterogeneity. This implies
-that the already substantial heterogeneity across direct replications
-may very well be even larger, if ES were corrected for (un)reliability.
-Understanding heterogeneity as an extent to which a theory or phenomenon
-is understood, deflated estimates of heterogeneity imply that the theory
-or phenomenon is understood even worse than initially assumed.
+did not serve as appropriate moderators or explanations of ES
+heterogeneity. Across most phenomena, we failed to reduce ES
+heterogeneity. If anything, we even increased ES heterogeneity, albeit
+to a very small degree. Generally, concerning the multi-site replication
+projects or collaborative research projects, accounting for differences
+in score reliability did not help us gain a better understanding of the
+phenomena.
 
@@
 empirical arguments, we are convinced that these differences in score
-reliability across replications oftentimes mask true heterogeneity. In
-realistic meta-analytic studies of conceptual replications, data stems
-from a variety of sources. Across different populations, different
-research settings or designs are used with varying or adjusted research
-instruments. It is unlikely that, across these replications, we can
-expect either the measuring quality or the true score variance to be
-stable. However, as demonstrated above, differences in score reliability
-and true score variance tend to lead to larger corrected ES
-heterogeneity, implying that observed ES heterogeneity was reduced by
-differences in score reliability. However, whether these results
-actually generalize beyond these data-sets remains to be seen.
-Unfortunately, the behavioural sciences are in dire need of open-data
-that resemble multi-site replications. As the majority of results
-discussed over the last years (in e.g. ManyLabs or Registered
-Replication Reports) employ single-indicator scales as dependent
-measures, score reliability can not be easily estimated in order to
-replicate our analyses.
-
-Paragraph on implications -\> data collected using no specific sampling
-schemes - little surprise that the radomness induced by sampling
-schemes, which manifests itself in true score variance heterogeneity,
-largely surpasses the heterogeneity induced by imperfect measurement.
+reliability across replications in the current state of Psychology and
+the behavioural sciences, do not serve as appropriate explanations of ES
+heterogeneity. However, whether these results actually generalize beyond
+these data-sets remains to be seen. Unfortunately, the behavioural
+sciences are in dire need of open-data that resemble multi-site
+replications. As the majority of results discussed over the last years
+(in e.g. ManyLabs or Registered Replication Reports) employ
+single-indicator scales as dependent measures, score reliability can not
+be easily estimated in order to replicate our analyses.
 
@@
 assumptions is defensible in this case, as the analytical arguments
-derived from the inequality should still hold up, nevertheless.
+derived from the inequality should still hold up, nevertheless. \[add to
+appendix?\] Additionally, we used the empirical examples available to
+assess in how far estimates derived using the delta-method
+approximations diverge from estimates derived without it. The
+differences identified are rather miniscule at the third decimal,
+reinforcing our belief that these assumption violations are defensible.
 
@@
 that have been neglected in the application of meta-analytic research
-over the last decades. We strongly agree with their ideas that
-underappreciated (differences in un-)reliability introduce substantial
-biases in meta-analytic estimates and tests in Psychology, which need to
-be corrected. Future applications of meta-analyses on standardized ES
-need to take score reliability into account to arrive at correct
-estimates and inferences. However, the meta-analyst should not expect
-that correcting ES for score reliability will reduce the extent of
-heterogeneity identified.
+over the last decades. We generally agree with their ideas that
+underappreciated (differences in un-)reliability can introduce
+substantial biases in meta-analytic estimates and tests in Psychology.
+However, based on our findings, we believe that the issues created by
+ignoring random sampling is much more dire and warrants immediate
+attention. Only once we can be sure that the variance in true scores or
+latent factors is actually meaningful for some population, we can turn
+towards imprecise measurements, expecting correction procedures to
+explain unwarranted heterogeneity. In the current state of Psychology,
+meta-analysts should not expect that correcting ES for score reliability
+will actually explain heterogeneity identified.

### 2025-03-27 Manuscript_MD_ScoreReLiability.qmd
minor fixed text arrange Results

@@
 zero or even in the negative space grows larger as well. If we know the
-mean size and heterogeneity of a phenomenon’s effect size, we would
+mean size and heterogeneity of a phenomenon’s effect, we would
 theoretically be able to establish an expected replication rate
@@
 replications in Psychology. In a separate re-analysis of large scale
-direct replications, Renkewitz et al. (2025) identify substantial
-heterogeneity in almost all projects where a non-zero effect could be
-identified. This aligns with the correlation found by Olsson-Collentine
-et al. (2020).
+direct replications, Renkewitz et al. (in preparation) identify
+substantial heterogeneity in almost all projects where a non-zero effect
+could be identified. This aligns with the correlation found by
+Olsson-Collentine et al. (2020).
 
@@
 replications on psychological phenomena (Fünderich et al. 2024). The
-data-sets of about 50 phenomena , largely stemming from the efforts of
+data-sets of about 50 phenomena, largely stemming from the efforts of
 the Many Labs studies, Registered Replication Reports or the
@@
 replications, but focus on original research or conceptual replications.
-However, as for each project, across all sites the same protocol is
-used, and the subjects are distributed across different countries, a
-data structure similar to that of a multi-site replication study
-emerges. What makes all these collaborative efforts valuable for this
-manuscript is the fact that all phenomena collected in the
-DRIPHT-repository are making use of two-group designs and are therefore
-easily assessed using standardised effect sizes.
+However, as for each project, across all sites the same protocol is used
+and the subjects are distributed across different countries, a data
+structure similar to that of a multi-site replication study emerges.
+What makes all these collaborative efforts valuable for this manuscript
+is the fact that all phenomena collected in the DRIPHT-repository are
+making use of two-group designs and are therefore easily assessed using
+standardised effect sizes.
 
@@
 \[osf-link\], while a brief summary can be found in Table 1.
-
-[Equations @eq-d], [-@eq-SEd], [-@eq-dc], and [-@eq-SEdc] were used to
-generate estimates of standardized ES, both uncorrected and corrected,
-with their corresponding standard errors. Generally, this leads to
-larger (absolute) ES and larger standard errors. As an example from the
-21 selected phenomena, the results from this procedure on Nosek’s
-Explicit Art sex differences phenomenon are displayed in the forest plot
-in @fig-forest. Here, grey dots represent the uncorrected ES for each
-sample, while black dots represent the corrected ES in each sample. The
-bars surrounding these dots show the respective 95%-Cis.
-
-In @fig-forest it becomes apparent that the reliability attenuation
-procedure leads to an increase in the individual absolute effect size,
-as the black dots are moved further away from zero. Simultaneously, the
-standard errors grew larger after the attenuation correction, which
-leads to larger 95%-CIs. This is most easily observed for the
-mturk-sample, which already had a rather large confidence interval to
-begin with. Additionally, in the diamond at the bottom of the figure, we
-can see that the meta-analytic estimate mean ES is also larger after the
-attenuation correction took place. In @tbl-meanES, the estimates and
-tests on the meta-analytic average ES, both corrected and uncorrected,
-can be found.
-
-```{r, echo = FALSE}
-#| label: fig-forest
-#| fig-cap: Forest Plot Nosek_Explicit_Art
-forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
-```
-
-```{r, echo = FALSE}
-#| label: tbl-meanES
-#| tbl-cap: Meta-analytic average  ES
-knitr::kable(x = read.csv(here("Tables/Mean_ES.csv")))
-```
-
-@tbl-meanES demonstrates that what we observed in Figure 1 holds across
-all 21 phenomena. All meta-analytic effects are larger in their absolute
-size after applying an attenuation correction procedure. Similarly, the
-uncertainty quantified in the estimates’ standard error is larger.
-Additionally, @tbl-meanES highlights which phenomena pass criterion (3)
-– the effect size must be statistically significantly distinguishable
-from zero. 12 phenomena pass this criterion, highlighted in boldface. It
-may be remarked that the attenuation correction procedure, in this data,
-appears to have no influence on whether a phenomenon passes the
-criterion. The differences in p-value are rather small and lead to no
-difference in conclusion, regarding a significance level of .05.
 
@@
 
-### Results
+### Results - Wich phenomena pass criterion 3)?
+
+[Equations @eq-d], [-@eq-SEd], [-@eq-dc], and [-@eq-SEdc] were used to
+generate estimates of standardized ES, both uncorrected and corrected,
+with their corresponding standard errors. Generally, this leads to
+larger (absolute) ES and larger standard errors. As an example from the
+21 selected phenomena, the results from this procedure on Nosek’s
+Explicit Art sex differences phenomenon are displayed in the forest plot
+in @fig-forest. Here, grey dots represent the uncorrected ES for each
+sample, while black dots represent the corrected ES in each sample. The
+bars surrounding these dots show the respective 95%-Cis.
 
 ```{r, echo = FALSE}
-#| label: tbl-tauES
-#| tbl-cap: Tests for ES Heterogeneity
-#knitr::kable(x = read.csv(here("Tables/Heterogeneity_ES.csv")))
-
-tab_hetES <- read.csv(here("Tables/Heterogeneity_ES.csv"))
-tab_hetrel <- read.csv(here("Tables/Results_RMA_alpha.csv"))
-tab_muES <- read.csv(here("Tables/Mean_ES.csv"))
-
-tab_both <- data.frame(tab_hetES$MASC,
-                       tab_hetrel[,c("mu_alpha_str", "tau_alpha", "QE_str", "QEp")],
-                       tab_hetES[,c("tau_raw", "QE_raw_str", "QEp_raw", "tau_cor", "QE_cor_str", "QEp_cor")])
-
-knitr::kable(tab_both,
-             col.names = c("MASC", 
-                           "$\\hat{\\mu}_{\\alpha}$ (95%-CI)", 
-                           "$\\hat{\\tau}_{\\alpha}$", 
-                           "QE (df)", "p", 
-                           "$\\hat{\\tau}_{\\delta_0}$", 
-                           "QE (df)", "p",
-                           "$\\hat{\\tau}_{\\delta}$", 
-                           "QE (df)", "p"),
-             escape = FALSE)
-# 
-# tab_one <- data.frame(tab_muES,
-#                       tab_hetrel[,c("mu_alpha_str", "tau_alpha", "QE_str", "QEp")])
+#| label: fig-forest
+#| fig-cap: Forest Plot Nosek_Explicit_Art
+forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
 ```
+
+In @fig-forest it becomes apparent that the reliability attenuation
+procedure leads to an increase in the individual absolute effect size,
+as the black dots are moved further away from zero. Simultaneously, the
+standard errors grew larger after the attenuation correction, which
+leads to larger 95%-CIs. This is most easily observed for the
+mturk-sample, which already had a rather large confidence interval to
+begin with. Additionally, in the diamond at the bottom of the figure, we
+can see that the meta-analytic estimate mean ES is also larger after the
+attenuation correction took place. In @tbl-meanES, the estimates and
+tests on the meta-analytic average ES, both corrected and uncorrected,
+can be found.
+
+@tbl-meanES demonstrates that what we observed in Figure 1 holds across
+all 21 phenomena. All meta-analytic effects are larger in their absolute
+size after applying an attenuation correction procedure. Similarly, the
+uncertainty quantified in the estimate's standard error is larger.
+Additionally, @tbl-meanES highlights which phenomena pass criterion (3)
+– the effect size must be statistically significantly distinguishable
+from zero. 12 phenomena pass this criterion, highlighted in boldface. It
+may be remarked that the attenuation correction procedure, in this data,
+appears to have no influence on whether a phenomenon passes the
+criterion. The differences in p-value are rather small and lead to no
+difference in conclusion, regarding a significance level of .05.
+
+```{r, echo = FALSE}
+#| label: tbl-meanES
+#| tbl-cap: Meta-analytic average  ES
+knitr::kable(x = read.csv(here("Tables/Mean_ES.csv")))
+```
+
+### Results - Estimates of heterogeneity
 
@@
 actually does reduce, and therefore explain, ES heterogeneity.
+
+```{r, echo = FALSE}
+#| label: tbl-tauES
+#| tbl-cap: Tests for ES Heterogeneity
+#knitr::kable(x = read.csv(here("Tables/Heterogeneity_ES.csv")))
+
+tab_hetES <- read.csv(here("Tables/Heterogeneity_ES.csv"))
+tab_hetrel <- read.csv(here("Tables/Results_RMA_alpha.csv"))
+
+tab_both <- data.frame(tab_hetES$MASC,
+                       tab_hetrel[,c("mu_alpha_str", "tau_alpha", "QE_str", "QEp")],
+                       tab_hetES[,c("tau_raw", "QE_raw_str", "QEp_raw", "tau_cor", "QE_cor_str", "QEp_cor")])
+
+knitr::kable(tab_both,
+             col.names = c("MASC", 
+                           "$\\hat{\\mu}_{\\rho_{XX'}}$ (95%-CI)", 
+                           "$\\hat{\\tau}_{\\rho_{XX'}}$", 
+                           "QE (df)", "p", 
+                           "$\\hat{\\tau}_{\\delta_0}$", 
+                           "QE (df)", "p",
+                           "$\\hat{\\tau}_{\\delta}$", 
+                           "QE (df)", "p"),
+             escape = FALSE)
+# 
+# tab_one <- data.frame(tab_muES,
+#                       tab_hetrel[,c("mu_alpha_str", "tau_alpha", "QE_str", "QEp")])
+```
+
+*Note:* $\hat{\mu}_{\rho_{XX'}}$ is the back-transformed meta-analytic
+estimate of mean score reliability; $\tau{\mu}_{\rho_{XX'}}$ is the
+back-transformed estimate of heterogeneity in score reliability; Note
+that, since in a meta-analysis, we attempt to estimate mean and variance
+of the parameter's true distribution, in this table
+$\hat{\mu}_{\delta_0}$ represents the meta-analytic estimate of mean
+uncorrected $d$ and $\hat{\tau}_{\delta_0}$ represents the estimate of
+its heterogeneity; accordingly, $\hat{\mu}_{\delta}$ represents the
+estimate of the meta-analytic mean of corrected $d_c$ and
+$\hat{\tau}_{\delta}$ the estimate of its heterogeneity.
 
@@
 
-## Alternative discussion of how score reliability affects ES heterogeneity
+## Relationship between score reliability and effect sizes in terms of heterogeneity
 
@@
 
-### Misinterpreted equation in Hunter & Schmidt (XXXX)
+### Hunter & Schmidt's misinterpreted equation
 
@@
 
-The metric $R_2$ on the other hand describes, in how far the attenuation
+The metric $R_2$ on the other hand describes in how far the attenuation
 correction procedure has successfully reduced heterogeneity in score
 variance $\tau^2_{\sigma^2_X}$. A value of 1 indicates that the
-heterogeneity in true score variance $\tau^2_{\sigma^2_T}$, the score
-variance after the attenuation correction, is essentially just as large
-as the heterogeneity initially observed. Smaller values indicate how
-much heterogeneity is "left", after applying an attenuation correction,
-relative to the initial heterogeneity. This means that a value of .7
-indicates that about 70% of heterogeneity in score variance remains,
-even after correcting for differences in score reliability. As only 30%
-of score variance heterogeneity could be removed, this would imply that
-differences in error score variance $\sigma^2_E$ were responsible for
-less than a third of the heterogeneity in score variances found. On the
-contrary, more than two thirds of heterogeneity could be attributed to
-actual differences in how the underlying true scores are distributed
-across samples.
+heterogeneity in true score variance $\sigma^2_T$ is essentially just as
+large as the heterogeneity initially observed. Smaller values indicate
+how much heterogeneity is "left", after applying an attenuation
+correction, relative to the initial heterogeneity. This means that a
+value of .7 indicates that about 70% of heterogeneity in score variance
+remains, even after correcting for differences in score reliability. As
+only 30% of score variance heterogeneity could be removed, this would
+imply that differences in error score variance $\sigma^2_E$ were
+responsible for less than a third of the heterogeneity in score
+variances found. On the contrary, more than two thirds of heterogeneity
+could be attributed to actual differences in how the underlying true
+scores are distributed across samples.
 
 Concerning the re-arranged inequality from Appendix A, we identify one
-crucial condition described in @eq-ineq_short.
+crucial condition or rule described in @eq-ineq_short.
 
@@
 is bound to hold true. This means that as long as the average score
-reliability, to the power of 3, does not exceed the reduction in score
-variance heterogeneity due to the attenuation correction, @eq-ineq_init
+reliability, to the power of 3, does not exceed the relative remaining
+variance heterogeneity after the attenuation correction, @eq-ineq_init
 is not violated. For example, assuming the average score reliability is
@@
 
-$R_1$ essentially describes the reduction in mean score variance due to
-the attenuation correction, and $R_2$ essentially described the
-reduction in heterogeneity of score variance due to the attenuation
+$R_1$ essentially describes the relative size of mean score variance
+after the attenuation correction, and $R_2$ essentially described the
+relative size of score variance heterogeneity after the attenuation
 correction. Therefore, @eq-ineq_init states that for such a correction
@@
 stronger weight, as $R_1$ is taken to the power of 3 in @eq-ineq_init.
+The smaller $R_1$ or $R_2$, the larger the reduction in heterogeneity or
+mean value.
 
@@
 
+It is important to note that the condition in @eq-ineq_short only
+describes under which conditions the inequality is guaranteed to hold,
+no matter what value other parameters like the mean value or
+heterogeneity in mean differences take on. This means that, even if the
+condition in @eq-ineq_short is violated, it is by no means guaranteed
+that an attenuation correction procedure would actually reduce
+heterogeneity in ES. This means that inference made from @eq-ineq_short
+is asymmetrical. While we know, if the inequality holds, heterogeneity
+in corrected ES will be larger than heterogeneity in uncorrected ES, we
+can't be certain that the opposite is true if the inequality does not
+hold.
+
 ***\[maybe move to discussion further below\]*** The majority of studies
@@
 boot-err (ref.) as an alternative method to test for and estimate the
-differences in error score variance $\sigma^2_E$. In this text, we
-deliver detailed instructions on how estimates for mean and
+differences in error score variance $\sigma^2_E$. In Beinhauer et al.
+(ref.), we deliver detailed instructions on how estimates for mean and
 heterogeneity of total score variance, and the score variance components
@@
 to the rest of the results. For the Shnabel_Willingness_Reconcile_RPP
-phenomenon, $R_2$ is very low, below .25. This makes it even lower than
-its $R_1^3 = .574$. This would imply that for this phenomenon, the
+phenomenon, $R_2$ is very low, below .25. This makes it lower than its
+$R_1^3 = .574$. This would imply that for this phenomenon, the
 conditions for a reduction in ES heterogeneity would be met. However, in
@@
 components, even more so than for ES, largely depends on the sample
-size. Beinhauer et al. (ref) come to the conclusion that at 10 items,
-which the WIllingness to Reconcile scale consists of, a sample size
-around 33 is required to achieve sufficient power for a statistical test
-of score variance heterogeneity. Most likely, the low size led to
-implausible estimates in both true and total score variance
+size. In Beinhauer et al. (ref) we come to the conclusion that at 10
+items, which the Willingness to Reconcile scale consists of, a sample
+size around 33 is required to achieve sufficient power for a statistical
+test of score variance heterogeneity. Most likely, the low sample size
+led to implausible estimates in both true and total score variance
 heterogeneity and subsequently to an implausible estimate of $R_2$.
@@
 
-Generally, @tbl-tauES implies that for almmost all measurements
+Generally, @tbl-tauES implies that for almost all measurements
 concerning these phenomena, the heterogeneity in observed score
@@
 #| tbl-cap: Re-analysis concerning metrics $R_1$ and $R_2$
-knitr::kable(read.csv(here("Tables/Variances_analysis.csv")))
+knitr::kable(read.csv(here("Tables/Variances_analysis.csv")),
+             col.names = c("MASC", 
+                           "$\\hat{\\mu}_{\\sigma^2_X}$", 
+                           "$\\hat{\\mu}_{\\sigma^2_T}$", 
+                           "$\\hat{\\tau}_{\\sigma^2_X}$", 
+                           "$\\hat{\\tau}_{\\sigma^2_T}$", 
+                           "$R_1$", 
+                           "$R_2$"))
 ```
@@
 score variance heterogeneity. At the same time, score reliability is a
-low-hanging fruit for researchers to optimise most researchers are aware
-and trained in assessing score reliability.
+low-hanging fruit to optimise as most researchers are aware and trained
+in assessing score reliability.
 
@@
 
-If we, in the space of behavioural sciences, keep making use of
-convenience samples simply because they are so convenient, we can not
-expect anything to change. Most likely, ES heterogeneity will remain a
+If we, in the space of Psychology, keep making use of convenience
+samples simply because they are so convenient, we can not expect
+anything to change. Most likely, ES heterogeneity will remain a
 substantial factor hampering the interpretability of original and
@@
 derived from the inequality should still hold up, nevertheless. \[add to
-appendix?\] Additionally, we used the empirical examples available to
+appendix!\] Additionally, we used the empirical examples available to
 assess in how far estimates derived using the delta-method

### 2025-03-27 Manuscript_MD_ScoreReLiability.qmd
fix inconsistent issue by more bootstrap

@@
 phenomena, the $R_2$ metric is larger than the $R_1$ metric.
-Additionally, we observe that across nine phenomena, $R_2$ is larger
+Additionally, we observe that across eight phenomena, $R_2$ is larger
 than .96. This implies that of the heterogeneity observed in score
@@
 
-For one of the phenomena, however, we observe a large disparity compared
-to the rest of the results. For the Shnabel_Willingness_Reconcile_RPP
-phenomenon, $R_2$ is very low, below .25. This makes it lower than its
-$R_1^3 = .574$. This would imply that for this phenomenon, the
-conditions for a reduction in ES heterogeneity would be met. However, in
-@tbl-tauES, we find no such reduction. What becomes apparent, however,
-is that for this phenomenon we are working with the low number of 9
-administration sites. Estimation precision for heterogeneity in variance
-components, even more so than for ES, largely depends on the sample
-size. In Beinhauer et al. (ref) we come to the conclusion that at 10
-items, which the Willingness to Reconcile scale consists of, a sample
-size around 33 is required to achieve sufficient power for a statistical
-test of score variance heterogeneity. Most likely, the low sample size
-led to implausible estimates in both true and total score variance
-heterogeneity and subsequently to an implausible estimate of $R_2$.
-Meta-analytic estimates, especially those concerning score variance
-components such as $R_2$, therefore should be interpreted with caution,
-if the sample size was particularly low.
-
-Similarly, for the two exceptional phenomena where the attenuation
-correction procedure actually led to an increase in ES heterogeneity
+```{r, echo = FALSE}
+#| label: tbl-R1R2
+#| tbl-cap: Re-analysis concerning metrics $R_1$ and $R_2$
+knitr::kable(read.csv(here("Tables/Variances_analysis.csv")),
+             col.names = c("MASC", 
+                           "$\\hat{\\mu}_{\\sigma^2_X}$", 
+                           "$\\hat{\\mu}_{\\sigma^2_T}$", 
+                           "$\\hat{\\tau}_{\\sigma^2_X}$", 
+                           "$\\hat{\\tau}_{\\sigma^2_T}$", 
+                           "$R_1$", 
+                           "$R_2$"))
+```
+
+For the two exceptional phenomena where the attenuation correction
+procedure actually led to an increase in ES heterogeneity
 (Srull_Behavior_Hostility and Shnabel_Willingness_Reconcile_RPP), the
@@
 that the majority of score variance heterogeneity was due to differences
-in true score variance $\sigma^2_T$. Both phenomena were examined using
-a rather low number of administration sites (22 and 9 respectively),
-implying that issues of estimation precision might explain the
-inconclusive results. Lastly, for one out of the 12 phenomena, the
-Albarracin_Priming_SAT phenomenon, we find a $\tau_{\sigma^2_X}$ of 0.
-Therefore, the $R_2$ metric could not be computed, as the denominator
-would have been zero. However, we also did not identify statistically
-significant ES heterogeneity in the first place. This phenomenon was
-assessed with a particularly low number of replications (8). As
-discussed, adequate power to detect heterogeneity in score reliability
-and observed score variance requires a larger number of replications
-than adequate power to detect heterogeneity.
+in true score variance $\sigma^2_T$. Since either the heterogeneity was
+so small or the sample size so low that the heterogeneity could not be
+statistically distinguished from zero, we should be cautious in the
+interpretation of a reduction in that estimate. Most likely, the
+reduction in heterogeneity is better explained by imprecise estimation
+than an actual explanation of unexplained heterogeneity.
+
+Lastly, for two out of the 12 phenomena, the Albarracin_Priming_SAT and
+the Shnabel_Willingness_Reconcile_Rev phenomena, we find
+$\tau_{\sigma^2_X}$ of 0. Therefore, the $R_2$ metric could not be
+computed, as the denominator would have been zero. However, we also did
+not identify statistically significant ES heterogeneity in the first
+place. The phenomena were assessed with a particularly low number of
+replications (8 in both cases). Estimation precision for heterogeneity
+in variance components, even more so than for ES, largely depends on the
+sample size. Most likely, the low sample size led to implausible
+estimates in both true and total score variance heterogeneity and
+subsequently to an implausible estimate of $R_2$. Meta-analytic
+estimates, especially those concerning score variance components such as
+$R_2$, therefore should be interpreted with caution, if the sample size
+was particularly low.
 
@@
 accompanied by a reduction in differences due to the correction.
-
-```{r, echo = FALSE}
-#| label: tbl-R1R2
-#| tbl-cap: Re-analysis concerning metrics $R_1$ and $R_2$
-knitr::kable(read.csv(here("Tables/Variances_analysis.csv")),
-             col.names = c("MASC", 
-                           "$\\hat{\\mu}_{\\sigma^2_X}$", 
-                           "$\\hat{\\mu}_{\\sigma^2_T}$", 
-                           "$\\hat{\\tau}_{\\sigma^2_X}$", 
-                           "$\\hat{\\tau}_{\\sigma^2_T}$", 
-                           "$R_1$", 
-                           "$R_2$"))
-```
 

### 2025-03-27 E2_Data-Analysis_explore.R

@@ -74,7 +74,7 @@ varE_est.L <- lapply(which(effect_index)[nn_eff_idx], FUN = function(x){
   # sometimes calculation of SE might lead to issues, as negative variances can not be log-transformed
   #  therefore, function is run within a tryCatch-environment, so the script does not crash
   # apply_Bootstrap_SE_Project.specific is the project-specific function
-  tryCatch(apply_Bootstrap_SE_Project.specific(na.omit(data.list[[x]]), component = "E", R = 100),
+  tryCatch(apply_Bootstrap_SE_Project.specific(na.omit(data.list[[x]]), component = "E", R = 3000),
            
            # print error to console
            error = function(e)(cat("ERROR: ", conditionMessage(e), " - ",
@@ -114,7 +114,7 @@ varX_est.L <- lapply(which(effect_index)[nn_eff_idx], FUN = function(x){
   # sometimes calculation of SE might lead to issues, as negative variances can not be log-transformed
   #  therefore, function is run within a tryCatch-environment, so the script does not crash
   # apply_Bootstrap_SE_Project.specific is the project-specific function
-  tryCatch(apply_Bootstrap_SE_Project.specific(na.omit(data.list[[x]]), component = "X", R = 100),
+  tryCatch(apply_Bootstrap_SE_Project.specific(na.omit(data.list[[x]]), component = "X", R = 3000),
            
            # print error to console
            error = function(e)(cat("ERROR: ", conditionMessage(e), " - ",
@@ -162,17 +162,19 @@ vars_rma_df <- do.call(rbind, vars_rma_L) %>%
   mutate(CV_E = sqrt(tau2_E)/mu_E,
          CV_X = sqrt(tau2_X)/mu_X,
          CV_T = sqrt(tau2_T)/mu_T) %>% 
+  mutate(tau_X = sqrt(tau2_X),
+         tau_T = sqrt(tau2_T)) %>% 
   mutate(R1 = mu_T / mu_X,
          R2 = tau2_T / tau2_X) %>% 
   mutate(MASC = ES_rma_df$MASC[ES_rma_df$corr == 0][nn_eff_idx]) %>% 
-  mutate_if(is.numeric, round, 3)
+  mutate_if(is.numeric, round, 3) 
 
 
 
 
 
 write.csv(vars_rma_df %>% 
[…]

### 2025-03-27 ReLiability_Function-library.R

@@ -258,11 +258,11 @@ forest_plot_rel <- function(rma_df, aggregates, ci.lvl = .975){
     arrange(desc(d_raw)) %>% 
     ggplot() +
     # point estimate of raw d
-    geom_point(aes(x = d_raw, y = 1:nrow(aggregates)), colour = "darkgrey") +
+    geom_point(aes(x = d_raw, y = 1:nrow(aggregates)-.3), colour = "darkgrey") +
     # CI of point estimate of raw d
-    geom_segment(aes(x = cil_raw, y = 1:nrow(aggregates), xend = ciu_raw, yend = 1:nrow(aggregates)), colour = "darkgrey") +
-    geom_segment(aes(x = cil_raw, xend = cil_raw, y = (1:nrow(aggregates))+.3, yend = (1:nrow(aggregates))-.3), colour = "darkgrey") +
-    geom_segment(aes(x = ciu_raw, xend = ciu_raw, y =( 1:nrow(aggregates))+.3, yend = (1:nrow(aggregates))-.3), colour = "darkgrey") +
+    geom_segment(aes(x = cil_raw, y = 1:nrow(aggregates)-.3, xend = ciu_raw, yend = 1:nrow(aggregates)-.3), colour = "darkgrey") +
+    geom_segment(aes(x = cil_raw, xend = cil_raw, y = (1:nrow(aggregates))+.3-.3, yend = (1:nrow(aggregates))-.3-.3), colour = "darkgrey") +
+    geom_segment(aes(x = ciu_raw, xend = ciu_raw, y =( 1:nrow(aggregates))+.3-.3, yend = (1:nrow(aggregates))-.3-.3), colour = "darkgrey") +
     # point estiamte of corrected d
     geom_point(aes(x = d_corr, y = 1:nrow(aggregates)), colour = "black") +
     # CI of point estimate of corrected d
@@ -273,7 +273,7 @@ forest_plot_rel <- function(rma_df, aggregates, ci.lvl = .975){
     geom_abline(slope = 0, intercept = -1, colour = "black") +
     # adding diamond showing rma-estimate of raw d
     geom_polygon(data = data.frame(x = c(rma_raw$ci.ub, rma_raw$mu, rma_raw$ci.lb, rma_raw$mu),
-                                   y = c(-3, -3-.7, -3, -3+.7)),
+                                   y = c(-3, -3-.7, -3, -3+.7)-.5),
                  aes(x = x, y = y),
[…]

### 2025-03-28 Manuscript_MD_ScoreReLiability.qmd
update MS add appendix

@@
 across two situations: first, heterogeneity of raw ES, as computed in
-equation 1 is assessed. Secondly, ES are corrected for imperfect
-reliability. Computing the heterogeneity of these corrected ES allows
-for an assessment, whether the heterogeneity did indeed shrink compared
-to the first situation, as predicted in W&D and H&S.
+@eq-d is assessed. Secondly, ES are corrected for imperfect reliability.
+Computing the heterogeneity of these corrected ES allows for an
+assessment, whether the heterogeneity did indeed shrink compared to the
+first situation, as predicted in W&D and H&S.
 
@@
 
-### Results - Wich phenomena pass criterion 3)?
+### Results - Which phenomena pass criterion 3)?
 
@@
 to estimate heterogeneity in raw ES $\delta_0$. Similarly, we can adjust
-Equation @eq-ratio_d0 to describe how, instead of total score standard
-deviation, mean and variance in true score standard deviation $\sigma_T$
-affect heterogeneity in corrected ES $\delta$.
+@eq-ratio_d0 to describe how, instead of total score standard deviation,
+mean and variance in true score standard deviation $\sigma_T$ affect
+heterogeneity in corrected ES $\delta$.
 
@@
 
-```{r, echo = FALSE}
-#| label: tbl-R1R2
-#| tbl-cap: Re-analysis concerning metrics $R_1$ and $R_2$
-knitr::kable(read.csv(here("Tables/Variances_analysis.csv")),
-             col.names = c("MASC", 
-                           "$\\hat{\\mu}_{\\sigma^2_X}$", 
-                           "$\\hat{\\mu}_{\\sigma^2_T}$", 
-                           "$\\hat{\\tau}_{\\sigma^2_X}$", 
-                           "$\\hat{\\tau}_{\\sigma^2_T}$", 
-                           "$R_1$", 
-                           "$R_2$"))
-```
-
 For the two exceptional phenomena where the attenuation correction
@@
 accompanied by a reduction in differences due to the correction.
+
+```{r, echo = FALSE}
+#| label: tbl-R1R2
+#| tbl-cap: Re-analysis concerning metrics $R_1$ and $R_2$
+knitr::kable(read.csv(here("Tables/Variances_analysis.csv")),
+             col.names = c("MASC", 
+                           "$\\hat{\\mu}_{\\sigma^2_X}$", 
+                           "$\\hat{\\mu}_{\\sigma^2_T}$", 
+                           "$\\hat{\\tau}_{\\sigma^2_X}$", 
+                           "$\\hat{\\tau}_{\\sigma^2_T}$", 
+                           "$R_1$", 
+                           "$R_2$"))
+```
 
@@
 
-Heterogeneity in ES even in direct replications, where experimental
-factors are held as constant as possible, is already substantial
-(Renkewitz, Fünderich, & Beinhauer, 2024). However, so far we have
-demonstrated that in several scenarios, differences in score reliability
-did not serve as appropriate moderators or explanations of ES
-heterogeneity. Across most phenomena, we failed to reduce ES
-heterogeneity. If anything, we even increased ES heterogeneity, albeit
-to a very small degree. Generally, concerning the multi-site replication
-projects or collaborative research projects, accounting for differences
-in score reliability did not help us gain a better understanding of the
-phenomena.
+### Limitations & constraints on generalizability
 
@@
 
-The analytical arguments presented following Equation 17 rested on the
-use of the delta-method, as Equation 17 contained parameters of the
-score standard deviations’ distributions, while the $R_1$ and $R_2$
-metrics were designed to discuss score variances. The delta-method rests
-on a number of assumptions: (1) approximations derived from the
-delta-method require for the transformed variable to follow a marginal
-normal distribution (ref). Since the variable under discussion here is
-the standard deviation, which is bounded to be larger than zero and has
-a non-normal sampling distribution (ref), this assumption is most likely
-violated; (2) the derivative of the transformation function needs to be
-available. As the transformation function is essentially just the square
-root, this assumption should hold up; (3) as the delta-method
-approximates the parameters of the transformed variable using the
-first-order Taylor-expansion, this implies that higher-order terms are
-required to be negligible. In practice, this implies that the variance
-of the untransformed variable needs to be small. However, as this entire
-procedure discusses the influence of variance in standard deviations on
-ES heterogeneity, introduced by differences in score reliability, this
-assumption is probably violated as well. The violation of these
-assumptions likely introduce a bias in estimates derived from these
-approximations (ref). However, the inequality derived with use of the
-delta-method is not used to actually derive any estimates, but to
-discuss how an attenuation correction procedure can affect ES
-heterogeneity. Therefore, we believe that the violation of these
-assumptions is defensible in this case, as the analytical arguments
-derived from the inequality should still hold up, nevertheless. \[add to
-appendix!\] Additionally, we used the empirical examples available to
-assess in how far estimates derived using the delta-method
+The analytical arguments leading to @eq-ineq_short rested on the use of
+the delta-method, as @eq-ineq_init contained parameters of the score
+standard deviations’ distributions, while the $R_1$ and $R_2$ metrics
+were designed to discuss score variances. The delta-method rests on a
+number of assumptions, at least one of which is likely violated in our
+analysis: the variance of the variable to be transformed needs to be
+small (ref.). As this entire procedure discusses the influence of
+variance in standard deviations on ES heterogeneity, introduced by
+differences in score reliability, assuming that this assumption holds
+seems somewhat off. Therefore, we used the empirical examples available
+to assess in how far estimates derived using the delta-method
 approximations diverge from estimates derived without it. The
-differences identified are rather miniscule at the third decimal,
-reinforcing our belief that these assumption violations are defensible.
-
-Similarly, all Equations following Equation 10 rest on the assumption
+differences identified are rather miniscule for the majority of
+phenomena, at the third decimal. Only for the PSACR002_behav_init
+phenomenon we find larger differences - therefore, conclusions for this
+phenomenon concerning the relation between $R_1$ and $R_2$ need to be
+drawn with caution. Generally, as the majority of phenomena did not
+display any bias in this assessment and since the delta-method is not
+actually used to generate estimates, we believe its use to be defensible
+nonetheless.
+
+Similarly, all equations following @eq-XYZ_cov rest on the assumption
 that mean difference ($MD$) and true or total score variance
@@
 whether the relationship was positive or negative, indicated by a
-positive or negative covariance in Equation 10, would most likely have
+positive or negative covariance in @eq-XYZ_cov, would most likely have
 an impact on how strongly ES heterogeneity is affected by the
@@
 like Cohen’s d itself, strong violations of this assumption would go far
-beyond invalidating the claims we derived in Equation 16. Instead,
-assumptions of the meta-analytic tests concerning the size ES and the
-presence of ES heterogeneity would be violated, making any research on
-change in ES heterogeneity due to attenuation correction obsolete.
+beyond invalidating the claims we derived in @eq-ineq_short. Most
+likely, assumptions of the meta-analytic tests concerning the
+meta-analytic mean size ES and the presence of ES heterogeneity would be
+violated, making any research on change in ES heterogeneity due to
+attenuation correction obsolete.
+
+### Conclusions
+
+Heterogeneity in ES even in direct replications, where experimental
+factors are held as constant as possible, is already substantial
+(Renkewitz, Fünderich, & Beinhauer, 2024). However, so far we have
+demonstrated that in several scenarios, differences in score reliability
+did not serve as appropriate moderators or explanations of ES
+heterogeneity. Across most phenomena, we failed to reduce ES
+heterogeneity. If anything, we even increased ES heterogeneity, albeit
+to a very small degree. Generally, concerning the multi-site replication
+projects or collaborative research projects, accounting for differences
+in score reliability did not help us gain a better understanding of the
+phenomena. We want to reinforce that, in light of uncontrolled variation
+in score variance due to the lack of appropriate sampling schemes,
 
@@
 will actually explain heterogeneity identified.
+
+## Appendix
+
+### A - Rearranging inequality
+
+Both metrics $R_1$ and $R_2$ make use of parameters of the distribution
+of true and total score variance, not standard deviations. This choice
+was made to facilitate the role of score reliability in the subsequent
+discussion, as score reliability itself is defined by variances, not
+standard deviations (see @eq-rel).
+
+Using the metrics $R_1$ and $R_2$, defined in [Equations @eq-R1] and
+[-@eq-R2], we attempt to understand under which circumstances the
+inequality defined in @eq-ineq_init holds true. However, while the
+metrics make use of mean and heterogeneity of the score variances,
+@eq-ineq_init makes use of mean and heterogeneity of the standard
+deviations. In order to incorporate the metrics, it is necessary to
+reparameterise the equation. Using the delta method to approximate mean
+value and heterogeneity of variance from those of the standard
+deviations, we know that (ref):
+$$\mu^2_{\sigma_X} \approx \mu_{\sigma^2_X} \quad \quad  and \quad \quad 
+\tau^2_{\sigma_X} \approx \frac{\tau^2_{\sigma_X^2}}{4\mu_{\sigma^2_X}}$$ {#eq-delta_method}
+
+While @eq-delta_method only explicitly contains parameters for the
+observed score variance/standard deviation, the same can be done using
+the true score variance/standard deviation. Using the approximates
+defined in @eq-delta_method for observed and true score variance, we
+arrive at a new inequality in @eq-ineq_CV.
+
+$$\frac{\tau_{MD}^2}{\mu_{\sigma^2_t}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_t^2}} \frac{\tau^2_{\sigma_t^2}}{4\mu_{\sigma^2_t}} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4\mu_{\sigma^2_x}}$$ {#eq-ineq_CV}
+
+@eq-ineq_CV describes that the heterogeneity in corrected ES is larger
+than the heterogeneity in uncorrected ES, using parameters of the
+distributions of mean difference MD, true score variance $\sigma^2_T$
+and observed score variance $\sigma^2_X$. By introducing the metrics
+$R_1$ and $R_2$ to this inequality, we can begin to disentangle which
+conditions need to be fulfilled for the inequality to hold. Rearranging
+the terms in [Equations @eq-R1] and [-@eq-R2], and subsequently entering
+these terms into @eq-ineq_CV leads to the following
+$$\frac{\tau_{MD}^2}{R_1 \mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} \frac{R_2 \tau^2_{\sigma_x^2}}{4 R_1 \mu_{\sigma^2_x}} > \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4\mu_{\sigma^2_x}}$$ {#eq-ineq_R}
+
+@eq-ineq_R, again, describes the inequality that heterogeneity in
+corrected ES is larger than the heterogeneity in uncorrected ES,
+incorporating the metrics $R_1$ and $R_2$. This equation alone is not
+sufficient to identify the relevant circumstances required for the
+inequality to hold. However, rearranging the terms from @eq-ineq_R leads
+to @eq-ineq_fin:
+
+$$ \tau_{MD}^2\left(\frac{1}{R_1} - 1\right) + \mu_{MD}^2 \frac{1}{4} CV^2_{\sigma^2_X} \left(\frac{R_2}{R_1^3}  - 1\right) > 0$$ {#eq-ineq_fin}
+
+The circumstances under which the inequality described in @eq-ineq_fin
+holds, are circumstances where the claims made in W&D and H&S are
+directly contradicted, as in those cases the ES heterogeneity is larger
+after applying an attenuation correction procedure. Concerning the terms
+in @eq-ineq_fin, we know that all terms left-hand of the inequality,
+outside of the brackets ($\tau^2_{MD}$, $\mu^2_{MD}$ and
+$CV^2_{\sigma^2_X}$), are bound to be positive. For the inequality to
+hold, we need the left-hand side of the equation to remain positive,
+larger than zero. We can distill two scenarios, under which this
+equation should hold: (a) one of the terms inside the brackets
+($\frac{1}{R_1} - 1$ and $\frac{R^2_2}{R_1}  - 1$) is positive and large
+enough, so that the second term not containing that same bracket is
+positively dominated by the first term; or (b) both terms inside the
+brackets need to be positive.
+
+Generally, we know that both $R_1$ and $R_2$ are bound to be positive,
+as both contain different, strictly positive parameters of the
+distributions of true and observed score variance. Additionally, we know
+that $R_1$ is equivalent to the average score reliability, and therefore
+bound between $0\le R_1\le1$. Therefore, the term inside the first
+bracket is bound to be positive ($\frac{1}{R_1} – 1$). Similarly, $R_2$,
+as the ratio of true and error score variance heterogeneity, is bound
+between $0\le R_2\le1$. If $R_1^3$ is smaller than $R_2$, then the term
+inside the second bracket turns negative. Generally, this means that the
+inequality can only be violated if the reduction in relative
+heterogeneity of score variance ($R_2$) is larger than the the average
+score reliability to the power of 3. Restated in mathematical terms,
+this leads us to the rule/condition described in @eq-ineq_short, in the
+main text.
+
+### B - Back-transformation log-transformed score variance
+
+Using the methdology underlying the Boot-Err approach (Beinhauer et al.
+ref), we generate estimates of meta-analytic mean and heterogeneity in
+log-transformed total score variance $\sigma^2_X$. To facilitate
+interpretation and allow for computation of the metrics $R_1$ and $R_2$,
+we need to back-transform those estimates to the original scale. Using
+@eq-mu_sigmaX we can generate an estimate of meta-analytic mean score
+variance on the scale of $\sigma^2_X$ and using @eq-tau_sigmaX we get an
+estimate of its heterogeneity.
+
+$$\hat{\mu}_{\sigma^2_X} = exp\left(\hat{\mu}_{ln[\sigma^2_X]} + .5\  \hat{\tau}^2_{ln[\sigma^2_X]}\right)$$ {#eq-mu_sigmaX}
+
+$$\hat{\tau^2}_{\sigma^2_X} = exp\left(\hat{\tau}^2_{ln[\sigma^2_X]} - 1\right) \cdot \exp\left(2\ \hat{\mu}_{ln[\sigma^2_X]} + \hat{\tau}^2_{ln[\sigma^2_X]}\right)$$ {#eq-tau_sigmaX}
+
+In [equations @eq-mu_sigmaX] and [-@eq-tau_sigmaX],
+$\hat{\mu}_{ln[\sigma^2_X]}$ refers to the mean estimate from the
+meta-analysis on log-transformed score variance,
+$\hat{\tau}^2_{ln[\sigma^2_X]}$ refers to the estimate of heterogeneity.
+
+In [equations @eq-mu_sigmaE] and [-@eq-tau_sigmaE] we generate estimates
+of meta-analytic mean and heterogeneity of error score variance
+$\sigma^2_E$. Just like for the total score variance, we make use of the
+estimates of mean and heterogeneity from the meta-analysis on
+log-transformed error score variance.
+
+$$\hat{\mu}_{\sigma^2_E} = exp\left(\hat{\mu}_{ln[\sigma^2_E]} + .5\  \hat{\tau}^2_{ln[\sigma^2_E]}\right)$$ {#eq-mu_sigmaE}
+
+$$\hat{\tau^2}_{\sigma^2_E} = exp\left(\hat{\tau}^2_{ln[\sigma^2_E]} - 1\right) \cdot \exp\left(2\ \hat{\mu}_{ln[\sigma^2_E]} + \hat{\tau}^2_{ln[\sigma^2_E]}\right)$$ {#eq-tau_sigmaE}
+
+### C - Examining bias due to delta-method
+
+To examine whether applying the delta-method to derive the conditions
+under which the inequality @eq-ineq_init holds, we compared estimates of
+mean and heterogeneity of total score standard deviation$\sigma_X$. To
+generate estimates that do not depend on use of the delta-method, we
+used bootstrapping and a log-transformation, similar as detailed in
+@sec-methods2 for the score variances $\sigma^2_X$. Using the natural
+log, we transformed estimates of score standard deviation and using
+bootstrapped samples we derived appropriate standard errors. From a
+random-effects meta-analysis we generated estimates of mean and
+heterogeneity and back-transformed those to the original scale of
+$\sigma_X$. Additionally, we used the delta-method as outlined in
+@eq-delta_method to estimate mean and heterogeneity in $\sigma_X$ from
+estimates derived from meta-analyses on $\sigma^2_X$. The estimates can
+be found in @tbl-delta.
+
+```{r, echo = FALSE}
+#| label: tbl-delta
+#| tbl-cap: Empirical assessment of appropriateness of delta-method
+knitr::kable(read.csv(here("Tables/delta_test.csv")),
+             col.names = c("MASC", 
+                           "$\\hat{\\mu}_{\\sigma_X}$", 
+                           "$\\hat{\\tau}^2_{\\sigma_X}$", 
+                           "$\\hat{\\mu}_{\\Delta \\sigma_X}$", 
+                           "$\\hat{\\tau}^2_{\\Delta \\sigma_X}$"))
+```
+
+Generally, for 11 out of 12 phenomena, we find miniscule differences
+between the two estimation strategies. For all but the
+PSACR002_behav_init phenomenon, the differences occur in the 3rd
+decimal, indicating that violations in the assumptions concerning the
+delta-method are negligible here. Only for the PSACR002_behav_init
+phenomenon we find larger differences. These are large enough to be
+concerning and pertain largely the estimate of heterogeneity.
+Accordingly, inferences regarding this phenomenon need to be interpreted
+with caution. Generally, since for the majority of phenomena we find no
+bias and since the delta-method is not used to generate estimates beyond
+this particular section, we believe its use to be defensible.

### 2025-04-09 Manuscript_MD_ScoreReLiability.qmd
init manuscript V6 update

@@
 parameter that needs to be assessed. In the meta-analytic context,
-heterogeneity describes the variation in effect sizes, free of sampling
-error, if these are not constant across replications. Therefore, the
-presence of heterogeneity implies that some replications of a single
-phenomenon may be successful, while others are not. If heterogeneity
-grows, meaning the phenomenon’s effect size varies more strongly for
-unexplained reasons, the probability of observing an effect size around
-zero or even in the negative space grows larger as well. If we know the
-mean size and heterogeneity of a phenomenon’s effect, we would
-theoretically be able to establish an expected replication rate
-(reference).
+heterogeneity describes the variation of effect sizes, free of sampling
+error. Therefore, the presence of heterogeneity can imply that some
+replications of a single phenomenon may be successful, while others are
+not. If heterogeneity increases, meaning the phenomenon’s effect size
+varies more strongly for unexplained reasons, the probability of
+observing an effect size around zero or even in the negative space grows
+larger as well. If we know the mean size and heterogeneity of a
+phenomenon’s effect, we would theoretically be able to establish an
+expected replication rate (reference).
 
@@
 reflects a more advanced understanding of the subject matter being
-studied” (2021, p.2). Alternatively, Schuetze and von Hippel (2023)
-argue that heterogeneity in effects is an indicator of a vague, poorly
+studied” (2021, p.2). Similarly, Schuetze and von Hippel (2023) argue
+that heterogeneity in effects is an indicator of a vague, poorly
 specified theory.
@@
 (references), for the first time allow researchers to estimate
-heterogeneity undistorted by selection processes induced by publication
-bias. In re-analyses of these studies, Olsson-Collentine et al. (2020)
+heterogeneity undistorted by selection processes like publication bias.
+In re-analyses of these studies, Olsson-Collentine et al. (2020)
 identify a positive correlation between a phenomenon’s effect size and
-its heterogeneity. Similarly, van Erp et al. (2017) or Stanley et al.
+its heterogeneity. Similarly, van Erp et al. (2017) and Stanley et al.
 (2018) estimate strong degrees of heterogeneity across conceptual
@@
 vary in their extent across replications. This is an argument put
-forward repeatedly by original researchers and replicators alike.
-Literature discussing the role of score reliability for standardized
-effect sizes claims that differences in score reliability should
-inflate, and therefore increase, heterogeneity. Over the following
-pages, we assess whether score reliability can in fact be made
-responsible for heterogeneity identified across phenomena. Even though
-only a limited number of phenomena allow for such an analysis, we find
-no empirical evidence for the argument that differences in score
-reliability inflate effect size heterogeneity. Therefore, in the last
-part of this text we turn towards an alternative discussion using
-analytical arguments on how differences in measurement quality affect
-standardized effect sizes and their differences.
+forward repeatedly by original researchers and replicators alike (refs).
+Hunter and Schmidt (xxx) and Wiernik and Dahlke (xxx), discussing the
+role of score reliability for standardized effect sizes, claim that
+differences in score reliability should inflate, and therefore increase,
+heterogeneity. This implies that standard corrections for imperfect
+score reliability, such as attenuation correction procedures, should
+remove that inflation, reducing heterogeneity in corrected effect sizes.
+Over the following pages, we assess whether score reliability can in
+fact be made responsible for heterogeneity identified across phenomena.
+We find no empirical evidence for the argument that correcting for
+differences in score reliability reduces effect size heterogeneity.
+Instead, to our great surprise, we find that heterogeneity might as well
+be larger after the correction procedure. Therefore, theoretical
+analyses regarding the relationship between score reliability and
+heterogeneity need to be false. In the second half of this text we make
+use of analytical arguments to identify the necessary conditions under
+which an attenuation correction inevitably leads to an increase in
+effect size heterogeneity.
 
@@
 
-Here, MD refers to the difference between two groups of interest, while
-$\sigma_X$ is the total pooled standard deviation. Since $d$ constitutes
-an observed effect size it is affected by sampling error. In terms of
-heterogeneity, we are not interested in the variance of $d$, but the
-variance of the true, underlying ES $\delta$. The standard error of
-Cohen’s $d$, which is used to weight individual estimates in a
-random-effects meta-analysis (reference) is computed as defined in
+Here, MD refers to the mean difference between two groups of interest,
+while $\sigma_X$ is the total pooled standard deviation. Since $d$
+constitutes an observed effect size it is affected by sampling error. In
+terms of heterogeneity, we are not interested in the variance of $d$,
+but the variance of the true, underlying ES $\delta$. In a
+random-effects meta-analytic model, we can use the standard error and an
+estimate of heterogeneity to understand how far off the individual
+observed $d$ is from the true $\delta$ in a weighting procedure
+(reference). The standard error of Cohen’s $d$ is computed as defined in
 @eq-SEd.
@@
 
-Therefore, in the following section we assess to what extent this
-intuition is supported by empirical evidence. The re-analysis of
-archival data serves to shed light on the following two research
-questions central to this text:
-
--   Can score reliability be used to explain heterogeneity in
-    standardized effect sizes?
-
--   How does score reliability affect ES heterogeneity?
+Therefore, we assess the extent to which this intuition is supported by
+empirical evidence in the following section. The re-analysis of archival
+data serves to shed light on the following research question:
+
+-   To what extent can heterogeneity in standardized effect sizes be
+    explained by means of an attenuation correction?
 
@@
 identical protocols. From five published projects, the data for the
-first three and the fifth Many Labs projects was made publicly available
-when the DRIPHT repository was set up (references). The Registered
-Replication Reports are similar collaborative efforts, with the sole
-distinction that for each report a single phenomenon was replicated
-several times. The Registered Reports 3-10 were added to the DRIPHT
-repository, as they employed experimental designs with at least two
-groups. Lastly, some projects from the Psychological Science Accelerator
-were added to the DRIPHT repository. Different to Many Labs or
-Registered Replication Projects, these collaborative efforts, while also
-distributed across the globe, do not formally attempt to perform direct
-replications, but focus on original research or conceptual replications.
-However, as for each project, across all sites the same protocol is used
-and the subjects are distributed across different countries, a data
-structure similar to that of a multi-site replication study emerges.
-What makes all these collaborative efforts valuable for this manuscript
-is the fact that all phenomena collected in the DRIPHT-repository are
-making use of two-group designs and are therefore easily assessed using
-standardised effect sizes.
+first three and the fifth Many Labs projects was harmonized and publicly
+stored in a standardized format when the DRIPHT repository was set up
+(references). The Registered Replication Reports are similar
+collaborative efforts, with the sole distinction that for each report a
+single phenomenon was replicated several times. The Registered Reports
+3-10 were added to the DRIPHT repository, as they employed experimental
+designs with at least two groups. Lastly, some projects from the
+Psychological Science Accelerator were added to the DRIPHT repository.
+In contrast to Many Labs or Registered Replication Projects, these
+collaborative efforts, while also distributed across the globe, do not
+formally attempt to perform direct replications, but focus on original
+research or conceptual replications. However, as each project used the
+same protocol and participants are distributed across different
+countries, a data structure similar to that of a multi-site replication
+study emerges. What makes all these collaborative efforts valuable for
+this manuscript is the fact that all phenomena collected in the
+DRIPHT-repository are making use of two-group designs and are therefore
+easily assessed using standardised mean differences.
 
@@
 to fulfil three conditions: (1) As we focus on the use of ES $d$, the
-phenomenon needs to be assessed using a two-group design. This is the
-case for all projects in the DRIPHT repository; (2) Score reliability
-estimates can be derived, this implies that the phenomenon needs to be
-measured using either several indicators forming a single scale or by
-repeatedly measuring across several timepoints; (3) Score reliability
-can only attenuate effect sizes that are not zero in the first place.
-Therefore, we focus on phenomena where meta-analytic mean ES can be
-statistically distinguished from zero.
+phenomenon needs to be assessed using a group design consisting of at
+least two groups. (2) Score reliability estimates can be derived, this
+implies that the phenomenon needs to be measured using either several
+indicators forming a single scale or by repeatedly measuring across
+several timepoints. (3) Score reliability can only attenuate effect
+sizes that are not zero in the first place. Therefore, we focus on
+phenomena where meta-analytic mean ES can be statistically distinguished
+from zero. While the first criterion is fulfilled by all phenomena found
+in the DRIPHT repository, the second and third criteria still need
+evaluation.
 
@@
 used to construct the effect size is measured by employing more than a
-single indicator. At the same time, however, this implies that the
-remaining 29 phenomena did not make use of more than a single indicator
-to measure the dependent variable. Therefore, for the majority of
-phenomena it is not possible to assess in how far measurement quality is
-sufficiently high in terms of e.g. internal consistency or low random
-error variance.
+single indicator. The number of indicators per scale varied between 3
+and 30. The majority of scales was made up of 3 to 6 items. At the same
+time, however, this implies that the remaining 29 phenomena did not make
+use of more than a single indicator to measure the dependent variable.
+Therefore, for the majority of phenomena it is not possible to assess in
+how far measurement quality is sufficiently high in terms of e.g.
+internal consistency or low random error variance.
 
@@
 designs make use of some form of control-treatment manipulation,
-randomly assigning participants to a condition (20 phenomena), some
+randomly assigning participants to a condition (19 phenomena), some
 phenomena refer to the effect of pre-existing differences, e.g.
@@
 using metafor version X.XX. To estimate meta-analytic mean ES and
-heterogeneity, the REML-estimator is used (ref). To identify which
+heterogeneity we make ues of the REML-estimator (ref). To identify which
 phenomena pass criterion (3), using a Wald-type significance test we
@@
 different from zero. For this hypothesis test, as all other hypothesis
-tests in this manuscript, a one-tailed significance level of .05 is
-used.
+tests in this manuscript, we employ a one-tailed significance level of
+.05.
 
@@
 it avoids a mismatch in assumptions between how the score is computed
-and how score reliability is estimated. Using these estimates of
-Cronbach’s Alpha, the individual estimates of Cohen’s d are corrected,
+and how score reliability is estimated. The individual estimates of
+Cohen’s d are corrected using these estimates of Cronbach's Alpha,
 according to @eq-dc, including the corrected standard errors (@eq-SEdc).
@@
 heterogeneity in uncorrected ES and corrected ES are compared. Using
-Cochran’s Q-test, we assess whether the estimates of heterogeneity are
+Cochran’s *Q*-test, we assess whether the estimates of heterogeneity are
 statistically significantly different from 0. Only for those projects
@@
 
-These procedures are followed separately for each phenomenon. The
-statistical programming language R, Version X.XXX (reference) is used
-for all data manipulation and statistical analysis.
-
-### Results - Which phenomena pass criterion 3)?
+We executed these steps for each phenomenon. The statistical programming
+language R, Version X.XXX (reference) is used for all data manipulation
+and statistical analysis.
+
+### Results
 
@@
 with their corresponding standard errors. Generally, this leads to
-larger (absolute) ES and larger standard errors. As an example from the
-21 selected phenomena, the results from this procedure on Nosek’s
-Explicit Art sex differences phenomenon are displayed in the forest plot
-in @fig-forest. Here, grey dots represent the uncorrected ES for each
-sample, while black dots represent the corrected ES in each sample. The
-bars surrounding these dots show the respective 95%-Cis.
-
-```{r, echo = FALSE}
+larger (absolute) ES and larger standard errors after the correction. As
+an example from the 21 selected phenomena, the results from this
+procedure on Nosek’s Explicit Art sex differences phenomenon are
+displayed in the forest plot in @fig-forest. Here, grey dots represent
+the uncorrected ES for each sample, while black dots represent the
+corrected ES in each sample. The bars surrounding these dots show the
+respective 95%-Cis.
+
+```{r, echo = FALSE, fig.height = 6}
 #| label: fig-forest 
-#| fig-cap: Forest Plot Nosek_Explicit_Art
+#| #| fig-cap: Forest Plot Nosek_Explicit_Art 
 forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
@@
 
-@tbl-meanES demonstrates that what we observed in Figure 1 holds across
-all 21 phenomena. All meta-analytic effects are larger in their absolute
-size after applying an attenuation correction procedure. Similarly, the
-uncertainty quantified in the estimate's standard error is larger.
-Additionally, @tbl-meanES highlights which phenomena pass criterion (3)
-– the effect size must be statistically significantly distinguishable
-from zero. 12 phenomena pass this criterion, highlighted in boldface. It
-may be remarked that the attenuation correction procedure, in this data,
-appears to have no influence on whether a phenomenon passes the
-criterion. The differences in p-value are rather small and lead to no
-difference in conclusion, regarding a significance level of .05.
+#### Which phenomena pass criterion 3)?
+
+@tbl-meanES demonstrates that what we observed in @fig-forest holds
+across all 21 phenomena. All meta-analytic effects are larger in their
+absolute size after applying an attenuation correction procedure.
+Similarly, the uncertainty quantified in the estimate's standard error
+is larger. Additionally, @tbl-meanES highlights which phenomena pass
+criterion (3)—the effect size must be statistically significantly
+distinguishable from zero. 12 phenomena pass this criterion, highlighted
+in boldface. It may be remarked that the attenuation correction
+procedure, in this data, appears to have no influence on whether a
+phenomenon passes the criterion. The differences in p-value are rather
+small and lead to no difference in conclusion, regarding a significance
+level
 
@@
 #| tbl-cap: Meta-analytic average  ES
-knitr::kable(x = read.csv(here("Tables/Mean_ES.csv")))
+knitr::kable(x = read.csv(here("Tables/Mean_ES.csv")),
+             col.names = c("MASC", 
+                           "$\\hat{\\mu}_{\\delta_0}$ (SE)", 
+                           "p", 
+                           "$\\hat{\\mu}_{\\delta}$ (SE)", 
+                           "p"),
+             escape = FALSE)
 ```
 
-### Results - Estimates of heterogeneity
+*Note:* Since in a meta-analysis, we attempt to estimate mean and
+variance of the parameter's true distribution, in this table
+$\hat{\mu}_{\delta_0}$ represents the meta-analytic estimate of mean
+uncorrected $d$ and $\hat{\mu}_{\delta}$ represents the estimate of the
+meta-analytic mean of corrected $d_c$.
+
+#### Estimates of heterogeneity
 
@@
 one of the phenomena, the attenuation correction procedure barely has an
-impact on the extent of ES heterogeneity. For all phenomena but
-Srull_Behaviour_Hostility, the heterogeneity in ES before and after the
-correction changes very little. This implies that for 11 out of 12
-phenomena, differences in score reliability can not be made responsible
-for heterogeneity in standardized effect sizes. Only for the
-Srull_Behaviour_Hostility phenomenon, accounting for score reliability
-actually does reduce, and therefore explain, ES heterogeneity.
+impact on the extent of ES heterogeneity. However, while the
+heterogeneity for the Srull_Behaviour_Hostility phenomenon seems reduced
+by about a third of its initial value, the ES heterogeneity for this
+phenomenon is rather small to begin with. Using a significance test, we
+can not distinguish this value confidently from zero. This implies that
+for the majority of phenomena, differences in score reliability can not
+be made responsible for heterogeneity in standardized effect sizes.
 
@@
 of the parameter's true distribution, in this table
-$\hat{\mu}_{\delta_0}$ represents the meta-analytic estimate of mean
-uncorrected $d$ and $\hat{\tau}_{\delta_0}$ represents the estimate of
-its heterogeneity; accordingly, $\hat{\mu}_{\delta}$ represents the
-estimate of the meta-analytic mean of corrected $d_c$ and
-$\hat{\tau}_{\delta}$ the estimate of its heterogeneity.
+$\hat{\tau}_{\delta_0}$ represents the estimate of heterogeneity in
+uncorrected $d$; accordingly, $\hat{\tau}_{\delta}$ represents the
+estimate of the heterogeneity in corrected $d_c$.
 
@@
 attenuation correction increased ES heterogeneity. Concerning the
-remaining two phenomena, for the Shnabel_Willingness_Reconcile_Rev, the
+remaining two phenomena, for the Shnabel_Willingness_Reconcile_RPP, the
 reduction in heterogeneity is miniscule and the estimate of
@@
 appears that differences in score reliability do not inflate
-heterogeneity in uncorrected effect sizes, but instead mask
-heterogeneity, that can be unveiled by accounting for score reliability.
+heterogeneity in uncorrected effect sizes. Instead these differences
+mask ES heterogeneity, that can be unveiled by accounting for score
+reliability.
 
@@
 
-In the previous section we demonstrated that assumptions formulated in
-the literature regarding meta-scientific methodology are false. We found
-a large number of phenomena where, contrary to popular claims, score
-reliability does not explain ES heterogeneity. Over the following
-paragraphs, we discuss how the true relationship between these two
-parameters might look like. Therein, we also identify under which
-circumstances an attenuation correction can actually reduce ES
-heterogeneity - and where it is more likely to inflate it.
-
-### Hunter & Schmidt's misinterpreted equation
-
-Hunter & Schmidt claim to have derived an equation which demonstrates
-how heterogeneity in score reliability inflates heterogeneity in ESs.
-“If the level of reliability is independent of the true effect size
-across studies, then, to a close approximation:” (p. 309). However, as
-can be easily verified, the equation presented in Hunter & Schmidt
-(ref.), Equation X on p. 309, does not imply that assumption across all
-conditions.
+In the re-analysis of archival data in the previous section we
+demonstrated that, at least for the multi-site direct replication and
+PSA-projects, that score reliability does not hold any noteworthy value
+as a potential moderator explaining ES heterogeneity. Additionally, we
+demonstrated that the current consensus formulated in the literature
+regarding how score reliability increases ES heterogeneity is false. We
+found a large number of phenomena where, contrary to popular claims,
+correcting for score reliability increases ES heterogeneity, instead of
+decreasing it.
+
+This implies that there is an error in the reasoning or intuition behind
+claims made in H&S and W&D, as our observations can not be explained by
+the current literature. Over the following paragraphs, we will briefly
+explore the intuition that might have led to the previous claims.
+Subsequently, we will use analytical arguments to clarify how score
+reliability actually affects ES heterogeneity. Therein, we also identify
+potential conditions under which an attenuation correction can actually
+reduce ES heterogeneity - and under which it is more likely to increase
+it.
+
+### Intuitive relationship
+
+It might help to think about linear models, when discussing the
+intuition that might underlie the idea, that controlling for score
+reliability should reduce ES heterogeneity. Heterogeneity describes
+unexplained variation in a parameter's true distribution, that we model
+using meta-analytical techniques. In that way, the idea of heterogeneity
+is closely related to residual variance, as we observe it in a simple
+linear model. Both describe variation that could not be explained by the
+predictive variables in the case of the simple linear model, or not
+explained by incorporating the Standard Errors in the case of a
+meta-analytic model.
+
+For the simple linear model, if we add an additional predictor to the
+model that we know shares variance with the parameter that we aim to
+predict, this should reduce the amount of residual variance that we
+observe across most instances. Generally, removing a source of variance
+by controlling for it, we expect any unexplained variation to shrink.
+Similarly, we know that differences in score reliability, by definition,
+are related to differences in standardized effect sizes. Therefore,
+intuitively, it seems sensible to expect that controlling for
+differences in score reliability should reduce the unexplained variance
+in standardized effect sizes, as described in H&S and W&D.
+
+Hunter & Schmidt have additionally derived an equation which
+demonstrates how heterogeneity in score reliability affects
+heterogeneity in ESs.\@eq-HS implies that heterogeneity in $\delta_0$ is
+essentially a function in which heterogeneity and mean value of $\delta$
+are weighted by mean score reliability $\rho_{xx‘}$ and its
+heterogeneity.
 
@@
 undistorted by sampling error and measurement error. Lastly,
-$\rho_{XX'}$ refers to score reliability.
-
-@eq-HS implies that heterogeneity in $\delta_0$ is essentially a
-function in which heterogeneity and mean value of $\delta$ are weighted
-by mean score reliability $\rho_{xx‘}$ and its heterogeneity. Using
-selected values for the parameters in @eq-HS, we demonstrate how the
-heterogeneity in standardized ES can be both inflated and deflated by
-the presence of differences in score reliability. In @fig-hs, the
-results of employing @eq-HS to compute the heterogeneity (variance) in
-uncorrected ES $\delta_0$ is highlighted. All else held equal, only the
-individual level of mean reliability $\mu_{\rho_{xx‘}}$ is varied from
-.5 to .8. Mean corrected ES $\mu_{\delta_0}$ is held constant at 1, its
-heterogeneity $\tau_{\delta_0}$ held constant at .2 and the
-heterogeneity of score reliability $\tau_{\rho_{xx‘}}$ held constant at
-.12.
-
-```{r, echo = FALSE}
-#| label: fig-HS
-#| fig-cap: Implications by Hunter & Schmidt's @eg-HS \n Note: The red line corresponds to the variance of uncorrected ES $\delta_0$.
-
-library(ggplot2)
-
-mu_delta0 <- 1
-var_delta0 <- .2^2
-mu_alpha <- seq(from = .5, to = .8, length.out = 100)
-var_alpha <- .12^2
-mu_delta <- mu_delta0/sqrt(mu_alpha)
-
-# var_delta0 <- mu_alpha^2 * var_delta + mu_delta^2 * var_alpha
-var_delta <- (var_delta0 - mu_delta^2 * var_alpha) / (mu_alpha^2)
-
-ggplot() +
-  geom_hline(yintercept = .04, colour = "red", linewidth = 1, linetype = "dashed") +
-  geom_line(aes(y = var_delta, x = mu_alpha), linewidth = 2) +
-  theme(panel.background = element_rect(fill = "transparent"), 
-        plot.background = element_rect(fill = "transparent", colour = "transparent"), 
-        panel.grid.major.y = element_line(colour = "grey"),
-        panel.grid.major.x = element_line(colour = "grey"),
-        panel.grid.minor = element_line(colour = "transparent"),
-        axis.ticks = element_line(colour = "grey"),
-        strip.background = element_rect(fill = "transparent")) +
-  labs(y = expression(tau^2~"["~delta~"]"),
-       x = expression(mu~"["~rho[xx]~"]"))
-```
-
-If everything else is held constant, the variance in corrected ES
-$\delta$ can be a function only of the mean level of score reliability
-$\mu_{\rho_{xx‘}}$, according to H&S equation. In that case, Figure 2
-demonstrates that correcting for imperfect score reliability does not
-inevitably lead to lower heterogeneity, as the line representing
-heterogeneity in corrected ES does cross the red line, representing the
-heterogeneity in uncorrected ES. Depending on the mean level of score
-reliability, heterogeneity in uncorrected ES can be higher or lower than
-heterogeneity in corrected ES.
-
-However, @eq-hs is additionally flawed, as it violates a crucial
-definition. As Hunter & Schmidt point out, @eq-hs is only valid “If the
-level of reliability is independent of the true effect size across
-studies \[…\]” (2014, p. 309). However, if we assume that true score
-variance $\sigma^2_T$ varies across administrations, it is unlikely that
-this assumption can hold. As demonstrated further above in @eg-dct, ES
-$\delta$ is standardized using the true standard deviation $\sigma_T$ in
-the sense of CTT. The fact that this parameter is found both in @eq-dct
-and @eq-rel demonstrates that ES $\delta$ and score reliability
-$\rho_{xx‘}$ are unlikely to constitute truly independent variables, as
-long as true score variance is not identical across administrations.
+$\rho_{XX'}$ refers to score reliability. Hunter & Schmidt (xxx) claim
+that @eq-HS demonstrates that differences in score reliability inflate
+ES heterogeneity, which means that an attenuation correction procedure
+would have to reduce ES heterogeneity. However, it can be easily
+verified that @eq-HS does not allow for such claims. Depending on both
+the mean value $\mu_{\rho_{xx‘}}$ and the heterogeneity
+$\tau^2_{\rho_{xx‘}}$ in score reliability, ES heterogeneity might be
+increased or decreased by the differences in score reliability.
+
+Most likely, Hunter and Schmidt did not fully explore the implications
+of @eq-HS, but followed the general intuition we lined out–"If a source
+of variation is controlled for, the amount of unexplained variance
+should be reduced." Wiernik and Dahlke (ref) refer to the work by Hunter
+& Schmidt (ref), likely following their intuition.
+
+In the re-analysis of archival data, we have demonstrated that this
+intution does not hold for the relationship of ES heterogeneity and
+score reliability. Instead, we will propose an alternative way that
+helps us discern, under which conditions we can expect an attenuation
+correction procedure to lead to an increase in ES heterogeneity.
 
@@
 
-However, realising that an ES, as defined in @eq-d, is actually a ratio,
-allows for a different analytical description of how differences in
-score reliability can affect ES heterogeneity. If the phenomenon is
+Realising that an ES, as defined in @eq-d, is actually a ratio, allows
+for a different analytical description of how differences in score
+reliability can affect ES heterogeneity. If the phenomenon is
 heterogeneous and score reliability varies across replications, it seems
 sensible to assume that both numerator and denominator (MD and
-$\sigma_X$) from @eq-d vary across replications. In that case, an ES can
-be described as a random variable stemming from a ratio distribution. A
-ratio distribution is a probability distribution constructed by dividing
-one random variable by a second random variable: $Z = \frac{X}{Y}$
-(reference). If the distribution of the components is known, first order
-taylor approximation may be used to generate estimates of mean and
-variance of the ratio variable (reference). Here we use the
+$\sigma_X$) from @eq-d vary across replications. In that case, the ES
+can be described as a random variable stemming from a ratio
+distribution. A ratio distribution is a probability distribution
+constructed by dividing one random variable by a second random variable:
+$Z = \frac{X}{Y}$ (reference). First order taylor approximation may be
+used to generate estimates of mean and variance of the ratio variable
+\[irrespective of distribution?\] (reference). Here we use the
 Taylor-estimator of the ratio’s variance to demonstrate how differences
@@
 
-$$\tau^2[\delta_0] \approx \frac{\tau^2[MD]}{\mu[\sigma_x]^2} + \frac{\mu[MD]^2}{\mu[\sigma_x]^4} var[\sigma_x]$$ {#eq-ratio_d0}
+$$\tau^2[\delta_0] \approx \frac{\tau^2[MD]}{\mu[\sigma_x]^2} + \frac{\mu[MD]^2}{\mu[\sigma_x]^4} \tau^2[\sigma_x]$$ {#eq-ratio_d0}
 
@@
 between zero and one, in the case of imperfect score reliability,
-$\sigma_T$ will always be smaller than total $\sigma_x$. Additionally,
-if the heterogeneity in $\sigma_X$ was introduced purely by differences
-in measurement precision – the error score variance $\sigma_E^2$, this
-would be removed by the attenuation correction. However, alternatively,
-the underlying latent variable we attempted to measure may not be
+$\sigma_T$ will always be smaller than total $\sigma_x$.
+
+In @eq-rel we see that total standard deviation $\sigma_X$ is
+constructed only from (the square of the sum of) error score variance
+$\sigma_E^2$ and true score variance $\sigma_T^2$. Measurement
+precision, since this should not pertain the true scores, should only
+affect the error score variance. Therefore, if the heterogeneity in
+$\sigma_X$ was introduced purely by differences in measurement
+precision, an attenuation correction procedure can remove that
+heterogeneity. Alternatively, the underlying latent variable may not be
 distributed identically across replications. In that case, true score
-variance $\sigma^2_T$ would vary across replications, leading to
-differences in both score reliability, corrected ES and raw ES.
-Importantly, this implies that even if measurements of perfect score
-reliability were taken across the different replications, heterogeneity
-in ES would still persist.
+variance $\sigma^2_T$ would vary across replications and induce
+heterogeneity in $\sigma_X$. This heterogeneity would lead to
+differences in standardized differences ES as well, but it can not be
+removed by means of an attenuation correction procedure.
 
@@
 $\tau^2\left[\delta_0\right]$, compared to the estimate of
-$\tau^2\left[\delta\right]$ from @eq-ratio_d. However, the equations
-shown throughout this text do not agree with this conclusion.
+$\tau^2\left[\delta\right]$ from @eq-ratio_d. However, as demonstrated
+in the re-analysis of archival data, we can already expect that this can
+not be an implication of these \[Equations @eq-ratio_d0\] and
+[-@eq-ratio_d].
 
@@
 function and the variance of either is placed in the numerator of the
-function, @eq-hs is not sufficient to explain how heterogeneity in ES
+function, @eq-HS is not sufficient to explain how heterogeneity in ES
 changes due to the attenuation correction.
@@
 heterogeneity. Conditions under which @eq-ineq_init does not hold are
-conditions under which the claims made in W&D and H&S are met. However,
-the inequality in this form does not convey sufficient information about
-the conditions under which it can potentially hold. In Appendix A, we
-rearranged @eq-ineq_init, so we end up with a form that we can use to
-actually derive conditions under which ES heterogeneity can be inflated
-or deflated by differences in score reliability.
+conditions under which the claims made in W&D and H&S are met.
+Therefore, this inequality is essentially the opposite of the claims
+made in W&D and H&S. Unfortunately, the inequality in this form does not
+convey sufficient information about the conditions under which it can
+potentially hold. In Appendix A, we rearranged @eq-ineq_init, so we end
+up with a form that we can use to actually derive conditions under which
+ES heterogeneity can be inflated or deflated by differences in score
+reliability.
 
@@
 
-#### Methods
+#### Methods {#sec-methods2}
 
@@
 bootstrapping to derive robust standard errors for these transformed
-observations. Using a random-efffects meta-analysis, we derive estimates
+observations. Using a random-effects meta-analysis, we derive estimates
 of mean and heterogeneity, which need to be backtransformed to the

### 2025-04-09 ReLiability_Function-library.R

@@ -290,7 +290,7 @@ forest_plot_rel <- function(rma_df, aggregates, ci.lvl = .975){
           panel.grid.minor = element_line(colour = "transparent"),
           axis.ticks = element_line(colour = "grey"),
           strip.background = element_rect(fill = "transparent"),
-          strip.text = element_text(size = 12)) +
+          axis.text.y = element_text(size = 12)) +
     scale_y_continuous(breaks = c(-3, 1:nrow(aggregates)), labels = c("RMA-estimate", aggregates$source)) +
     labs(x = "Cohen's d",
          y = "Country") 

### 2025-04-09 test_delta_assumptions.R

@@ -9,7 +9,7 @@ varX_est.L <- lapply(which(effect_index)[nn_eff_idx], FUN = function(x){
   # sometimes calculation of SE might lead to issues, as negative variances can not be log-transformed
   #  therefore, function is run within a tryCatch-environment, so the script does not crash
   # apply_Bootstrap_SE_Project.specific is the project-specific function
-  tryCatch(apply_Bootstrap_SE_Project.specific(na.omit(data.list[[x]]), component = "X", R = 100),
+  tryCatch(apply_Bootstrap_SE_Project.specific(na.omit(data.list[[x]]), component = "X", R = 3000),
            
            # print error to console
            error = function(e)(cat("ERROR: ", conditionMessage(e), " - ",
@@ -41,6 +41,43 @@ varX_rma.list <- lapply(seq_along(varX_est.L), FUN = function(x){
 })
 
 
+# generate estimates of ln-observed score variance for each sample & projects
+varE_est.L <- lapply(which(effect_index)[nn_eff_idx], FUN = function(x){
+  
+  # sometimes calculation of SE might lead to issues, as negative variances can not be log-transformed
+  #  therefore, function is run within a tryCatch-environment, so the script does not crash
+  # apply_Bootstrap_SE_Project.specific is the project-specific function
+  tryCatch(apply_Bootstrap_SE_Project.specific(na.omit(data.list[[x]]), component = "E", R = 3000),
+           
+           # print error to console
+           error = function(e)(cat("ERROR: ", conditionMessage(e), " - ",
+                                   substr(names(data.list), 
+                                          (regexpr("Project) Data/", names(data.list)) + 14), 
+                                          (nchar(names(data.list))-4))[x], 
+                                   " - ", x, "\n")))
+})
+
+
+
+
+# Perform random-effects meta-analysis on estimates of ln-observed score variance using metafor
+varE_rma.list <- lapply(seq_along(varE_est.L), FUN = function(x){
+  
+  # at times estimates of ln-varX & SE may be NA, as negative estimates can't be transformed
+  #  therefore, function is nested in tryCatch, so it doesn't break down with errors
+  tryCatch(metafor::rma(measure = "GEN", method = "REML", 
+                        yi = varX_est.L[[x]]$var.est, 
[…]

### 2025-04-11 Manuscript_MD_ScoreReLiability.qmd
update manuscript, init refs

@@
 title: "Manuscript"
+bibliography: ReLiability_references.bib
+csl: styles/apa-6th-edition.csl
 format: html
@@
 replication will replicate successfully, heterogeneity is a crucial
-parameter that needs to be assessed. In the meta-analytic context,
-heterogeneity describes the variation of effect sizes, free of sampling
-error. Therefore, the presence of heterogeneity can imply that some
-replications of a single phenomenon may be successful, while others are
-not. If heterogeneity increases, meaning the phenomenon’s effect size
-varies more strongly for unexplained reasons, the probability of
-observing an effect size around zero or even in the negative space grows
-larger as well. If we know the mean size and heterogeneity of a
+parameter that needs to be assessed [@stanley2018]. In the meta-analytic
+context, heterogeneity describes the variation of effect sizes, free of
+sampling error [@dersimonian1986; @higgins2002; @riley2011]. Therefore,
+the presence of heterogeneity can imply that some replications of a
+single phenomenon may be successful, while others are not. If
+heterogeneity increases, meaning the phenomenon’s effect size varies
+more strongly for unexplained reasons, the probability of observing an
+effect size around zero or even in the negative space grows larger as
+well [@kenny2019]. If we know the mean size and heterogeneity of a
 phenomenon’s effect, we would theoretically be able to establish an
-expected replication rate (reference).
+expected replication rate [e.g. @vu2024].
 
@@
 reflects a more advanced understanding of the subject matter being
-studied” (2021, p.2). Similarly, Schuetze and von Hippel (2023) argue
-that heterogeneity in effects is an indicator of a vague, poorly
-specified theory.
+studied” [-@LindenHönekopp, p. 2]. Similarly, von Hippel and Schuetze
+[-@vonHippel2023] argue that heterogeneity in effects is an indicator of
+a vague, poorly specified theory.
 
 Large-scale attempts of direct replications, using identical protocols,
-such as the Many Labs studies or Registered Replication Reports
-(references), for the first time allow researchers to estimate
+such as the Many Labs studies or Registered Replication Reports [e.g.
+@ManyLabs1; @RRHart], for the first time allow researchers to estimate
 heterogeneity undistorted by selection processes like publication bias.
-In re-analyses of these studies, Olsson-Collentine et al. (2020)
+In re-analyses of these studies, Olsson-Collentine et al. [-@olsson2020]
 identify a positive correlation between a phenomenon’s effect size and
-its heterogeneity. Similarly, van Erp et al. (2017) and Stanley et al.
-(2018) estimate strong degrees of heterogeneity across conceptual
-replications in Psychology. In a separate re-analysis of large scale
-direct replications, Renkewitz et al. (in preparation) identify
-substantial heterogeneity in almost all projects where a non-zero effect
-could be identified. This aligns with the correlation found by
-Olsson-Collentine et al. (2020).
+its heterogeneity. Similarly, van Erp et al. [-@vanErp2017] and Stanley
+et al. [-@stanley2018] estimate strong degrees of heterogeneity across
+conceptual replications in Psychology. In a separate re-analysis of
+large scale direct replications, Renkewitz et al. (in preparation)
+identify substantial heterogeneity in almost all projects where a
+non-zero effect could be identified. This aligns with the correlation
+found by Olsson-Collentine et al. [-@olsson2020].
 
@@
 to a specific theory, might be impairments in score reliability which
-vary in their extent across replications. This is an argument put
-forward repeatedly by original researchers and replicators alike (refs).
-Hunter and Schmidt (xxx) and Wiernik and Dahlke (xxx), discussing the
-role of score reliability for standardized effect sizes, claim that
-differences in score reliability should inflate, and therefore increase,
-heterogeneity. This implies that standard corrections for imperfect
-score reliability, such as attenuation correction procedures, should
-remove that inflation, reducing heterogeneity in corrected effect sizes.
-Over the following pages, we assess whether score reliability can in
-fact be made responsible for heterogeneity identified across phenomena.
-We find no empirical evidence for the argument that correcting for
-differences in score reliability reduces effect size heterogeneity.
-Instead, to our great surprise, we find that heterogeneity might as well
-be larger after the correction procedure. Therefore, theoretical
-analyses regarding the relationship between score reliability and
-heterogeneity need to be false. In the second half of this text we make
-use of analytical arguments to identify the necessary conditions under
-which an attenuation correction inevitably leads to an increase in
-effect size heterogeneity.
+vary in their extent across replications. \[This is an argument put
+forward repeatedly by original researchers and replicators alike (do we
+actually know of any?).\] Hunter and Schmidt [-@hunter2004] and Wiernik
+and Dahlke [-@wiernik2020], discussing the role of score reliability for
+standardized effect sizes, claim that differences in score reliability
+should inflate, and therefore increase, heterogeneity. This implies that
+standard corrections for imperfect score reliability, such as
+attenuation correction procedures, should remove that inflation,
+reducing heterogeneity in corrected effect sizes. Over the following
+pages, we assess whether score reliability can in fact be made
+responsible for heterogeneity identified across phenomena. We find no
+empirical evidence for the argument that correcting for differences in
+score reliability reduces effect size heterogeneity. Instead, to our
+great surprise, we find that heterogeneity might as well be larger after
+the correction procedure. Therefore, theoretical analyses regarding the
+relationship between score reliability and heterogeneity need to be
+false. In the second half of this text we make use of analytical
+arguments to identify the necessary conditions under which an
+attenuation correction inevitably leads to an increase in effect size
+heterogeneity.
 
@@
 In both the initial reports of large scale replication attempts, as well
-as the re-analyses by Olsson-Collentine et al. (2020), standardized
-effect sizes (from here-on abbreviated as ES) such as Cohen’s d or
-Hedge’s g were used. For the remainder of the article, Cohen’s d,
-defined in @eq-d, will be used as an exemplary estimate of ES, as it is
-used across a wide range of contexts and well understood by a broad
-audience.
+as the re-analyses by Olsson-Collentine et al. [-@olsson2020],
+standardized effect sizes (from here-on abbreviated as ES) such as
+Cohen’s d or Hedge’s g were used. For the remainder of the article,
+Cohen’s d, defined in @eq-d, will be used as an exemplary estimate of
+ES, as it is used across a wide range of contexts and well understood by
+a broad audience.
 
@@
 observed $d$ is from the true $\delta$ in a weighting procedure
-(reference). The standard error of Cohen’s $d$ is computed as defined in
-@eq-SEd.
+[@borenstein2009] The standard error of Cohen’s $d$ is computed as
+defined in @eq-SEd.
 
@@
 
-As is widely known, score reliability affects ES. In the context of
-classical test theory (CTT) score reliability $\rho_{XX’}$ is defined as
-the ratio of true to total score variance, as defined in @eq-rel.
+As is widely known, score reliability affects ES [e.g. @hunter2004;
+@wiernik2020]. In the context of classical test theory (CTT) score
+reliability $\rho_{XX’}$ is defined as the ratio of true to total score
+variance [@lord2008], as defined in @eq-rel.
 
@@
 reliability is an aspect of a measuring instrument applied to a
-population. In the meta-analytic context, as studies tend to be
-replicated in different populations, neither true nor error score
-variance can be expected to always be identical across replications.
-This most likely leads to heterogeneity in score reliability, which, as
-discussed, is bound to affect ES heterogeneity as well.
+population [@lord2008; @vacha1998]. In the meta-analytic context, as
+studies tend to be replicated in a variety of populations, neither true
+nor error score variance can be expected to always be identical across
+replications. This most likely leads to heterogeneity in score
+reliability, which, as discussed, is bound to affect ES heterogeneity as
+well.
 
@@
 heterogeneity and confounding moderator effects, and (c) confounding
-publication-bias and sensitivity analyses*” (p. 3, 2020). Additionally,
-more clearly, they state that “*If the studies included in a
-meta-analysis differ in their measures’ reliabilities, heterogeneity
+publication-bias and sensitivity analyses*” [p. 3, -@wiernik2020].
+Additionally, more clearly, they state that “*If the studies included in
+a meta-analysis differ in their measures’ reliabilities, heterogeneity
 estimates will be artifactually inflated, erroneously suggesting larger
-potential moderator effects*” (p. 4, Wiernik & Dahlke, 2020). They base
-their claims largely on work done by Hunter and Schmidt (2014), who
+potential moderator effects*” [p. 4, -@wiernik2020]. They base their
+claims largely on work done by Hunter and Schmidt [-@hunter2004], who
 claim that “*Variation in reliability across studies causes variation in
@@
 
-As both sources Hunter and Schmidt (2014) and Wiernik and Dahlke (2020)
-will be referred to repeatedly, we abbreviate these references as H&S
-and W&D respectively.
+As both sources Hunter and Schmidt [-@hunter2004] and Wiernik and Dahlke
+[-@wiernik2020] will be referred to repeatedly, we abbreviate these
+references as H&S and W&D respectively.
 
@@
 known as attenuation correction and, while not without its criticisms
-(reference), is a widespread practice in Psychology (reference). @eq-dc
-describes a simple attenuation correction procedure.
+[@muchinsky1996; @schmidt1996; @schennach2016], is a widespread practice
+in Psychology (reference). @eq-dc describes a simple attenuation
+correction procedure.
 
@@
 is corrected by dividing it by the estimate of score reliability
-$\hat{\rho}_{xx’}$. As described in W&D or Lord & Novick (references),
+$\hat{\rho}_{xx’}$. As described in W&D or Lord & Novick [-@lord2008],
 this is essentially the same, as if the corrected estimate of $d_c$ was
@@
 uncovered differences in score reliability across replications of the
-same phenomena (ref McShaw etc.). The claims made in W&D and H&S imply
-that the ES heterogeneity uncovered in these projects therefore could
+same phenomena [@shaw2020]. The claims made in W&D and H&S imply that
+the ES heterogeneity uncovered in these projects therefore could
 potentially be explained by these differences in score reliability.
@@
 authors as well as the multi-site replication authors claim that
-differences in score reliability might be responsible (references).
+differences in score reliability might be responsible (again, do we
+actually know of any?).
 
@@
 We have collected a large number of openly available multi-site direct
-replications on psychological phenomena (Fünderich et al. 2024). The
-data-sets of about 50 phenomena, largely stemming from the efforts of
-the Many Labs studies, Registered Replication Reports or the
-Psychological Science Accelerator, are made openly available in a
-standardised format at \[osf.io\], the DRIPHT repository.
+replications on psychological phenomena [@dripht2024]. The data-sets of
+about 50 phenomena, largely stemming from the efforts of the Many Labs
+studies, Registered Replication Reports or the Psychological Science
+Accelerator, are made openly available in a standardised format at
+\[<https://osf.io/7p8rk/>\], the DRIPHT repository.
 
@@
 Concerning the re-arranged inequality from Appendix A, we identify one
-crucial condition or rule described in @eq-ineq_short.
+necessary but not sufficient condition or rule described in
+@eq-ineq_short.
 
 $$R_1^3 \leq R_2$$ {#eq-ineq_short}
+
+With necessary but not sufficient we mean that, as long as
+@eq-ineq_short holds, an attenuation correction will inevitably increase
+ES heterogeneity. However, if the inequality does not hold, it is not
+guaranteed that the attenuation correction can decrease ES
+heterogeneity. Whether ES heterogeneity can be decreased depends on
+other parameters, that we left out of this short form, but can be found,
+accompanied by a more detailed description, in @eq-ineq_fin.
 
@@
 variance heterogeneity after the attenuation correction, @eq-ineq_init
-is not violated. For example, assuming the average score reliability is
-about .8, as long as $R_2$ is at least .512 or larger, the inequality
-holds true and ES heterogeneity grows larger as a consequence of the
-attenuation correction procedure. In such cases, the claims made in W&D
-and H&S could not hold.
+is not violated. For example, assuming average score reliability is
+about .8, the heterogeneity in $\sigma^2_T$ needs to make up less than
+51.2% ($R_2 \geq .8^3 = .512$) of the heterogeneity in total score
+variance $\sigma^2_X$. At $R_1 = .8$, as long as $R_2 \geq .512$, the
+claims made in W&D and H&S can not hold.
 
@@
 relative size of score variance heterogeneity after the attenuation
-correction. Therefore, @eq-ineq_init states that for such a correction
-procedure to truly reduce ES heterogeneity, the reduction in the
-heterogeneity of score variance needs to be substantially larger than
-the reduction in its mean values. The mean value reduction carries
-stronger weight, as $R_1$ is taken to the power of 3 in @eq-ineq_init.
-The smaller $R_1$ or $R_2$, the larger the reduction in heterogeneity or
-mean value.
+correction. Therefore, @eq-ineq_short states the following regarding the
+change in mean value and heterogeneity due to the correction: ES
+heterogeneity will always be reduced, as long as the remaining
+heterogeneity in score variance is not substantially smaller than the
+size of the remaining mean value.
 
@@
 across administrations, it is unlikely that an attenuation correction
-can actually produce such a low $R_2$. Generally speaking, $R_1$, and
-therefore the reduction in mean score variance, largely depends on the
-relative measurement precision. If measurements were made more
-precisely, average score reliability and $R_1$ would be higher. On the
-other hand $R_2$, and therefore the reduction in heterogeneity of score
-variance, depends on how different the measurement precision was at the
-individual administrations, but also how different the individual
-administrations sites are regarding their true score variance. If the
-administration sites carry large differences in true score variance
-$R_2$ is unlikely to be substantially smaller than $R_1$.
+can actually produce such a low $R_2$. Generally speaking, $R_1$ largely
+depends on the relative measurement precision. If measurements were made
+more precisely, average score reliability and $R_1$ would be higher. On
+the other hand, $R_2$ depends on how different the measurement precision
+was at the individual administrations, but also how different the
+individual administrations sites are regarding their true score
+variance. If the administration sites carry large differences in true
+score variance $R_2$ is unlikely to be substantially smaller than $R_1$.
 
@@
 describes under which conditions the inequality is guaranteed to hold,
-no matter what value other parameters like the mean value or
-heterogeneity in mean differences take on. This means that, even if the
-condition in @eq-ineq_short is violated, it is by no means guaranteed
-that an attenuation correction procedure would actually reduce
-heterogeneity in ES. This means that inference made from @eq-ineq_short
-is asymmetrical. While we know, if the inequality holds, heterogeneity
-in corrected ES will be larger than heterogeneity in uncorrected ES, we
-can't be certain that the opposite is true if the inequality does not
-hold.
-
-***\[maybe move to discussion further below\]*** The majority of studies
-in the social sciences make use of convenience sampling. Often,
-respondents are collected from a local student pool, where students
-themselves would choose which studies to partake in. In such cases, we
-believe there is little reason for us to expect homogeneity in true
-score variance across administration sites. As participants across
-administrations self-select into the study, no real sampling process
-takes place - implying that we have no reason to assume that the true
-score variance is stable across these administration sites. However,
-since researchers are well aware of the importance of measurement
-precision, which is also held stable more easily, we believe there is
-little reason for $R_1$ to be extremely low.
+no matter what value other parameters like the mean value of $MD$ or its
+heterogeneity take on. This means that, even if the condition in
+@eq-ineq_short is violated, it is by no means guaranteed that an
+attenuation correction procedure would actually reduce heterogeneity in
+ES. This means that inference made from @eq-ineq_short is asymmetrical.
+While we know, if the inequality holds, heterogeneity in corrected ES
+will be larger than heterogeneity in uncorrected ES, we can't be certain
+that the opposite is true if the inequality does not hold.
 
@@
 
-@tbl-tauES showed that for all phenomena, where we identified
+@tbl-tauES showed that for all phenomena, for which we identified
 statistically significant ES heterogeneity, the heterogeneity was even
-larger after the attenuation correction procedure. Subsequently, in
-Appendix A we demonstrate that if the $R_2$ metric is larger than
-$R_1^3$, ES heterogeneity is bound to be larger after applying said
-procedure. Here, we will re-analyse the data, estimating $R_1$ and $R_2$
-for each phenomenon respectively. This allows us to assess in how far
-the estimates of ES heterogeneity align with our claims regarding the
-$R_1$ and $R_2$ metrics.
+larger after the attenuation correction procedure. Here, we will
+re-analyse the data, estimating $R_1$ and $R_2$ for each phenomenon
+respectively. Therein, we not only show that our claims derived from
+analytical arguments hold up in empirical assessment, but also
+demonstrate how the metrics $R_1$ and $R_2$ may be used to inform how
+score reliability affected ES heterogeneity.
 
@@
 
-Estimation of $R_1$ and $R_2$ require estimates of meta-analytic mean
+Estimation of $R_1$ and $R_2$ requires estimates of meta-analytic mean
 and heterogeneity of both the total score variance $\sigma^2_X$ and the
 true score variance $\sigma^2_T$. Elsewhere, we proposed the method
-boot-err (ref.) as an alternative method to test for and estimate the
-differences in error score variance $\sigma^2_E$. In Beinhauer et al.
-(ref.), we deliver detailed instructions on how estimates for mean and
-heterogeneity of total score variance, and the score variance components
-can be derived. For the total score variance $\sigma^2_X$, all we need
-to do is to appropriately transform the observed estimates using the
-variance-stabilizing properties of the natural log. Subsequently, we use
-bootstrapping to derive robust standard errors for these transformed
-observations. Using a random-effects meta-analysis, we derive estimates
-of mean and heterogeneity, which need to be backtransformed to the
-original scale. Appropriate equation to back-transform estimates can be
-found in Appendix B.
-
-For the true score variance $\sigma^2_T$, we found that directly
-estimating its heterogeneity, even if an appropriate
-variance-stabilizing transformation is used, is severely biased.
-Instead, we can estimate the error score variance $\sigma^2_E$ using an
-estimate of score reliability. Performing the boot-err, which, again,
-implies transformations using the natural log, deriving standard errors
-using Bootstrapping, running a random-effects meta-analysis and
-back-transforming the estimates, nets us estimates of mean and
-heterogeneity of $\sigma^2_E$. Since we already have the same estimates
-available for the total score variance, we can estimate meta-analytic
-mean and heterogeneity of true score variance.
-
-$$\hat{\mu}_{\sigma^2_T} = \hat{\mu}_{\sigma^2_X} - \hat{\mu}_{\sigma^2_E} \ \ \ \ \ \ and \ \ \ \ \  \hat{\tau}^2_{\sigma^2_T} = \hat{\tau}^2_{\sigma^2_X} - \hat{\tau}^2_{\sigma^2_E}$$ {#eq-meantauT}
-
-@eq-meantauT provides estimates of meta-analytic mean of true score
-variance and heterogeneity. These estimates are largely unbiased, unlike
-estimates stemming from using the boot-err to directly run meta-analytic
-models on true score variance estimates. Interested readers are referred
-to Beinhauer et al. (ref.)
+Boot-Err (ref.) as an alternative method to test for and estimate the
+differences in error score variance $\sigma^2_E$. Among other things, in
+Beinhauer et al. (ref.), we deliver detailed instructions on how
+estimates for mean and heterogeneity of total score variance, and the
+score variance components can be derived. This means that Boot-Err can
+be used to generate all estimates required for computation of the
+metrics $R_1$ and $R_2$.
+
+For the following analysis, Boot-Err is used. As Boot-Err makes use of
+meta-analytic techniques and bootstrap sampling, we perform analyses
+using the packages *metafor*, version 4.8-0, and *boot*, version 1.3-31,
+in the statistical programming language *R*, version 4.4.2.
 
@@
 
-For the two exceptional phenomena where the attenuation correction
-procedure actually led to an increase in ES heterogeneity
+For the two exceptions, those phenomena where the attenuation correction
+procedure actually led to an decrease in ES heterogeneity
 (Srull_Behavior_Hostility and Shnabel_Willingness_Reconcile_RPP), the
@@
 reduction in heterogeneity is better explained by imprecise estimation
-than an actual explanation of unexplained heterogeneity.
+than an actual reduction of heterogeneity.
 
@@
 
-Generally, @tbl-tauES implies that for almost all measurements
-concerning these phenomena, the heterogeneity in observed score
-variances could not be substantially explained by differences in
-measurement quality. Only for one phenomenon, where differences in score
-variance could be found, $R_2$ was smaller than $R_1^3$. As discussed,
-for this phenomenon, the number of administration sites was too low to
-accurately estimate heterogeneity in score variance components. For all
-remaining phenomena, $R_2$ was larger than $R_1$, implying that an
-attenuation correction procedure could not reduce ES heterogeneity. The
-results in @tbl-tauES agree with this conclusion, as no phenomena could
-be identified where statistically significant ES heterogeneity was
-accompanied by a reduction in differences due to the correction.
+Generally, @tbl-R1R2 implies that for all measurements concerning these
+phenomena, the heterogeneity in observed score variances could not be
+substantially explained by differences in measurement quality. For all
+phenomena, $R_2$ was larger than $R_1$, implying that an attenuation
+correction procedure could not reduce ES heterogeneity in any of these
+cases. The results in @tbl-tauES agree with this conclusion, as no
+phenomena could be identified where statistically significant ES
+heterogeneity was accompanied by a reduction in differences due to the
+correction.
 
@@
 We have examined meta-analytic heterogeneity in standardized mean
-difference. Since score reliability is often discussed as a source for
+differences. Since score reliability is often discussed as a source for
 such heterogeneity, we estimated the heterogeneity before and after
@@
 differences in ES), the attenuation correction led to increased
-estimates of ES heterogeneity. This implies that the claims found in W&D
-and H&S stem from a general misunderstanding of how score reliability
-can affect ESs.
+estimates of ES heterogeneity. This means that the claims found in W&D
+and H&S, implying that an attenuation correction can reduce ES
+heterogeneity, stem from a general misunderstanding of how score
+reliability can affect ESs.
 
 Using analytical arguments, we derived a broad rule, outlining under
-which conditions we can expect an attenuation correction to reduce ES
+which conditions we can expect an attenuation correction to increase ES
 heterogeneity. In general, score reliability affects ESs by inflating
@@
 Since $R_1$ is carrying the power of 3, it implies that the reduction in
-mean score variance is weighted more heavily than the reduction in score
-variance heterogeneity. This is best described through an example: We
-know what an acceptable average score reliability looks like in the area
-of Psychology. Researchers largely target a score reliability of at
-least .7, preferably .8. An average score reliability of .7 implies that
-70% of the variance observed in scores can actually be attributed to
-true underlying differences. From our derived rule, we know that, as
-long as $R_2$ is larger or equal to $.7^3 = .343$, the attenuation
-correction procedure can only increase ES heterogeneity. If we want to
-explain ES heterogeneity through score reliability, even though only 30%
-of the variance observed in score can be attributed to random
-measurement error, at least 65.7%, almost two thirds, of the
+mean score variance is affecting the ES heterogeneity more strongly than
+the reduction in score variance heterogeneity. This is best described
+through an example: We know what an acceptable average score reliability
+looks like in the area of Psychology. Researchers largely target a score
+reliability of at least .7, preferably .8. An average score reliability
+of .7 implies that 70% of the variance observed in scores can actually
+be attributed to true underlying differences. From our derived rule, we
+know that, as long as $R_2$ is larger or equal to $.7^3 = .343$, the
+attenuation correction procedure can only increase ES heterogeneity. If
+we want to explain ES heterogeneity through score reliability, even
+though only 30% of the variance observed in score can be attributed to
+random measurement error, at least 65.7%, almost two thirds, of the
 heterogeneity in score variance would need to be attributable to
@@
 administration sites.
-
-This also means that ES heterogeneity is best reduced, if average score
-reliability ($R_1$) is substantially large. In such a scenario, the
-$R_2$ is more likely to be low enough for the inequality to no longer
-hold. The larger the average score reliability, the more likely it is
-that the attenuation correction procedure can actually reduce ES
-heterogeneity. However, as the correction would only be rather small in
-such cases, it remains to be seen whether the reduction in ES
-heterogeneity would be sufficient to claim that differences in score
-reliability actually posed a relevant moderator in such a case.
 
@@
 
-Taken together, these facets lead to what we observed throughout this
-text: Reliability coefficients are in an acceptable range, leading to a
-small, but constant reduction in mean score variance if we apply an
+The combination of these aspects can explain what we observed throughout
+this text: Reliability coefficients are in an acceptable range, leading
+to a small, but constant reduction in mean score variance if we apply an
 attenuation correction. At the same time, there is substantial variation
@@
 heterogeneity estimates in score variance. Therefore, as indicated by
-our rule, an attenuation correction in these setting can rarely deliver
-a reduction in ES heterogeneity. This also means that, in such settings,
-score reliability does not explain any ES heterogeneity. If anything, it
-masks it to a small degree.
-
-If we, in the space of Psychology, keep making use of convenience
-samples simply because they are so convenient, we can not expect
-anything to change. Most likely, ES heterogeneity will remain a
-substantial factor hampering the interpretability of original and
-replication studies. We could learn more about standardized effect sizes
-and why our phenomena vary across settings, populations or continents by
-not following the current status quo of convenience samples. Until this
-status quo changes, however, we will keep wasting valuable resources by
-attempts of discussing ES heterogeneity or assessing uninformative
-differences in score reliability.
-
-These results fit in with work recently published/preprinted by
-Olsson-Collentine et al. (2023). In a large simulation scheme, they
-demonstrate that differences in score reliability across administrations
-typically deflate heterogeneity in uncorrected correlations. Only as the
-true heterogeneity in correlations grows larger, score reliability
-differences actually inflate heterogeneity as claimed in W&D. While
-correlations and standardized effect sizes are not identical, the way
-score (un)reliability affects these parameters is highly similar.
+our inequality, an attenuation correction is likely to have very little
+impact on ES heterogeneity. If anything, it most likely increases it.
+This also means that, in such settings, score reliability does not serve
+as a moderator explaining ES heterogeneity.
+
+If we, in the field of Psychology and other social or behavioural
+sciences, keep making use of convenience samples simply because they are
+so convenient, we can not expect anything to change. Most likely, ES
+heterogeneity will remain a substantial factor hampering the
+interpretability of original and replication studies. Instead, we should
+focus our data collection efforts on the generation of samples, where we
+know that the variance in the sample means something. This is most
+easily achieved by proper sampling techniques, which ensure that the
+sample is representative for some population. In such cases, it might be
+that the true score variance varies less strongly than what we currently
+observe. And should the heterogeneity in true score variance remain as
+large, than we know that this is meaningful information from which we
+can draw meaningful inferences. Therein, we could learn more about
+standardized effect sizes and why our phenomena vary across settings,
+populations or continents by not following the current status quo of
+convenience samples. Until this status quo changes, however, we will
+keep wasting valuable resources by attempts of discussing ES
+heterogeneity or assessing uninformative differences in score
+reliability.
+
+Our results fit in well with recent work by Olsson-Collentine et al.
+(2023). In a large simulation scheme, they demonstrate that differences
+in score reliability across administrations typically deflate
+heterogeneity in uncorrected correlations. In such cases, an attenuation
+correction would increase correlation coefficient heterogeneity, as we
+observe for standardized mean differences. Only as the true
+heterogeneity in correlations grows larger would the correction
+procedure reduce heterogeneity as implied in W&D. While correlations and
+standardized effect sizes are not identical, the way score
+(un)reliability affects these parameters is highly similar.
 
@@
 non-representative data-sets. Based on the combination of analytical and
-empirical arguments, we are convinced that these differences in score
+empirical arguments, we are convinced that differences in score
 reliability across replications in the current state of Psychology and
@@
 
-The analytical arguments leading to @eq-ineq_short rested on the use of
-the delta-method, as @eq-ineq_init contained parameters of the score
-standard deviations’ distributions, while the $R_1$ and $R_2$ metrics
-were designed to discuss score variances. The delta-method rests on a
-number of assumptions, at least one of which is likely violated in our
-analysis: the variance of the variable to be transformed needs to be
-small (ref.). As this entire procedure discusses the influence of
-variance in standard deviations on ES heterogeneity, introduced by
-differences in score reliability, assuming that this assumption holds
-seems somewhat off. Therefore, we used the empirical examples available
-to assess in how far estimates derived using the delta-method
-approximations diverge from estimates derived without it. The
-differences identified are rather miniscule for the majority of
-phenomena, at the third decimal. Only for the PSACR002_behav_init
-phenomenon we find larger differences - therefore, conclusions for this
-phenomenon concerning the relation between $R_1$ and $R_2$ need to be
-drawn with caution. Generally, as the majority of phenomena did not
-display any bias in this assessment and since the delta-method is not
-actually used to generate estimates, we believe its use to be defensible
-nonetheless.
+The analytical arguments, detailed in Appendix X, leading to
+@eq-ineq_short, rest on the use of the delta-method. @eq-ineq_init
+contains parameters of the score standard deviations’ distributions,
+while the $R_1$ and $R_2$ metrics were designed to discuss score
+variances. The delta-method rests on a number of assumptions, at least
+one of which is likely violated in our analysis: the variance of the
+variable to be transformed needs to be small (ref.). This entire
+procedure discusses the influence of variance in standard deviations on
+ES heterogeneity, introduced by differences in score reliability.
+Therefore, assuming that this assumption holds seems naive. In the
+Appendix X, we explore the archival data, to understand in how far
+employing the delta-method on meta-analyses of standard deviations lead
+to biased results than the meta-analyses on score variances we performed
+using the Boot-Err technique. The differences identified are rather
+miniscule for the majority of phenomena, at the third decimal. Only for
+the PSACR002_behav_init phenomenon we find larger differences -
+therefore, conclusions for this phenomenon concerning the relation
+between $R_1$ and $R_2$ need to be drawn with caution. Generally, as the
+majority of phenomena did not display any bias in this assessment and
+since the delta-method is not actually used to generate estimates, we
+believe its use to be defensible nonetheless.
 
@@
 factors are held as constant as possible, is already substantial
-(Renkewitz, Fünderich, & Beinhauer, 2024). However, so far we have
-demonstrated that in several scenarios, differences in score reliability
-did not serve as appropriate moderators or explanations of ES
-heterogeneity. Across most phenomena, we failed to reduce ES
+(Renkewitz, Fünderich, & Beinhauer, in preparation). However, so far we
+have demonstrated that in several scenarios, differences in score
+reliability did not serve as appropriate moderators or explanations of
+ES heterogeneity. Across most phenomena, we failed to reduce ES
 heterogeneity. If anything, we even increased ES heterogeneity, albeit
@@
 in score reliability did not help us gain a better understanding of the
-phenomena. We want to reinforce that, in light of uncontrolled variation
-in score variance due to the lack of appropriate sampling schemes,
+phenomena.
 
@@
 
+## References
+
+::: {#refs}
+:::
+
 ## Appendix
@@
 sufficient to identify the relevant circumstances required for the
-inequality to hold. However, rearranging the terms from @eq-ineq_R leads
-to @eq-ineq_fin:
-
-$$ \tau_{MD}^2\left(\frac{1}{R_1} - 1\right) + \mu_{MD}^2 \frac{1}{4} CV^2_{\sigma^2_X} \left(\frac{R_2}{R_1^3}  - 1\right) > 0$$ {#eq-ineq_fin}
+inequality to hold. In a series of steps, we will rearrange @eq-ineq_R
+to facilitate interpretation.
+
+Firstly, for @eq-ineq_rearr1 we pulled all terms onto the left side of
+the inequality.
+
+$$\frac{\tau_{MD}^2}{R_1 \mu_{\sigma^2_x}} - \frac{\tau_{MD}^2}{\mu_{\sigma^2_x}} + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} \frac{R_2 \tau^2_{\sigma_x^2}}{4 R_1 \mu_{\sigma^2_x}} - \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4\mu_{\sigma^2_x}} >  0$$ {#eq-ineq_rearr1}
+
+Secondly, as all terms contain $\mu_{\sigma^2_x}$ in the denominator, to
+arrive at @eq-ineq_rearr2 we divided by $\mu_{\sigma^2_x}$.
+
+$$\frac{\tau_{MD}^2}{R_1} - \tau_{MD}^2 + \frac{\mu_{MD}^2}{R_1^2 \mu^2_{\sigma^2_x}} \frac{R_2 \tau^2_{\sigma_x^2}}{4 R_1} - \frac{\mu_{MD}^2}{\mu^2_{\sigma_x^2}} \frac{\tau^2_{\sigma_x^2}}{4} >  0$$ {#eq-ineq_rearr2}
+
+For @eq-ineq_rearr3 we properly combined the multiplied fractions.
+
+$$\frac{\tau_{MD}^2}{R_1} - \tau_{MD}^2 + \frac{\mu_{MD}^2 R_2 \tau^2_{\sigma_x^2}}{4R_1^3 \mu^2_{\sigma^2_x}} - \frac{\mu_{MD}^2\tau^2_{\sigma_x^2}}{4\mu^2_{\sigma_x^2}}  >  0$$ {#eq-ineq_rearr3}
+
+Taking $\frac{\tau_{\sigma^2_x}}{\tau_{\sigma^2_x}} = CV{\sigma^2_x}$
+allows to simplify for @eq-ineq_rearr4.
+
+$$\frac{\tau_{MD}^2}{R_1} - \tau_{MD}^2 + \frac{\mu_{MD}^2 R_2}{4R_1^3}CV{\sigma^2_x} - \frac{\mu_{MD}^2}{4}CV{\sigma^2_x}  >  0$$ {#eq-ineq_rearr4}
+
+As a last step, factoring out finalizes rearranging the terms from
+@eq-ineq_R to @eq-ineq_fin:
+
+$$ \tau_{MD}^2\left(\frac{1}{R_1} - 1\right) + \frac{\mu_{MD}^2 CV^2_{\sigma^2_X}}{4} \left(\frac{R_2}{R_1^3}  - 1\right) > 0$$ {#eq-ineq_fin}
 
@@
 equation should hold: (a) one of the terms inside the brackets
-($\frac{1}{R_1} - 1$ and $\frac{R^2_2}{R_1}  - 1$) is positive and large
+($\frac{1}{R_1} - 1$ or $\frac{R_2}{R^3_1}  - 1$) is positive and large
 enough, so that the second term not containing that same bracket is
-positively dominated by the first term; or (b) both terms inside the
-brackets need to be positive.
+positively dominated by the first term, meaning it is sufficiently large
+that the positive term cancels out the negative term; or (b) both terms
+inside the brackets need to be positive.
 
@@
 as the ratio of true and error score variance heterogeneity, is bound
-between $0\le R_2\le1$. If $R_1^3$ is smaller than $R_2$, then the term
+between $0\le R_2\le1$. If $R_1^3$ is larger than $R_2$, then the term
 inside the second bracket turns negative. Generally, this means that the
-inequality can only be violated if the reduction in relative
-heterogeneity of score variance ($R_2$) is larger than the the average
-score reliability to the power of 3. Restated in mathematical terms,
-this leads us to the rule/condition described in @eq-ineq_short, in the
-main text.
+inequality can only be violated if the relative remaining heterogeneity
+of score variance ($R_2$) is smaller than the the average score
+reliability to the power of 3. Restated in mathematical terms, this
+leads us to the rule/condition described in @eq-ineq_short, in the main
+text.
+
+However, it is crucial to note that violation of $R_1^3 \leq R_2$ is
+just a necessary, but not sufficient condition for the inequality not to
+hold anymore. Should $R_1^3 \leq R_2$ not hold, the negative term in
+@eq-ineq_fin still needs to be sufficiently large, that it can
+negatively dominate the positive term. That would mean that the
+combination of parameters $\mu_{MD}$, $\tau_{MD}$ and $CV_{\sigma^2_x}$
+need to take on values, that allow for such a scenario. Since we do not
+want to make any claims regarding the unstandardized mean difference and
+how it relates to the coefficient of variation of score variance, we can
+not distill a scenario in which it is guaranteed, that @eq-ineq_fin is
+violated. Therefore $R_1^3 \leq R_2$ only describes a condition, under
+which it is guaranteed that @eq-ineq_fin holds. If $R_1^3 \leq R_2$ is
+violated, it is a first necessary condition for the entire inequality to
+be violated, but it is not sufficient as it still depends on other
+parameters.
 
@@
 
-To examine whether applying the delta-method to derive the conditions
-under which the inequality @eq-ineq_init holds, we compared estimates of
-mean and heterogeneity of total score standard deviation$\sigma_X$. To
-generate estimates that do not depend on use of the delta-method, we
-used bootstrapping and a log-transformation, similar as detailed in
-@sec-methods2 for the score variances $\sigma^2_X$. Using the natural
-log, we transformed estimates of score standard deviation and using
-bootstrapped samples we derived appropriate standard errors. From a
-random-effects meta-analysis we generated estimates of mean and
-heterogeneity and back-transformed those to the original scale of
-$\sigma_X$. Additionally, we used the delta-method as outlined in
-@eq-delta_method to estimate mean and heterogeneity in $\sigma_X$ from
-estimates derived from meta-analyses on $\sigma^2_X$. The estimates can
-be found in @tbl-delta.
+We want to examine whether applying the delta-method sufficiently
+approximates the moments of the distribution of score variance to derive
+the conditions under which the inequality @eq-ineq_init holds. To do so,
+we compared estimates of mean and heterogeneity of total score variance
+$\sigma^2_X$ derived using Boot-Err, and by running a meta-analysis on
+total score standard deviation $\sigma_x$ and subsequently applying the
+delta-method to estimate mean and heterogeneity of $\sigma^2_X$ .
+Therein, we generated two sets of estimates of $\mu_{\sigma^2_x}$ and
+$\tau^2_{\sigma^2_x}$, one via Boot-Err and one via the delta-method. If
+there are systematic differences between the sets, this points towards
+bias in the delta-method approximation. The estimates can be found in
+@tbl-delta.
 
@@
              col.names = c("MASC", 
-                           "$\\hat{\\mu}_{\\sigma_X}$", 
-                           "$\\hat{\\tau}^2_{\\sigma_X}$", 
-                           "$\\hat{\\mu}_{\\Delta \\sigma_X}$", 
-                           "$\\hat{\\tau}^2_{\\Delta \\sigma_X}$"))
+                           "$\\hat{\\mu}_{\\sigma^2_X}$", 
+                           "$\\hat{\\tau}^2_{\\sigma^2_X}$", 
+                           "$\\hat{\\mu}_{\\Delta \\sigma^2_X}$", 
+                           "$\\hat{\\tau}^2_{\\Delta \\sigma^2_X}$"))
 ```
 
-Generally, for 11 out of 12 phenomena, we find miniscule differences
-between the two estimation strategies. For all but the
-PSACR002_behav_init phenomenon, the differences occur in the 3rd
-decimal, indicating that violations in the assumptions concerning the
-delta-method are negligible here. Only for the PSACR002_behav_init
-phenomenon we find larger differences. These are large enough to be
-concerning and pertain largely the estimate of heterogeneity.
-Accordingly, inferences regarding this phenomenon need to be interpreted
-with caution. Generally, since for the majority of phenomena we find no
-bias and since the delta-method is not used to generate estimates beyond
-this particular section, we believe its use to be defensible.
+Generally, for 11 out of 12 phenomena, we find miniscule, but mostly
+negative, differences between the two estimation strategies. This means
+that there is a slight downward bias in the estimates from the
+delta-method. However, for the majority of phenomena, the differences
+occur in the 2nd decimal, indicating that violations in the assumptions
+concerning the delta-method are negligible here. Only for the
+PSACR002_behav_init phenomenon we find larger differences. These are
+large enough to be concerning and pertain largely the estimate of
+heterogeneity. Accordingly, inferences regarding this phenomenon need to
+be interpreted with caution. Generally, since for the majority of
+phenomena we find very little bias and as the delta-method is not used
+to generate estimates beyond this particular section, we believe its use
+to be defensible.

### 2025-04-11 Manuscript_MD_ScoreReLiability.qmd
update manuscript - add bibliography

@@
 such as the Many Labs studies or Registered Replication Reports [e.g.
-@ManyLabs1; @RRHart], for the first time allow researchers to estimate
+@ManyLabs1; @RRRHart], for the first time allow researchers to estimate
 heterogeneity undistorted by selection processes like publication bias.
@@
 identical protocols. From five published projects, the data for the
-first three and the fifth Many Labs projects was harmonized and publicly
-stored in a standardized format when the DRIPHT repository was set up
-(references). The Registered Replication Reports are similar
-collaborative efforts, with the sole distinction that for each report a
-single phenomenon was replicated several times. The Registered Reports
-3-10 were added to the DRIPHT repository, as they employed experimental
-designs with at least two groups. Lastly, some projects from the
-Psychological Science Accelerator were added to the DRIPHT repository.
-In contrast to Many Labs or Registered Replication Projects, these
-collaborative efforts, while also distributed across the globe, do not
-formally attempt to perform direct replications, but focus on original
-research or conceptual replications. However, as each project used the
-same protocol and participants are distributed across different
-countries, a data structure similar to that of a multi-site replication
-study emerges. What makes all these collaborative efforts valuable for
-this manuscript is the fact that all phenomena collected in the
-DRIPHT-repository are making use of two-group designs and are therefore
-easily assessed using standardised mean differences.
+first three and the fifth Many Labs projects [@ManyLabs1; @ManyLabs2;
+@ManyLabs3; @ManyLabs5] was harmonized and publicly stored in a
+standardized format when the DRIPHT repository was set up [@dripht2024].
+The Registered Replication Reports are similar collaborative efforts,
+with the sole distinction that for each report a single phenomenon was
+replicated several times. The Registered Reports 3-10 were added to the
+DRIPHT repository, as they employed experimental designs with at least
+two groups [@RRRHart; @RRRSripada; @RRRFinkel; @RRRStrack; @RRRRand;
+@RRRDijksterhuis; @RRRSrull; @RRRMazar]. Lastly, some projects from the
+Psychological Science Accelerator were added to the DRIPHT repository,
+namely PSA006, PSACR001, PSACR002 and PSACR003 [@PSA006; @PSACR001;
+@PSACR002; @PSACR003]. In contrast to Many Labs or Registered
+Replication Projects, these collaborative efforts, while also
+distributed across the globe, do not formally attempt to perform direct
+replications, but focus on original research or conceptual replications.
+However, as each project used the same protocol and participants are
+distributed across different countries, a data structure similar to that
+of a multi-site replication study emerges. What makes all these
+collaborative efforts valuable for this manuscript is the fact that all
+phenomena collected in the DRIPHT-repository are making use of two-group
+designs and are therefore easily assessed using standardised mean
+differences.
 
@@
 (@eq-SEd). Subsequently, a random-effects meta-analysis is performed
-using metafor version X.XX. To estimate meta-analytic mean ES and
-heterogeneity we make ues of the REML-estimator (ref). To identify which
-phenomena pass criterion (3), using a Wald-type significance test we
-assess for which phenomenon the meta-analytic mean ES is significantly
-different from zero. For this hypothesis test, as all other hypothesis
-tests in this manuscript, we employ a one-tailed significance level of
-.05.
+using metafor version 4.8-0 [@metafor]. To estimate meta-analytic mean
+ES and heterogeneity we make use of the REML-estimator, as that
+estimator tends to perform best across a range of conditions
+[@hoenekopp2022]. To identify which phenomena pass criterion (3), using
+a Wald-type significance test we assess for which phenomenon the
+meta-analytic mean ES is significantly different from zero. For this
+hypothesis test, as all other hypothesis tests in this manuscript, we
+employ a one-tailed significance level of .05.
 
@@
 other estimates of score reliability, such as McDonald’s Omega
-(reference), Guttman’s Lambda 2 or 4 (reference), or the Greatest Lower
-Bound (reference) might be better suited. However, all scales employed
-in this project are scored by computing the simple mean or sum of
-responses across items for each individual. Computing the mean or sum of
-several items implicitly assumes tau-equivalence, as each item
+[@mcdonald1999], Guttman’s Lambda 2 or 4 [@guttman1945], or the Greatest
+Lower Bound [@tenBerge2004] might be better suited. However, all scales
+employed in this project are scored by computing the simple mean or sum
+of responses across items for each individual. Computing the mean or sum
+of several items implicitly assumes tau-equivalence, as each item
 contributes equally to the individual’s test score. Therefore,
@@
 indicators of heterogeneity are readily available, such as $I^2$, $H^2$
-or the coefficient of variation. However, the descriptions found in H&S
-and W&D discuss the absolute amount of heterogeneity in terms of
-variance ($\tau^2$) or standard deviation ($\tau$). These parameters
-discuss the variability of the standardized ES in the population in
-terms of variance or standard deviation. Therefore, we also make use of
-absolute heterogeneity here.
+or the coefficient of variation $CV$[@borenstein2009; @higgins2002].
+However, the descriptions found in H&S and W&D discuss the absolute
+amount of heterogeneity in terms of variance ($\tau^2$) or standard
+deviation ($\tau$). These parameters discuss the variability of the
+standardized ES in the population in terms of variance or standard
+deviation. Therefore, we also make use of absolute heterogeneity here.
 
@@
 heterogeneity in uncorrected ES and corrected ES are compared. Using
-Cochran’s *Q*-test, we assess whether the estimates of heterogeneity are
-statistically significantly different from 0. Only for those projects
-where we have sufficient confidence that the individual estimates of
-heterogeneity are larger than zero can we actually begin to interpret
-whether there was any change.
+Cochran’s *Q*-test [@cochran1954], we assess whether the estimates of
+heterogeneity are statistically significantly different from 0. Only for
+those projects where we have sufficient confidence that the individual
+estimates of heterogeneity are larger than zero can we actually begin to
+interpret whether there was any change.
 
@@
 reliability are assessed meta-analytically as well. To do so, a
-Reliability Generalization Meta-Analysis is performed (references). This
-entails that individual estimates of score reliability are adequately
-transformed using Bonett’s transformation:
+Reliability Generalization Meta-Analysis is performed [@vacha1998;
+@sanchez2013]. This entails that individual estimates of score
+reliability are adequately transformed using Bonett’s transformation:
 $T_{\hat{\rho}} = ln(1 - \hat{\rho})$. This transformation has
 variance-stabilising properties and therefore allows for adequate
-inferences concerning heterogeneity in score reliability (references).
+inferences concerning heterogeneity in score reliability [@bonett2002].
 Cochran’s Q-test is used to identify statistically significant
@@
 be back-transformed to the original score reliability scale
-(references):
+[@beinhauer2025]:
 
@@
 We executed these steps for each phenomenon. The statistical programming
-language R, Version X.XXX (reference) is used for all data manipulation
-and statistical analysis.
+language R, Version 4.4.2 [@RCore] is used for all data manipulation and
+statistical analysis. Graphics are constructed using the R package
+*ggplot2*, version 3.5.1 [@ggplot2].
 
@@
 ($\delta$), which are central to assess the research questions posed in
-section XXX. Most importantly, @tbl-tauES demonstrates that for all but
-one of the phenomena, the attenuation correction procedure barely has an
-impact on the extent of ES heterogeneity. However, while the
-heterogeneity for the Srull_Behaviour_Hostility phenomenon seems reduced
-by about a third of its initial value, the ES heterogeneity for this
-phenomenon is rather small to begin with. Using a significance test, we
-can not distinguish this value confidently from zero. This implies that
-for the majority of phenomena, differences in score reliability can not
-be made responsible for heterogeneity in standardized effect sizes.
+section *Attenuation Correction*. Most importantly, @tbl-tauES
+demonstrates that for all but one of the phenomena, the attenuation
+correction procedure barely has an impact on the extent of ES
+heterogeneity. However, while the heterogeneity for the
+Srull_Behaviour_Hostility phenomenon seems reduced by about a third of
+its initial value, the ES heterogeneity for this phenomenon is rather
+small to begin with. Using a significance test, we can not distinguish
+this value confidently from zero. This implies that for the majority of
+phenomena, differences in score reliability can not be made responsible
+for heterogeneity in standardized effect sizes.
 
@@
 as a potential moderator explaining ES heterogeneity. Additionally, we
-demonstrated that the current consensus formulated in the literature
-regarding how score reliability increases ES heterogeneity is false. We
-found a large number of phenomena where, contrary to popular claims,
-correcting for score reliability increases ES heterogeneity, instead of
-decreasing it.
+demonstrated that claims formulated in W&D and H&S regarding how score
+reliability increases ES heterogeneity is false. We found a large number
+of phenomena where, contrary to popular claims, correcting for score
+reliability increases ES heterogeneity, instead of decreasing it.
 
@@
 unexplained variation in a parameter's true distribution, that we model
-using meta-analytical techniques. In that way, the idea of heterogeneity
-is closely related to residual variance, as we observe it in a simple
-linear model. Both describe variation that could not be explained by the
-predictive variables in the case of the simple linear model, or not
-explained by incorporating the Standard Errors in the case of a
-meta-analytic model.
+using meta-analytical techniques [@borenstein2009; @higgins2002]. In
+that way, the idea of heterogeneity is closely related to residual
+variance, as we observe it in a simple linear model. Both describe
+variation that could not be explained by the predictive variables in the
+case of the simple linear model, or not explained by incorporating the
+Standard Errors in the case of a meta-analytic model.
 
@@
 demonstrates how heterogeneity in score reliability affects
-heterogeneity in ESs.\@eq-HS implies that heterogeneity in $\delta_0$ is
-essentially a function in which heterogeneity and mean value of $\delta$
-are weighted by mean score reliability $\rho_{xx‘}$ and its
-heterogeneity.
+heterogeneity in ESs [-@hunter2004, p. XX]. @eq-HS implies that
+heterogeneity in $\delta_0$ is essentially a function in which
+heterogeneity and mean value of $\delta$ are weighted by mean score
+reliability $\rho_{xx‘}$ and its heterogeneity.
 
@@
 undistorted by sampling error and measurement error. Lastly,
-$\rho_{XX'}$ refers to score reliability. Hunter & Schmidt (xxx) claim
-that @eq-HS demonstrates that differences in score reliability inflate
-ES heterogeneity, which means that an attenuation correction procedure
-would have to reduce ES heterogeneity. However, it can be easily
-verified that @eq-HS does not allow for such claims. Depending on both
-the mean value $\mu_{\rho_{xx‘}}$ and the heterogeneity
-$\tau^2_{\rho_{xx‘}}$ in score reliability, ES heterogeneity might be
-increased or decreased by the differences in score reliability.
+$\rho_{XX'}$ refers to score reliability. Hunter & Schmidt
+[-@hunter2004] claim that @eq-HS demonstrates that differences in score
+reliability inflate ES heterogeneity, which means that an attenuation
+correction procedure would have to reduce ES heterogeneity. However, it
+can be easily verified that @eq-HS does not allow for such claims.
+Depending on both the mean value $\mu_{\rho_{xx‘}}$ and the
+heterogeneity $\tau^2_{\rho_{xx‘}}$ in score reliability, ES
+heterogeneity might be increased or decreased by the differences in
+score reliability.
 
@@
 of variation is controlled for, the amount of unexplained variance
-should be reduced." Wiernik and Dahlke (ref) refer to the work by Hunter
-& Schmidt (ref), likely following their intuition.
+should be reduced." Wiernik and Dahlke [-@wiernik2020] refer to the work
+by Hunter & Schmidt [-@wiernik2020], likely following their intuition.
 
 In the re-analysis of archival data, we have demonstrated that this
-intution does not hold for the relationship of ES heterogeneity and
+intuition does not hold for the relationship of ES heterogeneity and
 score reliability. Instead, we will propose an alternative way that
@@
 constructed by dividing one random variable by a second random variable:
-$Z = \frac{X}{Y}$ (reference). First order taylor approximation may be
-used to generate estimates of mean and variance of the ratio variable
-\[irrespective of distribution?\] (reference). Here we use the
+$Z = \frac{X}{Y}$. First order taylor approximation may be used to
+generate estimates of mean and variance of the ratio variable
+\[irrespective of distribution?\] [@vanKempen1999]. Here we use the
 Taylor-estimator of the ratio’s variance to demonstrate how differences
@@
 $R_1$ essentially describes the relative size of mean score variance
-after the attenuation correction, and $R_2$ essentially described the
+after the attenuation correction, and $R_2$ essentially describes the
 relative size of score variance heterogeneity after the attenuation
@@
 
-It is important to note that the condition in @eq-ineq_short only
+It is important to reiterate that the condition in @eq-ineq_short only
 describes under which conditions the inequality is guaranteed to hold,
 no matter what value other parameters like the mean value of $MD$ or its
-heterogeneity take on. This means that, even if the condition in
+heterogeneity take on. This means that, if the condition in
 @eq-ineq_short is violated, it is by no means guaranteed that an
 attenuation correction procedure would actually reduce heterogeneity in
-ES. This means that inference made from @eq-ineq_short is asymmetrical.
-While we know, if the inequality holds, heterogeneity in corrected ES
-will be larger than heterogeneity in uncorrected ES, we can't be certain
-that the opposite is true if the inequality does not hold.
+ES. Thereby, inference made from @eq-ineq_short is asymmetrical. While
+we know, if the inequality holds, heterogeneity in corrected ES will be
+larger than heterogeneity in uncorrected ES, we can't be certain that
+the opposite is true if the inequality does not hold.
 
@@
 differences in error score variance $\sigma^2_E$. Among other things, in
-Beinhauer et al. (ref.), we deliver detailed instructions on how
-estimates for mean and heterogeneity of total score variance, and the
-score variance components can be derived. This means that Boot-Err can
-be used to generate all estimates required for computation of the
+Beinhauer et al. [-@beinhauer2025] we deliver detailed instructions on
+how estimates for mean and heterogeneity of total score variance, and
+the score variance components can be derived. This means that Boot-Err
+can be used to generate all estimates required for computation of the
 metrics $R_1$ and $R_2$.
@@
 meta-analytic techniques and bootstrap sampling, we perform analyses
-using the packages *metafor*, version 4.8-0, and *boot*, version 1.3-31,
-in the statistical programming language *R*, version 4.4.2.
+using the packages *metafor*, version 4.8-0 [@metafor], and *boot*,
+version 1.3-31 [@boot], in the statistical programming language *R*,
+version 4.4.2 [@RCore].
 
@@
 
-Since $R_1$ is carrying the power of 3, it implies that the reduction in
-mean score variance is affecting the ES heterogeneity more strongly than
-the reduction in score variance heterogeneity. This is best described
-through an example: We know what an acceptable average score reliability
-looks like in the area of Psychology. Researchers largely target a score
-reliability of at least .7, preferably .8. An average score reliability
-of .7 implies that 70% of the variance observed in scores can actually
-be attributed to true underlying differences. From our derived rule, we
-know that, as long as $R_2$ is larger or equal to $.7^3 = .343$, the
-attenuation correction procedure can only increase ES heterogeneity. If
-we want to explain ES heterogeneity through score reliability, even
-though only 30% of the variance observed in score can be attributed to
-random measurement error, at least 65.7%, almost two thirds, of the
-heterogeneity in score variance would need to be attributable to
-differences in the error score variance. In other words, at most 34.3%
-of heterogeneity in score variances may be due to actual differences in
-how the latent variable is distributed (in terms of spread) across the
-administration sites.
+Since $R_1$ is carrying the exponent of 3, it implies that the reduction
+in mean score variance is affecting the ES heterogeneity more strongly
+than the reduction in score variance heterogeneity. This is best
+described through an example: We know what an acceptable average score
+reliability looks like in the area of Psychology. Researchers largely
+target a score reliability of at least .7, preferably .8. An average
+score reliability of .7 implies that 70% of the variance observed in
+scores can actually be attributed to true underlying differences. From
+our derived rule, we know that, as long as $R_2$ is larger or equal to
+$.7^3 = .343$, the attenuation correction procedure can only increase ES
+heterogeneity. If we want to explain ES heterogeneity through score
+reliability, even though only 30% of the variance observed in score can
+be attributed to random measurement error, at least 65.7%, almost two
+thirds, of the heterogeneity in score variance would need to be
+attributable to differences in the error score variance. In other words,
+at most 34.3% of heterogeneity in score variances may be due to actual
+differences in how the latent variable is distributed (in terms of
+spread) across the administration sites.
 
 In the examples we re-analysed throughout this text, as well as the
-majority of research studies done in the behavioural sciences, no proper
-sampling techniques are employed. Most samples where these instruments
-are applied come to be through convenience sampling, where students
-self-select into studies. Without a valid sampling scheme, we
-essentially have have no basis to assume anything about the true score
-variance. Most importantly, however, we have no reason to assume that
-the true score variance could potentially be homogeneous or stable
-across administration sites. Instead, it is very likely that the
-convenience sampling procedures produce samples that vary extremely
-strongly in how they are put together, leading to large degrees of true
-score variance heterogeneity. At the same time, score reliability is a
-low-hanging fruit to optimise as most researchers are aware and trained
-in assessing score reliability.
+majority of research studies done in the behavioural sciences
+[@hultsch2002; @bornstein2013], no proper sampling techniques are
+employed. Most samples where these instruments are applied come to be
+through convenience sampling, where students self-select into studies.
+Without a valid sampling scheme, we essentially have have no basis to
+assume anything about the true score variance [@hultsch2002]. Most
+importantly, however, we have no reason to assume that the true score
+variance could potentially be homogeneous or stable across
+administration sites. Instead, it is very likely that the convenience
+sampling procedures produce samples that vary extremely strongly in how
+they are put together, leading to large degrees of true score variance
+heterogeneity. At the same time, score reliability is a low-hanging
+fruit to optimise as most researchers are aware and trained in assessing
+score reliability [@hussey2025].
 
@@
 Our results fit in well with recent work by Olsson-Collentine et al.
-(2023). In a large simulation scheme, they demonstrate that differences
-in score reliability across administrations typically deflate
-heterogeneity in uncorrected correlations. In such cases, an attenuation
-correction would increase correlation coefficient heterogeneity, as we
-observe for standardized mean differences. Only as the true
-heterogeneity in correlations grows larger would the correction
+[@olsson2023]. In a large simulation scheme, they demonstrate that
+differences in score reliability across administrations typically
+deflate heterogeneity in uncorrected correlations. In such cases, an
+attenuation correction would increase correlation coefficient
+heterogeneity, as we observe for standardized mean differences. Only as
+the true heterogeneity in correlations grows larger would the correction
 procedure reduce heterogeneity as implied in W&D. While correlations and
@@
 
-The analytical arguments, detailed in Appendix X, leading to
+The analytical arguments, detailed in *Appendix A*, leading to
 @eq-ineq_short, rest on the use of the delta-method. @eq-ineq_init
@@
 one of which is likely violated in our analysis: the variance of the
-variable to be transformed needs to be small (ref.). This entire
+variable to be transformed needs to be small [@cramer1999]. This entire
 procedure discusses the influence of variance in standard deviations on
@@
 Therefore, assuming that this assumption holds seems naive. In the
-Appendix X, we explore the archival data, to understand in how far
+*Appendix C*, we explore the archival data, to understand in how far
 employing the delta-method on meta-analyses of standard deviations lead
@@
 
-The work done by Hunter and Schmidt (2014) was crucial in informing and
-guiding methodology development in the field of meta-analyses.
-Similarly, Wiernik and Dahlke (2020) raise a number of important points
-that have been neglected in the application of meta-analytic research
-over the last decades. We generally agree with their ideas that
+The work done by Hunter and Schmidt [-@hunter2004] was crucial in
+informing and guiding methodology development in the field of
+meta-analyses. Similarly, Wiernik and Dahlke [-@wiernik2020] raise a
+number of important points that go far beyond attenuation correction and
+have been neglected in the application of meta-analytic research over
+the last decades. We generally agree with their ideas that
 underappreciated (differences in un-)reliability can introduce
@@
 
-Using the methdology underlying the Boot-Err approach (Beinhauer et al.
-ref), we generate estimates of meta-analytic mean and heterogeneity in
+Using the methoodlogy underlying the Boot-Err approach [@beinhauer2025],
+we generate estimates of meta-analytic mean and heterogeneity in
 log-transformed total score variance $\sigma^2_X$. To facilitate
@@
 
-In [equations @eq-mu_sigmaE] and [-@eq-tau_sigmaE] we generate estimates
+In [Equations @eq-mu_sigmaE] and [-@eq-tau_sigmaE] we generate estimates
 of meta-analytic mean and heterogeneity of error score variance

### 2025-04-11 test_delta_assumptions.R

@@ -1,3 +1,70 @@
+## Score ReLiability
+
+# E3
+
+# Clean & extract data
+
+
+# library loading and installing as necessary
+
+
+# relevant libraries required for this script
+packages <- c("magrittr", "dplyr", "boot", "here")
+
+# check, whether library already installed or not - install and load as needed:
+apply(as.matrix(packages), MARGIN = 1, FUN = function(x) {
+  
+  pkg_avail <- nzchar(system.file(package = x))   # check if library is installed on system
+  
+  if(pkg_avail){
+    require(x, character.only = TRUE)             # load the library, if already installed
+    
+  }else{
+    install.packages(x)                           # install the library, if missing
+    require(x, character.only = TRUE)             # load after installation
+  }
+})
+
+source(here("ReLiability_Function-library.R"))
+
+
+B_alpha_rma_df <- read.csv(here("Data/Processed/Reliability_analysis.csv"))
+
+ES_rma_df <- read.csv(here("Data/Processed/Aggregates_ES_analysis.csv"))
+
+agg_L <- readRDS(here("Data/Processed/Aggregates_simple.csv"))
+
+MASC_names <- substr(list.files(here("Data/Extracted (Project) Data")), 1, nchar(list.files(here("Data/Extracted (Project) Data")))-4) 
+
+
+
+
+data_files <- list.files(here("Data/Extracted (Project) Data"), full.names = TRUE)
+
+
+data.list <- lapply(data_files, read.csv)
+
+
+data.list[[4]]$source <- substr(data.list[[4]]$source, start = nchar(data.list[[4]]$source) - 2, stop = nchar(data.list[[4]]$source))
+
+
+effect_index <- MASC_names %in% c("Albarracin_Priming_SAT", 
+                                  "Carter_Flag_Priming", "Caruso_Currency_Priming",
+                                  "Dijksterhuis_trivia", "Finkel_Exit_Forgiveness", 
+                                  "Finkel_Neglect_Forgiveness",
+                                  "Giessner_Vertical_Position", "Hart_Criminal_Intentionality",    
+                                  "Hart_Detailed_Processing", "Hart_Intention_Attribution",       
+                                  "Husnu_Imagined_Contact", "Nosek_Explicit_Art",
+                                  "Nosek_Explicit_Math", "PSACR001_anxiety_int", 
[…]

### 2025-04-11 Manuscript_ScoreReLiability.qmd
text updates

@@
 title: "Score ReLiability? Exploring the role of score reliability for effect size heterogeneity"
+sbutitle: ""
 format: html
 editor: visual
+author: "Beinhauer, L.J., Fünderich, J.H., Renkewitz, F."
+engine: knitr
 ---
@@
 Large-scale attempts of direct replications, using identical protocols, such as the Many Labs studies or Registered Replication Reports (references), for the first time allow researchers to estimate heterogeneity undistorted by typical experimental factors.
-In re-analyses of these studies, Olsson-Collentine et al.
-(2020) identify a strong correlation between a studies effect size and its heterogeneity.
-Similarly, van Erp et al.
-(2017) or Stanley et al.
-(2018) estimate strong degrees of heterogeneity across psychological replications in general.
-In a separate re-analysis of large scale direct replications, Renkewitz et al.
-(2024) identify substantial heterogeneity in almost all projects where a non-zero effect could be identified.
-This aligns with the correlation found by Olsson-Collentine et al.
-(2020).
+In re-analyses of these studies, Olsson-Collentine et al. (2020) identify a strong correlation between a studies effect size and its heterogeneity. 
+Similarly, van Erp et al. (2017) or Stanley et al. (2018) estimate strong degrees of heterogeneity across psychological replications in general. 
+In a separate re-analysis of large scale direct replications, Renkewitz et al. (2024) identify substantial heterogeneity in almost all projects where a non-zero effect could be identified. 
+This aligns with the correlation found by Olsson-Collentine et al. (2020).
 
@@
 
+Re-arranging equation [-@eg-HS] leads to the following equation:
+
+$$Var(\delta) = \frac{Var(\delta_0) - [E(\delta)]^2 Var(a)}{E[a^2]}$$ {#eq-HS_r}
+
+where:
+$$E(\delta) = \frac{E(\delta_0)}{E(a)}$$ {#eq-HS_E}
+
+$$
+
 ## Analytical Arguments
@@
 If equation [-@eq-HS] were to be true, it would imply that heterogeneity in $\delta_0$ is essentially a function in which heterogeneity and mean value of $\delta$ are weighted by mean score reliability $a$ and its heterogeneity.
-In this scenario, it is easy to construct cases where heterogeneity in $\delta_0$ is inflated or deflated by heterogeneity in $a$.
-In table 1, two such cases are demonstrated.
+In this scenario, it is easy to construct cases where heterogeneity in $\delta_0$ is inflated or deflated by heterogeneity in $a$. In table 1, two such cases are demonstrated. 
 All else is held equal, only the mean score reliability $E[a]$ is changed from .7 to .8. 
@@
 
-Hunter & Schmidt point out that equation [-@eq-HS] is only valid “*If the level of reliability is independent of the true effect size across studies \[…\]”* (2014, p.
-309).
+Hunter & Schmidt point out that equation [-@eq-HS] is only valid “*If the level of reliability is independent of the true effect size across studies \[…\]”* (2014, p. 309). 
 However, both parameters, ES $\delta$ and score reliability $a$ in original notation can not, by definition, be independent parameters. 
@@
 Re-analyses of the ManyLabs and Registered Replication Reports have shown that such differences in true variance across replications are the norm, not the exception.
-In that case, some variance in $\sigma_T$ would remain, albeit smaller than the variance in $\sigma$.
-We reformulate equation [-@eq-TaylorES0] in terms of $\sigma_T$
+In that case, some variance in $\sigma_T$ would remain, albeit smaller than the variance in $\sigma$. We reformulate equation [-@eq-TaylorES0] in terms of $\sigma_T$
 
@@
 
-The implication of this equation is, again, not straightforward.
-The expected value of $\sigma_T$ is guaranteed to be smaller than the expected value of uncorrected $\sigma$.
+The implication of this equation is, again, not straightforward. The expected value of $\sigma_T$ is guaranteed to be smaller than the expected value of uncorrected $\sigma$. 
 Similarly, the variance in $\sigma_T$ is guaranteed to be smaller than the uncorrected variance in $\sigma$. 
@@
 The four groups may be collapsed into a single treatment group (reconstrual and repurposing) and a single control group (passive and active control). 
-In total, Wang et al.
-(2021) assessed 21,644 individuals from 87 countries/regions.
-These countries/regions served as the study-level grouping variable for our meta-analyses.
+In total, Wang et al. (2021) assessed 21,644 individuals from 87 countries/regions. These countries/regions served as the study-level grouping variable for our meta-analyses.
 
@@
 
+{{< pagebreak >}} 
 *Table 3*
@@
 Grey dots represent the uncorrected ESs for each country, with accompanying confidence interval represented by the bars surrounding the dots. 
-The overlying black dots on the other hand represent the corrected ESs, with additional confidence intervals.  The position of the grey and black diamonds on the bottom display the meta-analytical estimate of mean ES, uncorrected (grey) and corrected (black), whereas the width of the diamonds represent the respective confidence interval.
+The overlying black dots on the other hand represent the corrected ESs, with additional confidence intervals. 
+The position of the grey and black diamonds on the bottom display the meta-analytical estimate of mean ES, uncorrected (grey) and corrected (black), whereas the width of the diamonds represent the respective confidence interval.
 The forest plot shows that, as expected, the attenuation correction leads to larger ES and larger individual confidence intervals. 
@@
 
-These results fit in well with work recently published/preprinted by Olsson-Collentine et al.
-(2023).
+These results fit in well with work recently published/preprinted by Olsson-Collentine et al. (2023). 
 In a large simulation scheme, they demonstrate that differences in score reliability across administrations initially deflate heterogeneity in correlations. 

### 2025-04-11 temp_results.qmd

@@
 To assess whether actual differences in score reliability existed, heterogeneity in score reliability is assessed across all four projects, making use of Reliability Generalization Meta-Analysis.
+Table 1 briefly summarises the results concerning differences in score reliability.
+From Cochran's Q-test, a hypothesis test on whether heterogeneity is larger than zero, we can gather that for the scales measuring all four phenomena, score reliability differs across samples.
+The individual heterogeneity in score reliabiliy tau_alpha, is rather low for the phenomena of Nosek, PSACR001 Anxiety, and PSACR002 negative photo.
+This is likely due to the higher levels of score reliability, averaging at least `r round(min(B_alpha_rma_df$mu_alpha[B_alpha_rma_df$mu_alpha >= .7]), 2)`.
+However, even for the phenomenon PSACR001_behav_int, where true mean score reliability is below .7 and the heterogeneity is rather large with tau_alpha = `r round(B_alpha_rma_df$tau_alpha[B_alpha_rma_df$mu_alpha < .7], 2)`, correcting for differences score reliability does not reduce ES heterogeneity.
+The analysis of score reliability highlights that the lack of reduction in corrected ES heterogeneity can not be explained away by a lack of variation in reliability or "insufficiently imperfect" reliability.
+
+```{r, echo = FALSE}
+tab_df <- B_alpha_rma_df %>% 
+  select(MASC, mu_alpha, tau_alpha, QE, k, QEp) %>% 
+  mutate(QEp = ifelse(QEp < .001, "<.001", QEp))
+
+knitr::kable(tab_df, digits = 3)
+```

### 2025-09-14 Manuscript_MD_ScoreReLiability.qmd
update manuscript up to methods

@@
 csl: styles/apa-6th-edition.csl
-format: html
+format: pdf
 editor: 
@@
 # relevant libraries required for this script
-packages <- c("ggplot2", "here", "metafor", "gridExtra", "knitr", "magrittr", "dplyr")
+packages <- c("ggplot2", "here", "metafor", "gridExtra", "knitr", "magrittr", "dplyr", "gt")
 
@@
 # Liability or reliability? Exploring the role of effect size attenuation on meta-analytic heterogeneity
+
+Lukas J. Beinhauer *University of Erfurt* ORCID:
+https://orcid.org/0000-0001-9841-3089
+
+Jens H. Fünderich *Ludwig Maximilian University of Munich* ORCID:
+https://orcid.org/0000-0002-7185-9248
+
+Frank Renkewitz *University of Erfurt* ORCID:
+https://orcid.org/0000-0001-8072-6802
 
@@
 Similarly, it has been argued that heterogeneity in effect sizes is an
-indicator of the theory’s “completeness” surrounding the phenomenon.
-Linden and Hönekopp argue that “low (as opposed to high) heterogeneity
-reflects a more advanced understanding of the subject matter being
-studied” [-@LindenHönekopp, p. 2]. Similarly, von Hippel and Schuetze
-[-@vonHippel2023] argue that heterogeneity in effects is an indicator of
-a vague, poorly specified theory.
+indicator of the theory’s “completeness” surrounding the phenomenon. For
+example, Linden and Hönekopp argue that “low (as opposed to high)
+heterogeneity reflects a more advanced understanding of the subject
+matter being studied” [-@LindenHönekopp, p. 2]. Similarly, von Hippel
+and Schuetze [-@vonHippel2023] argue that heterogeneity in effects is an
+indicator of a vague, poorly specified theory.
 
 Large-scale attempts of direct replications, using identical protocols,
-such as the Many Labs studies or Registered Replication Reports [e.g.
-@ManyLabs1; @RRRHart], for the first time allow researchers to estimate
-heterogeneity undistorted by selection processes like publication bias.
-In re-analyses of these studies, Olsson-Collentine et al. [-@olsson2020]
-identify a positive correlation between a phenomenon’s effect size and
-its heterogeneity. Similarly, van Erp et al. [-@vanErp2017] and Stanley
-et al. [-@stanley2018] estimate strong degrees of heterogeneity across
-conceptual replications in Psychology. In a separate re-analysis of
-large scale direct replications, Renkewitz et al. (in preparation)
-identify substantial heterogeneity in almost all projects where a
-non-zero effect could be identified. This aligns with the correlation
-found by Olsson-Collentine et al. [-@olsson2020].
+such as the Many Labs studies or Registered Replication Reports
+[@ManyLabs1; e.g. @RRRHart], for the first time allow researchers to
+estimate heterogeneity undistorted by selection processes like
+publication bias. In re-analyses of these studies, Olsson-Collentine et
+al. [-@olsson2020] identify a positive correlation between a
+phenomenon’s effect size and its heterogeneity. Both van Erp et al.
+[-@vanErp2017] and Stanley et al. [-@stanley2018] estimate strong
+degrees of heterogeneity across conceptual replications in Psychology.
+In a separate re-analysis of large scale direct replications, Renkewitz
+et al. (in preparation) identify substantial heterogeneity in almost all
+projects where a non-zero effect could be identified. This aligns with
+the correlation found by Olsson-Collentine et al. [-@olsson2020].
 
@@
 vary in their extent across replications. \[This is an argument put
-forward repeatedly by original researchers and replicators alike (do we
-actually know of any?).\] Hunter and Schmidt [-@hunter2004] and Wiernik
-and Dahlke [-@wiernik2020], discussing the role of score reliability for
+forward repeatedly by original researchers and replicators alike, such
+as Olsson-Collentine et al. [-@olsson2020] (do we know any more?).\]
+Hunter and Schmidt [-@hunter2004] and Wiernik and Dahlke
+[-@wiernik2020], discussing the role of score reliability for
 standardized effect sizes, claim that differences in score reliability
@@
 attenuation correction procedures, should remove that inflation,
-reducing heterogeneity in corrected effect sizes. Over the following
-pages, we assess whether score reliability can in fact be made
-responsible for heterogeneity identified across phenomena. We find no
-empirical evidence for the argument that correcting for differences in
-score reliability reduces effect size heterogeneity. Instead, to our
-great surprise, we find that heterogeneity might as well be larger after
-the correction procedure. Therefore, theoretical analyses regarding the
-relationship between score reliability and heterogeneity need to be
-false. In the second half of this text we make use of analytical
-arguments to identify the necessary conditions under which an
-attenuation correction inevitably leads to an increase in effect size
-heterogeneity.
+reducing heterogeneity in corrected effect sizes.
+
+Over the following pages, we assess whether score reliability can in
+fact be made responsible for heterogeneity identified across several
+psychological phenomena. However, we find no empirical evidence for the
+argument that correcting for differences in score reliability reduces
+effect size heterogeneity. Instead, to our great surprise, we find that
+heterogeneity might as well be larger after the correction procedure.
+Therefore, previous analyses and assumptions regarding the relationship
+between score reliability and heterogeneity need to be false. In the
+second half of this text we make use of analytical arguments to identify
+the necessary conditions under which an attenuation correction
+inevitably leads to an increase in effect size heterogeneity \[Do we
+need any "Schleifen" in this paragraph - to reduce informational load /
+density\].
 
@@
 Here, MD refers to the mean difference between two groups of interest,
-while $\sigma_X$ is the total pooled standard deviation. Since $d$
-constitutes an observed effect size it is affected by sampling error. In
-terms of heterogeneity, we are not interested in the variance of $d$,
-but the variance of the true, underlying ES $\delta$. In a
-random-effects meta-analytic model, we can use the standard error and an
-estimate of heterogeneity to understand how far off the individual
-observed $d$ is from the true $\delta$ in a weighting procedure
-[@borenstein2009] The standard error of Cohen’s $d$ is computed as
-defined in @eq-SEd.
-
-$$SE_d=\sqrt{\frac{n_1+n_2}{n_1n_2}+\frac{d^2}{2\left(n_1+n_2\right)}}$$ {#eq-SEd}
-
-As is widely known, score reliability affects ES [e.g. @hunter2004;
-@wiernik2020]. In the context of classical test theory (CTT) score
-reliability $\rho_{XX’}$ is defined as the ratio of true to total score
-variance [@lord2008], as defined in @eq-rel.
+while $\sigma_X$ is the pooled standard deviation. As is widely known,
+score reliability affects ES [e.g. @hunter2004; @wiernik2020]. In the
+context of classical test theory (CTT) score reliability $\rho_{XX’}$ is
+defined as the ratio of true to total score variance [@lord2008], as
+defined in @eq-rel.
 
@@
 [@muchinsky1996; @schmidt1996; @schennach2016], is a widespread practice
-in Psychology (reference). @eq-dc describes a simple attenuation
-correction procedure.
+in Psychology (reference, alt.: repeatedly recommended + references).
+@eq-dc describes a simple attenuation correction procedure.
 
@@
 @eq-rel. Thereby, an estimate of ES is generated, under the premise that
-score reliability would have been perfect ($\rho_{xx'} = 1$). As the
-corrected estimate of Cohen’s $d$ is larger and requires an additional
-unknown parameter, an estimate of score reliability, the larger
-uncertainty in this parameter should be reflected in a larger standard
-error. The standard error for $d_c$ is defined in @eq-SEdc.
-
-$$SE_{d_c} = \sqrt{SE^2_d \frac{d_c}{d}}$$ {#eq-SEdc}
+score reliability would have been perfect ($\rho_{xx'} = 1$).
 
@@
 -   To what extent can heterogeneity in standardized effect sizes be
-    explained by means of an attenuation correction?
+    reduced by means of an attenuation correction?
 
@@
 replications on psychological phenomena [@dripht2024]. The data-sets of
-about 50 phenomena, largely stemming from the efforts of the Many Labs
+about 55 phenomena, largely stemming from the efforts of the Many Labs
 studies, Registered Replication Reports or the Psychological Science
 Accelerator, are made openly available in a standardised format at
-\[<https://osf.io/7p8rk/>\], the DRIPHT repository.
+\[<https://osf.io/7p8rk/>\], the DRIPHT repository. A documentation of
+the standardised format, including a web-application to easily explore
+the public data-sets, can be found in Fünderich et al.
+[-@fuenderich_metapipex2024].
 
@@
 first three and the fifth Many Labs projects [@ManyLabs1; @ManyLabs2;
-@ManyLabs3; @ManyLabs5] was harmonized and publicly stored in a
+@ManyLabs3; @ManyLabs5] were harmonized and publicly stored in a
 standardized format when the DRIPHT repository was set up [@dripht2024].
@@
 distributed across the globe, do not formally attempt to perform direct
-replications, but focus on original research or conceptual replications.
-However, as each project used the same protocol and participants are
-distributed across different countries, a data structure similar to that
-of a multi-site replication study emerges. What makes all these
+replications. Instead, these projects focus on original research or
+conceptual research done collaboratively in laboratories across the
+globe. However, as each project used the same protocol and participants
+are distributed across different countries, a data structure similar to
+that of a multi-site replication study emerges. What makes all these
 collaborative efforts valuable for this manuscript is the fact that all
@@
 While replication data on a decent number of psychological phenomena is
-available in the DRIPHT-repository, to demonstrate how incorporating
-score reliability affects heterogeneity in ES, selected phenomena need
-to fulfil three conditions: (1) As we focus on the use of ES $d$, the
+available in the DRIPHT-repository, selected phenomena need to fulfil
+three conditions to demonstrate how incorporating score reliability
+affects heterogeneity in ES: (1) As we focus on the use of ES $d$, the
 phenomenon needs to be assessed using a group design consisting of at
@@
 phenomena where meta-analytic mean ES can be statistically distinguished
-from zero. While the first criterion is fulfilled by all phenomena found
-in the DRIPHT repository, the second and third criteria still need
-evaluation.
+from zero. This aligns with the observation made in (Renkewitz,
+Fünderich & Beinhauer, in preparation) and Olsson-Collentine et al.
+[-@olsson2020]. As null-effects displayed little to no heterogeneity,
+there is no variability that might be reduced by incorporating
+differences in score reliability. While the first criterion is fulfilled
+by all phenomena found in the DRIPHT repository, the second and third
+criteria still need evaluation.
 
@@
 the different phenomena and how they were measured can be found at
-\[osf-link\], while a brief summary can be found in Table 1.
+\[osf-link\], while a brief summary can be found in @tbl-data.
 
@@
 
+In order to derive estimates of meta-analytic heterogeneity, we make use
+of a random-effects meta-analysis. Broken down, a random-effects
+meta-analysis essentially pools the sampling uncertainty of the
+different estimates that are to be aggregated. By removing this pooled
+estimate from the observed variability in the effect size estimates, we
+can generate an estimate of its heterogeneity - how strongly the effect
+sizes vary beyond their pooled sampling uncertainty.
+
+For the standardized mean difference $d$, we can compute the standard
+error as an estimate of its sampling uncertainty. The standard error of
+Cohen’s $d$ is computed as defined in @eq-SEd.
+
+$$SE_d=\sqrt{\frac{n_1+n_2}{n_1n_2}+\frac{d^2}{2\left(n_1+n_2\right)}}$$ {#eq-SEd}
+
+As the corrected estimate of Cohen’s $d$ is larger and requires an
+additional unknown parameter, an estimate of score reliability, the
+larger uncertainty in this parameter should be reflected in a larger
+standard error. The standard error for $d_c$ is defined in @eq-SEdc.
+
+$$SE_{d_c} = \sqrt{SE^2_d \frac{d_c}{d}}$$ {#eq-SEdc}
+
 ### Methods
@@
 estimator tends to perform best across a range of conditions
-[@hoenekopp2022]. To identify which phenomena pass criterion (3), using
+[@hoenekopp2022]. To identify which phenomena pass criterion (3) we use
 a Wald-type significance test we assess for which phenomenon the
@@
 an example from the 21 selected phenomena, the results from this
-procedure on Nosek’s Explicit Art sex differences phenomenon are
-displayed in the forest plot in @fig-forest. Here, grey dots represent
-the uncorrected ES for each sample, while black dots represent the
-corrected ES in each sample. The bars surrounding these dots show the
-respective 95%-Cis.
+procedure on replications on Nosek’s study on sex differences in
+preferences for the subject of Arts is displayed in the forest plot in
+@fig-forest. Here, grey dots represent the uncorrected ES for each
+sample, while black dots represent the corrected ES in each sample. The
+bars surrounding these dots show the respective 95%-Cis.
 
@@
 #| label: fig-forest 
-#| #| fig-cap: Forest Plot Nosek_Explicit_Art 
+#| fig-cap: Forest Plot ML1 - Nosek (Art)
 forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
@@
 criterion (3)—the effect size must be statistically significantly
-distinguishable from zero. 12 phenomena pass this criterion, highlighted
-in boldface. It may be remarked that the attenuation correction
-procedure, in this data, appears to have no influence on whether a
-phenomenon passes the criterion. The differences in p-value are rather
-small and lead to no difference in conclusion, regarding a significance
-level
+distinguishable from zero at $p < .05$. 12 phenomena pass this
+criterion. It may be remarked that the attenuation correction procedure,
+in this data, appears to have no influence on whether a phenomenon
+passes the criterion. The differences in p-value are rather small and
+lead to no difference in conclusion, regarding a significance level of
+5%.
 
@@
 #| tbl-cap: Meta-analytic average  ES
-knitr::kable(x = read.csv(here("Tables/Mean_ES.csv")),
-             col.names = c("MASC", 
-                           "$\\hat{\\mu}_{\\delta_0}$ (SE)", 
-                           "p", 
-                           "$\\hat{\\mu}_{\\delta}$ (SE)", 
-                           "p"),
-             escape = FALSE)
+
+gt::gt(read.csv(here("Tables/Mean_ES.csv")) %>% 
+         arrange(MASC)) %>% 
+  gt::cols_label(MASC = "MASC",
+                 mu_str_raw = gt::md("$\\hat{\\mu}_{\\delta_0}$ (SE)"),
+                 pval_raw = "p",
+                 mu_str_cor = gt::md("$\\hat{\\mu}_{\\delta}$ (SE)"),
+                 pval_cor = "p") %>% 
+  gt::tab_options(table.width = gt::pct(100),
+                  column_labels.font.weight = "bold") # %>% 
+  # gt::cols_width(pval_raw ~ gt::px(40),
+  #                pval_cor ~ gt::px(40))
 ```
@@
 correction procedure barely has an impact on the extent of ES
-heterogeneity. However, while the heterogeneity for the
-Srull_Behaviour_Hostility phenomenon seems reduced by about a third of
-its initial value, the ES heterogeneity for this phenomenon is rather
-small to begin with. Using a significance test, we can not distinguish
-this value confidently from zero. This implies that for the majority of
-phenomena, differences in score reliability can not be made responsible
-for heterogeneity in standardized effect sizes.
+heterogeneity. However, while the heterogeneity for the phenomenon from
+RRR9 (Behaviour) seems reduced by about a third of its initial value,
+the ES heterogeneity for this phenomenon is rather small to begin with.
+Using a significance test, we can not distinguish this value confidently
+from zero. This implies that for the majority of phenomena, differences
+in score reliability can not be made responsible for heterogeneity in
+standardized effect sizes.
 
@@
 
-knitr::kable(tab_both,
-             col.names = c("MASC", 
-                           "$\\hat{\\mu}_{\\rho_{XX'}}$ (95%-CI)", 
-                           "$\\hat{\\tau}_{\\rho_{XX'}}$", 
-                           "QE (df)", "p", 
-                           "$\\hat{\\tau}_{\\delta_0}$", 
-                           "QE (df)", "p",
-                           "$\\hat{\\tau}_{\\delta}$", 
-                           "QE (df)", "p"),
-             escape = FALSE)
-# 
+# knitr::kable(tab_both,
+#              col.names = c("MASC", 
+#                            "$\\hat{\\mu}_{\\rho_{XX'}}$ (95%-CI)", 
+#                            "$\\hat{\\tau}_{\\rho_{XX'}}$", 
+#                            "QE (df)", "p", 
+#                            "$\\hat{\\tau}_{\\delta_0}$", 
+#                            "QE (df)", "p",
+#                            "$\\hat{\\tau}_{\\delta}$", 
+#                            "QE (df)", "p"),
+#              escape = FALSE)
+
+gt::gt(tab_both  %>% 
+  arrange(tab_hetES.MASC)) %>% 
+  gt::cols_label(tab_hetES.MASC = "MASC",
+                 mu_alpha_str = gt::md("$\\hat{\\mu}_{\\rho_{XX'}}$"),
+                 tau_alpha = gt::md("$\\hat{\\tau}_{\\rho_{XX'}}$"),
+                 QE_str = "QE (df)",
+                 QEp = "p",
+                 tau_raw = gt::md("$\\hat{\\tau}_{\\delta_0}$"),
+                 QE_raw_str = "QE (df)",
+                 QEp_raw = "p",
+                 tau_raw = gt::md("$\\hat{\\tau}_{\\delta}$"),
+                 QE_raw_str = "QE (df)",
+                 QEp_raw = "p") %>% 
+  gt::tab_options(table.width = gt::pct(100),
+                  column_labels.font.weight = "bold") #%>% 
+  # gt::cols_width(pval_raw ~ gt::px(40),
+  #                pval_cor ~ gt::px(40))
+
 # tab_one <- data.frame(tab_muES,
@@
 attenuation correction increased ES heterogeneity. Concerning the
-remaining two phenomena, for the Shnabel_Willingness_Reconcile_RPP, the
-reduction in heterogeneity is miniscule and the estimate of
+remaining two phenomena, for the phenomenon described in ML5 - Shnabel
+(RPP), the reduction in heterogeneity is miniscule and the estimate of
 heterogeneity is not statistically significant in the first place. Most
@@
 issues of estimation, as only eight replications are available.
-Similarly, only for the Srull_Behaviour_Hostility phenomenon, we find a
-substantial reduction in heterogeneity due to the attenuation correction
-procedure. The estimates of ES heterogeneity are not statistically
-significant, however, which hampers our ability to meaningfully
-interpret and compare these estimates. For the majority of phenomena it
-appears that differences in score reliability do not inflate
-heterogeneity in uncorrected effect sizes. Instead these differences
-mask ES heterogeneity, that can be unveiled by accounting for score
-reliability.
+Similarly, only for the phenomenon described in RRR9 (Behaviour), we
+find a seemingly substantial reduction in heterogeneity due to the
+attenuation correction procedure. The estimates of ES heterogeneity are
+not statistically significant, however, which hampers our ability to
+meaningfully interpret and compare these estimates. For the majority of
+phenomena it appears that differences in score reliability do not
+inflate heterogeneity in uncorrected effect sizes. Instead these
+differences mask ES heterogeneity, that can be unveiled by accounting
+for score reliability.
 
@@
 demonstrated that, at least for the multi-site direct replication and
-PSA-projects, that score reliability does not hold any noteworthy value
-as a potential moderator explaining ES heterogeneity. Additionally, we
+PSA-projects, score reliability does not hold any noteworthy value as a
+potential moderator explaining ES heterogeneity. Additionally, we
 demonstrated that claims formulated in W&D and H&S regarding how score
-reliability increases ES heterogeneity is false. We found a large number
-of phenomena where, contrary to popular claims, correcting for score
-reliability increases ES heterogeneity, instead of decreasing it.
+reliability increases ES heterogeneity are false. We found a large
+number of phenomena where, contrary to popular claims, correcting for
+score reliability increases ES heterogeneity, instead of decreasing it.
 
@@
 reliability actually affects ES heterogeneity. Therein, we also identify
-potential conditions under which an attenuation correction can actually
-reduce ES heterogeneity - and under which it is more likely to increase
-it.
+potential conditions under which an attenuation correction will actually
+increase ES heterogeneity - and under which it is more likely to
+decrease it.
 
@@
 intuition does not hold for the relationship of ES heterogeneity and
-score reliability. Instead, we will propose an alternative way that
-helps us discern, under which conditions we can expect an attenuation
-correction procedure to lead to an increase in ES heterogeneity.
+score reliability. Instead, we propose an alternative way that helps us
+discern, under which conditions we can expect an attenuation correction
+procedure to lead to an increase in ES heterogeneity.
 
@@
 generate estimates of mean and variance of the ratio variable
-\[irrespective of distribution?\] [@vanKempen1999]. Here we use the
+[@vanKempen1999]. Taylor approximations typically require no specific
+assumptions regarding the distribution form, but that the variable (here
+a ratio) is a) differentiable, b) has existing moments and that c) its
+higher-order moments are negligible (e.g. limited variability and
+monotonicity) (still need a good reference here). Here we use the
 Taylor-estimator of the ratio’s variance to demonstrate how differences
@@
 
-To do so, we introduce two additional metrics.
+To do so, we had to introduce two additional metrics.
 
@@
 true score variance $\sigma^2_T$. Elsewhere, we proposed the method
-Boot-Err (ref.) as an alternative method to test for and estimate the
-differences in error score variance $\sigma^2_E$. Among other things, in
-Beinhauer et al. [-@beinhauer2025] we deliver detailed instructions on
-how estimates for mean and heterogeneity of total score variance, and
-the score variance components can be derived. This means that Boot-Err
-can be used to generate all estimates required for computation of the
-metrics $R_1$ and $R_2$.
+Boot-Err [@beinhauer2025] as an alternative method to test for and
+estimate the differences in error score variance $\sigma^2_E$. Among
+other things, in Beinhauer et al. [-@beinhauer2025] we deliver detailed
+instructions on how estimates for mean and heterogeneity of total score
+variance, and the score variance components can be derived. This means
+that Boot-Err can be used to generate all estimates required for
+computation of the metrics $R_1$ and $R_2$.
 
@@
 For the two exceptions, those phenomena where the attenuation correction
-procedure actually led to an decrease in ES heterogeneity
-(Srull_Behavior_Hostility and Shnabel_Willingness_Reconcile_RPP), the
-significance test for the ES heterogeneity was not significant, both
-before and after the correction. The results regarding $R_2$ indicate
-that the majority of score variance heterogeneity was due to differences
-in true score variance $\sigma^2_T$. Since either the heterogeneity was
-so small or the sample size so low that the heterogeneity could not be
-statistically distinguished from zero, we should be cautious in the
-interpretation of a reduction in that estimate. Most likely, the
-reduction in heterogeneity is better explained by imprecise estimation
-than an actual reduction of heterogeneity.
-
-Lastly, for two out of the 12 phenomena, the Albarracin_Priming_SAT and
-the Shnabel_Willingness_Reconcile_Rev phenomena, we find
+procedure actually led to a decrease in ES heterogeneity (RRR9
+(Behaviour) and ML5 - Shnabel (RPP)), the significance test for the ES
+heterogeneity was not significant, both before and after the correction.
+The results regarding $R_2$ indicate that the majority of score variance
+heterogeneity was due to differences in true score variance
+$\sigma^2_T$. Since either the heterogeneity was so small or the sample
+size so low that the heterogeneity could not be statistically
+distinguished from zero, we should be cautious in the interpretation of
+a reduction in that estimate. Most likely, the reduction in
+heterogeneity is better explained by imprecise estimation than an actual
+reduction of heterogeneity.
+
+Lastly, for two out of the 12 phenomena, the one described in ML5 -
+Albarracin and the other described in ML5 - Shnabel (Rev), we find
 $\tau_{\sigma^2_X}$ of 0. Therefore, the $R_2$ metric could not be
@@
 place. The phenomena were assessed with a particularly low number of
-replications (8 in both cases). Estimation precision for heterogeneity
-in variance components, even more so than for ES, largely depends on the
-sample size. Most likely, the low sample size led to implausible
-estimates in both true and total score variance heterogeneity and
-subsequently to an implausible estimate of $R_2$. Meta-analytic
-estimates, especially those concerning score variance components such as
-$R_2$, therefore should be interpreted with caution, if the sample size
-was particularly low.
+replications (9 and 8 respectively). Estimation precision for
+heterogeneity in variance components, even more so than for ES, largely
+depends on the sample size. Most likely, the low sample size led to
+implausible estimates in both true and total score variance
+heterogeneity and subsequently to an implausible estimate of $R_2$.
+Meta-analytic estimates, especially those concerning score variance
+components such as $R_2$, therefore should be interpreted with caution,
+if the sample size was particularly low.
 
@@
 #| tbl-cap: Re-analysis concerning metrics $R_1$ and $R_2$
-knitr::kable(read.csv(here("Tables/Variances_analysis.csv")),
-             col.names = c("MASC", 
-                           "$\\hat{\\mu}_{\\sigma^2_X}$", 
-                           "$\\hat{\\mu}_{\\sigma^2_T}$", 
-                           "$\\hat{\\tau}_{\\sigma^2_X}$", 
-                           "$\\hat{\\tau}_{\\sigma^2_T}$", 
-                           "$R_1$", 
-                           "$R_2$"))
+# knitr::kable(read.csv(here("Tables/Variances_analysis.csv")),
+#              col.names = c("MASC", 
+#                            "$\\hat{\\mu}_{\\sigma^2_X}$", 
+#                            "$\\hat{\\mu}_{\\sigma^2_T}$", 
+#                            "$\\hat{\\tau}_{\\sigma^2_X}$", 
+#                            "$\\hat{\\tau}_{\\sigma^2_T}$", 
+#                            "$R_1$", 
+#                            "$R_2$"))
+
+gt::gt(read.csv(here("Tables/Variances_analysis.csv")) %>% 
+         arrange(MASC)) %>% 
+  gt::cols_label(MASC = "MASC",
+                 mu_X = gt::md("$\\hat{\\mu}_{\\sigma^2_X}$"),
+                 mu_T = gt::md("$\\hat{\\mu}_{\\sigma^2_T}$"),
+                 tau_X = gt::md("$\\hat{\\tau}_{\\sigma^2_X}$"),
+                 tau_T = gt::md("$\\hat{\\tau}_{\\sigma^2_T}$"),
+                 R1 = gt::md("R_1"),
+                 R2 = gt::md("R_2")) %>% 
+  gt::tab_options(table.width = gt::pct(100),
+                  column_labels.font.weight = "bold") 
 ```
@@
 formulated in W&D and H&S. In almost all cases (every single case, if we
-restrict ourself to phenomena where we found statistically significant
+restrict ourselves to phenomena where we found statistically significant
 differences in ES), the attenuation correction led to increased
@@
 fraction of true score variance heterogeneity in the total score
-variance heterogeneity. We find that, as long as $R_1^3 < R_2$, an
+variance heterogeneity. We find that, as long as $R_1^3 \leq R_2$, an
 attenuation correction will increase ES heterogeneity.
@@
 Therefore, assuming that this assumption holds seems naive. In the
-*Appendix C*, we explore the archival data, to understand in how far
+*Appendix B*, we explore the archival data, to understand in how far
 employing the delta-method on meta-analyses of standard deviations lead
@@
 miniscule for the majority of phenomena, at the third decimal. Only for
-the PSACR002_behav_init phenomenon we find larger differences -
+the phenomenon in PSACR002 (Behaviour) we find larger differences -
 therefore, conclusions for this phenomenon concerning the relation
@@
 factors are held as constant as possible, is already substantial
-(Renkewitz, Fünderich, & Beinhauer, in preparation). However, so far we
+(Renkewitz, Fünderich & Beinhauer, in preparation). However, so far we
 have demonstrated that in several scenarios, differences in score
@@
 
+```{r, echo = FALSE}
+#| label: tbl-data
+#| tbl-cap: Brief summary of phenomena, designs and measures employed
+
+tab <- read.csv("Tables/data_summary.csv", fileEncoding = "UTF-8", sep = ";") %>% 
+  arrange(MASC)
+
+gt::gt(tab)
+```
+
 ## References
@@
 
-Taking $\frac{\tau_{\sigma^2_x}}{\tau_{\sigma^2_x}} = CV{\sigma^2_x}$
+Taking $\frac{\tau_{\sigma^2_x}}{\mu_{\sigma^2_x}} = CV{\sigma^2_x}$
 allows to simplify for @eq-ineq_rearr4.
 
-$$\frac{\tau_{MD}^2}{R_1} - \tau_{MD}^2 + \frac{\mu_{MD}^2 R_2}{4R_1^3}CV{\sigma^2_x} - \frac{\mu_{MD}^2}{4}CV{\sigma^2_x}  >  0$$ {#eq-ineq_rearr4}
+$$\frac{\tau_{MD}^2}{R_1} - \tau_{MD}^2 + \frac{\mu_{MD}^2 R_2}{4R_1^3}CV^2{\sigma^2_x} - \frac{\mu_{MD}^2}{4}CV^2{\sigma^2_x}  >  0$$ {#eq-ineq_rearr4}
 
@@
 
-### B - Back-transformation log-transformed score variance
-
-Using the methoodlogy underlying the Boot-Err approach [@beinhauer2025],
-we generate estimates of meta-analytic mean and heterogeneity in
-log-transformed total score variance $\sigma^2_X$. To facilitate
-interpretation and allow for computation of the metrics $R_1$ and $R_2$,
-we need to back-transform those estimates to the original scale. Using
-@eq-mu_sigmaX we can generate an estimate of meta-analytic mean score
-variance on the scale of $\sigma^2_X$ and using @eq-tau_sigmaX we get an
-estimate of its heterogeneity.
-
-$$\hat{\mu}_{\sigma^2_X} = exp\left(\hat{\mu}_{ln[\sigma^2_X]} + .5\  \hat{\tau}^2_{ln[\sigma^2_X]}\right)$$ {#eq-mu_sigmaX}
-
-$$\hat{\tau^2}_{\sigma^2_X} = exp\left(\hat{\tau}^2_{ln[\sigma^2_X]} - 1\right) \cdot \exp\left(2\ \hat{\mu}_{ln[\sigma^2_X]} + \hat{\tau}^2_{ln[\sigma^2_X]}\right)$$ {#eq-tau_sigmaX}
-
-In [equations @eq-mu_sigmaX] and [-@eq-tau_sigmaX],
-$\hat{\mu}_{ln[\sigma^2_X]}$ refers to the mean estimate from the
-meta-analysis on log-transformed score variance,
-$\hat{\tau}^2_{ln[\sigma^2_X]}$ refers to the estimate of heterogeneity.
-
-In [Equations @eq-mu_sigmaE] and [-@eq-tau_sigmaE] we generate estimates
-of meta-analytic mean and heterogeneity of error score variance
-$\sigma^2_E$. Just like for the total score variance, we make use of the
-estimates of mean and heterogeneity from the meta-analysis on
-log-transformed error score variance.
-
-$$\hat{\mu}_{\sigma^2_E} = exp\left(\hat{\mu}_{ln[\sigma^2_E]} + .5\  \hat{\tau}^2_{ln[\sigma^2_E]}\right)$$ {#eq-mu_sigmaE}
-
-$$\hat{\tau^2}_{\sigma^2_E} = exp\left(\hat{\tau}^2_{ln[\sigma^2_E]} - 1\right) \cdot \exp\left(2\ \hat{\mu}_{ln[\sigma^2_E]} + \hat{\tau}^2_{ln[\sigma^2_E]}\right)$$ {#eq-tau_sigmaE}
-
-### C - Examining bias due to delta-method
+### B - Examining bias due to delta-method
 
@@
 #| tbl-cap: Empirical assessment of appropriateness of delta-method
-knitr::kable(read.csv(here("Tables/delta_test.csv")),
-             col.names = c("MASC", 
-                           "$\\hat{\\mu}_{\\sigma^2_X}$", 
-                           "$\\hat{\\tau}^2_{\\sigma^2_X}$", 
-                           "$\\hat{\\mu}_{\\Delta \\sigma^2_X}$", 
-                           "$\\hat{\\tau}^2_{\\Delta \\sigma^2_X}$"))
+# knitr::kable(read.csv(here("Tables/delta_test.csv")),
+#              col.names = c("MASC", 
+#                            "$\\hat{\\mu}_{\\sigma^2_X}$", 
+#                            "$\\hat{\\tau}^2_{\\sigma^2_X}$", 
+#                            "$\\hat{\\mu}_{\\Delta \\sigma^2_X}$", 
+#                            "$\\hat{\\tau}^2_{\\Delta \\sigma^2_X}$"))
+
+gt::gt(read.csv(here("Tables/delta_test.csv")) %>% 
+         arrange(unique.ES_rma_df.MASC..nn_eff_idx.)) %>% 
+  gt::cols_label(unique.ES_rma_df.MASC..nn_eff_idx. = "MASC",
+                 vars.mu_X = gt::md("$\\hat{\\mu}_{\\sigma^2_X}$"),
+                 vars.tau2_X = gt::md("$\\hat{\\tau}^2_{\\sigma^2_X}$"),
+                 delta_mu_X = gt::md("$\\hat{\\mu}_{\\Delta \\sigma^2_X}$"),
+                 delta_tau2_X = gt::md("$\\hat{\\tau}^2_{\\Delta \\sigma^2_X}$")) %>% 
+  gt::tab_options(table.width = gt::pct(100),
+                  column_labels.font.weight = "bold") 
 ```
+
+*Note:* $\hat{\mu}_{\sigma^2_X}$ is the back-transformed meta-analytic
+estimate of mean total score variance, obtained via Boot-Err;
+$\hat{\mu}_{\sigma^2_X}$ is the back-transformed estimate of its
+heterogeneity. On the other hand, $\hat{\mu}_{\Delta\sigma^2_X}$ is the
+back-transformed meta-analytic estimate of mean total score variance,
+obtained from application of the delta method;
+$\hat{\mu}_{\Delta\sigma^2_X}$ is the back-transformed estimate of its
+heterogeneity.
 
@@
 occur in the 2nd decimal, indicating that violations in the assumptions
-concerning the delta-method are negligible here. Only for the
-PSACR002_behav_init phenomenon we find larger differences. These are
+concerning the delta-method are negligible here. Only for the phenomenon
+described in PSACR002 (Behaviour) we find larger differences. These are
 large enough to be concerning and pertain largely the estimate of

### 2025-09-15 Manuscript_MD_ScoreReLiability.qmd
update manuscript to estimates of heterogeneity

@@
 [@hoenekopp2022]. To identify which phenomena pass criterion (3) we use
-a Wald-type significance test we assess for which phenomenon the
+a Wald-type significance test, assessing for which phenomenon the
 meta-analytic mean ES is significantly different from zero. For this
@@
 Additionally, estimates of score reliability are derived. While not
-without criticism (references), Cronbach’s Alpha is used to estimate
-score reliability. It is often noted that tau-equivalence is an
+without criticism [e.g. @sijtsma2009], Cronbach’s Alpha is used to
+estimate score reliability. It is often noted that tau-equivalence is an
 unrealistic assumption to hold against real-world data, implying that
@@
 and how score reliability is estimated. The individual estimates of
-Cohen’s d are corrected using these estimates of Cronbach's Alpha,
-according to @eq-dc, including the corrected standard errors (@eq-SEdc).
-Thereby, estimates of ES are computed, which are corrected for imperfect
+Cohen’s $d$ are corrected using estimates of Cronbach's Alpha, according
+to @eq-dc, including the corrected standard errors (@eq-SEdc). Thereby,
+estimates of ES are computed, which are corrected for imperfect
 reliability. Again, a random-effects meta-analysis is run using metafor,
@@
 amount of heterogeneity in terms of variance ($\tau^2$) or standard
-deviation ($\tau$). These parameters discuss the variability of the
-standardized ES in the population in terms of variance or standard
-deviation. Therefore, we also make use of absolute heterogeneity here.
+deviation ($\tau$). Therefore, we also make use of absolute
+heterogeneity here.
 
@@
 estimates of heterogeneity are larger than zero can we actually begin to
-interpret whether there was any change.
+interpret whether there was any change. \[Short paragraph discussing
+that a test for difference in heterogeneity would be more informative?
+Maybe move it to the discussion section - hint at it here?\]
 
@@
 reliability are adequately transformed using Bonett’s transformation:
-$T_{\hat{\rho}} = ln(1 - \hat{\rho})$. This transformation has
-variance-stabilising properties and therefore allows for adequate
-inferences concerning heterogeneity in score reliability [@bonett2002].
-Cochran’s Q-test is used to identify statistically significant
-heterogeneity in transformed score reliability. Estimates derived from a
-Reliability Generalization Meta-Analysis using these transformations can
-be back-transformed to the original score reliability scale
-[@beinhauer2025]:
+
+$$T_{\hat{\rho}} = ln(1 - \hat{\rho})$$ {#eq-Bonett}
+
+The transformation in @eq-Bonett has variance-stabilising properties and
+therefore allows for adequate inferences concerning heterogeneity in
+score reliability [@bonett2002]. Cochran’s Q-test is used to identify
+statistically significant heterogeneity in transformed score
+reliability. Estimates derived from a Reliability Generalization
+Meta-Analysis using these transformations can be back-transformed to the
+original score reliability scale [@beinhauer2025]:
 
@@
 an example from the 21 selected phenomena, the results from this
-procedure on replications on Nosek’s study on sex differences in
+procedure on replications of Nosek’s study on sex differences in
 preferences for the subject of Arts is displayed in the forest plot in
@@
 #| label: fig-forest 
-#| fig-cap: Forest Plot ML1 - Nosek (Art)
+#| fig-cap: Forest Plot ML1 - replications of Nosek's studz on sex differences in subject preferences (Arts)
 forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
@@
 procedure leads to an increase in the individual absolute effect size,
-as the black dots are moved further away from zero. Simultaneously, the
+as the black dots are moved further from zero. Simultaneously, the
 standard errors grew larger after the attenuation correction, which
 leads to larger 95%-CIs. This is most easily observed for the
-mturk-sample, which already had a rather large confidence interval to
-begin with. Additionally, in the diamond at the bottom of the figure, we
-can see that the meta-analytic estimate mean ES is also larger after the
-attenuation correction took place. In @tbl-meanES, the estimates and
-tests on the meta-analytic average ES, both corrected and uncorrected,
-can be found.
+MTurk-sample, which had a rather large confidence interval to begin
+with. Additionally, in the diamond at the bottom of the figure, we can
+see that the meta-analytic mean ES is also larger after the attenuation
+correction took place. @tbl-meanES contains estimates and test results
+for all phenomena, both corrected and uncorrected.
 
@@
   gt::cols_label(MASC = "MASC",
-                 mu_str_raw = gt::md("$\\hat{\\mu}_{\\delta_0}$ (SE)"),
+                 #mu_str_raw = gt::md("$\\hat{\\mu}_{\\delta_0}$ (SE)"),
+                 mu_str_raw = gt::md("Mean $d$ (SE)"),
                  pval_raw = "p",
-                 mu_str_cor = gt::md("$\\hat{\\mu}_{\\delta}$ (SE)"),
+                 #mu_str_cor = gt::md("$\\hat{\\mu}_{\\delta}$ (SE)"),
+                 mu_str_cor = gt::md("Mean $d_c$ (SE)"),
                  pval_cor = "p") %>% 
@@
 
-*Note:* Since in a meta-analysis, we attempt to estimate mean and
-variance of the parameter's true distribution, in this table
-$\hat{\mu}_{\delta_0}$ represents the meta-analytic estimate of mean
-uncorrected $d$ and $\hat{\mu}_{\delta}$ represents the estimate of the
+*Note:* In this table *Mean* $d$ represents the meta-analytic estimate
+of mean uncorrected $d$ and *Mean* $d_c$ represents the estimate of the
 meta-analytic mean of corrected $d_c$.

### 2025-09-21 Manuscript_MD_ScoreReLiability.qmd
update manuscript up to appendix

@@
 
+The estimates concerning the meta-analytic mean and heterogeneity of
+these standardized effect sizes attempt to describe their population
+distribution. Therefore, over the following paragraphs, whenever we
+refer to estimates derived by means of random effects meta-analysis, we
+will use the Greek letter $\delta_0$ to denote estimates of uncorrected
+$d$'s population distribution and $\delta$ to denote estimates of
+corrected $d_c$'s population distribution.
+
 ### Methods
@@
 
+In @eq-rel_back_mu, $\mu_{\rho_{XX'}}$ is the back-transformed
+meta-analytic mean score reliability - on the original coefficient's
+scale between 0 and 1. Similarly, $\tau_{\mu_{\rho_{XX'}}}$ is the
+back-transformed heterogeneity in score reliability:
+
 $$\tau_{\rho_{xx'}}^2 \approx exp(\mu_{T\left[\rho_{xx'}\right]})2 \tau_{T\left[\rho_{xx'}\right]}^2+exp(\mu_{T\left[\rho_{xx'}\right]})^2\tau_{T\left[\rho_{xx'}\right]}^4$$ {#eq-rel_back_tau}
@@
 heterogeneity for both uncorrected ES ($\delta_0$) and corrected ES
-($\delta$), which are central to assess the research questions posed in
+($\delta$), which are central to assess the research question posed in
 section *Attenuation Correction*. Most importantly, @tbl-tauES
@@
 heterogeneity. However, while the heterogeneity for the phenomenon from
-RRR9 (Behaviour) seems reduced by about a third of its initial value,
-the ES heterogeneity for this phenomenon is rather small to begin with.
-Using a significance test, we can not distinguish this value confidently
-from zero. This implies that for the majority of phenomena, differences
-in score reliability can not be made responsible for heterogeneity in
-standardized effect sizes.
+RRR9 (Behaviour) seems to be reduced by about a third of its initial
+value, the ES heterogeneity for this phenomenon is rather small to begin
+with. Using a significance test, we can not distinguish this value
+confidently from zero. This implies that for the majority of phenomena,
+differences in score reliability can not be made responsible for
+heterogeneity in standardized effect sizes.
 
@@
 *Note:* $\hat{\mu}_{\rho_{XX'}}$ is the back-transformed meta-analytic
-estimate of mean score reliability; $\tau{\mu}_{\rho_{XX'}}$ is the
-back-transformed estimate of heterogeneity in score reliability; Note
-that, since in a meta-analysis, we attempt to estimate mean and variance
-of the parameter's true distribution, in this table
+estimate of mean score reliability; $\tau_{\mu_{\rho_{XX'}}}$ is the
+back-transformed estimate of heterogeneity in score reliability;
 $\hat{\tau}_{\delta_0}$ represents the estimate of heterogeneity in
-uncorrected $d$; accordingly, $\hat{\tau}_{\delta}$ represents the
-estimate of the heterogeneity in corrected $d_c$.
+uncorrected $d$; $\hat{\tau}_{\delta}$ represents the estimate of the
+heterogeneity in corrected $d_c$.
 
@@
 heterogeneity grows as a result from the attenuation correction
-procedure. Even though these differences tend to be small, it is clear
-that in all those cases, not only did accounting for score reliability
-not explain any ES heterogeneity, we find even more than initially
-expected. This is incompatible with claims made in W&D and H&S, the
-attenuation correction increased ES heterogeneity. Concerning the
+procedure. However, it needs to be noted that these differences are small
+and could not be distinguished from sampling uncertainty in a statistical
+test. For this reason, we do not mean to interpret these differences, while 
+we note that the direction of differences are clearly inconsistent with
+expectations formulated in the literature discussed so far.
+These results are incompatible with claims made in W&D and H&S.
+
+Besides the direction of these differences, it is noteworthy that considering
+differences in score reliability barely affects heterogeneity in standardized
+effect sizes. [ausführen, oder kommt das noch einmal wieder?]
+
+Concerning the
 remaining two phenomena, for the phenomenon described in ML5 - Shnabel
@@
 
-## Relationship between score reliability and effect sizes in terms of heterogeneity
+## Relationship between score reliability and effect size heterogeneity
 
@@
 demonstrated that claims formulated in W&D and H&S regarding how score
-reliability increases ES heterogeneity are false. We found a large
+reliability heterogeneity increases ES heterogeneity are false. We found a large
 number of phenomena where, contrary to popular claims, correcting for
@@
 It might help to think about linear models, when discussing the
-intuition that might underlie the idea, that controlling for score
+intuitive idea that controlling for score
 reliability should reduce ES heterogeneity. Heterogeneity describes
@@
 variance, as we observe it in a simple linear model. Both describe
-variation that could not be explained by the predictive variables in the
-case of the simple linear model, or not explained by incorporating the
-Standard Errors in the case of a meta-analytic model.
-
-For the simple linear model, if we add an additional predictor to the
-model that we know shares variance with the parameter that we aim to
-predict, this should reduce the amount of residual variance that we
-observe across most instances. Generally, removing a source of variance
+variation that could not be explained by the model specifications, be it a model that 
+employs predictive variables like a simple linear model or a meta-analytic
+model incorporating uncertainty in terms of standard errors.
+
+The residual variance that remains in a simple linear model will be reduced
+in all instances, where we use a predictor that we know shares 
+variance with the predicted parameter. Generally, removing a source of variance
 by controlling for it, we expect any unexplained variation to shrink.
@@
 $\delta_0$, carrying a 0 in the index, refers to the uncorrected ES,
-undistorted by sampling error but not corrected for measuring error. On
-the other hand $\delta$, no index, refers to the corrected ES,
+undistorted by sampling error but not corrected for measuring error. Without an 
+index, $\delta$, on the other hand, refers to the corrected ES,
 undistorted by sampling error and measurement error. Lastly,
@@
 higher-order moments are negligible (e.g. limited variability and
-monotonicity) (still need a good reference here). Here we use the
+monotonicity) (Asymptotic statistics, A.W. van der Vaart - find book online). Here we use the
 Taylor-estimator of the ratio’s variance to demonstrate how differences
@@
 In @eq-rel we see that total standard deviation $\sigma_X$ is
-constructed only from (the square of the sum of) error score variance
+constructed only from error score variance
 $\sigma_E^2$ and true score variance $\sigma_T^2$. Measurement
-precision, since this should not pertain the true scores, should only
+precision should generally only
 affect the error score variance. Therefore, if the heterogeneity in
@@
 in the re-analysis of archival data, we can already expect that this can
-not be an implication of these \[Equations @eq-ratio_d0\] and
-[-@eq-ratio_d].
+not be an implication of these Equations.
 
 Instead, what is guaranteed is that the expected value of $\sigma_T$ is
-smaller than the expected value of $\sigma_X$. Similarly, the
-heterogeneity in $\sigma_T$ is guaranteed to be smaller than the
+smaller than or equal to the expected value of $\sigma_X$. Similarly, the
+heterogeneity in $\sigma_T$ is guaranteed to be smaller than or equal to the
 uncorrected heterogeneity in $\sigma_X$. However, since the expected
@@
 conditions under which the claims made in W&D and H&S are met.
-Therefore, this inequality is essentially the opposite of the claims
-made in W&D and H&S. Unfortunately, the inequality in this form does not
+Unfortunately, the inequality in this form does not
 convey sufficient information about the conditions under which it can
@@
 $\mu_{\rho_{xx'}}$ do not share not the same value, in practice we
-observe little to no differences between the two parameters. Large
+observe little to no differences between the two parameters. [add to appendix]. Large
 values indicate that correcting the individual score variances (or
@@
 larger amount of random error score variance, leading to a stronger
-correction.
+correction. The metric $R_2$ on the other hand describes in how far the attenuation
+correction procedure has successfully reduced heterogeneity in score
+variance $\tau^2_{\sigma^2_X}$.
 
@@
 
-The metric $R_2$ on the other hand describes in how far the attenuation
-correction procedure has successfully reduced heterogeneity in score
-variance $\tau^2_{\sigma^2_X}$. A value of 1 indicates that the
-heterogeneity in true score variance $\sigma^2_T$ is essentially just as
+A value of 1 indicates that the
+heterogeneity in true score variance $\sigma^2_T$ is exactly as
 large as the heterogeneity initially observed. Smaller values indicate
@@
 Concerning the re-arranged inequality from Appendix A, we identify one
-necessary but not sufficient condition or rule described in
+sufficient but not necessary condition or rule described in
 @eq-ineq_short.
@@
 
-With necessary but not sufficient we mean that, as long as
+With sufficient but not necessary we mean that, as long as
 @eq-ineq_short holds, an attenuation correction will inevitably increase
@@
 heterogeneity. Whether ES heterogeneity can be decreased depends on
-other parameters, that we left out of this short form, but can be found,
-accompanied by a more detailed description, in @eq-ineq_fin.
+additional parameters, that we left out of this short form, but can be found,
+accompanied by a more detailed description, in @eq-ineq_fin, Appendix A.
 
@@
 change in mean value and heterogeneity due to the correction: ES
-heterogeneity will always be reduced, as long as the remaining
+heterogeneity will always be increased, as long as the remaining proportion of
 heterogeneity in score variance is not substantially smaller than the
-size of the remaining mean value.
+remaining proportion of its mean value.
 
@@
 score variance $R_2$ is unlikely to be substantially smaller than $R_1$.
+[Na ja, eine Reliabilität von .9 bedeutet erstmal, dass es 9mal so viel wahre Varianz wie Fehlervarianz gibt.
+In diesem Fall müsste R2 unter 0.73 liegen.
+Das impliziert in der Heterogenität ein Verhältnis von ungefähr 3 : 1.
+Die 9mal größere wahre Varianz müsste also für weniger als ¾ der Heterogenität verantwortlich sein.
+In Deinem Textbeispiel (Reliabilität von .8) müsste die 4mal größere wahre Varianz für weniger als die Hälfte der Heterogenität verantwortlich sein.
+– Das erscheint mir beides per se ziemlich unplausibel… Also auch dann, wenn die „administration sites do NOT carry large differences in true score variance”.
+– In anderen Worten: Die Frage ist, ob die absolute Größe der Varianz überhaupt irgendeine Rolle spielt oder ob alles lediglich an den Verhältnissen hängt, die durch die Reliabilität vorgegeben sind.
+Ich habe dazu aber auch keine eindeutige Antwort… Wenn ich überzeugt wäre, dass sich *homogene* true score variances durch planvolle, kontrollierte Stichprobenziehung herstellen lassen, wäre das ein Gegenargument.]
 
@@
 analytical arguments hold up in empirical assessment, but also
-demonstrate how the metrics $R_1$ and $R_2$ may be used to inform how
-score reliability affected ES heterogeneity.
+demonstrate how the metrics $R_1$ and $R_2$ help understand how
+score reliability affected ES heterogeneity in practice.
 
@@
 
-For the following analysis, Boot-Err is used. As Boot-Err makes use of
+As Boot-Err makes use of
 meta-analytic techniques and bootstrap sampling, we perform analyses
@@
 variances, including their heterogeneity and metrics $R_1$ and $R_2$
-across all 12 phenomena. In @tbl-R1R2 we see that for ten out of 12
+across all 12 phenomena. We see that for ten out of 12
 phenomena, the $R_2$ metric is larger than the $R_1$ metric.
@@
 
-For the two exceptions, those phenomena where the attenuation correction
+For the two exceptions in @tbl-tauES, those phenomena where the attenuation correction
 procedure actually led to a decrease in ES heterogeneity (RRR9
-(Behaviour) and ML5 - Shnabel (RPP)), the significance test for the ES
+(Behaviour) and ML5 - Shnabel (RPP)), we find that $R_2$ is larger than $R_1$.
+This clearly contradicts our interpretation of @eq-ineq_int. As long as $R_2$ is
+larger than $R_1^3$, we expect a correction procedure to increase effect 
+size heterogeneity. However, we recommend to be cautious when interpreting both the change and
+the direction of said change for these two projects. 
+The significance test for the ES
 heterogeneity was not significant, both before and after the correction.
-The results regarding $R_2$ indicate that the majority of score variance
-heterogeneity was due to differences in true score variance
-$\sigma^2_T$. Since either the heterogeneity was so small or the sample
+Since either the heterogeneity was so small or the sample
 size so low that the heterogeneity could not be statistically
-distinguished from zero, we should be cautious in the interpretation of
-a reduction in that estimate. Most likely, the reduction in
+distinguished from zero, most likely, the reduction in
 heterogeneity is better explained by imprecise estimation than an actual
@@
 
-Lastly, for two out of the 12 phenomena, the one described in ML5 -
-Albarracin and the other described in ML5 - Shnabel (Rev), we find
+Lastly, for two out of the 12 phenomena, ML5 -
+Albarracin and ML5 - Shnabel (Rev), we find
 $\tau_{\sigma^2_X}$ of 0. Therefore, the $R_2$ metric could not be
 computed, as the denominator would have been zero. However, we also did
-not identify statistically significant ES heterogeneity in the first
-place. The phenomena were assessed with a particularly low number of
+not identify statistically significant heterogeneity in score reliability
+across replications, nor ES heterogeneity. The phenomena were assessed with a 
+particularly low number of
 replications (9 and 8 respectively). Estimation precision for
 heterogeneity in variance components, even more so than for ES, largely
-depends on the sample size. Most likely, the low sample size led to
+depends on the sample size. Two alternative explanations for this observation
+are available: either, there is no heterogeneity in score variances across 
+replications. In that case, there is also no heterogeneity that could be
+broken down into true and error score variance heterogeneity. Alternatively,
+the low sample size led to
 implausible estimates in both true and total score variance
@@
 Meta-analytic estimates, especially those concerning score variance
-components such as $R_2$, therefore should be interpreted with caution,
+components such as $R_2$, should generally be interpreted with caution,
 if the sample size was particularly low.
 
-Generally, @tbl-R1R2 implies that for all measurements concerning these
-phenomena, the heterogeneity in observed score variances could not be
+Generally, @tbl-R1R2 implies that for none of the measurements concerning these
+phenomena, the heterogeneity in observed score variances could be
 substantially explained by differences in measurement quality. For all
@@
 correcting for differences in reliability coefficients using an
-attenuation correction procedure. We found that a) in the majority of
-cases, the change in ES heterogeneity is small to non-existent. This
-implies that, across the data-sets we got to examine, differences in
+attenuation correction procedure. While we initially observed heterogeneity
+in score reliability across almost all measuring scales, the extent of this
+heterogeneity was very small across all scales. This initial finding already
+implies that differences in score reliability can only have a tiny impact
+on ES heterogeneity, if any.
+
+This implication fits our subsequent finding, as we found that a) in the majority of
+cases, the change in ES heterogeneity is small to non-existent. Across the 
+data-sets we got to examine, differences in
 score reliability do not serve as an appropriate explanation for ES
@@
 estimates of ES heterogeneity. This means that the claims found in W&D
-and H&S, implying that an attenuation correction can reduce ES
+and H&S, implying that an attenuation correction will reduce ES
 heterogeneity, stem from a general misunderstanding of how score
@@
 differences in how the latent variable is distributed (in terms of
-spread) across the administration sites.
-
-In the examples we re-analysed throughout this text, as well as the
-majority of research studies done in the behavioural sciences
-[@hultsch2002; @bornstein2013], no proper sampling techniques are
-employed. Most samples where these instruments are applied come to be
+spread) across the administration sites. This disparity grows even larger, 
+the higher the average score reliability is.
+
+In the 
+majority of research studies done in the social and behavioural sciences
+[@hultsch2002; @bornstein2013; @ashraf2017; @sears1986; @zhao2021], no 
+proper sampling techniques beyond convenience sampling are
+employed. Most samples re-analysed throughout this text came to be
 through convenience sampling, where students self-select into studies.
-Without a valid sampling scheme, we essentially have have no basis to
+Without a valid sampling scheme, we essentially have no basis to
 assume anything about the true score variance [@hultsch2002]. Most
@@
 corrected for, the attenuation correction has a very small effect on
-heterogeneity estimates in score variance. Therefore, as indicated by
-our inequality, an attenuation correction is likely to have very little
+heterogeneity estimates in score variance. Therefore, an attenuation 
+correction is likely to have very little
 impact on ES heterogeneity. If anything, it most likely increases it.
@@
 as a moderator explaining ES heterogeneity.
-
-If we, in the field of Psychology and other social or behavioural
-sciences, keep making use of convenience samples simply because they are
-so convenient, we can not expect anything to change. Most likely, ES
-heterogeneity will remain a substantial factor hampering the
-interpretability of original and replication studies. Instead, we should
-focus our data collection efforts on the generation of samples, where we
-know that the variance in the sample means something. This is most
-easily achieved by proper sampling techniques, which ensure that the
-sample is representative for some population. In such cases, it might be
-that the true score variance varies less strongly than what we currently
-observe. And should the heterogeneity in true score variance remain as
-large, than we know that this is meaningful information from which we
-can draw meaningful inferences. Therein, we could learn more about
-standardized effect sizes and why our phenomena vary across settings,
-populations or continents by not following the current status quo of
-convenience samples. Until this status quo changes, however, we will
-keep wasting valuable resources by attempts of discussing ES
-heterogeneity or assessing uninformative differences in score
-reliability.
 
@@
 
-### Conclusions
+## Conclusions
 
 Heterogeneity in ES even in direct replications, where experimental
-factors are held as constant as possible, is already substantial
-(Renkewitz, Fünderich & Beinhauer, in preparation). However, so far we
+factors are held as constant as possible, is substantial
+(Renkewitz, Fünderich & Beinhauer, in preparation). However, we
 have demonstrated that in several scenarios, differences in score
@@
 phenomena.
+
+### Where to go from here
+
+This manuscript has demonstrated that, even though differences in
+measurement precision are present across most phenomena we could re-analyse, 
+these are strongly outweighed by the differences in true score variance.
+However, what a true score variance actually means, strongly depends on 
+how that number came to be. A sample that is very diverse in its composition
+likely comes with a large true score variance. On the other hand, a sample
+that is highly homogeneous in its composition comes with a smaller true score
+variance. Importantly, the same is not true for random measurement error. 
+Error variance should always be the same size, if the errors are truly random,
+no matter how the sample is put together. Unfortunately, 
+most phenomena we re-analysed here came to be through means of convenience sampling. 
+Therefore, it is not clear what the large differences in true score variance 
+actually imply - are they expressions of
+genuine differences between the populations, or just expressions of different
+self-selection procedures across the samples? As the variance observed in
+a convenience sample is not representative to some *specificied* population, 
+we simply can not know.
+
+Issues of generalizability
+due to the lack of sampling techniques have been, and are still, repeatedly
+called out for the same reasons. Our findings therein reiterate these calls. 
+We believe that we should
+focus our data collection efforts on the generation of samples, where we
+know that the variance in a sample is actually meaningful. This is most
+easily achieved by proper sampling techniques, which ensure that the
+sample is representative for some specified population. In such cases, it might be
+that the true score variance varies less strongly than what we currently
+observe. And should the heterogeneity in true score variance remain as
+large, than we know that this is crucial information from which we
+can draw meaningful inferences. We could learn more about
+standardized effect sizes and why our phenomena vary across settings, 
+populations or continents. We strongly believe that the valuable resources 
+invested into
+research output could be put towards more informative use than what seems
+to be the status quo by ensuring that the variances observed in different
+samples are representative for specified populations.
 
@@
 However, based on our findings, we believe that the issues created by
-ignoring random sampling is much more dire and warrants immediate
+ignoring random sampling are much more dire and warrant immediate
 attention. Only once we can be sure that the variance in true scores or

### 2025-10-20 Manuscript_MD_ScoreReLiability.qmd
update manuscript - add abstracts

@@
 ---
-title: "Manuscript"
+title: "Differences in score reliability do not explain meta-analytic heterogeneity in standardised effect sizes"
 bibliography: ReLiability_references.bib
 csl: styles/apa-6th-edition.csl
-format: pdf
+format: docx
 editor: 
@@
 
-# Liability or reliability? Exploring the role of effect size attenuation on meta-analytic heterogeneity
+#
 
@@
 https://orcid.org/0000-0001-8072-6802
+
+#### Abstract
+
+Heterogeneity in standardized effect sizes, such as Cohen’s d, is informative because it indicates that a psychological phenomenon is not yet fully understood.
+Prior work by Hunter and Schmidt as well as Wiernik and Dahlke has argued that such heterogeneity partly reflects differences in measurement precision and that correcting for these differences using attenuation correction should therefore reduce heterogeneity.
+We reanalyze data from large-scale collaborative projects, including the ManyLabs studies, Registered Replication Reports, and the Psychological Science Accelerator, and demonstrate that this assumption does not hold universally.
+Treating standardized effect sizes as ratio variables, we offer an alternative account of how differences in measurement precision relate to observed heterogeneity.
+Across many psychological research contexts, score reliability tends to be high and measurement error relatively stable.
+Under these conditions, when samples differ substantially in their composition or diversity, attenuation correction procedures are more likely to increase rather than decrease heterogeneity.
+We conclude by discussing the limited relevance of random measurement error for explaining effect size heterogeneity in the context of convenience sampling.
+
+#### Non-technical abstract
+Researchers often find that studies on the same psychological effect report different results, known as differences in effect sizes.
+Such variation suggests that the phenomenon is not yet fully understood.
+Some have argued that part of this variation comes from differences in measurement precision—how reliably a study’s tools or tests measure what they intend to measure.
+Based on this reasoning, it has been suggested that adjusting or “correcting” for these differences should make results more consistent across studies.
+We revisited this assumption by looking at large collaborative projects in psychology, such as ManyLabs, the Registered Replication Reports, and the Psychological Science Accelerator.
+We found that the expected reduction in variation does not generally occur.
+In fact, when measurement precision is already quite high - as it usually is in psychology, where researchers carefully select reliable measures - but samples differ widely in who participates, such corrections can actually increase differences in reported effect sizes.
+In simpler terms, differences in how diverse the samples of participants are across studies (which comes about from convenience sampling) matters more for explaining variation in results than small differences in how precisely a measure captures a construct.
+We introduce a simple rule that identifies when correcting for measurement differences is likely to backfire, increasing rather than reducing heterogeneity.
+We conclude by discussing why random measurement error plays a smaller role in explaining effect size differences than the use of diverse, convenience-based samples.
 
@@
 to a specific theory, might be impairments in score reliability which
-vary in their extent across replications. \[This is an argument put
-forward repeatedly by original researchers and replicators alike, such
-as Olsson-Collentine et al. [-@olsson2020] (do we know any more?).\]
-Hunter and Schmidt [-@hunter2004] and Wiernik and Dahlke
-[-@wiernik2020], discussing the role of score reliability for
-standardized effect sizes, claim that differences in score reliability
-should inflate, and therefore increase, heterogeneity. This implies that
-standard corrections for imperfect score reliability, such as
-attenuation correction procedures, should remove that inflation,
-reducing heterogeneity in corrected effect sizes.
+vary in their extent across replications [@olsson2020]. Hunter and
+Schmidt [-@hunter2004] and Wiernik and Dahlke [-@wiernik2020],
+discussing the role of score reliability for standardized effect sizes,
+claim that differences in score reliability should inflate, and
+therefore increase, heterogeneity. This implies that standard
+corrections for imperfect score reliability, such as attenuation
+correction procedures, should remove that inflation, reducing
+heterogeneity in corrected effect sizes.
 
@@
 known as attenuation correction and, while not without its criticisms
-[@muchinsky1996; @schmidt1996; @schennach2016], is a widespread practice
-in Psychology (reference, alt.: repeatedly recommended + references).
-@eq-dc describes a simple attenuation correction procedure.
+[@muchinsky1996; @schmidt1996; @schennach2016], is a repeatedly
+recommended practice in Psychology [@hunter2004, @wiernik2020] . @eq-dc
+describes a simple attenuation correction procedure.
 
@@
 potentially be explained by these differences in score reliability.
-Similarly, attempting to explain the substantial extent of ES
-heterogeneity identified in psychological phenomena, both the original
-authors as well as the multi-site replication authors claim that
-differences in score reliability might be responsible (again, do we
-actually know of any?).
+Similarly, in a re-analysis of multi-site replication projects,
+Olsson-Collention et al. [-@olsson2020] discuss differences in score
+reliability as a potential moderator for ES heterogeneity.
 
@@
 the different phenomena and how they were measured can be found at
-\[osf-link\], while a brief summary can be found in @tbl-data.
+\[https://osf.io/g2fzq\], while a brief summary can be found in appendix C.
 
@@
 estimates of heterogeneity are larger than zero can we actually begin to
-interpret whether there was any change. \[Short paragraph discussing
-that a test for difference in heterogeneity would be more informative?
-Maybe move it to the discussion section - hint at it here?\]
+interpret whether there was any change.
 
@@
 sample, while black dots represent the corrected ES in each sample. The
-bars surrounding these dots show the respective 95%-Cis.
+bars surrounding these dots show the respective 95%-CIs.
 
@@
 #| label: fig-forest 
-#| fig-cap: Forest Plot ML1 - replications of Nosek's studz on sex differences in subject preferences (Arts)
+#| fig-cap: Forest Plot ML1 - replications of Nosek's study on sex differences in subject preferences (Arts)
 forest_plot_rel(ES_rma_df[which(ES_rma_df$MASC == "Nosek_Explicit_Art"),], agg_L[[which(MASC_names == "Nosek_Explicit_Art")]])
 ```
+
+*Note:* Grey dots represent uncorrected estimates of $d$, while black
+dots represent corrected estimates of $d_c$. Similarly, grey bars
+visualise the 95%-CI for uncorrected estimates, black bars the 95%-CIs
+for corrected estimates. The grey diamond represents the meta-analytic
+estimate of uncorrected ES $\mu_{d}$, its width represents the 95%-CI of
+said estimate. The black diamond represents the same parameters for the
+meta-analytic estimate of corrected ES $\mu_{d_c}$,.
 
@@
   gt::cols_label(MASC = "MASC",
-                 #mu_str_raw = gt::md("$\\hat{\\mu}_{\\delta_0}$ (SE)"),
-                 mu_str_raw = gt::md("Mean $d$ (SE)"),
+                 mu_str_raw = gt::md("$\\hat{\\mu}_{d}$ (SE)"),
+                 #mu_str_raw = gt::md("Mean $d$ (SE)"),
                  pval_raw = "p",
-                 #mu_str_cor = gt::md("$\\hat{\\mu}_{\\delta}$ (SE)"),
-                 mu_str_cor = gt::md("Mean $d_c$ (SE)"),
+                 mu_str_cor = gt::md("$\\hat{\\mu}_{d_c}$ (SE)"),
+                 #mu_str_cor = gt::md("Mean $d_c$ (SE)"),
                  pval_cor = "p") %>% 
@@
 
-*Note:* In this table *Mean* $d$ represents the meta-analytic estimate
-of mean uncorrected $d$ and *Mean* $d_c$ represents the estimate of the
-meta-analytic mean of corrected $d_c$.
+*Note:* In this table $\hat{\mu}_{d}$ represents the meta-analytic
+estimate of uncorrected $d$ and $\hat{\mu}_{d_c}$ represents the
+estimate of the meta-analytic mean of corrected $d_c$.
 
@@
 remaining phenomena. Additionally, it contains estimates of
-heterogeneity for both uncorrected ES ($\delta_0$) and corrected ES
-($\delta$), which are central to assess the research question posed in
-section *Attenuation Correction*. Most importantly, @tbl-tauES
-demonstrates that for all but one of the phenomena, the attenuation
-correction procedure barely has an impact on the extent of ES
-heterogeneity. However, while the heterogeneity for the phenomenon from
-RRR9 (Behaviour) seems to be reduced by about a third of its initial
-value, the ES heterogeneity for this phenomenon is rather small to begin
-with. Using a significance test, we can not distinguish this value
-confidently from zero. This implies that for the majority of phenomena,
-differences in score reliability can not be made responsible for
-heterogeneity in standardized effect sizes.
+heterogeneity for both uncorrected ES ($d$) and corrected ES ($d_c$),
+which are central to assess the research question posed in section
+*Attenuation Correction*. Most importantly, @tbl-tauES demonstrates that
+for all but one of the phenomena, the attenuation correction procedure
+barely has an impact on the extent of ES heterogeneity. However, while
+the heterogeneity for the phenomenon from RRR9 (Behaviour) seems to be
+reduced by about a third of its initial value, the ES heterogeneity for
+this phenomenon is rather small to begin with. Using a significance
+test, we can not distinguish this value confidently from zero. This
+implies that for the majority of phenomena, differences in score
+reliability can not be made responsible for heterogeneity in
+standardized effect sizes.
 
@@
                  QEp = "p",
-                 tau_raw = gt::md("$\\hat{\\tau}_{\\delta_0}$"),
+                 tau_raw = gt::md("$\\hat{\\tau}_{d}$"),
                  QE_raw_str = "QE (df)",
                  QEp_raw = "p",
-                 tau_raw = gt::md("$\\hat{\\tau}_{\\delta}$"),
+                 tau_raw = gt::md("$\\hat{\\tau}_{d_c}$"),
                  QE_raw_str = "QE (df)",
@@
 back-transformed estimate of heterogeneity in score reliability;
-$\hat{\tau}_{\delta_0}$ represents the estimate of heterogeneity in
-uncorrected $d$; $\hat{\tau}_{\delta}$ represents the estimate of the
-heterogeneity in corrected $d_c$.
+$\hat{\tau}_{d}$ represents the estimate of heterogeneity in uncorrected
+$d$; $\hat{\tau}_{d_c}$ represents the estimate of the heterogeneity in
+corrected $d_c$.
 
@@
 heterogeneity grows as a result from the attenuation correction
-procedure. However, it needs to be noted that these differences are small
-and could not be distinguished from sampling uncertainty in a statistical
-test. For this reason, we do not mean to interpret these differences, while 
-we note that the direction of differences are clearly inconsistent with
-expectations formulated in the literature discussed so far.
-These results are incompatible with claims made in W&D and H&S.
-
-Besides the direction of these differences, it is noteworthy that considering
-differences in score reliability barely affects heterogeneity in standardized
-effect sizes. [ausführen, oder kommt das noch einmal wieder?]
-
-Concerning the
-remaining two phenomena, for the phenomenon described in ML5 - Shnabel
-(RPP), the reduction in heterogeneity is miniscule and the estimate of
-heterogeneity is not statistically significant in the first place. Most
-likely, the difference in ES heterogeneity estimates can be explained by
-issues of estimation, as only eight replications are available.
-Similarly, only for the phenomenon described in RRR9 (Behaviour), we
-find a seemingly substantial reduction in heterogeneity due to the
-attenuation correction procedure. The estimates of ES heterogeneity are
-not statistically significant, however, which hampers our ability to
-meaningfully interpret and compare these estimates. For the majority of
-phenomena it appears that differences in score reliability do not
-inflate heterogeneity in uncorrected effect sizes. Instead these
-differences mask ES heterogeneity, that can be unveiled by accounting
-for score reliability.
+procedure. However, it needs to be noted that these differences are
+small and could not be distinguished from sampling uncertainty in a
+statistical test. For this reason, we do not mean to interpret these
+differences, while we note that the direction of differences are clearly
+inconsistent with expectations formulated in the literature discussed so
+far. These results are incompatible with claims made in W&D and H&S.
+
+Generally, however, it is noteworthy that removing differences in score
+reliability only had a very little impact on ES heterogeneity. It
+appears that, across the phenomena assessed in this project, the
+heterogeneity before and after the correction procedure barely changes.
+At least for the projects re-analysed here, differences in score
+reliability do not affect ES heterogeneity as strongly as assumed.
+
+Concerning the remaining two phenomena, for the phenomenon described in
+ML5 - Shnabel (RPP), the reduction in heterogeneity is miniscule and the
+estimate of heterogeneity is not statistically significant in the first
+place. Most likely, the difference in ES heterogeneity estimates can be
+explained by issues of estimation, as only eight replications are
+available. Similarly, only for the phenomenon described in RRR9
+(Behaviour), we find a seemingly substantial reduction in heterogeneity
+due to the attenuation correction procedure. The estimates of ES
+heterogeneity are not statistically significant, however, which hampers
+our ability to meaningfully interpret and compare these estimates. For
+the majority of phenomena it appears that differences in score
+reliability do not inflate heterogeneity in uncorrected effect sizes.
+Instead these differences mask ES heterogeneity, that can be unveiled by
+accounting for score reliability.
 
@@
 demonstrated that claims formulated in W&D and H&S regarding how score
-reliability heterogeneity increases ES heterogeneity are false. We found a large
-number of phenomena where, contrary to popular claims, correcting for
-score reliability increases ES heterogeneity, instead of decreasing it.
+reliability heterogeneity increases ES heterogeneity are false. We found
+a large number of phenomena where, contrary to popular claims,
+correcting for score reliability increases ES heterogeneity, instead of
+decreasing it.
 
@@
 It might help to think about linear models, when discussing the
-intuitive idea that controlling for score
-reliability should reduce ES heterogeneity. Heterogeneity describes
-unexplained variation in a parameter's true distribution, that we model
-using meta-analytical techniques [@borenstein2009; @higgins2002]. In
-that way, the idea of heterogeneity is closely related to residual
-variance, as we observe it in a simple linear model. Both describe
-variation that could not be explained by the model specifications, be it a model that 
-employs predictive variables like a simple linear model or a meta-analytic
-model incorporating uncertainty in terms of standard errors.
-
-The residual variance that remains in a simple linear model will be reduced
-in all instances, where we use a predictor that we know shares 
-variance with the predicted parameter. Generally, removing a source of variance
-by controlling for it, we expect any unexplained variation to shrink.
-Similarly, we know that differences in score reliability, by definition,
-are related to differences in standardized effect sizes. Therefore,
-intuitively, it seems sensible to expect that controlling for
+intuitive idea that controlling for score reliability should reduce ES
+heterogeneity. Heterogeneity describes unexplained variation in a
+parameter's true distribution, that we model using meta-analytical
+techniques [@borenstein2009; @higgins2002]. In that way, the idea of
+heterogeneity is closely related to residual variance, as we observe it
+in a simple linear model. Both describe variation that could not be
+explained by the model specifications, be it a model that employs
+predictive variables like a simple linear model or a meta-analytic model
+incorporating uncertainty in terms of standard errors.
+
+The residual variance that remains in a simple linear model will be
+reduced in all instances, where we use a predictor that we know shares
+variance with the predicted parameter. Generally, removing a source of
+variance by controlling for it, we expect any unexplained variation to
+shrink. Similarly, we know that differences in score reliability, by
+definition, are related to differences in standardized effect sizes.
+Therefore, intuitively, it seems sensible to expect that controlling for
 differences in score reliability should reduce the unexplained variance
@@
 demonstrates how heterogeneity in score reliability affects
-heterogeneity in ESs [-@hunter2004, p. XX]. @eq-HS implies that
+heterogeneity in ESs [-@hunter2004, p. 309]. @eq-HS implies that
 heterogeneity in $\delta_0$ is essentially a function in which
@@
 $\delta_0$, carrying a 0 in the index, refers to the uncorrected ES,
-undistorted by sampling error but not corrected for measuring error. Without an 
-index, $\delta$, on the other hand, refers to the corrected ES,
-undistorted by sampling error and measurement error. Lastly,
+undistorted by sampling error but not corrected for measuring error.
+Without an index, $\delta$, on the other hand, refers to the corrected
+ES, undistorted by sampling error and measurement error. Lastly,
 $\rho_{XX'}$ refers to score reliability. Hunter & Schmidt
@@
 higher-order moments are negligible (e.g. limited variability and
-monotonicity) (Asymptotic statistics, A.W. van der Vaart - find book online). Here we use the
-Taylor-estimator of the ratio’s variance to demonstrate how differences
-in score reliability may affect heterogeneity in ES.
+monotonicity) (Asymptotic statistics, A.W. van der Vaart - find book
+online). Here we use the Taylor-estimator of the ratio’s variance to
+demonstrate how differences in score reliability may affect
+heterogeneity in ES.
 
@@
 In @eq-rel we see that total standard deviation $\sigma_X$ is
-constructed only from error score variance
-$\sigma_E^2$ and true score variance $\sigma_T^2$. Measurement
-precision should generally only
+constructed only from error score variance $\sigma_E^2$ and true score
+variance $\sigma_T^2$. Measurement precision should generally only
 affect the error score variance. Therefore, if the heterogeneity in
@@
 Instead, what is guaranteed is that the expected value of $\sigma_T$ is
-smaller than or equal to the expected value of $\sigma_X$. Similarly, the
-heterogeneity in $\sigma_T$ is guaranteed to be smaller than or equal to the
-uncorrected heterogeneity in $\sigma_X$. However, since the expected
-value of either standard deviation is placed in the denominator of the
-function and the variance of either is placed in the numerator of the
-function, @eq-HS is not sufficient to explain how heterogeneity in ES
-changes due to the attenuation correction.
+smaller than or equal to the expected value of $\sigma_X$. Similarly,
+the heterogeneity in $\sigma_T$ is guaranteed to be smaller than or
+equal to the uncorrected heterogeneity in $\sigma_X$. However, since the
+expected value of either standard deviation is placed in the denominator
+of the function and the variance of either is placed in the numerator of
+the function, @eq-HS is not sufficient to explain how heterogeneity in
+ES changes due to the attenuation correction.
 
@@
 conditions under which the claims made in W&D and H&S are met.
-Unfortunately, the inequality in this form does not
-convey sufficient information about the conditions under which it can
-potentially hold. In Appendix A, we rearranged @eq-ineq_init, so we end
-up with a form that we can use to actually derive conditions under which
-ES heterogeneity can be inflated or deflated by differences in score
-reliability.
+Unfortunately, the inequality in this form does not convey sufficient
+information about the conditions under which it can potentially hold. In
+Appendix A, we rearranged @eq-ineq_init, so we end up with a form that
+we can use to actually derive conditions under which ES heterogeneity
+can be inflated or deflated by differences in score reliability.
 
@@
 mean total score variance can be attributed to mean true score variance.
-Therefore, this metric is close to the average score reliability. Even
-though formally $R_1$ and the meta-analytic mean score reliability
-$\mu_{\rho_{xx'}}$ do not share not the same value, in practice we
-observe little to no differences between the two parameters. [add to appendix]. Large
+Therefore, this metric is close to the average score reliability. Large
 values indicate that correcting the individual score variances (or
@@
 larger amount of random error score variance, leading to a stronger
-correction. The metric $R_2$ on the other hand describes in how far the attenuation
-correction procedure has successfully reduced heterogeneity in score
-variance $\tau^2_{\sigma^2_X}$.
+correction. The metric $R_2$ on the other hand describes in how far the
+attenuation correction procedure has successfully reduced heterogeneity
+in score variance $\tau^2_{\sigma^2_X}$.
 
@@
 
-A value of 1 indicates that the
-heterogeneity in true score variance $\sigma^2_T$ is exactly as
-large as the heterogeneity initially observed. Smaller values indicate
-how much heterogeneity is "left", after applying an attenuation
-correction, relative to the initial heterogeneity. This means that a
-value of .7 indicates that about 70% of heterogeneity in score variance
-remains, even after correcting for differences in score reliability. As
-only 30% of score variance heterogeneity could be removed, this would
-imply that differences in error score variance $\sigma^2_E$ were
-responsible for less than a third of the heterogeneity in score
-variances found. On the contrary, more than two thirds of heterogeneity
-could be attributed to actual differences in how the underlying true
-scores are distributed across samples.
+A value of 1 indicates that the heterogeneity in true score variance
+$\sigma^2_T$ is exactly as large as the heterogeneity initially
+observed. Smaller values indicate how much heterogeneity is "left",
+after applying an attenuation correction, relative to the initial
+heterogeneity. This means that a value of .7 indicates that about 70% of
+heterogeneity in score variance remains, even after correcting for
+differences in score reliability. As only 30% of score variance
+heterogeneity could be removed, this would imply that differences in
+error score variance $\sigma^2_E$ were responsible for less than a third
+of the heterogeneity in score variances found. On the contrary, more
+than two thirds of heterogeneity could be attributed to actual
+differences in how the underlying true scores are distributed across
+samples.
 
@@
 heterogeneity. Whether ES heterogeneity can be decreased depends on
-additional parameters, that we left out of this short form, but can be found,
-accompanied by a more detailed description, in @eq-ineq_fin, Appendix A.
+additional parameters, that we left out of this short form, but can be
+found, accompanied by a more detailed description, in @eq-ineq_fin,
+Appendix A.
 
@@
 change in mean value and heterogeneity due to the correction: ES
-heterogeneity will always be increased, as long as the remaining proportion of
-heterogeneity in score variance is not substantially smaller than the
-remaining proportion of its mean value.
+heterogeneity will always be increased, as long as the remaining
+proportion of heterogeneity in score variance is not substantially
+smaller than the remaining proportion of its mean value.
 
@@
 individual administrations sites are regarding their true score
-variance. If the administration sites carry large differences in true
-score variance $R_2$ is unlikely to be substantially smaller than $R_1$.
-[Na ja, eine Reliabilität von .9 bedeutet erstmal, dass es 9mal so viel wahre Varianz wie Fehlervarianz gibt.
-In diesem Fall müsste R2 unter 0.73 liegen.
-Das impliziert in der Heterogenität ein Verhältnis von ungefähr 3 : 1.
-Die 9mal größere wahre Varianz müsste also für weniger als ¾ der Heterogenität verantwortlich sein.
-In Deinem Textbeispiel (Reliabilität von .8) müsste die 4mal größere wahre Varianz für weniger als die Hälfte der Heterogenität verantwortlich sein.
-– Das erscheint mir beides per se ziemlich unplausibel… Also auch dann, wenn die „administration sites do NOT carry large differences in true score variance”.
-– In anderen Worten: Die Frage ist, ob die absolute Größe der Varianz überhaupt irgendeine Rolle spielt oder ob alles lediglich an den Verhältnissen hängt, die durch die Reliabilität vorgegeben sind.
-Ich habe dazu aber auch keine eindeutige Antwort… Wenn ich überzeugt wäre, dass sich *homogene* true score variances durch planvolle, kontrollierte Stichprobenziehung herstellen lassen, wäre das ein Gegenargument.]
+variance.
+
+In our example of $R_1 = .8$, the true score variance is 4 times as large
+as the error score variance. However, for the attenuation correction to
+reduce ES heterogeneity, the smaller error score variance would need to
+account for almost 50 % of heterogeneity in score variance. This asymmetry
+grows even larger, as the average score reliability grows larger.
+This demonstrated that the individual size of either
+mean score variance component or its heterogeneity is not of importance.
+Instead, the ratio of these components determine, whether heterogeneity
+can decrease after an attenuation correction.
 
@@
 analytical arguments hold up in empirical assessment, but also
-demonstrate how the metrics $R_1$ and $R_2$ help understand how
-score reliability affected ES heterogeneity in practice.
+demonstrate how the metrics $R_1$ and $R_2$ help understand how score
+reliability affected ES heterogeneity in practice.
 
@@
 
-As Boot-Err makes use of
-meta-analytic techniques and bootstrap sampling, we perform analyses
-using the packages *metafor*, version 4.8-0 [@metafor], and *boot*,
-version 1.3-31 [@boot], in the statistical programming language *R*,
-version 4.4.2 [@RCore].
+As Boot-Err makes use of meta-analytic techniques and bootstrap
+sampling, we perform analyses using the packages *metafor*, version
+4.8-0 [@metafor], and *boot*, version 1.3-31 [@boot], in the statistical
+programming language *R*, version 4.4.2 [@RCore].
 
@@
 variances, including their heterogeneity and metrics $R_1$ and $R_2$
-across all 12 phenomena. We see that for ten out of 12
-phenomena, the $R_2$ metric is larger than the $R_1$ metric.
-Additionally, we observe that across eight phenomena, $R_2$ is larger
-than .96. This implies that of the heterogeneity observed in score
-variances, 97% or more could not be explained by differences in random
-error introduced by the measuring instrument, but by genuine differences
-in the underlying distribution.
-
-For the two exceptions in @tbl-tauES, those phenomena where the attenuation correction
-procedure actually led to a decrease in ES heterogeneity (RRR9
-(Behaviour) and ML5 - Shnabel (RPP)), we find that $R_2$ is larger than $R_1$.
-This clearly contradicts our interpretation of @eq-ineq_int. As long as $R_2$ is
-larger than $R_1^3$, we expect a correction procedure to increase effect 
-size heterogeneity. However, we recommend to be cautious when interpreting both the change and
-the direction of said change for these two projects. 
-The significance test for the ES
-heterogeneity was not significant, both before and after the correction.
-Since either the heterogeneity was so small or the sample
+across all 12 phenomena. We see that for ten out of 12 phenomena, the
+$R_2$ metric is larger than the $R_1$ metric. Additionally, we observe
+that across eight phenomena, $R_2$ is larger than .96. This implies that
+of the heterogeneity observed in score variances, 97% or more could not
+be explained by differences in random error introduced by the measuring
+instrument, but by genuine differences in the underlying distribution.
+
+For the two exceptions in @tbl-tauES, those phenomena where the
+attenuation correction procedure actually led to a decrease in ES
+heterogeneity (RRR9 (Behaviour) and ML5 - Shnabel (RPP)), we find that
+$R_2$ is larger than $R_1$. This clearly contradicts our interpretation
+of @eq-ineq_init. As long as $R_2$ is larger than $R_1^3$, we expect a
+correction procedure to increase effect size heterogeneity. However, we
+recommend to be cautious when interpreting both the change and the
+direction of said change for these two projects. The significance test
+for the ES heterogeneity was not significant, both before and after the
+correction. Since either the heterogeneity was so small or the sample
 size so low that the heterogeneity could not be statistically
-distinguished from zero, most likely, the reduction in
-heterogeneity is better explained by imprecise estimation than an actual
-reduction of heterogeneity.
-
-Lastly, for two out of the 12 phenomena, ML5 -
-Albarracin and ML5 - Shnabel (Rev), we find
-$\tau_{\sigma^2_X}$ of 0. Therefore, the $R_2$ metric could not be
-computed, as the denominator would have been zero. However, we also did
-not identify statistically significant heterogeneity in score reliability
-across replications, nor ES heterogeneity. The phenomena were assessed with a 
-particularly low number of
-replications (9 and 8 respectively). Estimation precision for
+distinguished from zero, most likely, the reduction in heterogeneity is
+better explained by imprecise estimation than an actual reduction of
+heterogeneity.
+
+Lastly, for two out of the 12 phenomena, ML5 - Albarracin and ML5 -
+Shnabel (Rev), we find $\tau_{\sigma^2_X}$ of 0. Therefore, the $R_2$
+metric could not be computed, as the denominator would have been zero.
+However, we also did not identify statistically significant
+heterogeneity in score reliability across replications, nor ES
+heterogeneity. The phenomena were assessed with a particularly low
+number of replications (9 and 8 respectively). Estimation precision for
 heterogeneity in variance components, even more so than for ES, largely
-depends on the sample size. Two alternative explanations for this observation
-are available: either, there is no heterogeneity in score variances across 
-replications. In that case, there is also no heterogeneity that could be
-broken down into true and error score variance heterogeneity. Alternatively,
-the low sample size led to
+depends on the sample size. Two alternative explanations for this
+observation are available: either, there is no heterogeneity in score
+variances across replications. In that case, there is also no
+heterogeneity that could be broken down into true and error score
+variance heterogeneity. Alternatively, the low sample size led to
 implausible estimates in both true and total score variance
@@
 
-Generally, @tbl-R1R2 implies that for none of the measurements concerning these
-phenomena, the heterogeneity in observed score variances could be
-substantially explained by differences in measurement quality. For all
-phenomena, $R_2$ was larger than $R_1$, implying that an attenuation
-correction procedure could not reduce ES heterogeneity in any of these
-cases. The results in @tbl-tauES agree with this conclusion, as no
-phenomena could be identified where statistically significant ES
-heterogeneity was accompanied by a reduction in differences due to the
-correction.
+Generally, @tbl-R1R2 implies that for none of the measurements
+concerning these phenomena, the heterogeneity in observed score
+variances could be substantially explained by differences in measurement
+quality. For all phenomena, $R_2$ was larger than $R_1$, implying that
+an attenuation correction procedure could not reduce ES heterogeneity in
+any of these cases. The results in @tbl-tauES agree with this
+conclusion, as no phenomena could be identified where statistically
+significant ES heterogeneity was accompanied by a reduction in
+differences due to the correction.
 
@@
 
+*Note:* $\hat{\mu}_{\sigma^2_X}$ is the back-transformed meta-analytic
+estimate of score variance; $\hat{\mu}_{\sigma^2_T}$ is the
+back-transformed meta-analytic estimate of score variance;
+$\tau_{\mu_{\sigma^2_X}}$ is the back-transformed estimate of
+heterogeneity in score variance; $\hat{\tau}_{{\sigma^2_X}}$ represents
+the estimate of heterogeneity in true score variance.
+
 ## Discussion
@@
 correcting for differences in reliability coefficients using an
-attenuation correction procedure. While we initially observed heterogeneity
-in score reliability across almost all measuring scales, the extent of this
-heterogeneity was very small across all scales. This initial finding already
-implies that differences in score reliability can only have a tiny impact
-on ES heterogeneity, if any.
-
-This implication fits our subsequent finding, as we found that a) in the majority of
-cases, the change in ES heterogeneity is small to non-existent. Across the 
-data-sets we got to examine, differences in
+attenuation correction procedure. While we initially observed
+heterogeneity in score reliability across almost all measuring scales,
+the extent of this heterogeneity was very small across all scales. This
+initial finding already implies that differences in score reliability
+can only have a tiny impact on ES heterogeneity, if any.
+
+This implication fits our subsequent finding, as we found that a) in the
+majority of cases, the change in ES heterogeneity is small to
+non-existent. Across the data-sets we got to examine, differences in
 score reliability do not serve as an appropriate explanation for ES
@@
 differences in how the latent variable is distributed (in terms of
-spread) across the administration sites. This disparity grows even larger, 
-the higher the average score reliability is.
-
-In the 
-majority of research studies done in the social and behavioural sciences
-[@hultsch2002; @bornstein2013; @ashraf2017; @sears1986; @zhao2021], no 
-proper sampling techniques beyond convenience sampling are
-employed. Most samples re-analysed throughout this text came to be
+spread) across the administration sites. This disparity grows even
+larger, the higher the average score reliability is.
+
+In the majority of research studies done in the social and behavioural
+sciences [@hultsch2002; @bornstein2013; @ashraf2017; @sears1986;
+@zhao2021], no proper sampling techniques beyond convenience sampling
+are employed. Most samples re-analysed throughout this text came to be
 through convenience sampling, where students self-select into studies.
-Without a valid sampling scheme, we essentially have no basis to
-assume anything about the true score variance [@hultsch2002]. Most
-importantly, however, we have no reason to assume that the true score
-variance could potentially be homogeneous or stable across
-administration sites. Instead, it is very likely that the convenience
-sampling procedures produce samples that vary extremely strongly in how
-they are put together, leading to large degrees of true score variance
-heterogeneity. At the same time, score reliability is a low-hanging
-fruit to optimise as most researchers are aware and trained in assessing
-score reliability [@hussey2025].
+Without a valid sampling scheme, we essentially have no basis to assume
+anything about the true score variance [@hultsch2002]. Most importantly,
+however, we have no reason to assume that the true score variance could
+potentially be homogeneous or stable across administration sites.
+Instead, it is very likely that the convenience sampling procedures
+produce samples that vary extremely strongly in how they are put
+together, leading to large degrees of true score variance heterogeneity.
+At the same time, score reliability is a low-hanging fruit to optimise
+as most researchers are aware and trained in assessing score reliability
+[@hussey2025].
 
@@
 heterogeneity estimates in score variance. Therefore, an attenuation
-correction is likely to have very little
-impact on ES heterogeneity. If anything, it most likely increases it.
-This also means that, in such settings, score reliability does not serve
-as a moderator explaining ES heterogeneity.
+correction is likely to have very little impact on ES heterogeneity. If
+anything, it most likely increases it. This also means that, in such
+settings, score reliability does not serve as a moderator explaining ES
+heterogeneity.
 
@@
 
-The analytical arguments, detailed in *Appendix A*, leading to
-@eq-ineq_short, rest on the use of the delta-method. @eq-ineq_init
-contains parameters of the score standard deviations’ distributions,
-while the $R_1$ and $R_2$ metrics were designed to discuss score
-variances. The delta-method rests on a number of assumptions, at least
-one of which is likely violated in our analysis: the variance of the
-variable to be transformed needs to be small [@cramer1999]. This entire
-procedure discusses the influence of variance in standard deviations on
-ES heterogeneity, introduced by differences in score reliability.
-Therefore, assuming that this assumption holds seems naive. In the
-*Appendix B*, we explore the archival data, to understand in how far
-employing the delta-method on meta-analyses of standard deviations lead
-to biased results than the meta-analyses on score variances we performed
-using the Boot-Err technique. The differences identified are rather
-miniscule for the majority of phenomena, at the third decimal. Only for
-the phenomenon in PSACR002 (Behaviour) we find larger differences -
-therefore, conclusions for this phenomenon concerning the relation
-between $R_1$ and $R_2$ need to be drawn with caution. Generally, as the
-majority of phenomena did not display any bias in this assessment and
-since the delta-method is not actually used to generate estimates, we
-believe its use to be defensible nonetheless.
+The analytical arguments presented here, represented by @eq-ineq_fin,
+hold as far as the use of the delta method does not lead to widely
+erroneous results. In *Appendix C*, we compare estimates derived from
+the delta-method and Boot-Err to assess in how far the delta-method
+leads to biased estimates. Across all but one data-set, differences were
+miniscule, leading us to believe that its use here was defensible. Only
+for the PSACR002-data-set, we identify larger differences. While we
+believe that $R_1$ and $R_2$ were estimated sufficiently well,
+@eq-ineq_fin does not necessarily hold for how score reliability
+heterogeneity affected ES heterogeneity here.
 
@@
 is in line with assumptions underlying CTT, it is by no means guaranteed
-that it holds in these data-sets. If this assumption was violated,
-whether the relationship was positive or negative, indicated by a
-positive or negative covariance in @eq-XYZ_cov, would most likely have
-an impact on how strongly ES heterogeneity is affected by the
-attenuation correction procedure, possibly even the direction of how it
-is affected. However, currently we have no reason to assume that there
-is a systematic relationship between mean difference and standard
-deviation. As this relationship would also affect the distribution of ES
-like Cohen’s d itself, strong violations of this assumption would go far
-beyond invalidating the claims we derived in @eq-ineq_short. Most
-likely, assumptions of the meta-analytic tests concerning the
-meta-analytic mean size ES and the presence of ES heterogeneity would be
-violated, making any research on change in ES heterogeneity due to
-attenuation correction obsolete.
+that it holds in these data-sets. If this assumption was violated, the
+impact of an attenuation correction procedure on ES heterogeneity would
+likely be affected. Whether this could lead to larger or lower ES
+heterogeneity after the procedure than we currently assume depends on
+whether the covariance in @eq-XYZ_cov is positive or negative. However,
+generally we have no reason to assume that there is a systematic
+relationship between mean difference and standard deviation. As this
+relationship would also affect the distribution of ES like Cohen’s d
+itself, strong violations of this assumption would go far beyond
+invalidating the claims we derived in @eq-ineq_short. Most likely,
+assumptions of the tests concerning the meta-analytic mean ES and the
+presence of ES heterogeneity would be violated, making any research on
+change in ES heterogeneity due to attenuation correction obsolete.
 
@@
 Heterogeneity in ES even in direct replications, where experimental
-factors are held as constant as possible, is substantial
-(Renkewitz, Fünderich & Beinhauer, in preparation). However, we
-have demonstrated that in several scenarios, differences in score
-reliability did not serve as appropriate moderators or explanations of
-ES heterogeneity. Across most phenomena, we failed to reduce ES
-heterogeneity. If anything, we even increased ES heterogeneity, albeit
-to a very small degree. Generally, concerning the multi-site replication
-projects or collaborative research projects, accounting for differences
-in score reliability did not help us gain a better understanding of the
+factors are held as constant as possible, is substantial (Renkewitz,
+Fünderich & Beinhauer, in preparation). However, we have demonstrated
+that in several scenarios, differences in score reliability did not
+serve as appropriate moderators or explanations of ES heterogeneity.
+Across most phenomena, we failed to reduce ES heterogeneity. If
+anything, we even increased ES heterogeneity, albeit to a very small
+degree. Generally, concerning the multi-site replication projects or
+collaborative research projects, accounting for differences in score
+reliability did not help us gain a better understanding of the
 phenomena.
@@
 This manuscript has demonstrated that, even though differences in
-measurement precision are present across most phenomena we could re-analyse, 
-these are strongly outweighed by the differences in true score variance.
-However, what a true score variance actually means, strongly depends on 
-how that number came to be. A sample that is very diverse in its composition
-likely comes with a large true score variance. On the other hand, a sample
-that is highly homogeneous in its composition comes with a smaller true score
-variance. Importantly, the same is not true for random measurement error. 
-Error variance should always be the same size, if the errors are truly random,
-no matter how the sample is put together. Unfortunately, 
-most phenomena we re-analysed here came to be through means of convenience sampling. 
-Therefore, it is not clear what the large differences in true score variance 
-actually imply - are they expressions of
-genuine differences between the populations, or just expressions of different
-self-selection procedures across the samples? As the variance observed in
-a convenience sample is not representative to some *specificied* population, 
-we simply can not know.
-
-Issues of generalizability
-due to the lack of sampling techniques have been, and are still, repeatedly
-called out for the same reasons. Our findings therein reiterate these calls. 
-We believe that we should
-focus our data collection efforts on the generation of samples, where we
-know that the variance in a sample is actually meaningful. This is most
+measurement precision are present across most phenomena we could
+re-analyse, these are strongly outweighed by the differences in true
+score variance. However, what a true score variance actually means,
+strongly depends on how that number came to be. A sample that is very
+diverse in its composition likely comes with a large true score
+variance. On the other hand, a sample that is highly homogeneous in its
+composition comes with a smaller true score variance. Importantly, the
+same is not true for random measurement error. Error variance should
+always be the same size, if the errors are truly random, no matter how
+the sample is put together. Unfortunately, most phenomena we re-analysed
+here came to be through means of convenience sampling. Therefore, it is
+not clear what the large differences in true score variance actually
+imply - are they expressions of genuine differences between the
+populations, or just expressions of different self-selection procedures
+across the samples? As the variance observed in a convenience sample is
+not representative to some *specificied* population, we simply can not
+know.
+
+Issues of generalizability due to the lack of sampling techniques have
+been, and are still, repeatedly called out for the same reasons. Our
+findings therein reiterate these calls. We believe that we should focus
+our data collection efforts on the generation of samples, where we know
+that the variance in a sample is actually meaningful. This is most
 easily achieved by proper sampling techniques, which ensure that the
-sample is representative for some specified population. In such cases, it might be
-that the true score variance varies less strongly than what we currently
-observe. And should the heterogeneity in true score variance remain as
-large, than we know that this is crucial information from which we
-can draw meaningful inferences. We could learn more about
+sample is representative for some specified population. In such cases,
+it might be that the true score variance varies less strongly than what
+we currently observe. And should the heterogeneity in true score
+variance remain as large, than we know that this is crucial information
+from which we can draw meaningful inferences. We could learn more about
 standardized effect sizes and why our phenomena vary across settings,
-populations or continents. We strongly believe that the valuable resources 
-invested into
-research output could be put towards more informative use than what seems
-to be the status quo by ensuring that the variances observed in different
-samples are representative for specified populations.
+populations or continents. We strongly believe that the valuable
+resources invested into research output could be put towards more
+informative use than what seems to be the status quo by ensuring that
+the variances observed in different samples are representative for
+specified populations.
 
@@
 
-```{r, echo = FALSE}
-#| label: tbl-data
-#| tbl-cap: Brief summary of phenomena, designs and measures employed
-
-tab <- read.csv("Tables/data_summary.csv", fileEncoding = "UTF-8", sep = ";") %>% 
-  arrange(MASC)
-
-gt::gt(tab)
-```
-
 ## References
@@
 
-### A - Rearranging inequality
-
-Both metrics $R_1$ and $R_2$ make use of parameters of the distribution
+### A - Rearranging the inequality
+
+Both metrics $R_1$ and $R_2$ make use of parameters of the distributions
 of true and total score variance, not standard deviations. This choice
-was made to facilitate the role of score reliability in the subsequent
-discussion, as score reliability itself is defined by variances, not
-standard deviations (see @eq-rel).
+was made to facilitate the discussion of score reliability, as score
+reliability itself is defined by variances, not standard deviations (see
+@eq-rel).
 
@@
 value and heterogeneity of variance from those of the standard
-deviations, we know that (ref):
+deviations, we know that [@casella2002]:
 $$\mu^2_{\sigma_X} \approx \mu_{\sigma^2_X} \quad \quad  and \quad \quad 
@@
 
-While @eq-delta_method only explicitly contains parameters for the
-observed score variance/standard deviation, the same can be done using
-the true score variance/standard deviation. Using the approximates
-defined in @eq-delta_method for observed and true score variance, we
-arrive at a new inequality in @eq-ineq_CV.
+While @eq-delta_method only contains parameters for the observed score
+variance/standard deviation, the same can be done using the true score
+variance/standard deviation. Using the approximates defined in
+@eq-delta_method for observed and true score variance, we arrive at a
+new inequality in @eq-ineq_CV.
 
@@
 Secondly, as all terms contain $\mu_{\sigma^2_x}$ in the denominator, to
-arrive at @eq-ineq_rearr2 we divided by $\mu_{\sigma^2_x}$.
+arrive at @eq-ineq_rearr2 we multiplied by $\mu_{\sigma^2_x}$.
 
@@
 
-Taking $\frac{\tau_{\sigma^2_x}}{\mu_{\sigma^2_x}} = CV{\sigma^2_x}$
-allows to simplify for @eq-ineq_rearr4.
+We define $CV{\sigma^2_x}$ as
+$CV{\sigma^2_x} = \frac{\tau_{\sigma^2_x}}{\mu_{\sigma^2_x}}$, thereby
+allowing to simplify for @eq-ineq_rearr4.
 
@@
 $CV^2_{\sigma^2_X}$), are bound to be positive. For the inequality to
-hold, we need the left-hand side of the equation to remain positive,
-larger than zero. We can distill two scenarios, under which this
-equation should hold: (a) one of the terms inside the brackets
-($\frac{1}{R_1} - 1$ or $\frac{R_2}{R^3_1}  - 1$) is positive and large
-enough, so that the second term not containing that same bracket is
-positively dominated by the first term, meaning it is sufficiently large
-that the positive term cancels out the negative term; or (b) both terms
-inside the brackets need to be positive.
+hold, we need the left-hand side to remain positive, larger than zero.
+We can distil two scenarios, under which this equation should hold: (a)
+one of the terms inside the brackets ($\frac{1}{R_1} - 1$ or
+$\frac{R_2}{R^3_1}  - 1$) is positive and large enough, so that the
+second term not containing that same bracket is positively dominated by
+the first term, meaning it is sufficiently large that the positive term
+cancels out the negative term; or (b) both terms inside the brackets
+need to be positive.
 
@@
 need to take on values, that allow for such a scenario. Since we do not
-want to make any claims regarding the unstandardized mean difference and
+want to make any claims regarding the unstandardised mean difference and
 how it relates to the coefficient of variation of score variance, we can
-not distill a scenario in which it is guaranteed, that @eq-ineq_fin is
+not distil a scenario in which it is guaranteed, that @eq-ineq_fin is
 violated. Therefore $R_1^3 \leq R_2$ only describes a condition, under
@@
 ### B - Examining bias due to delta-method
+
+The analytical arguments, detailed in *Appendix A*, leading to
+@eq-ineq_short, rest on the use of the delta-method [@vandervaart2000,
+@cramer1999, @casella2002]. @eq-ineq_init contains parameters of
+distributions of the standard deviations of scores, while the $R_1$ and
+$R_2$ metrics were designed to discuss score variances. The delta-method
+rests on a number of assumptions, at least one of which is likely
+violated in our analysis: the variance of the variable to be transformed
+needs to be small [@cramer1999]. This entire procedure discusses the
+influence of variance in standard deviations on ES heterogeneity,
+introduced by differences in score reliability. Therefore, assuming that
+this assumption holds seems naive. Here, we explore the archival data,
+to understand in how far employing the delta-method on meta-analyses of
+standard deviations lead to results biased more strongly than estimates
+from the meta-analyses on score variances we performed using the
+Boot-Err technique.
 
@@
 estimate of mean total score variance, obtained via Boot-Err;
-$\hat{\mu}_{\sigma^2_X}$ is the back-transformed estimate of its
+$\hat{\tau}_{\sigma^2_X}$ is the back-transformed estimate of its
 heterogeneity. On the other hand, $\hat{\mu}_{\Delta\sigma^2_X}$ is the
@@
 obtained from application of the delta method;
-$\hat{\mu}_{\Delta\sigma^2_X}$ is the back-transformed estimate of its
+$\hat{\tau}_{\Delta\sigma^2_X}$ is the back-transformed estimate of its
 heterogeneity.
@@
 to be defensible.
+
+### C - Summary of Phenomena, designs and measures
+
+```{r, echo = FALSE}
+#| label: tbl-data
+#| tbl-cap: Brief summary of phenomena, designs and measures employed
+
+tab <- read.csv("Tables/data_summary.csv", fileEncoding = "UTF-8", sep = ";") %>% 
+  arrange(MASC)
+
+gt::gt(tab)
+```

### 2025-12-23 Manuscript_MD_ScoreReLiability.qmd
Update - FR comments

@@
 bibliography: ReLiability_references.bib
-csl: styles/apa-6th-edition.csl
+csl: styles/apa-7th-edition.csl
 format: docx
@@
 
-Heterogeneity in standardized effect sizes, such as Cohen’s d, is informative because it indicates that a psychological phenomenon is not yet fully understood.
-Prior work by Hunter and Schmidt as well as Wiernik and Dahlke has argued that such heterogeneity partly reflects differences in measurement precision and that correcting for these differences using attenuation correction should therefore reduce heterogeneity.
-We reanalyze data from large-scale collaborative projects, including the ManyLabs studies, Registered Replication Reports, and the Psychological Science Accelerator, and demonstrate that this assumption does not hold universally.
-Treating standardized effect sizes as ratio variables, we offer an alternative account of how differences in measurement precision relate to observed heterogeneity.
-Across many psychological research contexts, score reliability tends to be high and measurement error relatively stable.
-Under these conditions, when samples differ substantially in their composition or diversity, attenuation correction procedures are more likely to increase rather than decrease heterogeneity.
-We conclude by discussing the limited relevance of random measurement error for explaining effect size heterogeneity in the context of convenience sampling.
+Heterogeneity in standardized effect sizes, such as Cohen’s d, is
+informative because it indicates that a psychological phenomenon is not
+yet fully understood. Prior work by Hunter and Schmidt as well as
+Wiernik and Dahlke has argued that such heterogeneity partly reflects
+differences in measurement precision and that correcting for these
+differences using attenuation correction should therefore reduce
+heterogeneity. We reanalyze data from large-scale collaborative
+projects, including the ManyLabs studies, Registered Replication
+Reports, and the Psychological Science Accelerator, and demonstrate that
+this assumption does not hold universally. Treating standardized effect
+sizes as ratio variables, we offer an alternative account of how
+differences in measurement precision relate to observed heterogeneity.
+Across many psychological research contexts, score reliability tends to
+be high and measurement error relatively stable. Under these conditions,
+when samples differ substantially in their composition or diversity,
+attenuation correction procedures are more likely to increase rather
+than decrease heterogeneity. We conclude by discussing the limited
+relevance of random measurement error for explaining effect size
+heterogeneity in the context of convenience sampling.
 
 #### Non-technical abstract
-Researchers often find that studies on the same psychological effect report different results, known as differences in effect sizes.
-Such variation suggests that the phenomenon is not yet fully understood.
-Some have argued that part of this variation comes from differences in measurement precision—how reliably a study’s tools or tests measure what they intend to measure.
-Based on this reasoning, it has been suggested that adjusting or “correcting” for these differences should make results more consistent across studies.
-We revisited this assumption by looking at large collaborative projects in psychology, such as ManyLabs, the Registered Replication Reports, and the Psychological Science Accelerator.
-We found that the expected reduction in variation does not generally occur.
-In fact, when measurement precision is already quite high - as it usually is in psychology, where researchers carefully select reliable measures - but samples differ widely in who participates, such corrections can actually increase differences in reported effect sizes.
-In simpler terms, differences in how diverse the samples of participants are across studies (which comes about from convenience sampling) matters more for explaining variation in results than small differences in how precisely a measure captures a construct.
-We introduce a simple rule that identifies when correcting for measurement differences is likely to backfire, increasing rather than reducing heterogeneity.
-We conclude by discussing why random measurement error plays a smaller role in explaining effect size differences than the use of diverse, convenience-based samples.
+
+Researchers often find that studies on the same psychological effect
+report different results, known as differences in effect sizes. Such
+variation suggests that the phenomenon is not yet fully understood. Some
+have argued that part of this variation comes from differences in
+measurement precision—how reliably a study’s tools or tests measure what
+they intend to measure. Based on this reasoning, it has been suggested
+that adjusting or “correcting” for these differences should make results
+more consistent across studies. We revisited this assumption by looking
+at large collaborative projects in psychology, such as ManyLabs, the
+Registered Replication Reports, and the Psychological Science
+Accelerator. We found that the expected reduction in variation does not
+generally occur. In fact, when measurement precision is already quite
+high - as it usually is in psychology, where researchers carefully
+select reliable measures - but samples differ widely in who
+participates, such corrections can actually increase differences in
+reported effect sizes. In simpler terms, differences in how diverse the
+samples of participants are across studies (which comes about from
+convenience sampling) matters more for explaining variation in results
+than small differences in how precisely a measure captures a construct.
+We introduce a simple rule that identifies when correcting for
+measurement differences is likely to increase rather than reduce
+heterogeneity. We conclude by discussing why random measurement error
+plays a smaller role in explaining effect size differences than the use
+of diverse, convenience-based samples.
 
@@
 
-One source for this heterogeneity, which is not a typical moderator tied
-to a specific theory, might be impairments in score reliability which
-vary in their extent across replications [@olsson2020]. Hunter and
-Schmidt [-@hunter2004] and Wiernik and Dahlke [-@wiernik2020],
-discussing the role of score reliability for standardized effect sizes,
-claim that differences in score reliability should inflate, and
-therefore increase, heterogeneity. This implies that standard
-corrections for imperfect score reliability, such as attenuation
-correction procedures, should remove that inflation, reducing
-heterogeneity in corrected effect sizes.
+The search for moderators that might explain said heterogeneity across
+several phenomena has not been particularly successful [e.g. @ManyLabs3,
+@ManyLabs2]. Another source for this heterogeneity, which is not a
+typical moderator tied to a specific theory, might be impairments in
+score reliability which vary in their extent across replications
+[@olsson2020]. Hunter and Schmidt [-@hunter2004] and Wiernik and Dahlke
+[-@wiernik2020], discussing the role of score reliability for
+standardized effect sizes, claim that differences in score reliability
+should inflate, and therefore increase, heterogeneity. This implies that
+standard corrections for imperfect score reliability, such as
+attenuation correction procedures, should remove that inflation,
+reducing heterogeneity in corrected effect sizes.
 
@@
 argument that correcting for differences in score reliability reduces
-effect size heterogeneity. Instead, to our great surprise, we find that
-heterogeneity might as well be larger after the correction procedure.
-Therefore, previous analyses and assumptions regarding the relationship
-between score reliability and heterogeneity need to be false. In the
-second half of this text we make use of analytical arguments to identify
-the necessary conditions under which an attenuation correction
-inevitably leads to an increase in effect size heterogeneity \[Do we
-need any "Schleifen" in this paragraph - to reduce informational load /
-density\].
+effect size heterogeneity. Instead, we find that heterogeneity might as
+well be larger after the correction procedure. Therefore, previous
+analyses and assumptions regarding the relationship between score
+reliability and heterogeneity need to be false. In the second half of
+this text we make use of analytical arguments to identify the necessary
+conditions under which an attenuation correction inevitably leads to an
+increase in effect size heterogeneity.
 
@@
 error.*” (p. 302). Overall, both references imply that differences in
-score (un)reliability inevitably lead to an inflated heterogeneity in
-ES.
+score reliability inevitably lead to an inflated heterogeneity in ES.
 
@@
 [@muchinsky1996; @schmidt1996; @schennach2016], is a repeatedly
-recommended practice in Psychology [@hunter2004, @wiernik2020] . @eq-dc
+recommended practice in Psychology [@hunter2004; @wiernik2020]. @eq-dc
 describes a simple attenuation correction procedure.
@@
 use of more than a single indicator to measure the dependent variable.
-Therefore, for the majority of phenomena it is not possible to assess in
-how far measurement quality is sufficiently high in terms of e.g.
-internal consistency or low random error variance.
+Therefore, for the majority of phenomena it is not possible to assess
+whether measurement quality is sufficiently high in terms of internal
+consistency or low random error variance.
 
@@
 the different phenomena and how they were measured can be found at
-\[https://osf.io/g2fzq\], while a brief summary can be found in appendix C.
+\[https://osf.io/g2fzq\], while a brief summary can be found in appendix
+C.
 
@@
 refer to estimates derived by means of random effects meta-analysis, we
-will use the Greek letter $\delta_0$ to denote estimates of uncorrected
-$d$'s population distribution and $\delta$ to denote estimates of
+will use the Greek letter $\delta$ to denote estimates of uncorrected
+$d$'s population distribution and $\delta_c$ to denote estimates of
 corrected $d_c$'s population distribution.
@@
 indicators of heterogeneity are readily available, such as $I^2$, $H^2$
-or the coefficient of variation $CV$[@borenstein2009; @higgins2002].
-However, the descriptions found in H&S and W&D discuss the absolute
-amount of heterogeneity in terms of variance ($\tau^2$) or standard
-deviation ($\tau$). Therefore, we also make use of absolute
-heterogeneity here.
+[@borenstein2009; @higgins2002]. A relative heterogeneity measure less
+common in Psychology, but easily computed is the coefficient of
+variation $CV = \frac{\tau}{\mu}$. However, the descriptions found in
+H&S and W&D discuss the absolute amount of heterogeneity in terms of
+variance ($\tau^2$) or standard deviation ($\tau$). Therefore, we also
+make use of absolute heterogeneity here.
 
@@
 
-[Equations @eq-d], [-@eq-SEd], [-@eq-dc], and [-@eq-SEdc] were used to
+[Equations @eq-d], [-@eq-dc], [-@eq-SEd], and [-@eq-SEdc] were used to
 generate estimates of standardized ES, both uncorrected and corrected,
@@
 
-#### Which phenomena pass criterion 3)?
+#### Which phenomena pass criterion (3)?
 
@@
 criterion (3)—the effect size must be statistically significantly
-distinguishable from zero at $p < .05$. 12 phenomena pass this
+distinguishable from zero at $p < .05$. Twelve phenomena pass this
 criterion. It may be remarked that the attenuation correction procedure,
@@
 *Note:* In this table $\hat{\mu}_{d}$ represents the meta-analytic
-estimate of uncorrected $d$ and $\hat{\mu}_{d_c}$ represents the
-estimate of the meta-analytic mean of corrected $d_c$.
+estimate of uncorrected $d$ and $\hat{\mu}_{d_c}$ represents the the
+meta-analytic estimate of corrected $d_c$.
 
@@
 test, we can not distinguish this value confidently from zero. This
-implies that for the majority of phenomena, differences in score
-reliability can not be made responsible for heterogeneity in
-standardized effect sizes.
+implies that for none of phenomena, differences in score reliability can
+be made responsible for heterogeneity in standardized effect sizes.
 
@@
                  QEp_raw = "p",
-                 tau_raw = gt::md("$\\hat{\\tau}_{d_c}$"),
-                 QE_raw_str = "QE (df)",
-                 QEp_raw = "p") %>% 
+                 tau_cor = gt::md("$\\hat{\\tau}_{d_c}$"),
+                 QE_cor_str = "QE (df)",
+                 QEp_cor = "p") %>% 
   gt::tab_options(table.width = gt::pct(100),
@@
 *Note:* $\hat{\mu}_{\rho_{XX'}}$ is the back-transformed meta-analytic
-estimate of mean score reliability; $\tau_{\mu_{\rho_{XX'}}}$ is the
-back-transformed estimate of heterogeneity in score reliability;
+estimate of mean score reliability; $\hat{\tau}_{\mu_{\rho_{XX'}}}$ is
+the back-transformed estimate of heterogeneity in score reliability;
 $\hat{\tau}_{d}$ represents the estimate of heterogeneity in uncorrected
@@
 procedure. However, it needs to be noted that these differences are
-small and could not be distinguished from sampling uncertainty in a
-statistical test. For this reason, we do not mean to interpret these
+rather small. For this reason, we do not mean to interpret these
 differences, while we note that the direction of differences are clearly
@@
 reliability do not inflate heterogeneity in uncorrected effect sizes.
-Instead these differences mask ES heterogeneity, that can be unveiled by
+Instead these differences mask ES heterogeneity that can be unveiled by
 accounting for score reliability.
@@
 heterogeneity. Heterogeneity describes unexplained variation in a
-parameter's true distribution, that we model using meta-analytical
+parameter's true distribution, that we model using meta-analytic
 techniques [@borenstein2009; @higgins2002]. In that way, the idea of
@@
 
-The residual variance that remains in a simple linear model will be
-reduced in all instances, where we use a predictor that we know shares
-variance with the predicted parameter. Generally, removing a source of
-variance by controlling for it, we expect any unexplained variation to
-shrink. Similarly, we know that differences in score reliability, by
-definition, are related to differences in standardized effect sizes.
-Therefore, intuitively, it seems sensible to expect that controlling for
+The residual variance that remains in a simple linear model is reduced
+in all instances, where we use a predictor that we know shares variance
+with the predicted parameter. Generally, removing a source of variance
+by controlling for it, we expect any unexplained variation to shrink.
+Similarly, we know that differences in score reliability, by definition,
+are related to differences in standardized effect sizes. Therefore,
+intuitively, it seems sensible to expect that controlling for
 differences in score reliability should reduce the unexplained variance
@@
 heterogeneity in ESs [-@hunter2004, p. 309]. @eq-HS implies that
-heterogeneity in $\delta_0$ is essentially a function in which
-heterogeneity and mean value of $\delta$ are weighted by mean score
-reliability $\rho_{xx‘}$ and its heterogeneity.
-
-$$\tau^2_{\delta_0} = [\mu_{\rho_{xx‘}}]^2 \tau^2_{\delta} + [\mu_{\delta}]^2 \tau^2_{\rho_{xx‘}}$$ {#eq-HS}
-
-We adjusted @eq-HS using the notation used throughout this text. Here,
-$\delta_0$, carrying a 0 in the index, refers to the uncorrected ES,
-undistorted by sampling error but not corrected for measuring error.
-Without an index, $\delta$, on the other hand, refers to the corrected
-ES, undistorted by sampling error and measurement error. Lastly,
-$\rho_{XX'}$ refers to score reliability. Hunter & Schmidt
-[-@hunter2004] claim that @eq-HS demonstrates that differences in score
-reliability inflate ES heterogeneity, which means that an attenuation
-correction procedure would have to reduce ES heterogeneity. However, it
-can be easily verified that @eq-HS does not allow for such claims.
-Depending on both the mean value $\mu_{\rho_{xx‘}}$ and the
-heterogeneity $\tau^2_{\rho_{xx‘}}$ in score reliability, ES
-heterogeneity might be increased or decreased by the differences in
-score reliability.
+heterogeneity in $\delta$ is essentially a function in which
+heterogeneity and mean value of $\delta_c$ are weighted by mean score
+reliability $\rho_{xx‘}$ and its heterogeneity. Here, we adjusted the
+notation from Hunter & Schmidt [-@hunter2004] to the notation used
+through this text.
+
+$$\tau^2_{\delta} = [\mu_{\rho_{xx‘}}]^2 \tau^2_{\delta_c} + [\mu_{\delta_c}]^2 \tau^2_{\rho_{xx‘}}$$ {#eq-HS}
+
+Here, $\delta$ refers to the uncorrected ES, undistorted by sampling
+error but not corrected for measuring error. With the index $c$,
+$\delta_c$, on the other hand, refers to the corrected ES, undistorted
+by sampling error and measurement error. Lastly, $\rho_{XX'}$ refers to
+score reliability. Hunter & Schmidt [-@hunter2004] claim that @eq-HS
+demonstrates that differences in score reliability inflate ES
+heterogeneity, which means that an attenuation correction procedure
+would have to reduce ES heterogeneity. However, it can be easily
+verified that @eq-HS does not allow for such claims. Depending on both
+the mean value $\mu_{\rho_{xx‘}}$ and the heterogeneity
+$\tau^2_{\rho_{xx‘}}$ in score reliability, ES heterogeneity might be
+increased or decreased by the differences in score reliability.
 
 Most likely, Hunter and Schmidt did not fully explore the implications
-of @eq-HS, but followed the general intuition we lined out–"If a source
-of variation is controlled for, the amount of unexplained variance
-should be reduced." Wiernik and Dahlke [-@wiernik2020] refer to the work
-by Hunter & Schmidt [-@wiernik2020], likely following their intuition.
+of @eq-HS, but followed the general intuition we lined out – "If a
+source of variation is controlled for, the amount of unexplained
+variance should be reduced." Wiernik and Dahlke [-@wiernik2020] refer to
+the work by Hunter & Schmidt [-@hunter2004], likely following their
+intuition.
 
@@
 score reliability. Instead, we propose an alternative way that helps us
-discern, under which conditions we can expect an attenuation correction
+discern under which conditions we can expect an attenuation correction
 procedure to lead to an increase in ES heterogeneity.
@@
 Realising that an ES, as defined in @eq-d, is actually a ratio, allows
-for a different analytical description of how differences in score
-reliability can affect ES heterogeneity. If the phenomenon is
+for an analytical description alternative to @eq-HS of how differences
+in score reliability can affect ES heterogeneity. If the phenomenon is
 heterogeneous and score reliability varies across replications, it seems
@@
 constructed by dividing one random variable by a second random variable:
-$Z = \frac{X}{Y}$. First order taylor approximation may be used to
+$Z = \frac{X}{Y}$. First order Taylor approximation may be used to
 generate estimates of mean and variance of the ratio variable
@@
 a ratio) is a) differentiable, b) has existing moments and that c) its
-higher-order moments are negligible (e.g. limited variability and
-monotonicity) (Asymptotic statistics, A.W. van der Vaart - find book
-online). Here we use the Taylor-estimator of the ratio’s variance to
-demonstrate how differences in score reliability may affect
-heterogeneity in ES.
+higher-order moments are negligible [e.g. @vandervaart2000]. Here we use
+the Taylor-estimator of the ratio’s variance to demonstrate how
+differences in score reliability may affect heterogeneity in ES.
 
@@
 
-$$\tau^2[\delta_0] \approx \frac{\tau^2[MD]}{\mu[\sigma_x]^2} + \frac{\mu[MD]^2}{\mu[\sigma_x]^4} \tau^2[\sigma_x]$$ {#eq-ratio_d0}
+$$\tau^2[\delta] \approx \frac{\tau^2[MD]}{\mu[\sigma_x]^2} + \frac{\mu[MD]^2}{\mu[\sigma_x]^4} \tau^2[\sigma_x]$$ {#eq-ratio_d0}
 
@@
 In that case, some heterogeneity in $\sigma_T$ would remain, albeit
-heterogeneity found in $\sigma_x$ was reduced. @eq-ratio_d0 essentially
+heterogeneity found in $\sigma_X$ was reduced. @eq-ratio_d0 essentially
 demonstrates how mean and heterogeneity of MD, paired with mean and
-heterogeneity of total score standard deviation $\sigma_x$ can be used
-to estimate heterogeneity in raw ES $\delta_0$. Similarly, we can adjust
+heterogeneity of total score standard deviation $\sigma_X$ can be used
+to estimate heterogeneity in raw ES $\delta$. Similarly, we can adjust
 @eq-ratio_d0 to describe how, instead of total score standard deviation,
 mean and variance in true score standard deviation $\sigma_T$ affect
-heterogeneity in corrected ES $\delta$.
-
-$$\tau^2[\delta] \approx \frac{\tau^2[MD]}{\mu[\sigma_T]^2} + \frac{\mu[MD]^2}{\mu[\sigma_T]^4} \tau^2[\sigma_T]$$ {#eq-ratio_d}
+heterogeneity in corrected ES $\delta_c$.
+
+$$\tau^2[\delta_c] \approx \frac{\tau^2[MD]}{\mu[\sigma_T]^2} + \frac{\mu[MD]^2}{\mu[\sigma_T]^4} \tau^2[\sigma_T]$$ {#eq-ratio_d}
 
 The differences between the two [Equations @eq-ratio_d0] and
-[-@eq-ratio_d] demonstrates how ES heterogeneity is affected by an
+[-@eq-ratio_d] demonstrate how ES heterogeneity is affected by an
 attenuation correction procedure. According to the quotes in H&S and
 W&D, @eq-ratio_d0 should lead to a larger value of
-$\tau^2\left[\delta_0\right]$, compared to the estimate of
-$\tau^2\left[\delta\right]$ from @eq-ratio_d. However, as demonstrated
+$\tau^2\left[\delta\right]$, compared to the estimate of
+$\tau^2\left[\delta_c\right]$ from @eq-ratio_d. However, as demonstrated
 in the re-analysis of archival data, we can already expect that this can
@@
 
-$$\tau^2_{\delta} > \tau^2_{\delta_0} \quad  \quad \quad \quad \quad  \frac{\tau_{MD}^2}{\mu_{\sigma_T}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_T}^4} \tau^2_{\sigma_T} > \frac{\tau_{MD}^2}{\mu_{\sigma_x}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_x}^4} \tau^2_{\sigma_x}$$ {#eq-ineq_init}
+$$\tau^2_{\delta_c} > \tau^2_{\delta} \quad  \quad \quad \quad \quad  \frac{\tau_{MD}^2}{\mu_{\sigma_T}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_T}^4} \tau^2_{\sigma_T} > \frac{\tau_{MD}^2}{\mu_{\sigma_x}^2} + \frac{\mu_{MD}^2}{\mu_{\sigma_x}^4} \tau^2_{\sigma_x}$$ {#eq-ineq_init}
 
@@
 
-In our example of $R_1 = .8$, the true score variance is 4 times as large
-as the error score variance. However, for the attenuation correction to
-reduce ES heterogeneity, the smaller error score variance would need to
-account for almost 50 % of heterogeneity in score variance. This asymmetry
-grows even larger, as the average score reliability grows larger.
-This demonstrated that the individual size of either
-mean score variance component or its heterogeneity is not of importance.
-Instead, the ratio of these components determine, whether heterogeneity
-can decrease after an attenuation correction.
+In our example of $R_1 = .8$, the true score variance is 4 times as
+large as the error score variance. However, for the attenuation
+correction to reduce ES heterogeneity, the smaller error score variance
+would need to account for almost 50 % of heterogeneity in score
+variance. This asymmetry grows even larger, as the average score
+reliability grows larger. This demonstrates that the individual size of
+either mean score variance component or its heterogeneity is not of
+importance. Instead, the ratio of these components determine whether
+heterogeneity can decrease after an attenuation correction.
 
@@
 no matter what value other parameters like the mean value of $MD$ or its
-heterogeneity take on. This means that, if the condition in
+heterogeneity assume. This means that, if the condition in
 @eq-ineq_short is violated, it is by no means guaranteed that an
@@
 correction procedure to increase effect size heterogeneity. However, we
-recommend to be cautious when interpreting both the change and the
-direction of said change for these two projects. The significance test
-for the ES heterogeneity was not significant, both before and after the
-correction. Since either the heterogeneity was so small or the sample
-size so low that the heterogeneity could not be statistically
-distinguished from zero, most likely, the reduction in heterogeneity is
-better explained by imprecise estimation than an actual reduction of
-heterogeneity.
+recommend caution when interpreting both the change and the direction of
+said change for these two projects. The significance test for the ES
+heterogeneity was not significant, both before and after the correction.
+Since either the heterogeneity was so small or the sample size so low
+that the heterogeneity could not be statistically distinguished from
+zero, most likely, the reduction in heterogeneity is better explained by
+imprecise estimation than an actual reduction of heterogeneity.
 
@@
 Meta-analytic estimates, especially those concerning score variance
-components such as $R_2$, should generally be interpreted with caution,
+components such as $R_2$, should generally be interpreted with caution
 if the sample size was particularly low.
@@
 back-transformed meta-analytic estimate of score variance;
-$\tau_{\mu_{\sigma^2_X}}$ is the back-transformed estimate of
+$\hat{\tau}_{\mu_{\sigma^2_X}}$ is the back-transformed estimate of
 heterogeneity in score variance; $\hat{\tau}_{{\sigma^2_X}}$ represents
@@
 
-This implication fits our subsequent finding, as we found that a) in the
+This implication fits our subsequent results, as we found that a) in the
 majority of cases, the change in ES heterogeneity is small to
@@
 heterogeneity. We also found that b) in the majority of cases, the
-direction of change in ES heterogeneity does not align with assumptions
+direction of change in ES heterogeneity did not align with assumptions
 formulated in W&D and H&S. In almost all cases (every single case, if we
@@
 and H&S, implying that an attenuation correction will reduce ES
-heterogeneity, stem from a general misunderstanding of how score
-reliability can affect ESs.
+heterogeneity, stem from a misunderstanding of how score reliability can
+affect ESs.
 
@@
 
-In the majority of research studies done in the social and behavioural
-sciences [@hultsch2002; @bornstein2013; @ashraf2017; @sears1986;
-@zhao2021], no proper sampling techniques beyond convenience sampling
-are employed. Most samples re-analysed throughout this text came to be
-through convenience sampling, where students self-select into studies.
-Without a valid sampling scheme, we essentially have no basis to assume
-anything about the true score variance [@hultsch2002]. Most importantly,
-however, we have no reason to assume that the true score variance could
-potentially be homogeneous or stable across administration sites.
-Instead, it is very likely that the convenience sampling procedures
-produce samples that vary extremely strongly in how they are put
-together, leading to large degrees of true score variance heterogeneity.
-At the same time, score reliability is a low-hanging fruit to optimise
-as most researchers are aware and trained in assessing score reliability
-[@hussey2025].
+In the majority of studies done in the social and behavioural sciences
+[@hultsch2002; @bornstein2013; @ashraf2017; @sears1986; @zhao2021], no
+proper sampling techniques beyond convenience sampling are employed.
+Most samples re-analysed throughout this text were collected by means of
+convenience sampling, where students self-select into studies. Without a
+valid sampling scheme, we essentially have no basis to assume anything
+about the true score variance [@hultsch2002]. Most importantly, however,
+we have no reason to assume that the true score variance is homogeneous
+or stable across administration sites. Instead, it is very likely that
+the convenience sampling procedures produce samples that vary extremely
+strongly in how they are put together, leading to large degrees of true
+score variance heterogeneity. At the same time, score reliability is a
+low-hanging fruit to optimise as most researchers are aware and trained
+in assessing score reliability [@hussey2025].
 
@@
 Our results fit in well with recent work by Olsson-Collentine et al.
-[@olsson2023]. In a large simulation scheme, they demonstrate that
+[-@olsson2023]. In a large simulation scheme, they demonstrate that
 differences in score reliability across administrations typically
@@
 
-The analytical arguments presented here, represented by @eq-ineq_fin,
-hold as far as the use of the delta method does not lead to widely
-erroneous results. In *Appendix C*, we compare estimates derived from
-the delta-method and Boot-Err to assess in how far the delta-method
-leads to biased estimates. Across all but one data-set, differences were
-miniscule, leading us to believe that its use here was defensible. Only
-for the PSACR002-data-set, we identify larger differences. While we
-believe that $R_1$ and $R_2$ were estimated sufficiently well,
-@eq-ineq_fin does not necessarily hold for how score reliability
-heterogeneity affected ES heterogeneity here.
+The analytical arguments presented here, represented by @eq-ineq_fin in
+*Appendix A*, hold as far as the use of the delta method does not lead
+to widely erroneous results. In *Appendix C*, we compare estimates
+derived from the delta-method and Boot-Err to assess in how far the
+delta-method leads to biased estimates. Across all but one data-set,
+differences were miniscule, leading us to believe that its use here was
+defensible. Only for the PSACR002-data-set, we identify larger
+differences. While we believe that $R_1$ and $R_2$ were estimated
+sufficiently well, @eq-ineq_fin does not necessarily hold for how score
+reliability heterogeneity affected ES heterogeneity for that particular
+study.
 
@@
 
-Heterogeneity in ES even in direct replications, where experimental
-factors are held as constant as possible, is substantial (Renkewitz,
-Fünderich & Beinhauer, in preparation). However, we have demonstrated
-that in several scenarios, differences in score reliability did not
-serve as appropriate moderators or explanations of ES heterogeneity.
-Across most phenomena, we failed to reduce ES heterogeneity. If
-anything, we even increased ES heterogeneity, albeit to a very small
-degree. Generally, concerning the multi-site replication projects or
-collaborative research projects, accounting for differences in score
-reliability did not help us gain a better understanding of the
+Heterogeneity in ES, even in direct replications where experimental
+factors are held as constant as possible, is substantial (Renkewitz et
+al., in preparation). However, we have demonstrated that in several
+scenarios, differences in score reliability did not serve as appropriate
+moderators or explanations of ES heterogeneity. Contrary to popular
+expectations, correction for differences in score reliability did not
+reduce ES heterogeneity. Instead, ES heterogeneity was increased, albeit
+to a very small degree. Generally, concerning the multi-site replication
+projects or collaborative research projects, accounting for differences
+in score reliability did not help us gain a better understanding of the
 phenomena.
@@
 
-This manuscript has demonstrated that, even though differences in
-measurement precision are present across most phenomena we could
-re-analyse, these are strongly outweighed by the differences in true
-score variance. However, what a true score variance actually means,
-strongly depends on how that number came to be. A sample that is very
-diverse in its composition likely comes with a large true score
-variance. On the other hand, a sample that is highly homogeneous in its
-composition comes with a smaller true score variance. Importantly, the
-same is not true for random measurement error. Error variance should
-always be the same size, if the errors are truly random, no matter how
-the sample is put together. Unfortunately, most phenomena we re-analysed
-here came to be through means of convenience sampling. Therefore, it is
-not clear what the large differences in true score variance actually
-imply - are they expressions of genuine differences between the
-populations, or just expressions of different self-selection procedures
-across the samples? As the variance observed in a convenience sample is
-not representative to some *specificied* population, we simply can not
-know.
+We have demonstrated that, even though differences in measurement
+precision are present across most phenomena we could re-analyse, these
+are strongly outweighed by the differences in true score variance.
+However, what a true score variance actually means strongly depends on
+how that number came to be. A sample that is very diverse in its
+composition likely comes with a large true score variance. On the other
+hand, a sample that is highly homogeneous in its composition comes with
+a smaller true score variance. Importantly, the same is not true for
+random measurement error. Error variance should always be the same size,
+if the errors are truly random, no matter how the sample is put
+together. Unfortunately, most phenomena we re-analysed here came to be
+through means of convenience sampling. Therefore, it is not clear what
+the large differences in true score variance actually imply - are they
+expressions of genuine differences between the populations, or just
+expressions of different self-selection procedures across the samples?
+As the variance observed in a convenience sample is not representative
+to some *specified* population, we simply can not know.
 
@@
 the last decades. We generally agree with their ideas that
-underappreciated (differences in un-)reliability can introduce
-substantial biases in meta-analytic estimates and tests in Psychology.
-However, based on our findings, we believe that the issues created by
-ignoring random sampling are much more dire and warrant immediate
-attention. Only once we can be sure that the variance in true scores or
-latent factors is actually meaningful for some population, we can turn
-towards imprecise measurements, expecting correction procedures to
-explain unwarranted heterogeneity. In the current state of Psychology,
+underappreciated (differences in) reliability can introduce substantial
+biases in meta-analytic estimates and tests in Psychology. However,
+based on our findings, we believe that the issues created by ignoring
+random sampling are much more dire and warrant immediate attention. Only
+once we can be sure that the variance in true scores or latent factors
+is actually meaningful for some population, we can turn towards
+imprecise measurements, expecting correction procedures to explain
+unwarranted heterogeneity. In the current state of Psychology,
 meta-analysts should not expect that correcting ES for score reliability
@@
 inequality can only be violated if the relative remaining heterogeneity
-of score variance ($R_2$) is smaller than the the average score
-reliability to the power of 3. Restated in mathematical terms, this
-leads us to the rule/condition described in @eq-ineq_short, in the main
-text.
+of score variance ($R_2$) is smaller than the average score reliability
+to the power of 3. Restated in mathematical terms, this leads us to the
+rule/condition described in @eq-ineq_short, in the main text.
 
@@
 The analytical arguments, detailed in *Appendix A*, leading to
-@eq-ineq_short, rest on the use of the delta-method [@vandervaart2000,
-@cramer1999, @casella2002]. @eq-ineq_init contains parameters of
+@eq-ineq_short, rest on the use of the delta-method [@vandervaart2000;
+@cramer1999; @casella2002]. @eq-ineq_init contains parameters of
 distributions of the standard deviations of scores, while the $R_1$ and
@@
 
-Generally, for 11 out of 12 phenomena, we find miniscule, but mostly
-negative, differences between the two estimation strategies. This means
+Generally, for 11 out of 12 phenomena, we find miniscule but mostly
+negative differences between the two estimation strategies. This means
 that there is a slight downward bias in the estimates from the
